---
title: "Samhouri et al tradeoff analysis clean data"
author: "Jameal Samhouri"
date: "10/28/2020"
output: 
  html_document:
      toc: true
      toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This document cleans up whale and fishing data for the analysis included in the Samhouri et al. tradeoff manuscript.

See pseudocode in "Full workflow for risk assessment and tradeoff analysis" google doc. https://docs.google.com/document/d/1T8S5rEOACMCeGa5efS5BnEQY4gOw-PbxmvILJvShsQo/edit

1. PROCESS WHALE MODEL OUTPUTS. [turn off if no changes to whale predictions]
2. PREP FISHING DATA
Note: user must run Grid5km_raimbow_prep.Rmd separately, after chunk 2 but before chunk 3 
3. PREP GRIDS AND JOIN TO FISHING DATA
4. PREP ENTANGLEMENT DATA

## Prep libraries and file paths

```{r prep, message=FALSE}

# just getting libraries loaded and file paths sorted

library(tidyverse)
library(foreign) # read.dbf()
library(here)
library(lubridate)
library(sf)
library(viridis)
library(ggrepel)
library(magrittr)
library(ggerr)
library(scales)
library(maps)
library(rnaturalearth)
library(gridExtra)
library(ggpubr)
library(knitr)
library(readr)

source(here::here("User_script_local.R"))
if (user == "JS") {
  #flag.save = TRUE # option to save all outputs or not
  
  path_entanglement_file <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Input_Data/Entanglement Data/2000_19/ca_confirmed_dcrb_entanglements.csv" # from Lauren Saez

} else if (user == "SMW") {
  NULL
} else {
  stop("Invalid user")
}


```

# 1) PROCESS WHALE MODEL OUTPUTS

## Humpbacks
```{r whales, echo=FALSE}

## Humpbacks
# overlay the predictions on the 5km grid
# source(paste0(here::here(), "/humpback_risk/","1a_Mn_preds_overlay5km.R"))

# aggregate the predictions by month and make them long.
# source(paste0(here::here(), "/humpback_risk/","1b_Mn_preds_aggr.R"))

```


## Blues
```{r whales2, echo=FALSE}

## Blues
# overlay the predictions on the 5km grid
# source(paste0(here::here(), "/blue_whale","BW_preds_overlay5km.Rmd"))

# aggregate the predictions by month
# source(paste0(here::here(), "/blue_whale","BW_preds_aggr.Rmd"))

```

# 2) PREP FISHING DATA

Vms matched to fish tickets
```{r fishing, echo=FALSE}

source(paste0(here::here(),"/tradeoffs/full_analysis_samhourietal/prep_data_for_scenario_df_function.R"))

# add period

```

All fish tickets for DCRB [not yet included]
```{r fishing2, echo=FALSE}


```

# 3) PREP GRIDS AND JOIN TO FISHING DATA
```{r grids, echo=FALSE}

#############################################
### Prep grids using Grid5km_raimbow_prep.Rmd
#############################################

# create keys and other data frames related to the base 5km grid used in RAIMBOW analyses. serves as a single 'crank to turn' when new fishing data or whale predictions are added to the analyses to ensure that specific data are formatted as expected for the various functions used in the analyses. Specifically it:
# + Determines the study area, aka which 5km grid cells have (ever) had fishing effort
# + Determines the associated CDFW large block number and region for each 5km grid cell in the study area
# + Extracts the depth value for each grid cell in the study area
# + Prepares whale data to be passed as an argument to risk_mgmt(), etc.

# Note Grid5km_landerase.R must be run prior to running Grid5km_raimbow_prep.Rmd if any changes have been made to the 5km grid from Blake. This part was left out of the Rmd doc b/c 1) it takes a while to run and 2) it's a one time run, i.e. the grid-cell land won't (hopefully doesn't) change.

# 110920. I could not figure out how to call the Rmd within an Rmd. So user must run Grid5km_raimbow_prep.Rmd separately, after chunk 2 but before chunk 3 

# source(paste0(here::here(),"/grid-prep/Grid5km_raimbow_prep.Rmd"), local = knitr::knit_global())
# sys.source(paste0(here::here(),"/grid-prep/Grid5km_raimbow_prep.Rmd"), 
#            envir = knitr::knit_global())
# read_chunk(paste0(here::here(),"/grid-prep/Grid5km_raimbow_prep.Rmd"))

```

```{r join_grids, echo=FALSE}
#############################################
# Join fishing data to region and depth keys
#############################################

source(paste0(here::here(),"/tradeoffs/join_fishing_data_regions_depth.R"))

# Note this Relies on Grid5km_key.rds: key(s) for grid cell ID, Region and CA_OFFShOR (CDFW large blocks), and AWM depth. These have their own dependencies, handled in Grid5km_raimbow_prep.Rmd


```

# 4) PREP ENTANGLEMENT DATA
```{r entanglements, echo=FALSE}

# all data based on "entanglement reports from lauren 100920.xlsx"
# focus on confirmed CA commercial Dungeness crab entanglements by year

# most code from Make time series of entanglements.Rmd
#Initial processing
entanglement_df <- readr::read_csv(path_entanglement_file)
glimpse(entanglement_df)

#unique(entanglement_df$Common_Nam)

entanglement_df_annual <- entanglement_df %>%
  pivot_longer(
    !Year, names_to = "common_nam", values_to = "count"
  ) %>%
#group_by(Year) %>% #, Common_Nam
  #tally() %>%
  mutate(
    Year.as.date = as.Date(paste(Year, 1, 1, sep = "-")) 
    ) %>%
  #complete(Year.as.date = seq.Date(min(Year.as.date), max(Year.as.date), by="year"),
  #         fill = list(n=0)) %>%
  #complete(Year, nesting(Common_Nam)) %>%
  mutate(
    species = case_when(common_nam == "Gray" | common_nam == "Killer" | common_nam == "Unidentified"  ~ "Other / Unidentified",
                        common_nam == "Humpback" ~ "Humpback Whale",
                        common_nam == "Blue" ~ "Blue Whale",
                        TRUE ~ "ERROR")
  )
glimpse(entanglement_df_annual)

# complete the df with zeroes for each species category

#complete_species_year <- expand.grid(species = unique(entanglement_df_annual$species), Year.as.date = seq.Date(as.Date(paste(1982, 1, 1, sep = "-")), max(entanglement_df_annual$Year.as.date), by="year"))

entanglement_df_annual_complete <- entanglement_df_annual %>%
  group_by(Year, Year.as.date, species) %>%
  summarise(
    count = sum(count), 
    .groups = 'drop'
  )
  #tally() #%>%
  #full_join(complete_species_year)
  
#entanglement_df_annual_complete$Year <- ifelse(is.na(entanglement_df_annual_complete$Year),year(entanglement_df_annual_complete$Year.as.date),entanglement_df_annual_complete$Year)

#entanglement_df_annual_complete$n <- ifelse(is.na(entanglement_df_annual_complete$n),0,entanglement_df_annual_complete$n)

glimpse(entanglement_df_annual_complete)

write_csv(entanglement_df_annual_complete, 
          paste0(here::here(), "/tradeoffs/full_analysis_samhourietal/output_dfs/entanglement_df_annual_complete.csv")
          )

```

