---
title: "Samhouri et al tradeoff analysis make figures"
author: "Jameal Samhouri"
date: "10/27/2020"
output: 
  html_document:
      toc: true
      toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This document reflects the figures included in the Samhouri et al. tradeoff manuscript.

See pseudocode in "Full workflow for risk assessment and tradeoff analysis" google doc. https://docs.google.com/document/d/1T8S5rEOACMCeGa5efS5BnEQY4gOw-PbxmvILJvShsQo/edit

* Figure 1. Maps of whales and DCRB pings in 3 time periods on 5km grid. Made outside of this doc.

* Figure 2. Changes over time (2009-2019) in (a) the number of confirmed whale entanglements along the US West Coast, (b) expected risk to blue and humpback whales, measured as spatial overlap (see Methods), from the California Dungeness crab fishery, and (c) revenue to the California Dungeness crab fishery. whale risk is calculated in the run_analysis RUN SCENARIOS step

* Figure 3. Comparison of expected (a) risk to blue whales, (b) risk to humpback whales, and (c) revenue to the Dungeness crab fishery under a range of status quo and hypothetical alternative management scenarios affecting the entire state of California, during each of three time periods representing before (2009-2014), during (2014-2018), and after (2018-2019) the Northeast Pacific Marine Heatwave. Whale risk and DCRB revenue by scenario calculated in the run_analysis RUN SCENARIOS step

* Figure 4. Tradeoff plots indicating the expected (a) risk to blue whales or (b) risk to humpback whales in comparison to expected revenue to the California Dungeness crab fishery under a range of hypothetical alternative management scenarios, during each of three time periods representing before (2009-2014), during (2014-2018), and after (2018-2019) the Northeast Pacific Marine Heatwave. Larger points represent the median ±1SE for each scenario across years, smaller points represent values for individual years. Whale risk and DCRB revenue by scenario calculated in the run_analysis RUN SCENARIOS step

SI Figures follow from there.
Entanglement ts by gear type (updated version of Fig14 in Saez et al. 2020)

## Prep libraries and file paths

```{r prep, include=FALSE}

# just getting libraries loaded and file paths sorted

library(tidyverse)
library(foreign) # read.dbf()
library(here)
library(lubridate)
library(sf)
library(viridis)
library(ggrepel)
library(magrittr)
library(ggerr)
library(scales)
library(maps)
library(rnaturalearth)
library(gridExtra)
library(ggpubr)
library(knitr)
library(rphylopic)
library(png)
library(RCurl)

source(here::here("User_script_local.R"))
if (user == "JS") {
  #flag.save = TRUE # option to save all outputs or not
  
  path_crabyear_summaries <- here::here(
            "tradeoffs",
            "full_analysis_samhourietal",
            "output_dfs",
            "crabyear_summaries.csv"
          )
    
  path_entanglement_file2 <- paste0(here::here(), "/tradeoffs/full_analysis_samhourietal/output_dfs/entanglement_df_annual_complete.csv")
  
  path_figures <- "/Users/jameal.samhouri/Dropbox/Projects/In progress/RAIMBOW/raimbow/tradeoffs/full_analysis_samhourietal/figures"
  
  path_df_tradeoff_focal_scenarios <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/df_tradeoff_focal_scenarios_2020-11-30.rds"
  
  path_annual_statewide_df_focal_scenarios <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/annual_statewide_df_focal_scenarios_2020-11-30.rds"
  
  path_Grid_5km_landerased <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/5x5 Grid/Grid_5km_landerased.RDATA"
  
  path_entanglement_file <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Input_Data/Entanglement Data/2000_19/2000_19.dbf" # from Lauren Saez
  
  path_mappingdf <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/Data frame for Figure 1 Samhouri et al.csv"
  
  path_rds <- "/Users/jameal.samhouri/Dropbox/Projects/In progress/RAIMBOW/raimbow/tradeoffs/Management scenarios"
  
  path_maps <- "/Users/jameal.samhouri/Dropbox/Projects/In progress/RAIMBOW/raimbow/tradeoffs/maps"
  
  path_status_quo <- paste0("/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/scenario_output_dataframes/risk_5km_yr_mth_sq_all_2021-01-06.rds") # status_quo_risk_2009_2019_yr_mth_2020-06-14.rds was from extract_status_quo_scenario_summary.R
  
  path_status_quo_small_vessels <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/scenario_output_dataframes/risk_5km_yr_mth_sq_small_2021-01-14.rds" # from run_analysis_sm_vessels.Rmd risk_5km_yr_mth_sq_all, could also get from extract_status_quo_scenario_summary_small.R which writes out status_quo_risk_small_vessels_2009_2019_yr_mth_2020-06-17.rds
  
  path_whales <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/5x5 Grid/Grid5km_whale.rds"
  
  path_Humpback_5km_long_monthly <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Input_Data/Humpback whale data/Forney et al./Humpback_5km_long_monthly.rds" 
  
  path_BlueWhale_5km_long_monthly <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Input_Data/Blue whale data/Overlay on 5km Grid/BlueWhale_5km_long_monthly.rds"
    
  path_effort_shift_scenariocomparisons_annual_statewide_df_n <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/annual_statewide_effort_shift_scenariocomparisons_n_2020-05-19.rds" # from effort_shift_comparison.R
  
  path_tradeoff_df_effort_shift_scenariocomparisons_n <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/tradeoff_df_effort_shift_scenariocomparisons_n_2020-05-19.rds"
  
  path_scenario_table_all <- "/Users/jameal.samhouri/Dropbox/Projects/In progress/RAIMBOW/raimbow/tradeoffs/Management scenarios/scenario_table_all_2020-06-09.rds"
  
  path_scenario_table_focal_scenarios <- "/Users/jameal.samhouri/Dropbox/Projects/In progress/RAIMBOW/raimbow/tradeoffs/Management scenarios/scenario_table_focal_scenarios_2020-06-12.rds"
  
  path_annual_statewide_df_focal_scenarios <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/annual_statewide_df_focal_scenarios_2020-06-12.rds" 
  
  path_annual_statewide_df_focal_scenarios_small_vessels <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/annual_statewide_df_focal_scenarios_small_vessels_2020-06-16.rds"
  
  path_df_tradeoff_focal_scenarios_small_vessels <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/df_tradeoff_focal_scenarios_small_vessels_2020-06-16.rds"
  
  path_prioritizr_all <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/prioritizr/prioritzr_outputs_all.rds"
  

} else if (user == "SMW") {
  NULL
} else {
  stop("Invalid user")
}


```

# Figure 1

Make some maps even though Blake will make nicer ones
```{r fig1, echo=FALSE}

# load df if not already in the env
# sq_maps_grid_noNA <- read_csv(path_mappingdf)

# from sam:
# Mn_monthly_multipanel.R has an example of using the tmap package, which I've had success with for paper-ready maps, https://github.com/karinforney/Mn3kmSDM
# https://github.com/mtennekes/tmap (with links to vignettes) 
# https://github.com/smwoodman/eSDM/blob/master/data-raw/figure_plot.R and https://github.com/smwoodman/eSDM/blob/master/data-raw/figure5.R

# plot median humpback densities 

map_hump <- ggplot() + 
  geom_sf(data=sq_maps_grid_noNA, 
          aes(fill=Humpback_dens,
              col=Humpback_dens
          )
  ) +
  facet_wrap(~time_period, nrow=1) +
  geom_sf(data=rmap.base,col=NA,fill='gray50') +
  scale_fill_viridis(na.value=NA,option="D",name="Humpback Density") + # ,breaks=seq(0,1,by=0.25),limits=c(0,1)
  scale_color_viridis(na.value=NA,option="D",name="Humpback Density") + # ,breaks=seq(0,1,by=0.25),limits=c(0,1)
  ggtitle("2009-2019 Median humpback densities\non Dungeness crab fishing grounds\nduring the fishing season") +
  coord_sf(xlim=c(sq_bbox[1],sq_bbox[3]),ylim=c(sq_bbox[2],sq_bbox[4])) + 
  theme_minimal() + #theme_classic() +
  theme(text=element_text(family="sans",size=10,color="black"),
        legend.text = element_text(size=10),
        axis.title=element_text(family="sans",size=14,color="black"),
        axis.text=element_text(family="sans",size=8,color="black"),
        panel.grid.major = element_line(color="gray50",linetype=3),
        axis.text.x.bottom = element_text(angle=45, vjust = 0.5),
        strip.text = element_text(size=14),
        title=element_text(size=16)
  )
map_hump

png(paste0(path_maps, "/map_hump_by_time_period.png"), width = 14, height = 10, units = "in", res = 300)
map_hump
invisible(dev.off())
#ggsave(here::here('tradeoffs','map_hump_by_time_period.png'),map_hump,h=8,w=6)

# plot median blue occurrence 

map_blue <- ggplot() + 
  geom_sf(data=sq_maps_grid_noNA, 
          aes(fill=Blue_occurrence,
              col=Blue_occurrence
          )
  ) +
  facet_wrap(~time_period, nrow=1) +
  geom_sf(data=rmap.base,col=NA,fill='gray50') +
  scale_fill_viridis(na.value=NA,option="D",name="Blue whale occurrence") + # ,breaks=seq(0,1,by=0.25),limits=c(0,1)
  scale_color_viridis(na.value=NA,option="D",name="Blue whale occurrence") + # ,breaks=seq(0,1,by=0.25),limits=c(0,1)
  ggtitle("2009-2019 Median blue whale occurrence\non Dungeness crab fishing grounds\nduring the fishing season") +
  coord_sf(xlim=c(sq_bbox[1],sq_bbox[3]),ylim=c(sq_bbox[2],sq_bbox[4])) + 
  theme_minimal() + #theme_classic() +
  theme(text=element_text(family="sans",size=10,color="black"),
        legend.text = element_text(size=10),
        axis.title=element_text(family="sans",size=14,color="black"),
        axis.text=element_text(family="sans",size=8,color="black"),
        panel.grid.major = element_line(color="gray50",linetype=3),
        axis.text.x.bottom = element_text(angle=45, vjust = 0.5),
        strip.text = element_text(size=14),
        title=element_text(size=16)
  )
map_blue

png(paste0(path_maps, "/map_blue_by_time_period.png"), width = 14, height = 10, units = "in", res = 300)
map_blue
invisible(dev.off())
#ggsave(here::here('tradeoffs','map_hump_by_time_period.png'),map_hump,h=8,w=6)

# plot median Num_DCRB_VMS_pings 
# Num_DCRB_VMS_pings is highly skewed
sq_maps_grid_noNA %>% 
  filter(time_period == "2018-2019") %>% 
  summarise(quantile = scales::percent(c(0.025, 0.25, 0.5, 0.75, 0.975)),
            Num_DCRB_VMS_pings = quantile(Num_DCRB_VMS_pings, c(0.025, 0.25, 0.5, 0.75, 0.975), na.rm=TRUE)
            )

# sq_maps_grid_noNA %>% 
#   filter(time_period == "2014-2018") %>%
#   #filter(Num_DCRB_VMS_pings > 0) %>%
#   select(Num_DCRB_VMS_pings) %>%
#   ggplot(aes(x=Num_DCRB_VMS_pings)) +
#   geom_histogram(fill='#33638DFF', alpha=0.8, binwidth=5)
# 
# sq_maps_grid_noNA %>% 
#   filter(time_period == "2014-2018") %>%
#   filter(Num_DCRB_VMS_pings > 0) %>%
#   select(Num_DCRB_VMS_pings) %>%
#   ggplot(aes(x=log10(Num_DCRB_VMS_pings+10^-6))) +
#   geom_histogram(fill='#33638DFF', alpha=0.8, binwidth=0.5)


map_pings <- ggplot() + 
  geom_sf(data=sq_maps_grid_noNA %>% filter(Num_DCRB_VMS_pings > 0), 
          aes(fill=log10(Num_DCRB_VMS_pings),#+10^-6
              col=log10(Num_DCRB_VMS_pings) )
          # aes(fill = Num_DCRB_VMS_pings,
          #     col = Num_DCRB_VMS_pings)
  ) +
  facet_wrap(~time_period, nrow=1) +
  geom_sf(data=rmap.base,col=NA,fill='gray50') +
  scale_fill_viridis(na.value=NA,option="D",name="log10\nVMS Pings") + # ,breaks=seq(0,1,by=0.25),limits=c(0,1)
  scale_color_viridis(na.value=NA,option="D",name="log10\nVMS Pings") + # ,breaks=seq(0,1,by=0.25),limits=c(0,1)
  ggtitle("2009-2019 Median VMS pings\nduring the Dungeness crab\nfishing season") +
  coord_sf(xlim=c(sq_bbox[1],sq_bbox[3]),ylim=c(sq_bbox[2],sq_bbox[4])) + 
  theme_minimal() + #theme_classic() +
  theme(text=element_text(family="sans",size=10,color="black"),
        legend.text = element_text(size=10),
        axis.title=element_text(family="sans",size=14,color="black"),
        axis.text=element_text(family="sans",size=8,color="black"),
        panel.grid.major = element_line(color="gray50",linetype=3),
        axis.text.x.bottom = element_text(angle=45, vjust = 0.5),
        strip.text = element_text(size=14),
        title=element_text(size=16)
  )
map_pings

png(paste0(path_maps, "/map_pings_by_time_period.png"), width = 14, height = 10, units = "in", res = 300)
map_pings
invisible(dev.off())
#ggsave(here::here('tradeoffs','map_hump_by_time_period.png'),map_hump,h=8,w=6)

### This is Figure 1 in Samhouri et al
ggarrange(map_blue,
          map_hump,
          map_pings,
          ncol=3,
          nrow=1,
          legend="top",
          labels="auto",
          vjust=8,
          hjust=0
          )
#,font.label=list(size=12)

png(paste0(path_figures, "/fig1_maps.png"), width = 14, height = 10, units = "in", res = 300)
ggarrange(map_blue,
          map_hump,
          map_pings,
          ncol=3,
          nrow=1,
          legend="top",
          labels="auto",
          vjust=8,
          hjust=0
          )
invisible(dev.off())


```

# Figure 2. 

Changes over time (2009-2019) in (a) the number of confirmed whale entanglements along the US West Coast, (b) expected risk to blue and humpback whales, measured as spatial overlap (see Methods), from the California Dungeness crab fishery, and (c) revenue to the California Dungeness crab fishery.

## Entanglements
TS of ONLY CA commercial DCRB entanglements
```{r fig2a, echo=FALSE}

entanglement_df_annual_complete <- read_csv(path_entanglement_file2)

entanglements_ts <- ggplot(entanglement_df_annual_complete %>% filter (Year > 2009), 
                           aes(x=Year, y=count, group=species, colour=species)) + # group=1 tells ggplot that there is only 1 group
  geom_point(size=2, position = position_dodge(0.4)) +
  geom_line(position = position_dodge(0.4)) + 
  geom_vline(xintercept = c(2014.5, 2018.5), linetype=2) +
  ylab("Number of\nentanglements") +
  xlab("") + # "Year
  #labs(subtitle = "Data from NMFS WRO") +
  #ggtitle("Number of Whale\nEntanglements,US West Coast") +
  scale_colour_viridis_d(begin=0.1, end=0.9) + #, option = "plasma") +
  scale_x_continuous(breaks=seq(2010, 2019 , 1),limits=c(2009.5,2019.5))+
  scale_y_continuous(breaks=seq(0, 20, 5),limits=c(0,20))+
  theme_classic() +
  #scale_fill_manual(values=c("lightskyblue", "coral")) +
  theme(legend.title = element_blank(),
        #title = element_text(size = 32),
        legend.text = element_text(size=12),
        axis.text.x = element_blank(),#element_text(hjust = 1,size = 12, angle = 90),
        axis.text.y = element_text(size = 12),
        axis.title = element_text(size = 12),
        legend.position = c(0.2, 0.8)
        )
entanglements_ts 

png(here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  "fig2a_entanglements_ts.png"), 
  width = 7.5, height = 2.5, units = "in", res = 300)
entanglements_ts
invisible(dev.off())


```

## Risk
TS of cumulative monthly risk to blue and humpback whales
```{r fig2b, echo=FALSE}

risk_out_sq <- risk_5km_yr_mth_sq_all
#risk_out_sq <- read_rds(path_status_quo)

# plot normalized blues

ts_sq_blue <- ggplot(
  data = annual_risk_out_sq %>% 
    filter(response_metric != "DCRB_rev" & response_metric != "n_risk_humpback") %>%
    filter(quantile != "2%" & quantile != "98%") %>%
    pivot_wider(
      names_from = quantile,
      values_from = response
    ), 
  aes(
    x = plotting_year, 
    y = `50%`
  )
) +
  geom_point(size=4) +
  geom_line() + 
  # geom_ribbon(aes(ymin = `25%`,
  #                 ymax = `75%`),
  #             alpha = 0.2
  #             ) +
  geom_pointrange(aes(ymin = `25%`,
                  ymax = `75%`)
              ) +
  # geom_errorbar(aes(ymin = `25%`,
  #                 ymax = `75%`)
  #             ) +
  geom_vline(xintercept = c(2014.5, 2018.5), linetype=2) +
  # facet_grid(rows = vars(response_metric),
  #            #scales = "free_y",
  #            labeller = as_labeller ( c(
  #              #`DCRB_rev` = "Dungeness crab\nfishery revenue\n($, x10^5)",
  #              `n_risk_blue` = ""#, # to\nblue whales",
  #              #`n_risk_humpback` = "" # "Risk to\nhumpback whales\n(normalized)"
  #              ) ),
  #            switch = "y"
  #            ) + 
  scale_x_continuous(breaks = seq(2010, 2019,1),
                     limits = c(2009.5,2019.5)
                     #labels = str_replace(unique(annual_risk_out_sq$crab_year),"_","-")
                     ) + 
  #ylim(0,0.08) +
  ylab("Normalized risk") + 
  xlab("") + # "Dungeness crab season"
  theme_classic() +
  theme(legend.title = element_blank(),
        #title = element_text(size = 26),
        legend.text = element_text(size = 20),
        legend.position = c(.15, .85),
        axis.text.x = element_blank(), #element_text(hjust = 1,size = 12, angle = 60),
        axis.text.y = element_text(size = 12),
        axis.title = element_text(size = 12),
        strip.text = element_text(size=12),
        strip.background = element_blank(),
        strip.placement = "left"
        )

png(here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  "fig2b_plot_ts_sq_blue.png"), 
  width = 7.5, height = 2.5, units = "in", res = 300)
ts_sq_blue
invisible(dev.off())

ts_sq_blue

# plot normalized humpbacks

ts_sq_hump <- ggplot(
  data = annual_risk_out_sq %>% 
    filter(response_metric != "DCRB_rev" & response_metric != "n_risk_blue") %>%
    filter(quantile != "2%" & quantile != "98%") %>%
    pivot_wider(
      names_from = quantile,
      values_from = response
    ), 
  aes(
    x = plotting_year, 
    y = `50%`
  )
) +
  geom_point(size=4) +
  geom_line() + 
  # geom_ribbon(aes(ymin = `25%`,
  #                 ymax = `75%`),
  #             alpha = 0.2
  #             ) +
  geom_pointrange(aes(ymin = `25%`,
                  ymax = `75%`)
              ) +
  # geom_errorbar(aes(ymin = `25%`,
  #                 ymax = `75%`)
  #             ) +
  geom_vline(xintercept = c(2014.5, 2018.5), linetype=2) +
  # facet_grid(rows = vars(response_metric),
  #            #scales = "free_y",
  #            labeller = as_labeller ( c(
  #              #`DCRB_rev` = "Dungeness crab\nfishery revenue\n($, x10^5)",
  #              #`n_risk_blue` = ""#, # to\nblue whales",
  #              `n_risk_humpback` = "" # "Risk to\nhumpback whales\n(normalized)"
  #              ) ),
  #            switch = "y"
  #            ) + 
  scale_x_continuous(breaks = seq(2010, 2019,1),
                     limits = c(2009.5,2019.5)
                     #labels = str_replace(unique(annual_risk_out_sq$crab_year),"_","-")
                     ) + 
  #ylim(0,0.08) +
  ylab("Normalized risk") + 
  xlab("") + # "Dungeness crab season"
  theme_classic() +
  theme(legend.title = element_blank(),
        #title = element_text(size = 26),
        legend.text = element_text(size = 20),
        legend.position = c(.15, .85),
        axis.text.x = element_blank(), #element_text(hjust = 1,size = 12, angle = 60),
        axis.text.y = element_text(size = 12),
        axis.title = element_text(size = 12),
        strip.text = element_text(size=12),
        strip.background = element_blank(),
        strip.placement = "left"
        )

png(here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  "fig2c_plot_ts_sq_hump.png"), 
  width = 7.5, height = 2.5, units = "in", res = 300)
ts_sq_hump
invisible(dev.off())

ts_sq_hump


# plot normalized humpbacks, normalized blues

ts_sq_whales <- ggplot(
  data = annual_risk_out_sq %>% 
    filter(response_metric != "DCRB_rev") %>%
    filter(quantile != "2%" & quantile != "98%") %>%
    pivot_wider(
      names_from = quantile,
      values_from = response
    ), 
  aes(
    x = plotting_year, 
    y = `50%`
  )
) +
  geom_point(size=4) +
  geom_line() + 
  # geom_ribbon(aes(ymin = `25%`,
  #                 ymax = `75%`),
  #             alpha = 0.2
  #             ) +
  geom_pointrange(aes(ymin = `25%`,
                  ymax = `75%`)
              ) +
  # geom_errorbar(aes(ymin = `25%`,
  #                 ymax = `75%`)
  #             ) +
  geom_vline(xintercept = c(2014.5, 2018.5), linetype=2) +
  facet_grid(rows = vars(response_metric),
             #scales = "free_y",
             labeller = as_labeller ( c(
               #`DCRB_rev` = "Dungeness crab\nfishery revenue\n($, x10^5)",
               `n_risk_blue` = "", # to\nblue whales",
               `n_risk_humpback` = "" # "Risk to\nhumpback whales\n(normalized)"
               ) ),
             switch = "y"
             ) + 
  scale_x_continuous(breaks = seq(2010, 2019,1),
                     limits = c(2009.5,2019.5)
                     #labels = str_replace(unique(annual_risk_out_sq$crab_year),"_","-")
                     ) + 
  #ylim(0,0.08) +
  ylab("Normalized risk") + 
  xlab("") + # "Dungeness crab season"
  theme_classic() +
  theme(legend.title = element_blank(),
        #title = element_text(size = 26),
        legend.text = element_text(size = 20),
        legend.position = c(.15, .85),
        axis.text.x = element_blank(), #element_text(hjust = 1,size = 12, angle = 60),
        axis.text.y = element_text(size = 12),
        axis.title = element_text(size = 12),
        strip.text = element_text(size=12),
        strip.background = element_blank(),
        strip.placement = "left"
        )

png(here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  "fig2bc_plot_ts_sq_whales.png"), 
  width = 7.5, height = 5, units = "in", res = 300)
ts_sq_whales
invisible(dev.off())

ts_sq_whales

# to add silhouettes check out https://cran.r-project.org/web/packages/rphylopic/readme/README.html and https://scrogster.wordpress.com/2014/06/02/adding-phylopic-org-silhouettes-to-r-plots/

# # humpback
# hump_url <- "http://phylopic.org/assets/images/submissions/ce70490a-79a5-47fc-afb9-834e45803ab4.512.png"
# hump_logo <-  readPNG(getURLContent(hump_url))
# 
# logoing_func<-function(logo, x, y, size){
#   dims<-dim(logo)[1:2] #number of x-y pixels for the logo (aspect ratio)
#   AR<-dims[1]/dims[2]
#   par(usr=c(0, 1, 0, 1))
#   rasterImage(logo, x-(size/2), y-(AR*size/2), x+(size/2), y+(AR*size/2), interpolate=TRUE)
# }
# 
# ts_sq_whales + logoing_func(hump_logo, x=0.10, y=0.10, size=0.15)

```

## Revenue
TS of total DCRB revenue each season
```{r fig2c, echo=FALSE}

#crabyear_summaries <- read_csv(path_crabyear_summaries)

# plot $

ts_sq_DCRBrev <- ggplot(
  data = crabyear_summaries, 
  aes(
    x = plotting_year, 
    y= sum_DCRB_rev/10^7
  )
) +
  geom_point(size=4) +
  geom_line() + 
  geom_vline(xintercept = c(2014.5, 2018.5), linetype=2) +
  ylab(expression(Fishery~revenue~("$,"~x10^7))) + 
  scale_x_continuous(breaks = seq(2010, 2019,1),
                     limits = c(2009.5,2019.5)
                     #labels = str_replace(unique(crabyear_summaries$crab_year),"_","-")
                     ) + 
  xlab("") + # "Dungeness crab season"
  theme_classic() +
  theme(legend.title = element_blank(),
        #title = element_text(size = 26),
        legend.text = element_text(size = 20),
        legend.position = c(.15, .85),
        axis.text.x = element_text(hjust = 1,size = 12, angle = 60),
        axis.text.y = element_text(size = 12),
        axis.title = element_text(size = 12),
        strip.text = element_text(size=14),
        strip.background = element_blank(),
        strip.placement = "left"
        )

png(here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  "fig2d_ts_sq_DCRBrev.png"), 
  width = 7.5, height = 2.5, units = "in", res = 300)
ts_sq_DCRBrev
invisible(dev.off())

ts_sq_DCRBrev


```

## Figure 2abcd

```{r fig2abcd, echo=FALSE}

# Figure 1 Samhouri et al.
 
ggarrange(entanglements_ts,
          ts_sq_whales,
          ts_sq_DCRBrev,
          ncol=1,
          #labels="auto",
          heights = c(1,2,1),
          hjust=0
          )

png(here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  "all_ts.png"), 
  width = 10, height = 14, units = "in", res = 300)
ggarrange(entanglements_ts,
          ts_sq_whales,
          ts_sq_DCRBrev,
          ncol=1,
          #labels="auto",
          heights = c(1,2,1),
          hjust=0
          )
invisible(dev.off())

```

# Figure 3. 

Figure 3. Comparison of expected (a) risk to blue whales, (b) risk to humpback whales, and (c) revenue to the Dungeness crab fishery under a range of status quo and hypothetical alternative management scenarios affecting the entire state of California, during each of three time periods representing before (2009-2014), during (2014-2018), and after (2018-2019) the Northeast Pacific Marine Heatwave. Whale risk and DCRB revenue by scenario calculated in the run_analysis RUN SCENARIOS step

## Risk to blue whales
```{r fig3a, echo=FALSE}

#normalized blues, statewide scenarios only, boxplot
# try using annual_df_focal_scenarios_statewide_ranks instead of statewide_annual_rank_bp

annual_df_focal_scenarios_statewide_ranks_fig <- 
  annual_df_focal_scenarios_statewide_ranks %>%
  ungroup() %>%
  mutate(
    scenario_rank = ifelse(
      pretty_scenario_names == "Status Quo", 0,
      rank_hump
      ),
    scenario_names_fig = c(
      rep("Status Quo", 10),
      rep("Delay", 10),
      rep("Spring Closure", 10),
      rep("Delay + Spring Closure", 10),
      rep("Spring Effort Reduction", 10),
      rep("Spring Depth Restriction", 10),
      rep("Spring Effort Reduction + Depth Restriction", 10)
    )
  )

plot_blue_risk_prepost_statewide_bp <- ggplot(
  data = annual_df_focal_scenarios_statewide_ranks_fig,
  aes(
    x = reorder(scenario_names_fig,-scenario_rank),
    y = n_risk_blue,
    colour = factor(pre_post),
    fill = factor(pre_post)
    )
  )  +
  geom_boxplot(position=position_dodge(width=0.5)) +#, width=0.1, color="grey", alpha=0.2) +
  geom_point(position=position_dodge(width=0.5), alpha=0.6)+
  xlab("") + # "Scenario"
  ylab("Normalized risk") +
  ylim(0,0.4) +
  scale_colour_viridis_d(option="cividis", begin=0.2, end=0.8, direction = -1) +
  scale_fill_viridis_d(option="cividis", begin=0.2, end=0.8, direction = -1, alpha=0.2) +
  scale_x_discrete(labels = wrap_format(30)) + 
  coord_flip() +
  theme_classic() +
  theme(
    legend.title = element_blank(),
    #title = element_text(size = 26),
    legend.text = element_text(size = 14),
    #legend.position = c(.1, .95),
    axis.text.x = element_text(size=12), #hjust = 1, size = 14, angle = 60
    axis.text.y = element_text(size = 12),
    axis.title = element_text(size = 14)
    #legend.position = "none"
    )

png(here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  "fig3a_plot_blue_risk_prepost_statewide_bp.png"), 
  width = 7.5, height = 2.5, units = "in", res = 300)
plot_blue_risk_prepost_statewide_bp
invisible(dev.off())

plot_blue_risk_prepost_statewide_bp

```


## Risk to blue whales geom_pointrange 
```{r fig3aa, echo=FALSE}

#normalized blues, statewide scenarios only, geom_pointrange
# try using annual_df_focal_scenarios_statewide_ranks instead of statewide_annual_rank_bp


annual_df_focal_scenarios_statewide_ranks_fig <- 
  annual_df_focal_scenarios_statewide_ranks %>%
  ungroup() %>%
  mutate(
    scenario_rank = ifelse(
      pretty_scenario_names == "Status Quo", 0,
      rank_hump
      ),
    scenario_names_fig = c(
      rep("Status Quo", 10),
      rep("Delay", 10),
      rep("Spring Closure", 10),
      rep("Delay + Spring Closure", 10),
      rep("Spring Effort Reduction", 10),
      rep("Spring Depth Restriction", 10),
      rep("Spring Effort Reduction + Depth Restriction", 10)
    )
  )

annual_df_focal_scenarios_statewide_ranks_pointrangefig <- annual_df_focal_scenarios_statewide_ranks_fig %>%
    group_by(pre_post, scenario_names_fig, scenario_number) %>%
    summarise(
      mean_n_risk_blue = mean(n_risk_blue, na.rm=TRUE),
      se_n_risk_blue = sd(n_risk_blue, na.rm=TRUE)/sqrt(n()),
      mean_n_risk_humpback = mean(n_risk_humpback, na.rm=TRUE),
      se_n_risk_humpback = sd(n_risk_humpback, na.rm=TRUE)/sqrt(n()),
      mean_DCRB_rev = mean(DCRB_rev, na.rm=TRUE),
      se_DCRB_rev = sd(DCRB_rev, na.rm=TRUE)/sqrt(n()),
      
      .groups = "drop"
    ) %>%
  arrange(
    -mean_n_risk_humpback
  ) %>%
  mutate(
    rank_hump = row_number(),
    scenario_rank = ifelse(
      scenario_names_fig == "Status Quo", 0,
      rank_hump
      )
  )

plot_blue_risk_prepost_statewide_pointrange <- ggplot(
  data = annual_df_focal_scenarios_statewide_ranks_pointrangefig,
  aes(
    x = reorder(scenario_names_fig,-scenario_rank),
    y = mean_n_risk_blue,
    colour = factor(pre_post, levels = c("2018-2019", "2014-2018", "2009-2014")),
    fill = factor(pre_post, levels = c("2018-2019", "2014-2018", "2009-2014"))
    )
  )  +
  geom_pointrange(aes(ymin = mean_n_risk_blue - se_n_risk_blue,
                  ymax = mean_n_risk_blue + se_n_risk_blue),
                  position=position_dodge(width=0.5), alpha=0.6) +
  geom_point(position=position_dodge(width=0.5), alpha=0.6) +
  # geom_line(aes(group=scenario_names_fig),
  #           position=position_dodge(width=0.5), alpha=0.6) +
  xlab("") + # "Scenario"
  ylab("Normalized risk") + #\n (mean +/- SE)") +
  ylim(0,0.4) +
  scale_colour_viridis_d(option="cividis", begin=0.2, end=0.8, direction = -1) +
  scale_fill_viridis_d(option="cividis", begin=0.2, end=0.8, direction = -1, alpha=0.2) +
  scale_x_discrete(labels = wrap_format(30)) + 
  coord_flip() +
  guides(fill = guide_legend(reverse = TRUE), colour = guide_legend(reverse = TRUE)) + 
  theme_classic() +
  theme(
    legend.title = element_blank(),
    #title = element_text(size = 26),
    legend.text = element_text(size = 14),
    #legend.position = c(.1, .95),
    axis.text.x = element_text(size=12), #hjust = 1, size = 14, angle = 60
    axis.text.y = element_text(size = 12),
    axis.title = element_text(size = 14)
    #legend.position = "none"
    )
plot_blue_risk_prepost_statewide_pointrange

png(here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  "fig3a_blue_risk_prepost_statewide_pointrange.png"), 
  width = 7.5, height = 2.5, units = "in", res = 300)
plot_blue_risk_prepost_statewide_pointrange
invisible(dev.off())

# this version does not use coord_flip. js thinks it may be better. but the rank of the scenarios is slightly off (delay shoudl be 3rd from right)

# plot_blue_risk_prepost_statewide_pointrange <- ggplot(
#   data = annual_df_focal_scenarios_statewide_ranks_pointrangefig,
#   aes(
#     x = reorder(scenario_names_fig,scenario_rank),
#     y = mean_n_risk_blue,
#     colour = factor(pre_post, levels = c("2018-2019", "2014-2018", "2009-2014")),
#     fill = factor(pre_post, levels = c("2018-2019", "2014-2018", "2009-2014"))
#     )
#   )  +
#   geom_pointrange(aes(ymin = mean_n_risk_blue - se_n_risk_blue,
#                   ymax = mean_n_risk_blue + se_n_risk_blue),
#                   position=position_dodge(width=0.5), alpha=0.6) +
#   geom_point(position=position_dodge(width=0.5), alpha=0.6) +
#   # geom_line(aes(group=scenario_names_fig),
#   #           position=position_dodge(width=0.5), alpha=0.6) +
#   xlab("") + # "Scenario"
#   ylab("Normalized risk") + #\n (mean +/- SE)") +
#   ylim(0,0.4) +
#   scale_colour_viridis_d(option="cividis", begin=0.2, end=0.8, direction = -1) +
#   scale_fill_viridis_d(option="cividis", begin=0.2, end=0.8, direction = -1, alpha=0.2) +
#   scale_x_discrete(labels = wrap_format(30)) + 
#   #coord_flip() +
#   guides(fill = guide_legend(reverse = TRUE), colour = guide_legend(reverse = TRUE)) + 
#   theme_classic() +
#   theme(
#     legend.title = element_blank(),
#     #title = element_text(size = 26),
#     legend.text = element_text(size = 14),
#     #legend.position = c(.1, .95),
#     axis.text.x = element_text(size=12, angle = 60, hjust = 1), #hjust = 1, size = 14, angle = 60
#     axis.text.y = element_text(size = 12),
#     axis.title = element_text(size = 14)
#     #legend.position = "none"
#     )
# plot_blue_risk_prepost_statewide_pointrange

```

## Risk to humpback whales
```{r fig3b, echo=FALSE}

#normalized humps, statewide scenarios only, boxplot
# try using annual_df_focal_scenarios_statewide_ranks instead of statewide_annual_rank_bp

plot_hump_risk_prepost_statewide_bp <- ggplot(
  data = annual_df_focal_scenarios_statewide_ranks_fig,
  aes(
    x = reorder(scenario_names_fig,-scenario_rank),
    y = n_risk_humpback,
    colour = factor(pre_post),
    fill = factor(pre_post)
    )
  )  +
  geom_boxplot(position=position_dodge(width=0.5)) +#, width=0.1, color="grey", alpha=0.2) +
  geom_point(position=position_dodge(width=0.5), alpha=0.6)+
  xlab("") + # "Scenario"
  ylab("Normalized risk") +
  ylim(0,0.4) +
  scale_colour_viridis_d(option="cividis", begin=0.2, end=0.8, direction = -1) +
  scale_fill_viridis_d(option="cividis", begin=0.2, end=0.8, direction = -1, alpha=0.2) +
  scale_x_discrete(labels = wrap_format(30)) + 
  coord_flip() +
  theme_classic() +
  theme(
    legend.title = element_blank(),
    #title = element_text(size = 26),
    legend.text = element_text(size = 14),
    #legend.position = c(.1, .95),
    axis.text.x = element_text(size=12), #hjust = 1, size = 14, angle = 60
    axis.text.y = element_text(size = 12),
    axis.title = element_text(size = 14)
    #legend.position = "none"
    )

png(here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  "fig3b_plot_hump_risk_prepost_statewide_bp.png"), 
  width = 7.5, height = 2.5, units = "in", res = 300)
plot_hump_risk_prepost_statewide_bp
invisible(dev.off())

plot_hump_risk_prepost_statewide_bp

```

## Risk to humpback whales geom_pointrange 
```{r fig3bb, echo=FALSE}

#normalized humps, statewide scenarios only, geom_pointrange
# try using annual_df_focal_scenarios_statewide_ranks instead of statewide_annual_rank_bp


# annual_df_focal_scenarios_statewide_ranks_fig <- 
#   annual_df_focal_scenarios_statewide_ranks %>%
#   ungroup() %>%
#   mutate(
#     scenario_rank = ifelse(
#       pretty_scenario_names == "Status Quo", 0,
#       rank_hump
#       ),
#     scenario_names_fig = c(
#       rep("Status Quo", 10),
#       rep("Delay", 10),
#       rep("Spring Closure", 10),
#       rep("Delay + Spring Closure", 10),
#       rep("Spring Effort Reduction", 10),
#       rep("Spring Depth Restriction", 10),
#       rep("Spring Effort Reduction + Depth Restriction", 10)
#     )
#   )
# 
# annual_df_focal_scenarios_statewide_ranks_pointrangefig <- annual_df_focal_scenarios_statewide_ranks_fig %>%
#     group_by(pre_post, scenario_names_fig, scenario_number) %>%
#     summarise(
#       mean_n_risk_blue = mean(n_risk_blue),
#       se_n_risk_blue = sd(n_risk_blue)/sqrt(n()),
#       mean_n_risk_humpback = mean(n_risk_humpback),
#       se_n_risk_humpback = sd(n_risk_humpback)/sqrt(n()),
#       mean_DCRB_rev = mean(DCRB_rev),
#       se_DCRB_rev = sd(DCRB_rev)/sqrt(n()),
#       
#       .groups = "drop"
#     ) %>%
#   arrange(
#     -mean_n_risk_humpback
#   ) %>%
#   mutate(
#     rank_hump = row_number(),
#     scenario_rank = ifelse(
#       scenario_names_fig == "Status Quo", 0,
#       rank_hump
#       )
#   )

plot_hump_risk_prepost_statewide_pointrange <- ggplot(
  data = annual_df_focal_scenarios_statewide_ranks_pointrangefig,
  aes(
    x = reorder(scenario_names_fig,-scenario_rank),
    y = mean_n_risk_humpback,
    colour = factor(pre_post, levels = c("2018-2019", "2014-2018", "2009-2014")),
    fill = factor(pre_post, levels = c("2018-2019", "2014-2018", "2009-2014"))
    )
  )  +
  geom_pointrange(aes(ymin = mean_n_risk_humpback - se_n_risk_humpback,
                  ymax = mean_n_risk_humpback + se_n_risk_humpback),
                  position=position_dodge(width=0.5), alpha=0.6) +
  geom_point(position=position_dodge(width=0.5), alpha=0.6) +
  # geom_line(aes(group=scenario_names_fig),
  #           position=position_dodge(width=0.5), alpha=0.6) +
  xlab("") + # "Scenario"
  ylab("Normalized risk") + #\n (mean +/- SE)") +
  ylim(0,0.4) +
  scale_colour_viridis_d(option="cividis", begin=0.2, end=0.8, direction = -1) +
  scale_fill_viridis_d(option="cividis", begin=0.2, end=0.8, direction = -1, alpha=0.2) +
  scale_x_discrete(labels = wrap_format(30)) + 
  coord_flip() +
  guides(fill = guide_legend(reverse = TRUE), colour = guide_legend(reverse = TRUE)) + 
  theme_classic() +
  theme(
    legend.title = element_blank(),
    #title = element_text(size = 26),
    legend.text = element_text(size = 14),
    #legend.position = c(.1, .95),
    axis.text.x = element_text(size=12), #hjust = 1, size = 14, angle = 60
    axis.text.y = element_text(size = 12),
    axis.title = element_text(size = 14)
    #legend.position = "none"
    )
plot_hump_risk_prepost_statewide_pointrange

png(here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  "fig3b_hump_risk_prepost_statewide_pointrange.png"), 
  width = 7.5, height = 2.5, units = "in", res = 300)
plot_hump_risk_prepost_statewide_pointrange
invisible(dev.off())

```

## DCRB revenue
```{r fig3c, echo=FALSE}

# DCRB revenue, statewide scenarios only, boxplot
# try using annual_df_focal_scenarios_statewide_ranks instead of statewide_annual_rank_bp

plot_DCRB_rev_prepost_statewide_bp <- ggplot(
  data = annual_df_focal_scenarios_statewide_ranks_fig,
  aes(
    x = reorder(scenario_names_fig,-scenario_rank),
    y = DCRB_rev/(10^7),
    colour = factor(pre_post),
    fill = factor(pre_post)
    )
  )  +
  geom_boxplot(position=position_dodge(width=0.5)) +#, width=0.1, color="grey", alpha=0.2) +
  geom_point(position=position_dodge(width=0.5), alpha=0.6)+
  xlab("") + # "Scenario"
  ylab(expression(Fishery~revenue~("$,"~x10^7))) +
  #ylim(0,0.4) +
  scale_colour_viridis_d(option="cividis", begin=0.2, end=0.8, direction = -1) +
  scale_fill_viridis_d(option="cividis", begin=0.2, end=0.8, direction = -1, alpha=0.2) +
  scale_x_discrete(labels = wrap_format(30)) + 
  coord_flip() +
  theme_classic() +
  theme(
    legend.title = element_blank(),
    #title = element_text(size = 26),
    legend.text = element_text(size = 14),
    #legend.position = c(.1, .95),
    axis.text.x = element_text(size=12), #hjust = 1, size = 14, angle = 60
    axis.text.y = element_text(size = 12),
    axis.title = element_text(size = 14)
    #legend.position = "none"
    )

png(here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  "fig3c_plot_DCRB_rev_prepost_statewide_bp.png"), 
  width = 7.5, height = 2.5, units = "in", res = 300)
plot_DCRB_rev_prepost_statewide_bp
invisible(dev.off())

plot_DCRB_rev_prepost_statewide_bp

```

## DCRB revenue geom_pointrange 
```{r fig3cc, echo=FALSE}

#DCRB_rev, statewide scenarios only, geom_pointrange
# try using annual_df_focal_scenarios_statewide_ranks instead of statewide_annual_rank_bp


# annual_df_focal_scenarios_statewide_ranks_fig <- 
#   annual_df_focal_scenarios_statewide_ranks %>%
#   ungroup() %>%
#   mutate(
#     scenario_rank = ifelse(
#       pretty_scenario_names == "Status Quo", 0,
#       rank_hump
#       ),
#     scenario_names_fig = c(
#       rep("Status Quo", 10),
#       rep("Delay", 10),
#       rep("Spring Closure", 10),
#       rep("Delay + Spring Closure", 10),
#       rep("Spring Effort Reduction", 10),
#       rep("Spring Depth Restriction", 10),
#       rep("Spring Effort Reduction + Depth Restriction", 10)
#     )
#   )
# 
# annual_df_focal_scenarios_statewide_ranks_pointrangefig <- annual_df_focal_scenarios_statewide_ranks_fig %>%
#     group_by(pre_post, scenario_names_fig, scenario_number) %>%
#     summarise(
#       mean_n_risk_blue = mean(n_risk_blue),
#       se_n_risk_blue = sd(n_risk_blue)/sqrt(n()),
#       mean_n_risk_humpback = mean(n_risk_humpback),
#       se_n_risk_humpback = sd(n_risk_humpback)/sqrt(n()),
#       mean_DCRB_rev = mean(DCRB_rev),
#       se_DCRB_rev = sd(DCRB_rev)/sqrt(n()),
#       
#       .groups = "drop"
#     ) %>%
#   arrange(
#     -mean_n_risk_humpback
#   ) %>%
#   mutate(
#     rank_hump = row_number(),
#     scenario_rank = ifelse(
#       scenario_names_fig == "Status Quo", 0,
#       rank_hump
#       )
#   )

plot_DCRB_rev_prepost_statewide_pointrange <- ggplot(
  data = annual_df_focal_scenarios_statewide_ranks_pointrangefig,
  aes(
    x = reorder(scenario_names_fig,-scenario_rank),
    y = mean_DCRB_rev/10^7,
    colour = factor(pre_post, levels = c("2018-2019", "2014-2018", "2009-2014")),
    fill = factor(pre_post, levels = c("2018-2019", "2014-2018", "2009-2014"))
    )
  )  +
  geom_pointrange(aes(ymin = (mean_DCRB_rev - se_DCRB_rev)/10^7,
                  ymax = (mean_DCRB_rev + se_DCRB_rev)/10^7),
                  position=position_dodge(width=0.5), alpha=0.6) +
  geom_point(position=position_dodge(width=0.5), alpha=0.6) +
  # geom_line(aes(group=scenario_names_fig),
  #           position=position_dodge(width=0.5), alpha=0.6) +
  xlab("") + # "Scenario"
  ylab(expression(Fishery~revenue~("$,"~x10^7))) +
  #ylim(0,0.4) +
  scale_colour_viridis_d(option="cividis", begin=0.2, end=0.8, direction = -1) +
  scale_fill_viridis_d(option="cividis", begin=0.2, end=0.8, direction = -1, alpha=0.2) +
  scale_x_discrete(labels = wrap_format(30)) + 
  coord_flip() +
  guides(fill = guide_legend(reverse = TRUE), colour = guide_legend(reverse = TRUE)) + 
  theme_classic() +
  theme(
    legend.title = element_blank(),
    #title = element_text(size = 26),
    legend.text = element_text(size = 14),
    #legend.position = c(.1, .95),
    axis.text.x = element_text(size=12), #hjust = 1, size = 14, angle = 60
    axis.text.y = element_text(size = 12),
    axis.title = element_text(size = 14)
    #legend.position = "none"
    )
plot_DCRB_rev_prepost_statewide_pointrange

png(here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  "fig3c_plot_DCRB_rev_prepost_statewide_pointrange.png"), 
  width = 7.5, height = 2.5, units = "in", res = 300)
plot_DCRB_rev_prepost_statewide_pointrange
invisible(dev.off())

```

# Figure 4

Figure 4. Tradeoff plots indicating the expected (a) risk to blue whales or (b) risk to humpback whales in comparison to expected revenue to the California Dungeness crab fishery under a range of hypothetical alternative management scenarios, during each of three time periods representing before (2009-2014), during (2014-2018), and after (2018-2019) the Northeast Pacific Marine Heatwave. Larger points represent the median ±1SE for each scenario across years, smaller points represent values for individual years. Whale risk and DCRB revenue by scenario calculated in the run_analysis RUN SCENARIOS step

## Blue whales only
```{r tradeoff1, echo = FALSE}

### normalized blues, by pre_post
# still getting warnings aroudn geom_errorbarh. that is because the 2018-19 period has no var. ok to proceed

df_tradeoff_focal_scenarios <- read_rds(path_df_tradeoff_focal_scenarios)

df_tradeoff_focal_scenarios %>%
  group_by(pretty_scenario_names, scenario_number, pre_post) %>%
  summarise(
    mean_relative_blwh_risk_n = mean(relative_blwh_risk_n),
    sd_relative_blwh_risk_n = sd(relative_blwh_risk_n),
    mean_relative_hump_risk_n = mean(relative_hump_risk_n),
    sd_relative_hump_risk_n = sd(relative_hump_risk_n),
    mean_relative_dollars = mean(relative_dollars),
    sd_relative_dollars = sd(relative_dollars)
  )

# 113020 need to update other to figure to reflect changes here

to_blue_n_pre_post <- ggplot(
  df_tradeoff_focal_scenarios, #%>% filter(scenario_number != "15" & scenario_number != "14" & scenario_number != "9"), #df
  aes(
    x=relative_dollars,
    y=relative_blwh_risk_n,
    group=pretty_scenario_names,
    fill=scenario_type,
    label = scenario_number,
    colour=scenario_type,
    shape = spatial_domain
  )
) + 
  geom_point(size=1, alpha=0.2) + #,position=position_dodge(width = 0.9)
  #geom_text(aes(label=scenario_number),size=1, alpha=0.5, check_overlap = TRUE) + #aes(label=plotting_year, colour=scenario_type), #str_replace(crab_year,"_","-"), _repel
  stat_err(spread = "se", mult = 1, width=.1, alpha = 0.5) + 
  stat_err(geom="point", size=3, alpha = 0.8) +
  #stat_err(geom="label", size=3, alpha = 0.8) + #
  stat_err(spread = "se", mult = 1, geom = "errorbarh", height = .1,alpha = 0.5) +
  #facet_grid(rows = vars(pre_post) ) + 
  facet_wrap(nrow = 3,  vars(pre_post) ) + 
  scale_colour_viridis_d(option="inferno", begin=0.2, end=0.8, direction = -1, alpha=0.8,
                         labels = function(x) str_wrap(x, width = 20)) +
  scale_fill_viridis_d(option="inferno", begin=0.2, end=0.8, direction = -1, alpha=0.8,
                       labels = function(x) str_wrap(x, width = 20)) +
  scale_shape_manual(values=c(21:25, 4, 8) ) +
  ylab("Relative reduction in risk") + #  to blues\n(normalized)
  xlab("Relative fishery revenue") + #  to the Dungeness crab fishery
  scale_y_continuous(breaks=seq(-25, 100, 25),limits=c(-30,100)) +
  guides(fill = guide_legend("Scenario type", order = 1), 
         colour = guide_legend("Scenario type", order = 1),
         shape = guide_legend("Spatial domain", order = 2)
         ) + 
  theme_classic(base_size = 12) +
  theme(
    #title = element_text(size = 26),
    legend.position = "right",
    legend.box="vertical", 
    #legend.margin=margin(),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    #axis.title = element_text(size = 14),
    legend.text = element_text(size=10),
    #strip.text = element_text(size=14),
    ) 
to_blue_n_pre_post

# need to figure out how to shrink to fit margins with font sizes

png(
  here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  paste0("Tradeoff plot - normalized blues only by pre_post_",today(),".png")), 
    width = 3.75, height = 6, units = "in", res = 300)
to_blue_n_pre_post
invisible(dev.off())

to_blue_n_pre_post_nolegend <- ggplot(
  df_tradeoff_focal_scenarios, #%>% filter(scenario_number != "15" & scenario_number != "14" & scenario_number != "9"), #df
  aes(
    x=relative_dollars,
    y=relative_blwh_risk_n,
    group=pretty_scenario_names,
    fill=scenario_type,
    #label = scenario_number,
    colour=scenario_type,
    shape = spatial_domain
  )
) + 
  geom_point(size=1, alpha=0.2) + #,position=position_dodge(width = 0.9)
  #geom_text(aes(label=plotting_year, colour=scenario_type), size=2, alpha=0.5, check_overlap = TRUE) + #str_replace(crab_year,"_","-"), _repel
  stat_err(spread = "se", mult = 1, width=.1, alpha = 0.5) + 
  stat_err(geom="point", size=3, alpha = 0.8) + #
  stat_err(spread = "se", mult = 1, geom = "errorbarh", height = .1,alpha = 0.5) +
  #facet_grid(rows = vars(pre_post) ) + 
  facet_wrap(nrow = 3,  vars(pre_post) ) + 
  scale_colour_viridis_d(option="inferno", begin=0.2, end=0.8, direction = -1, alpha=0.8,
                         labels = function(x) str_wrap(x, width = 10)) +
  scale_fill_viridis_d(option="inferno", begin=0.2, end=0.8, direction = -1, alpha=0.8,
                       labels = function(x) str_wrap(x, width = 10)) +
  scale_shape_manual(values=c(21:25, 4, 8) ) +
  ylab("Relative reduction in risk") + #  to blues\n(normalized)
  xlab("Relative fishery revenue") + #  to the Dungeness crab fishery
  scale_y_continuous(breaks=seq(-25, 100, 25),limits=c(-30,100)) +
  guides(fill = guide_legend("Scenario type", order = 1), 
         colour = guide_legend("Scenario type", order = 1),
         shape = guide_legend("Spatial domain", order = 2)
         ) + 
  theme_classic(base_size = 12) +
  theme(
    #title = element_text(size = 26),
    legend.position = "none",
    legend.box="vertical", 
    #legend.margin=margin(),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    #axis.title = element_text(size = 14),
    legend.text = element_text(size=10),
    #strip.text = element_text(size=14),
    ) 
to_blue_n_pre_post_nolegend

png(
  here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  paste0("Tradeoff plot - normalized blues only by pre_post_nolegend_",today(),".png")), 
    width = 3.75, height = 6, units = "in", res = 300)
to_blue_n_pre_post_nolegend
invisible(dev.off())


```


## Humpback whales only
```{r tradeoff2, echo = FALSE}

### normalized humps, by pre_post
# still getting warnings aroudn geom_errorbarh. that is because the 2018-19 period has no var. ok to proceed

to_hump_n_pre_post <- ggplot(
  df_tradeoff_focal_scenarios, #%>% filter(scenario_number != "15" & scenario_number != "14" & scenario_number != "9"), #df
  aes(
    x=relative_dollars,
    y=relative_hump_risk_n,
    group=pretty_scenario_names,
    fill=scenario_type,
    #label = scenario_number,
    colour=scenario_type,
    shape = spatial_domain
  )
) + 
  geom_point(size=1, alpha=0.2) + #,position=position_dodge(width = 0.9)
  #geom_text(aes(label=plotting_year, colour=scenario_type), size=2, alpha=0.5, check_overlap = TRUE) + #str_replace(crab_year,"_","-"), _repel
  stat_err(spread = "se", mult = 1, width=.1, alpha = 0.5) + 
  stat_err(geom="point", size=3, alpha = 0.8) + #
  stat_err(spread = "se", mult = 1, geom = "errorbarh", height = .1,alpha = 0.5) +
  #facet_grid(rows = vars(pre_post) ) + 
  facet_wrap(nrow = 3,  vars(pre_post) ) + 
  scale_colour_viridis_d(option="inferno", begin=0.2, end=0.8, direction = -1, alpha=0.8,
                         labels = function(x) str_wrap(x, width = 10)) +
  scale_fill_viridis_d(option="inferno", begin=0.2, end=0.8, direction = -1, alpha=0.8,
                       labels = function(x) str_wrap(x, width = 10)) +
  scale_shape_manual(values=c(21:25, 4, 8) ) +
  ylab("Relative reduction in risk") + #  to humps\n(normalized)
  xlab("Relative fishery revenue") + #  to the Dungeness crab fishery
  scale_y_continuous(breaks=seq(-25, 100, 25),limits=c(-30,100)) +
  guides(fill = guide_legend("Scenario type", order = 1), 
         colour = guide_legend("Scenario type", order = 1),
         shape = guide_legend("Spatial domain", order = 2)
         ) + 
  theme_classic(base_size = 12) +
  theme(
    #title = element_text(size = 26),
    legend.position = "right",
    legend.box="vertical", 
    #legend.margin=margin(),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    #axis.title = element_text(size = 14),
    legend.text = element_text(size=10),
    #strip.text = element_text(size=14),
    ) 
to_hump_n_pre_post

# need to figure out how to shrink to fit margins with font sizes

png(
  here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  paste0("Tradeoff plot - normalized humps only by pre_post_",today(),".png")), 
    width = 3.75, height = 6, units = "in", res = 300)
to_hump_n_pre_post
invisible(dev.off())

to_hump_n_pre_post_nolegend <- ggplot(
  df_tradeoff_focal_scenarios, #%>% filter(scenario_number != "15" & scenario_number != "14" & scenario_number != "9"), #df
  aes(
    x=relative_dollars,
    y=relative_hump_risk_n,
    group=pretty_scenario_names,
    fill=scenario_type,
    #label = scenario_number,
    colour=scenario_type,
    shape = spatial_domain
  )
) + 
  geom_point(size=1, alpha=0.2) + #,position=position_dodge(width = 0.9)
  #geom_text(aes(label=plotting_year, colour=scenario_type), size=2, alpha=0.5, check_overlap = TRUE) + #str_replace(crab_year,"_","-"), _repel
  stat_err(spread = "se", mult = 1, width=.1, alpha = 0.5) + 
  stat_err(geom="point", size=3, alpha = 0.8) + #
  stat_err(spread = "se", mult = 1, geom = "errorbarh", height = .1,alpha = 0.5) +
  #facet_grid(rows = vars(pre_post) ) + 
  facet_wrap(nrow = 3,  vars(pre_post) ) + 
  scale_colour_viridis_d(option="inferno", begin=0.2, end=0.8, direction = -1, alpha=0.8,
                         labels = function(x) str_wrap(x, width = 10)) +
  scale_fill_viridis_d(option="inferno", begin=0.2, end=0.8, direction = -1, alpha=0.8,
                       labels = function(x) str_wrap(x, width = 10)) +
  scale_shape_manual(values=c(21:25, 4, 8) ) +
  ylab("Relative reduction in risk") + #  to humps\n(normalized)
  xlab("Relative fishery revenue") + #  to the Dungeness crab fishery
  scale_y_continuous(breaks=seq(-25, 100, 25),limits=c(-30,100)) +
  guides(fill = guide_legend("Scenario type", order = 1), 
         colour = guide_legend("Scenario type", order = 1),
         shape = guide_legend("Spatial domain", order = 2)
         ) + 
  theme_classic(base_size = 12) +
  theme(
    #title = element_text(size = 26),
    legend.position = "none",
    legend.box="vertical", 
    #legend.margin=margin(),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    #axis.title = element_text(size = 14),
    legend.text = element_text(size=10),
    #strip.text = element_text(size=14),
    ) 
to_hump_n_pre_post_nolegend

png(
  here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  paste0("Tradeoff plot - normalized humps only by pre_post_nolegend_",today(),".png")), 
    width = 3.75, height = 6, units = "in", res = 300)
to_hump_n_pre_post_nolegend
invisible(dev.off())


```


# SUPPLEMENTARY INFORMATION

## Compare effects across spatial scales

Blues
```{r scales1, echo = FALSE}

plot_blwh_risk_scale_compare <- ggplot(
  data = scale_compare_focal_scenarios,
  aes(
    x = scale,
    y = mean_relative_blwh_risk_n,
    colour = scenario_type,
    fill = scenario_type,
    shape = scenario_type
    )
  )  +
  geom_point(alpha = 0.8, position=position_dodge(width=0.5)) +
  geom_errorbar(aes(
    ymin = mean_relative_blwh_risk_n - (2*se_relative_blwh_risk_n), ymax = mean_relative_blwh_risk_n + 2*(se_relative_blwh_risk_n)
  ),
  position=position_dodge(width=0.5), size=0.5) + 
  geom_path(aes(colour=scenario_type, group= scenario_type), position=position_dodge(width=0.5)) +
  facet_wrap(pre_post~.) +
  scale_colour_viridis_d(option="inferno", begin=0.2, end=0.8, direction = -1) +
  scale_fill_viridis_d(option="inferno", begin=0.2, end=0.8, direction = -1) +
  scale_shape_manual(values=c(21:25, 4) ) +
  scale_x_discrete(labels = wrap_format(30)) + 
  guides(colour = guide_legend("Scenario"), fill = guide_legend("Scenario"), shape = guide_legend("Scenario")) +
  xlab("Spatial Scale") +
  ylab("Relative risk reduction\n(mean +/- 2SE)") +
  theme_classic(base_size = 12) +
  theme(
    legend.title = element_blank(),
    #title = element_text(size = 26),
    #legend.text = element_text(size = 18),
    #legend.position = c(.1, .95), #"none"
    axis.text.x = element_text(size = 12, hjust = 1, angle=60), #hjust = 1, size = 14, angle = 60
    axis.text.y = element_text(size = 12),
    #axis.title = element_text(size = 14),
    legend.text = element_text(size=10),
    #axis.text.y = element_text(size = 12),
    #axis.title = element_text(size = 14),
    #strip.text = element_text(size=12)
    #legend.position = "none"
    )
plot_blwh_risk_scale_compare
 
png(
  here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  "plot_blwh_risk_scale_compare.png"), 
  width = 7.5, height = 10, units = "in", res = 300)
plot_blwh_risk_scale_compare
invisible(dev.off())

```

Humps
```{r scales2, echo = FALSE}

plot_hump_risk_scale_compare <- ggplot(
  data = scale_compare_focal_scenarios,
  aes(
    x = scale,
    y = mean_relative_hump_risk_n,
    colour = scenario_type,
    fill = scenario_type,
    shape = scenario_type
    )
  )  +
  geom_point(alpha = 0.8, position=position_dodge(width=0.5)) +
  geom_errorbar(aes(
    ymin = mean_relative_hump_risk_n - (2*se_relative_hump_risk_n), ymax = mean_relative_hump_risk_n + 2*(se_relative_hump_risk_n)
  ),
  position=position_dodge(width=0.5), size=0.5) + 
  geom_path(aes(colour=scenario_type, group= scenario_type), position=position_dodge(width=0.5)) +
  facet_wrap(pre_post~.) +
  scale_colour_viridis_d(option="inferno", begin=0.2, end=0.8, direction = -1) +
  scale_fill_viridis_d(option="inferno", begin=0.2, end=0.8, direction = -1) +
  scale_shape_manual(values=c(21:25, 4) ) +
  scale_x_discrete(labels = wrap_format(30)) + 
  guides(colour = guide_legend("Scenario"), fill = guide_legend("Scenario"), shape = guide_legend("Scenario")) +
  xlab("Spatial Scale") +
  ylab("Relative risk reduction\n(mean +/- 2SE)") +
  theme_classic(base_size = 12) +
  theme(
    legend.title = element_blank(),
    #title = element_text(size = 26),
    #legend.text = element_text(size = 18),
    #legend.position = c(.1, .95), #"none"
    axis.text.x = element_text(size = 12, hjust = 1, angle=60), #hjust = 1, size = 14, angle = 60
    axis.text.y = element_text(size = 12),
    #axis.title = element_text(size = 14),
    legend.text = element_text(size=10),
    #axis.text.y = element_text(size = 12),
    #axis.title = element_text(size = 14),
    #strip.text = element_text(size=12)
    #legend.position = "none"
    )
plot_hump_risk_scale_compare
 
png(
  here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  "plot_hump_risk_scale_compare.png"), 
  width = 7.5, height = 10, units = "in", res = 300)
plot_hump_risk_scale_compare
invisible(dev.off())


```

DCRB
```{r scales3, echo = FALSE}

plot_DCRB_rev_scale_compare <- ggplot(
  data = scale_compare_focal_scenarios,
  aes(
    x = scale,
    y = mean_relative_dollars,
    colour = scenario_type,
    fill = scenario_type,
    shape = scenario_type
    )
  )  +
  geom_point(alpha = 0.8, position=position_dodge(width=0.5)) +
  geom_errorbar(aes(
    ymin = mean_relative_dollars - (2*se_relative_dollars), ymax = mean_relative_dollars + 2*(se_relative_dollars)
  ),
  position=position_dodge(width=0.5), size=0.5) + 
  geom_path(aes(colour=scenario_type, group= scenario_type), position=position_dodge(width=0.5)) +
  facet_wrap(pre_post~.) +
  scale_colour_viridis_d(option="inferno", begin=0.2, end=0.8, direction = -1) +
  scale_fill_viridis_d(option="inferno", begin=0.2, end=0.8, direction = -1) +
  scale_shape_manual(values=c(21:25, 4) ) +
  scale_x_discrete(labels = wrap_format(30)) + 
  guides(colour = guide_legend("Scenario"), fill = guide_legend("Scenario"), shape = guide_legend("Scenario")) +
  xlab("Spatial Scale") +
  ylab("Relative fishery revenue\n(mean +/- 2SE)") +
  theme_classic(base_size = 12) +
  theme(
    legend.title = element_blank(),
    #title = element_text(size = 26),
    #legend.text = element_text(size = 18),
    #legend.position = c(.1, .95), #"none"
    axis.text.x = element_text(size = 12, hjust = 1, angle=60), #hjust = 1, size = 14, angle = 60
    axis.text.y = element_text(size = 12),
    #axis.title = element_text(size = 14),
    legend.text = element_text(size=10),
    #axis.text.y = element_text(size = 12),
    #axis.title = element_text(size = 14),
    #strip.text = element_text(size=12)
    #legend.position = "none"
    )
plot_DCRB_rev_scale_compare
 
png(
  here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  "plot_DCRB_rev_scale_compare.png"), 
  width = 7.5, height = 10, units = "in", res = 300)
plot_DCRB_rev_scale_compare
invisible(dev.off())


```

Altogether
```{r scales4, echo = FALSE}

png(
  here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  "p_scale_compare.png"), 
  width = 7.5, height = 10, units = "in", res = 300)
ggarrange(plot_blwh_risk_scale_compare,
             plot_hump_risk_scale_compare, 
          plot_DCRB_rev_scale_compare,
             nrow = 2,
              ncol=2,
             common.legend = TRUE, 
             legend="bottom",
          labels="auto",
          hjust=0
             )
invisible(dev.off())

ggarrange(plot_blwh_risk_scale_compare,
             plot_hump_risk_scale_compare, 
          plot_DCRB_rev_scale_compare,
             nrow = 2,
              ncol=2,
             common.legend = TRUE, 
             legend="bottom",
          labels="auto",
          hjust=0
             )


```


## Cost effectiveness

### CE by period and scenario

Blues
```{r ce1, echo = FALSE}

# ce for blues, 
# did not remove scenarios where whale risk is higher than SQ on avg

ce_blue_pre_post <- ggplot(
  annual_df_ranks_period_focal_hump %>% 
    mutate(
      pretty_scenario_names_label = case_when(
        pre_post == "2009-2014" ~ pretty_scenario_names,
        TRUE ~ ""
      )
    ), # %>% filter(mean_relative_blwh_risk_n < 0 | mean_relative_hump_risk_n < 0),
  aes(
    x=pre_post,
    y=mean_cbblwh,
    group=pretty_scenario_names,
    fill=pretty_scenario_names,
    colour=pretty_scenario_names,
    label=pretty_scenario_names
  )
) + 
  geom_text_repel(size=1, alpha=0.8) + 
  geom_line() + 
  scale_colour_viridis_d(option="inferno", begin=0.2, end=0.8, direction = -1, alpha=0.8,
                         labels = function(x) str_wrap(x, width = 10)) +
  scale_fill_viridis_d(option="inferno", begin=0.2, end=0.8, direction = -1, alpha=0.8,
                       labels = function(x) str_wrap(x, width = 10)) +
  #scale_shape_manual(values=c(21:25, 4, 8) ) +
  ylab("Cost effectiveness") + #  to humps\n(normalized)
  xlab("Period") + #  to the Dungeness crab fishery
  #scale_y_continuous(breaks=seq(-25, 100, 25),limits=c(-30,100)) +
  guides(fill = guide_legend("Scenario type", order = 1), 
         colour = guide_legend("Scenario type", order = 1)
         ) + 
  theme_classic(base_size = 12) +
  theme(
    #title = element_text(size = 26),
    legend.position = "none",
    legend.box="vertical", 
    #legend.margin=margin(),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    #axis.title = element_text(size = 14),
    legend.text = element_text(size=10),
    #strip.text = element_text(size=14),
    ) 
ce_blue_pre_post

png(
  here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  paste0("Cost effectiveness plot - blues by pre_post_nolegend_",today(),".png")), 
    width = 3.75, height = 5, units = "in", res = 300)
ce_blue_pre_post
invisible(dev.off())

```

Humpbacks
```{r ce2, echo = FALSE}

# ce for humps
# did not remove scenarios where whale risk is higher than SQ on avg

ce_hump_pre_post <- ggplot(
  annual_df_ranks_period_focal_hump %>% 
    mutate(
      pretty_scenario_names_label = case_when(
        pre_post == "2009-2014" ~ pretty_scenario_names,
        TRUE ~ ""
      )
    ), # %>% filter(mean_relative_blwh_risk_n < 0 | mean_relative_hump_risk_n < 0),
  aes(
    x=pre_post,
    y=mean_cbhump,
    group=pretty_scenario_names,
    fill=pretty_scenario_names,
    colour=pretty_scenario_names,
    label=pretty_scenario_names
  )
) + 
  geom_text_repel(size=1, alpha=0.8) + 
  geom_line() + 
  scale_colour_viridis_d(option="inferno", begin=0.2, end=0.8, direction = -1, alpha=0.8,
                         labels = function(x) str_wrap(x, width = 10)) +
  scale_fill_viridis_d(option="inferno", begin=0.2, end=0.8, direction = -1, alpha=0.8,
                       labels = function(x) str_wrap(x, width = 10)) +
  #scale_shape_manual(values=c(21:25, 4, 8) ) +
  ylab("Cost effectiveness") + #  to humps\n(normalized)
  xlab("Period") + #  to the Dungeness crab fishery
  #scale_y_continuous(breaks=seq(-25, 100, 25),limits=c(-30,100)) +
  guides(fill = guide_legend("Scenario type", order = 1), 
         colour = guide_legend("Scenario type", order = 1)
         ) + 
  theme_classic(base_size = 12) +
  theme(
    #title = element_text(size = 26),
    legend.position = "none",
    legend.box="vertical", 
    #legend.margin=margin(),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    #axis.title = element_text(size = 14),
    legend.text = element_text(size=10),
    #strip.text = element_text(size=14),
    ) 
ce_hump_pre_post

png(
  here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  paste0("Cost effectiveness plot - humps by pre_post_nolegend_",today(),".png")), 
    width = 3.75, height = 5, units = "in", res = 300)
ce_hump_pre_post
invisible(dev.off())

```

## Rank changes

### Rank changes by period and scenario

Blues
```{r ce1, echo = FALSE}

# rel risk for blues, 
# did not remove scenarios where whale risk is higher than SQ on avg

risk_blue_pre_post <- ggplot(
  annual_df_ranks_period_focal_hump %>% 
    mutate(
      pretty_scenario_names_label = case_when(
        pre_post == "2009-2014" ~ pretty_scenario_names,
        TRUE ~ ""
      )
    ), # %>% filter(mean_relative_blwh_risk_n < 0 | mean_relative_hump_risk_n < 0),
  aes(
    x=pre_post,
    y=mean_relative_blwh_risk_n,
    group=pretty_scenario_names,
    fill=pretty_scenario_names,
    colour=pretty_scenario_names,
    label=pretty_scenario_names
  )
) + 
  geom_text_repel(size=1, alpha=0.8) + 
  geom_line() + 
  scale_colour_viridis_d(option="inferno", begin=0.2, end=0.8, direction = -1, alpha=0.8,
                         labels = function(x) str_wrap(x, width = 10)) +
  scale_fill_viridis_d(option="inferno", begin=0.2, end=0.8, direction = -1, alpha=0.8,
                       labels = function(x) str_wrap(x, width = 10)) +
  #scale_shape_manual(values=c(21:25, 4, 8) ) +
  ylab("Relative risk reduction") + #  to humps\n(normalized)
  xlab("Period") + #  to the Dungeness crab fishery
  #scale_y_continuous(breaks=seq(-25, 100, 25),limits=c(-30,100)) +
  guides(fill = guide_legend("Scenario type", order = 1), 
         colour = guide_legend("Scenario type", order = 1)
         ) + 
  theme_classic(base_size = 12) +
  theme(
    #title = element_text(size = 26),
    legend.position = "none",
    legend.box="vertical", 
    #legend.margin=margin(),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    #axis.title = element_text(size = 14),
    legend.text = element_text(size=10),
    #strip.text = element_text(size=14),
    ) 
risk_blue_pre_post

png(
  here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  "risk_blue_pre_post.png"), 
    width = 3.75, height = 5, units = "in", res = 300)
risk_blue_pre_post
invisible(dev.off())

rank_blue_pre_post <- ggplot(
  annual_df_ranks_period_focal_hump %>% 
    mutate(
      pretty_scenario_names_label = case_when(
        pre_post == "2009-2014" ~ pretty_scenario_names,
        TRUE ~ ""
      )
    ), # %>% filter(mean_relative_blwh_risk_n < 0 | mean_relative_hump_risk_n < 0),
  aes(
    x=pre_post,
    y=rank_blwh,
    group=pretty_scenario_names,
    fill=pretty_scenario_names,
    colour=pretty_scenario_names,
    label=pretty_scenario_names
  )
) + 
  geom_text_repel(size=1, alpha=0.8) + 
  geom_line() + 
  scale_colour_viridis_d(option="inferno", begin=0.2, end=0.8, direction = -1, alpha=0.8,
                         labels = function(x) str_wrap(x, width = 10)) +
  scale_fill_viridis_d(option="inferno", begin=0.2, end=0.8, direction = -1, alpha=0.8,
                       labels = function(x) str_wrap(x, width = 10)) +
  #scale_shape_manual(values=c(21:25, 4, 8) ) +
  ylab("Rank risk reduction") + #  to humps\n(normalized)
  xlab("Period") + #  to the Dungeness crab fishery
  #scale_y_continuous(breaks=seq(-25, 100, 25),limits=c(-30,100)) +
  guides(fill = guide_legend("Scenario type", order = 1), 
         colour = guide_legend("Scenario type", order = 1)
         ) + 
  theme_classic(base_size = 12) +
  theme(
    #title = element_text(size = 26),
    legend.position = "none",
    legend.box="vertical", 
    #legend.margin=margin(),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    #axis.title = element_text(size = 14),
    legend.text = element_text(size=10),
    #strip.text = element_text(size=14),
    ) 
rank_blue_pre_post

png(
  here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  "rank_blue_pre_post.png"), 
    width = 3.75, height = 5, units = "in", res = 300)
rank_blue_pre_post
invisible(dev.off())

```


Humpbacks
```{r ce1, echo = FALSE}

# rel risk for humps, 
# did not remove scenarios where whale risk is higher than SQ on avg

risk_hump_pre_post <- ggplot(
  annual_df_ranks_period_focal_hump %>% 
    mutate(
      pretty_scenario_names_label = case_when(
        pre_post == "2009-2014" ~ pretty_scenario_names,
        TRUE ~ ""
      )
    ), # %>% filter(mean_relative_blwh_risk_n < 0 | mean_relative_hump_risk_n < 0),
  aes(
    x=pre_post,
    y=mean_relative_hump_risk_n,
    group=pretty_scenario_names,
    fill=pretty_scenario_names,
    colour=pretty_scenario_names,
    label=pretty_scenario_names
  )
) + 
  geom_text_repel(size=1, alpha=0.8) + 
  geom_line() + 
  scale_colour_viridis_d(option="inferno", begin=0.2, end=0.8, direction = -1, alpha=0.8,
                         labels = function(x) str_wrap(x, width = 10)) +
  scale_fill_viridis_d(option="inferno", begin=0.2, end=0.8, direction = -1, alpha=0.8,
                       labels = function(x) str_wrap(x, width = 10)) +
  #scale_shape_manual(values=c(21:25, 4, 8) ) +
  ylab("Relative risk reduction") + #  to humps\n(normalized)
  xlab("Period") + #  to the Dungeness crab fishery
  #scale_y_continuous(breaks=seq(-25, 100, 25),limits=c(-30,100)) +
  guides(fill = guide_legend("Scenario type", order = 1), 
         colour = guide_legend("Scenario type", order = 1)
         ) + 
  theme_classic(base_size = 12) +
  theme(
    #title = element_text(size = 26),
    legend.position = "none",
    legend.box="vertical", 
    #legend.margin=margin(),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    #axis.title = element_text(size = 14),
    legend.text = element_text(size=10),
    #strip.text = element_text(size=14),
    ) 
risk_hump_pre_post

png(
  here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  "risk_hump_pre_post.png"), 
    width = 3.75, height = 5, units = "in", res = 300)
risk_hump_pre_post
invisible(dev.off())

rank_hump_pre_post <- ggplot(
  annual_df_ranks_period_focal_hump %>% 
    mutate(
      pretty_scenario_names_label = case_when(
        pre_post == "2009-2014" ~ pretty_scenario_names,
        TRUE ~ ""
      )
    ), # %>% filter(mean_relative_hump_risk_n < 0 | mean_relative_hump_risk_n < 0),
  aes(
    x=pre_post,
    y=rank_hump,
    group=pretty_scenario_names,
    fill=pretty_scenario_names,
    colour=pretty_scenario_names,
    label=pretty_scenario_names
  )
) + 
  geom_text_repel(size=1, alpha=0.8) + 
  geom_line() + 
  scale_colour_viridis_d(option="inferno", begin=0.2, end=0.8, direction = -1, alpha=0.8,
                         labels = function(x) str_wrap(x, width = 10)) +
  scale_fill_viridis_d(option="inferno", begin=0.2, end=0.8, direction = -1, alpha=0.8,
                       labels = function(x) str_wrap(x, width = 10)) +
  #scale_shape_manual(values=c(21:25, 4, 8) ) +
  ylab("Rank risk reduction") + #  to humps\n(normalized)
  xlab("Period") + #  to the Dungeness crab fishery
  #scale_y_continuous(breaks=seq(-25, 100, 25),limits=c(-30,100)) +
  guides(fill = guide_legend("Scenario type", order = 1), 
         colour = guide_legend("Scenario type", order = 1)
         ) + 
  theme_classic(base_size = 12) +
  theme(
    #title = element_text(size = 26),
    legend.position = "none",
    legend.box="vertical", 
    #legend.margin=margin(),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    #axis.title = element_text(size = 14),
    legend.text = element_text(size=10),
    #strip.text = element_text(size=14),
    ) 
rank_hump_pre_post

png(
  here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  "rank_hump_pre_post.png"), 
    width = 3.75, height = 5, units = "in", res = 300)
rank_hump_pre_post
invisible(dev.off())


```


## Risk due to small vessels
TS of cumulative monthly risk to blue and humpback whales
```{r fig2b_sm, echo=FALSE}

#risk_out_sq_sm <- risk_5km_yr_mth_sq_all
risk_out_sq_sm <- read_rds(path_status_quo_small_vessels)

# Make data frame for Figure 2 time series Samhouri et al, sm vessels
# summarise cumulative sum of $ and risk by crab_year
annual_risk_out_sq_sm <- risk_out_sq_sm %>%
  mutate(
  #month = as.numeric(substr(year_month,6,7)),
  crab_year = func_crab_year(year_month),
  plotting_year = as.numeric(substr(crab_year,6,9))
  ) %>%
  group_by(crab_year, plotting_year, year_month) %>%
  summarise(
    sum_DCRB_rev = sum(DCRB_rev, na.rm=TRUE)/10^5,
    sum_n_risk_humpback = sum(n_risk_humpback, na.rm=TRUE),
    sum_n_risk_blue = sum(n_risk_blue, na.rm=TRUE),
    
    .groups = "drop"
    ) %>%
  group_by(crab_year, plotting_year) %>%
  summarise(
    quantile = scales::percent(c(0.025, 0.25, 0.5, 0.75, 0.975)),
    DCRB_rev = quantile(sum_DCRB_rev, c(0.025, 0.25, 0.5, 0.75, 0.975), na.rm=TRUE),
    n_risk_humpback = quantile(sum_n_risk_humpback, c(0.025, 0.25, 0.5, 0.75, 0.975), na.rm=TRUE),
    n_risk_blue = quantile(sum_n_risk_blue, c(0.025, 0.25, 0.5, 0.75, 0.975), na.rm=TRUE),
    
    .groups = "drop"
    ) %>%
  pivot_longer(
    cols = c(DCRB_rev, n_risk_humpback, n_risk_blue), 
    names_to = "response_metric", 
    values_to = "response"
  )
glimpse(annual_risk_out_sq_sm)

# plot normalized blues

ts_sq_blue_sm <- ggplot(
  data = annual_risk_out_sq_sm %>% 
    filter(response_metric != "DCRB_rev" & response_metric != "n_risk_humpback") %>%
    filter(quantile != "2%" & quantile != "98%") %>%
    pivot_wider(
      names_from = quantile,
      values_from = response
    ), 
  aes(
    x = plotting_year, 
    y = `50%`
  )
) +
  geom_point(size=4) +
  geom_line() + 
  # geom_ribbon(aes(ymin = `25%`,
  #                 ymax = `75%`),
  #             alpha = 0.2
  #             ) +
  geom_pointrange(aes(ymin = `25%`,
                  ymax = `75%`)
              ) +
  # geom_errorbar(aes(ymin = `25%`,
  #                 ymax = `75%`)
  #             ) +
  geom_vline(xintercept = c(2014.5, 2018.5), linetype=2) +
  # facet_grid(rows = vars(response_metric),
  #            #scales = "free_y",
  #            labeller = as_labeller ( c(
  #              #`DCRB_rev` = "Dungeness crab\nfishery revenue\n($, x10^5)",
  #              `n_risk_blue` = ""#, # to\nblue whales",
  #              #`n_risk_humpback` = "" # "Risk to\nhumpback whales\n(normalized)"
  #              ) ),
  #            switch = "y"
  #            ) + 
  scale_x_continuous(breaks = seq(2010, 2019,1),
                     limits = c(2009.5,2019.5)
                     #labels = str_replace(unique(annual_risk_out_sq_sm$crab_year),"_","-")
                     ) + 
  #ylim(0,0.08) +
  ylab("Normalized risk") + 
  xlab("") + # "Dungeness crab season"
  theme_classic() +
  theme(legend.title = element_blank(),
        #title = element_text(size = 26),
        legend.text = element_text(size = 20),
        legend.position = c(.15, .85),
        axis.text.x = element_blank(), #element_text(hjust = 1,size = 12, angle = 60),
        axis.text.y = element_text(size = 12),
        axis.title = element_text(size = 12),
        strip.text = element_text(size=12),
        strip.background = element_blank(),
        strip.placement = "left"
        )

png(here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  "fig2b_plot_ts_sq_blue_sm_vessels.png"), 
  width = 7.5, height = 2.5, units = "in", res = 300)
ts_sq_blue_sm
invisible(dev.off())

ts_sq_blue_sm

# plot normalized humpbacks

ts_sq_hump_sm <- ggplot(
  data = annual_risk_out_sq_sm %>% 
    filter(response_metric != "DCRB_rev" & response_metric != "n_risk_blue") %>%
    filter(quantile != "2%" & quantile != "98%") %>%
    pivot_wider(
      names_from = quantile,
      values_from = response
    ), 
  aes(
    x = plotting_year, 
    y = `50%`
  )
) +
  geom_point(size=4) +
  geom_line() + 
  # geom_ribbon(aes(ymin = `25%`,
  #                 ymax = `75%`),
  #             alpha = 0.2
  #             ) +
  geom_pointrange(aes(ymin = `25%`,
                  ymax = `75%`)
              ) +
  # geom_errorbar(aes(ymin = `25%`,
  #                 ymax = `75%`)
  #             ) +
  geom_vline(xintercept = c(2014.5, 2018.5), linetype=2) +
  # facet_grid(rows = vars(response_metric),
  #            #scales = "free_y",
  #            labeller = as_labeller ( c(
  #              #`DCRB_rev` = "Dungeness crab\nfishery revenue\n($, x10^5)",
  #              #`n_risk_blue` = ""#, # to\nblue whales",
  #              `n_risk_humpback` = "" # "Risk to\nhumpback whales\n(normalized)"
  #              ) ),
  #            switch = "y"
  #            ) + 
  scale_x_continuous(breaks = seq(2010, 2019,1),
                     limits = c(2009.5,2019.5)
                     #labels = str_replace(unique(annual_risk_out_sq_sm$crab_year),"_","-")
                     ) + 
  #ylim(0,0.08) +
  ylab("Normalized risk") + 
  xlab("") + # "Dungeness crab season"
  theme_classic() +
  theme(legend.title = element_blank(),
        #title = element_text(size = 26),
        legend.text = element_text(size = 20),
        legend.position = c(.15, .85),
        axis.text.x = element_blank(), #element_text(hjust = 1,size = 12, angle = 60),
        axis.text.y = element_text(size = 12),
        axis.title = element_text(size = 12),
        strip.text = element_text(size=12),
        strip.background = element_blank(),
        strip.placement = "left"
        )

png(here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  "fig2c_plot_ts_sq_hump_sm_vessels.png"), 
  width = 7.5, height = 2.5, units = "in", res = 300)
ts_sq_hump_sm
invisible(dev.off())

ts_sq_hump_sm


# plot normalized humpbacks, normalized blues

ts_sq_whales_sm <- ggplot(
  data = annual_risk_out_sq_sm %>% 
    filter(response_metric != "DCRB_rev") %>%
    filter(quantile != "2%" & quantile != "98%") %>%
    pivot_wider(
      names_from = quantile,
      values_from = response
    ), 
  aes(
    x = plotting_year, 
    y = `50%`
  )
) +
  geom_point(size=4) +
  geom_line() + 
  # geom_ribbon(aes(ymin = `25%`,
  #                 ymax = `75%`),
  #             alpha = 0.2
  #             ) +
  geom_pointrange(aes(ymin = `25%`,
                  ymax = `75%`)
              ) +
  # geom_errorbar(aes(ymin = `25%`,
  #                 ymax = `75%`)
  #             ) +
  geom_vline(xintercept = c(2014.5, 2018.5), linetype=2) +
  facet_grid(rows = vars(response_metric),
             #scales = "free_y",
             labeller = as_labeller ( c(
               #`DCRB_rev` = "Dungeness crab\nfishery revenue\n($, x10^5)",
               `n_risk_blue` = "", # to\nblue whales",
               `n_risk_humpback` = "" # "Risk to\nhumpback whales\n(normalized)"
               ) ),
             switch = "y"
             ) + 
  scale_x_continuous(breaks = seq(2010, 2019,1),
                     limits = c(2009.5,2019.5)
                     #labels = str_replace(unique(annual_risk_out_sq_sm$crab_year),"_","-")
                     ) + 
  #ylim(0,0.08) +
  ylab("Normalized risk") + 
  xlab("") + # "Dungeness crab season"
  theme_classic() +
  theme(legend.title = element_blank(),
        #title = element_text(size = 26),
        legend.text = element_text(size = 20),
        legend.position = c(.15, .85),
        axis.text.x = element_blank(), #element_text(hjust = 1,size = 12, angle = 60),
        axis.text.y = element_text(size = 12),
        axis.title = element_text(size = 12),
        strip.text = element_text(size=12),
        strip.background = element_blank(),
        strip.placement = "left"
        )

png(here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  "fig2bc_plot_ts_sq_whales_sm_vessels.png"), 
  width = 7.5, height = 5, units = "in", res = 300)
ts_sq_whales_sm
invisible(dev.off())

ts_sq_whales_sm

# to add silhouettes check out https://cran.r-project.org/web/packages/rphylopic/readme/README.html and https://scrogster.wordpress.com/2014/06/02/adding-phylopic-org-silhouettes-to-r-plots/

# # humpback
# hump_url <- "http://phylopic.org/assets/images/submissions/ce70490a-79a5-47fc-afb9-834e45803ab4.512.png"
# hump_logo <-  readPNG(getURLContent(hump_url))
# 
# logoing_func<-function(logo, x, y, size){
#   dims<-dim(logo)[1:2] #number of x-y pixels for the logo (aspect ratio)
#   AR<-dims[1]/dims[2]
#   par(usr=c(0, 1, 0, 1))
#   rasterImage(logo, x-(size/2), y-(AR*size/2), x+(size/2), y+(AR*size/2), interpolate=TRUE)
# }
# 
# ts_sq_whales + logoing_func(hump_logo, x=0.10, y=0.10, size=0.15)

```

## Figure 3 but for small vessels. 

NOTE: THE CODE BELOW HAS NOT YET BEEN MODIFIED TO FOCUS ON SM VESSELS ONLY. NEED TO ADD FILE PATHS IN 1ST CHUNK ABOVE. PICK UP HERE. 011421

Figure 3. Comparison of expected (a) risk to blue whales, (b) risk to humpback whales, and (c) revenue to the Dungeness crab fishery under a range of status quo and hypothetical alternative management scenarios affecting the entire state of California, during each of three time periods representing before (2009-2014), during (2014-2018), and after (2018-2019) the Northeast Pacific Marine Heatwave. Whale risk and DCRB revenue by scenario calculated in the run_analysis RUN SCENARIOS step

## Risk to blue whales
```{r fig3a_sm, echo=FALSE}

#normalized blues, statewide scenarios only, boxplot
# try using annual_df_focal_scenarios_statewide_ranks instead of statewide_annual_rank_bp

annual_df_focal_scenarios_statewide_ranks_fig <- 
  annual_df_focal_scenarios_statewide_ranks %>%
  ungroup() %>%
  mutate(
    scenario_rank = ifelse(
      pretty_scenario_names == "Status Quo", 0,
      rank_hump
      ),
    scenario_names_fig = c(
      rep("Status Quo", 10),
      rep("Delay", 10),
      rep("Spring Closure", 10),
      rep("Delay + Spring Closure", 10),
      rep("Spring Effort Reduction", 10),
      rep("Spring Depth Restriction", 10),
      rep("Spring Effort Reduction + Depth Restriction", 10)
    )
  )

plot_blue_risk_prepost_statewide_bp <- ggplot(
  data = annual_df_focal_scenarios_statewide_ranks_fig,
  aes(
    x = reorder(scenario_names_fig,-scenario_rank),
    y = n_risk_blue,
    colour = factor(pre_post),
    fill = factor(pre_post)
    )
  )  +
  geom_boxplot(position=position_dodge(width=0.5)) +#, width=0.1, color="grey", alpha=0.2) +
  geom_point(position=position_dodge(width=0.5), alpha=0.6)+
  xlab("") + # "Scenario"
  ylab("Normalized risk") +
  ylim(0,0.4) +
  scale_colour_viridis_d(option="cividis", begin=0.2, end=0.8, direction = -1) +
  scale_fill_viridis_d(option="cividis", begin=0.2, end=0.8, direction = -1, alpha=0.2) +
  scale_x_discrete(labels = wrap_format(30)) + 
  coord_flip() +
  theme_classic() +
  theme(
    legend.title = element_blank(),
    #title = element_text(size = 26),
    legend.text = element_text(size = 14),
    #legend.position = c(.1, .95),
    axis.text.x = element_text(size=12), #hjust = 1, size = 14, angle = 60
    axis.text.y = element_text(size = 12),
    axis.title = element_text(size = 14)
    #legend.position = "none"
    )

png(here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  "fig3a_plot_blue_risk_prepost_statewide_bp.png"), 
  width = 7.5, height = 2.5, units = "in", res = 300)
plot_blue_risk_prepost_statewide_bp
invisible(dev.off())

plot_blue_risk_prepost_statewide_bp

```


## Risk to blue whales geom_pointrange 
```{r fig3aa_sm, echo=FALSE}

#normalized blues, statewide scenarios only, geom_pointrange
# try using annual_df_focal_scenarios_statewide_ranks instead of statewide_annual_rank_bp


annual_df_focal_scenarios_statewide_ranks_fig <- 
  annual_df_focal_scenarios_statewide_ranks %>%
  ungroup() %>%
  mutate(
    scenario_rank = ifelse(
      pretty_scenario_names == "Status Quo", 0,
      rank_hump
      ),
    scenario_names_fig = c(
      rep("Status Quo", 10),
      rep("Delay", 10),
      rep("Spring Closure", 10),
      rep("Delay + Spring Closure", 10),
      rep("Spring Effort Reduction", 10),
      rep("Spring Depth Restriction", 10),
      rep("Spring Effort Reduction + Depth Restriction", 10)
    )
  )

annual_df_focal_scenarios_statewide_ranks_pointrangefig <- annual_df_focal_scenarios_statewide_ranks_fig %>%
    group_by(pre_post, scenario_names_fig, scenario_number) %>%
    summarise(
      mean_n_risk_blue = mean(n_risk_blue, na.rm=TRUE),
      se_n_risk_blue = sd(n_risk_blue, na.rm=TRUE)/sqrt(n()),
      mean_n_risk_humpback = mean(n_risk_humpback, na.rm=TRUE),
      se_n_risk_humpback = sd(n_risk_humpback, na.rm=TRUE)/sqrt(n()),
      mean_DCRB_rev = mean(DCRB_rev, na.rm=TRUE),
      se_DCRB_rev = sd(DCRB_rev, na.rm=TRUE)/sqrt(n()),
      
      .groups = "drop"
    ) %>%
  arrange(
    -mean_n_risk_humpback
  ) %>%
  mutate(
    rank_hump = row_number(),
    scenario_rank = ifelse(
      scenario_names_fig == "Status Quo", 0,
      rank_hump
      )
  )

plot_blue_risk_prepost_statewide_pointrange <- ggplot(
  data = annual_df_focal_scenarios_statewide_ranks_pointrangefig,
  aes(
    x = reorder(scenario_names_fig,-scenario_rank),
    y = mean_n_risk_blue,
    colour = factor(pre_post, levels = c("2018-2019", "2014-2018", "2009-2014")),
    fill = factor(pre_post, levels = c("2018-2019", "2014-2018", "2009-2014"))
    )
  )  +
  geom_pointrange(aes(ymin = mean_n_risk_blue - se_n_risk_blue,
                  ymax = mean_n_risk_blue + se_n_risk_blue),
                  position=position_dodge(width=0.5), alpha=0.6) +
  geom_point(position=position_dodge(width=0.5), alpha=0.6) +
  # geom_line(aes(group=scenario_names_fig),
  #           position=position_dodge(width=0.5), alpha=0.6) +
  xlab("") + # "Scenario"
  ylab("Normalized risk") + #\n (mean +/- SE)") +
  ylim(0,0.4) +
  scale_colour_viridis_d(option="cividis", begin=0.2, end=0.8, direction = -1) +
  scale_fill_viridis_d(option="cividis", begin=0.2, end=0.8, direction = -1, alpha=0.2) +
  scale_x_discrete(labels = wrap_format(30)) + 
  coord_flip() +
  guides(fill = guide_legend(reverse = TRUE), colour = guide_legend(reverse = TRUE)) + 
  theme_classic() +
  theme(
    legend.title = element_blank(),
    #title = element_text(size = 26),
    legend.text = element_text(size = 14),
    #legend.position = c(.1, .95),
    axis.text.x = element_text(size=12), #hjust = 1, size = 14, angle = 60
    axis.text.y = element_text(size = 12),
    axis.title = element_text(size = 14)
    #legend.position = "none"
    )
plot_blue_risk_prepost_statewide_pointrange

png(here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  "fig3a_blue_risk_prepost_statewide_pointrange.png"), 
  width = 7.5, height = 2.5, units = "in", res = 300)
plot_blue_risk_prepost_statewide_pointrange
invisible(dev.off())

# this version does not use coord_flip. js thinks it may be better. but the rank of the scenarios is slightly off (delay shoudl be 3rd from right)

# plot_blue_risk_prepost_statewide_pointrange <- ggplot(
#   data = annual_df_focal_scenarios_statewide_ranks_pointrangefig,
#   aes(
#     x = reorder(scenario_names_fig,scenario_rank),
#     y = mean_n_risk_blue,
#     colour = factor(pre_post, levels = c("2018-2019", "2014-2018", "2009-2014")),
#     fill = factor(pre_post, levels = c("2018-2019", "2014-2018", "2009-2014"))
#     )
#   )  +
#   geom_pointrange(aes(ymin = mean_n_risk_blue - se_n_risk_blue,
#                   ymax = mean_n_risk_blue + se_n_risk_blue),
#                   position=position_dodge(width=0.5), alpha=0.6) +
#   geom_point(position=position_dodge(width=0.5), alpha=0.6) +
#   # geom_line(aes(group=scenario_names_fig),
#   #           position=position_dodge(width=0.5), alpha=0.6) +
#   xlab("") + # "Scenario"
#   ylab("Normalized risk") + #\n (mean +/- SE)") +
#   ylim(0,0.4) +
#   scale_colour_viridis_d(option="cividis", begin=0.2, end=0.8, direction = -1) +
#   scale_fill_viridis_d(option="cividis", begin=0.2, end=0.8, direction = -1, alpha=0.2) +
#   scale_x_discrete(labels = wrap_format(30)) + 
#   #coord_flip() +
#   guides(fill = guide_legend(reverse = TRUE), colour = guide_legend(reverse = TRUE)) + 
#   theme_classic() +
#   theme(
#     legend.title = element_blank(),
#     #title = element_text(size = 26),
#     legend.text = element_text(size = 14),
#     #legend.position = c(.1, .95),
#     axis.text.x = element_text(size=12, angle = 60, hjust = 1), #hjust = 1, size = 14, angle = 60
#     axis.text.y = element_text(size = 12),
#     axis.title = element_text(size = 14)
#     #legend.position = "none"
#     )
# plot_blue_risk_prepost_statewide_pointrange

```

## Risk to humpback whales
```{r fig3b_sm, echo=FALSE}

#normalized humps, statewide scenarios only, boxplot
# try using annual_df_focal_scenarios_statewide_ranks instead of statewide_annual_rank_bp

plot_hump_risk_prepost_statewide_bp <- ggplot(
  data = annual_df_focal_scenarios_statewide_ranks_fig,
  aes(
    x = reorder(scenario_names_fig,-scenario_rank),
    y = n_risk_humpback,
    colour = factor(pre_post),
    fill = factor(pre_post)
    )
  )  +
  geom_boxplot(position=position_dodge(width=0.5)) +#, width=0.1, color="grey", alpha=0.2) +
  geom_point(position=position_dodge(width=0.5), alpha=0.6)+
  xlab("") + # "Scenario"
  ylab("Normalized risk") +
  ylim(0,0.4) +
  scale_colour_viridis_d(option="cividis", begin=0.2, end=0.8, direction = -1) +
  scale_fill_viridis_d(option="cividis", begin=0.2, end=0.8, direction = -1, alpha=0.2) +
  scale_x_discrete(labels = wrap_format(30)) + 
  coord_flip() +
  theme_classic() +
  theme(
    legend.title = element_blank(),
    #title = element_text(size = 26),
    legend.text = element_text(size = 14),
    #legend.position = c(.1, .95),
    axis.text.x = element_text(size=12), #hjust = 1, size = 14, angle = 60
    axis.text.y = element_text(size = 12),
    axis.title = element_text(size = 14)
    #legend.position = "none"
    )

png(here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  "fig3b_plot_hump_risk_prepost_statewide_bp.png"), 
  width = 7.5, height = 2.5, units = "in", res = 300)
plot_hump_risk_prepost_statewide_bp
invisible(dev.off())

plot_hump_risk_prepost_statewide_bp

```

## Risk to humpback whales geom_pointrange 
```{r fig3bb_sm, echo=FALSE}

#normalized humps, statewide scenarios only, geom_pointrange
# try using annual_df_focal_scenarios_statewide_ranks instead of statewide_annual_rank_bp


# annual_df_focal_scenarios_statewide_ranks_fig <- 
#   annual_df_focal_scenarios_statewide_ranks %>%
#   ungroup() %>%
#   mutate(
#     scenario_rank = ifelse(
#       pretty_scenario_names == "Status Quo", 0,
#       rank_hump
#       ),
#     scenario_names_fig = c(
#       rep("Status Quo", 10),
#       rep("Delay", 10),
#       rep("Spring Closure", 10),
#       rep("Delay + Spring Closure", 10),
#       rep("Spring Effort Reduction", 10),
#       rep("Spring Depth Restriction", 10),
#       rep("Spring Effort Reduction + Depth Restriction", 10)
#     )
#   )
# 
# annual_df_focal_scenarios_statewide_ranks_pointrangefig <- annual_df_focal_scenarios_statewide_ranks_fig %>%
#     group_by(pre_post, scenario_names_fig, scenario_number) %>%
#     summarise(
#       mean_n_risk_blue = mean(n_risk_blue),
#       se_n_risk_blue = sd(n_risk_blue)/sqrt(n()),
#       mean_n_risk_humpback = mean(n_risk_humpback),
#       se_n_risk_humpback = sd(n_risk_humpback)/sqrt(n()),
#       mean_DCRB_rev = mean(DCRB_rev),
#       se_DCRB_rev = sd(DCRB_rev)/sqrt(n()),
#       
#       .groups = "drop"
#     ) %>%
#   arrange(
#     -mean_n_risk_humpback
#   ) %>%
#   mutate(
#     rank_hump = row_number(),
#     scenario_rank = ifelse(
#       scenario_names_fig == "Status Quo", 0,
#       rank_hump
#       )
#   )

plot_hump_risk_prepost_statewide_pointrange <- ggplot(
  data = annual_df_focal_scenarios_statewide_ranks_pointrangefig,
  aes(
    x = reorder(scenario_names_fig,-scenario_rank),
    y = mean_n_risk_humpback,
    colour = factor(pre_post, levels = c("2018-2019", "2014-2018", "2009-2014")),
    fill = factor(pre_post, levels = c("2018-2019", "2014-2018", "2009-2014"))
    )
  )  +
  geom_pointrange(aes(ymin = mean_n_risk_humpback - se_n_risk_humpback,
                  ymax = mean_n_risk_humpback + se_n_risk_humpback),
                  position=position_dodge(width=0.5), alpha=0.6) +
  geom_point(position=position_dodge(width=0.5), alpha=0.6) +
  # geom_line(aes(group=scenario_names_fig),
  #           position=position_dodge(width=0.5), alpha=0.6) +
  xlab("") + # "Scenario"
  ylab("Normalized risk") + #\n (mean +/- SE)") +
  ylim(0,0.4) +
  scale_colour_viridis_d(option="cividis", begin=0.2, end=0.8, direction = -1) +
  scale_fill_viridis_d(option="cividis", begin=0.2, end=0.8, direction = -1, alpha=0.2) +
  scale_x_discrete(labels = wrap_format(30)) + 
  coord_flip() +
  guides(fill = guide_legend(reverse = TRUE), colour = guide_legend(reverse = TRUE)) + 
  theme_classic() +
  theme(
    legend.title = element_blank(),
    #title = element_text(size = 26),
    legend.text = element_text(size = 14),
    #legend.position = c(.1, .95),
    axis.text.x = element_text(size=12), #hjust = 1, size = 14, angle = 60
    axis.text.y = element_text(size = 12),
    axis.title = element_text(size = 14)
    #legend.position = "none"
    )
plot_hump_risk_prepost_statewide_pointrange

png(here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  "fig3b_hump_risk_prepost_statewide_pointrange.png"), 
  width = 7.5, height = 2.5, units = "in", res = 300)
plot_hump_risk_prepost_statewide_pointrange
invisible(dev.off())

```


## Figure 4 but for small vessels

Figure 4. Tradeoff plots indicating the expected (a) risk to blue whales or (b) risk to humpback whales in comparison to expected revenue to the California Dungeness crab fishery under a range of hypothetical alternative management scenarios, during each of three time periods representing before (2009-2014), during (2014-2018), and after (2018-2019) the Northeast Pacific Marine Heatwave. Larger points represent the median ±1SE for each scenario across years, smaller points represent values for individual years. Whale risk and DCRB revenue by scenario calculated in the run_analysis RUN SCENARIOS step

## Blue whales only
```{r tradeoff1_sm, echo = FALSE}

### normalized blues, by pre_post
# still getting warnings aroudn geom_errorbarh. that is because the 2018-19 period has no var. ok to proceed

df_tradeoff_focal_scenarios <- read_rds(path_df_tradeoff_focal_scenarios)

df_tradeoff_focal_scenarios %>%
  group_by(pretty_scenario_names, scenario_number, pre_post) %>%
  summarise(
    mean_relative_blwh_risk_n = mean(relative_blwh_risk_n),
    sd_relative_blwh_risk_n = sd(relative_blwh_risk_n),
    mean_relative_hump_risk_n = mean(relative_hump_risk_n),
    sd_relative_hump_risk_n = sd(relative_hump_risk_n),
    mean_relative_dollars = mean(relative_dollars),
    sd_relative_dollars = sd(relative_dollars)
  )

# 113020 need to update other to figure to reflect changes here

to_blue_n_pre_post <- ggplot(
  df_tradeoff_focal_scenarios, #%>% filter(scenario_number != "15" & scenario_number != "14" & scenario_number != "9"), #df
  aes(
    x=relative_dollars,
    y=relative_blwh_risk_n,
    group=pretty_scenario_names,
    fill=scenario_type,
    label = scenario_number,
    colour=scenario_type,
    shape = spatial_domain
  )
) + 
  geom_point(size=1, alpha=0.2) + #,position=position_dodge(width = 0.9)
  #geom_text(aes(label=scenario_number),size=1, alpha=0.5, check_overlap = TRUE) + #aes(label=plotting_year, colour=scenario_type), #str_replace(crab_year,"_","-"), _repel
  stat_err(spread = "se", mult = 1, width=.1, alpha = 0.5) + 
  stat_err(geom="point", size=3, alpha = 0.8) +
  #stat_err(geom="label", size=3, alpha = 0.8) + #
  stat_err(spread = "se", mult = 1, geom = "errorbarh", height = .1,alpha = 0.5) +
  #facet_grid(rows = vars(pre_post) ) + 
  facet_wrap(nrow = 3,  vars(pre_post) ) + 
  scale_colour_viridis_d(option="inferno", begin=0.2, end=0.8, direction = -1, alpha=0.8,
                         labels = function(x) str_wrap(x, width = 20)) +
  scale_fill_viridis_d(option="inferno", begin=0.2, end=0.8, direction = -1, alpha=0.8,
                       labels = function(x) str_wrap(x, width = 20)) +
  scale_shape_manual(values=c(21:25, 4, 8) ) +
  ylab("Relative reduction in risk") + #  to blues\n(normalized)
  xlab("Relative fishery revenue") + #  to the Dungeness crab fishery
  scale_y_continuous(breaks=seq(-25, 100, 25),limits=c(-30,100)) +
  guides(fill = guide_legend("Scenario type", order = 1), 
         colour = guide_legend("Scenario type", order = 1),
         shape = guide_legend("Spatial domain", order = 2)
         ) + 
  theme_classic(base_size = 12) +
  theme(
    #title = element_text(size = 26),
    legend.position = "right",
    legend.box="vertical", 
    #legend.margin=margin(),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    #axis.title = element_text(size = 14),
    legend.text = element_text(size=10),
    #strip.text = element_text(size=14),
    ) 
to_blue_n_pre_post

# need to figure out how to shrink to fit margins with font sizes

png(
  here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  paste0("Tradeoff plot - normalized blues only by pre_post_",today(),".png")), 
    width = 3.75, height = 6, units = "in", res = 300)
to_blue_n_pre_post
invisible(dev.off())

to_blue_n_pre_post_nolegend <- ggplot(
  df_tradeoff_focal_scenarios, #%>% filter(scenario_number != "15" & scenario_number != "14" & scenario_number != "9"), #df
  aes(
    x=relative_dollars,
    y=relative_blwh_risk_n,
    group=pretty_scenario_names,
    fill=scenario_type,
    #label = scenario_number,
    colour=scenario_type,
    shape = spatial_domain
  )
) + 
  geom_point(size=1, alpha=0.2) + #,position=position_dodge(width = 0.9)
  #geom_text(aes(label=plotting_year, colour=scenario_type), size=2, alpha=0.5, check_overlap = TRUE) + #str_replace(crab_year,"_","-"), _repel
  stat_err(spread = "se", mult = 1, width=.1, alpha = 0.5) + 
  stat_err(geom="point", size=3, alpha = 0.8) + #
  stat_err(spread = "se", mult = 1, geom = "errorbarh", height = .1,alpha = 0.5) +
  #facet_grid(rows = vars(pre_post) ) + 
  facet_wrap(nrow = 3,  vars(pre_post) ) + 
  scale_colour_viridis_d(option="inferno", begin=0.2, end=0.8, direction = -1, alpha=0.8,
                         labels = function(x) str_wrap(x, width = 10)) +
  scale_fill_viridis_d(option="inferno", begin=0.2, end=0.8, direction = -1, alpha=0.8,
                       labels = function(x) str_wrap(x, width = 10)) +
  scale_shape_manual(values=c(21:25, 4, 8) ) +
  ylab("Relative reduction in risk") + #  to blues\n(normalized)
  xlab("Relative fishery revenue") + #  to the Dungeness crab fishery
  scale_y_continuous(breaks=seq(-25, 100, 25),limits=c(-30,100)) +
  guides(fill = guide_legend("Scenario type", order = 1), 
         colour = guide_legend("Scenario type", order = 1),
         shape = guide_legend("Spatial domain", order = 2)
         ) + 
  theme_classic(base_size = 12) +
  theme(
    #title = element_text(size = 26),
    legend.position = "none",
    legend.box="vertical", 
    #legend.margin=margin(),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    #axis.title = element_text(size = 14),
    legend.text = element_text(size=10),
    #strip.text = element_text(size=14),
    ) 
to_blue_n_pre_post_nolegend

png(
  here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  paste0("Tradeoff plot - normalized blues only by pre_post_nolegend_",today(),".png")), 
    width = 3.75, height = 6, units = "in", res = 300)
to_blue_n_pre_post_nolegend
invisible(dev.off())


```

## Humpback whales only
```{r tradeoff2_sm, echo = FALSE}

### normalized humps, by pre_post
# still getting warnings aroudn geom_errorbarh. that is because the 2018-19 period has no var. ok to proceed

to_hump_n_pre_post <- ggplot(
  df_tradeoff_focal_scenarios, #%>% filter(scenario_number != "15" & scenario_number != "14" & scenario_number != "9"), #df
  aes(
    x=relative_dollars,
    y=relative_hump_risk_n,
    group=pretty_scenario_names,
    fill=scenario_type,
    #label = scenario_number,
    colour=scenario_type,
    shape = spatial_domain
  )
) + 
  geom_point(size=1, alpha=0.2) + #,position=position_dodge(width = 0.9)
  #geom_text(aes(label=plotting_year, colour=scenario_type), size=2, alpha=0.5, check_overlap = TRUE) + #str_replace(crab_year,"_","-"), _repel
  stat_err(spread = "se", mult = 1, width=.1, alpha = 0.5) + 
  stat_err(geom="point", size=3, alpha = 0.8) + #
  stat_err(spread = "se", mult = 1, geom = "errorbarh", height = .1,alpha = 0.5) +
  #facet_grid(rows = vars(pre_post) ) + 
  facet_wrap(nrow = 3,  vars(pre_post) ) + 
  scale_colour_viridis_d(option="inferno", begin=0.2, end=0.8, direction = -1, alpha=0.8,
                         labels = function(x) str_wrap(x, width = 10)) +
  scale_fill_viridis_d(option="inferno", begin=0.2, end=0.8, direction = -1, alpha=0.8,
                       labels = function(x) str_wrap(x, width = 10)) +
  scale_shape_manual(values=c(21:25, 4, 8) ) +
  ylab("Relative reduction in risk") + #  to humps\n(normalized)
  xlab("Relative fishery revenue") + #  to the Dungeness crab fishery
  scale_y_continuous(breaks=seq(-25, 100, 25),limits=c(-30,100)) +
  guides(fill = guide_legend("Scenario type", order = 1), 
         colour = guide_legend("Scenario type", order = 1),
         shape = guide_legend("Spatial domain", order = 2)
         ) + 
  theme_classic(base_size = 12) +
  theme(
    #title = element_text(size = 26),
    legend.position = "right",
    legend.box="vertical", 
    #legend.margin=margin(),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    #axis.title = element_text(size = 14),
    legend.text = element_text(size=10),
    #strip.text = element_text(size=14),
    ) 
to_hump_n_pre_post

# need to figure out how to shrink to fit margins with font sizes

png(
  here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  paste0("Tradeoff plot - normalized humps only by pre_post_",today(),".png")), 
    width = 3.75, height = 6, units = "in", res = 300)
to_hump_n_pre_post
invisible(dev.off())

to_hump_n_pre_post_nolegend <- ggplot(
  df_tradeoff_focal_scenarios, #%>% filter(scenario_number != "15" & scenario_number != "14" & scenario_number != "9"), #df
  aes(
    x=relative_dollars,
    y=relative_hump_risk_n,
    group=pretty_scenario_names,
    fill=scenario_type,
    #label = scenario_number,
    colour=scenario_type,
    shape = spatial_domain
  )
) + 
  geom_point(size=1, alpha=0.2) + #,position=position_dodge(width = 0.9)
  #geom_text(aes(label=plotting_year, colour=scenario_type), size=2, alpha=0.5, check_overlap = TRUE) + #str_replace(crab_year,"_","-"), _repel
  stat_err(spread = "se", mult = 1, width=.1, alpha = 0.5) + 
  stat_err(geom="point", size=3, alpha = 0.8) + #
  stat_err(spread = "se", mult = 1, geom = "errorbarh", height = .1,alpha = 0.5) +
  #facet_grid(rows = vars(pre_post) ) + 
  facet_wrap(nrow = 3,  vars(pre_post) ) + 
  scale_colour_viridis_d(option="inferno", begin=0.2, end=0.8, direction = -1, alpha=0.8,
                         labels = function(x) str_wrap(x, width = 10)) +
  scale_fill_viridis_d(option="inferno", begin=0.2, end=0.8, direction = -1, alpha=0.8,
                       labels = function(x) str_wrap(x, width = 10)) +
  scale_shape_manual(values=c(21:25, 4, 8) ) +
  ylab("Relative reduction in risk") + #  to humps\n(normalized)
  xlab("Relative fishery revenue") + #  to the Dungeness crab fishery
  scale_y_continuous(breaks=seq(-25, 100, 25),limits=c(-30,100)) +
  guides(fill = guide_legend("Scenario type", order = 1), 
         colour = guide_legend("Scenario type", order = 1),
         shape = guide_legend("Spatial domain", order = 2)
         ) + 
  theme_classic(base_size = 12) +
  theme(
    #title = element_text(size = 26),
    legend.position = "none",
    legend.box="vertical", 
    #legend.margin=margin(),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    #axis.title = element_text(size = 14),
    legend.text = element_text(size=10),
    #strip.text = element_text(size=14),
    ) 
to_hump_n_pre_post_nolegend

png(
  here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "figures",
  paste0("Tradeoff plot - normalized humps only by pre_post_nolegend_",today(),".png")), 
    width = 3.75, height = 6, units = "in", res = 300)
to_hump_n_pre_post_nolegend
invisible(dev.off())


```