---
title: "Samhouri et al tradeoff analysis make figures"
author: "Jameal Samhouri"
date: "10/27/2020"
output: 
  html_document:
      toc: true
      toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This document reflects the figures included in the Samhouri et al. tradeoff manuscript.

See pseudocode in "Full workflow for risk assessment and tradeoff analysis" google doc. https://docs.google.com/document/d/1T8S5rEOACMCeGa5efS5BnEQY4gOw-PbxmvILJvShsQo/edit

* Figure 1. Maps of whales and DCRB pings in 3 time periods on 5km grid. Made outside of this doc.

* Figure 2. Changes over time (2009-2019) in (a) the number of confirmed whale entanglements along the US West Coast, (b) expected risk to blue and humpback whales, measured as spatial overlap (see Methods), from the California Dungeness crab fishery, and (c) revenue to the California Dungeness crab fishery. whale risk is calculated in the run_analysis RUN SCENARIOS step

* Figure 3. Comparison of expected (a) risk to blue whales, (b) risk to humpback whales, and (c) revenue to the Dungeness crab fishery under a range of status quo and hypothetical alternative management scenarios affecting the entire state of California, during each of three time periods representing before (2009-2014), during (2014-2018), and after (2018-2019) the Northeast Pacific Marine Heatwave. Whale risk and DCRB revenue by scenario calculated in the run_analysis RUN SCENARIOS step

* Figure 4. Tradeoff plots indicating the expected (a) risk to blue whales or (b) risk to humpback whales in comparison to expected revenue to the California Dungeness crab fishery under a range of hypothetical alternative management scenarios, during each of three time periods representing before (2009-2014), during (2014-2018), and after (2018-2019) the Northeast Pacific Marine Heatwave. Larger points represent the median ±1SE for each scenario across years, smaller points represent values for individual years. Whale risk and DCRB revenue by scenario calculated in the run_analysis RUN SCENARIOS step

SI Figures follow from there.
Entanglement ts by gear type (updated version of Fig14 in Saez et al. 2020)

## Prep libraries and file paths

```{r prep, include=FALSE}

# just getting libraries loaded and file paths sorted

library(tidyverse)
library(foreign) # read.dbf()
library(here)
library(lubridate)
library(sf)
library(viridis)
library(ggrepel)
library(magrittr)
library(ggerr)
library(scales)
library(maps)
library(rnaturalearth)
library(gridExtra)
library(ggpubr)
library(knitr)

source(here::here("User_script_local.R"))
if (user == "JS") {
  #flag.save = TRUE # option to save all outputs or not
  
  path_entanglement_file2 <- paste0(here::here(), "/tradeoffs/full_analysis_samhourietal/output_dfs/entanglement_df_annual_complete.csv")
  
path_figures <- "/Users/jameal.samhouri/Dropbox/Projects/In progress/RAIMBOW/raimbow/tradeoffs/full_analysis_samhourietal/figures"
  
  path_Grid_5km_landerased <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/5x5 Grid/Grid_5km_landerased.RDATA"
  
  path_entanglement_file <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Input_Data/Entanglement Data/2000_19/2000_19.dbf" # from Lauren Saez
  

  
  path_rds <- "/Users/jameal.samhouri/Dropbox/Projects/In progress/RAIMBOW/raimbow/tradeoffs/Management scenarios"
  
  path_maps <- "/Users/jameal.samhouri/Dropbox/Projects/In progress/RAIMBOW/raimbow/tradeoffs/maps"
  
  path_status_quo <- paste0("/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/scenario_output_dataframes/status_quo_risk_2009_2019_yr_mth_2020-06-14.rds") # from extract_status_quo_scenario_summary.R
  
  path_status_quo_small_vessels <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/scenario_output_dataframes/status_quo_risk_small_vessels_2009_2019_yr_mth_2020-06-17.rds" # from extract_status_quo_scenario_summary_small.R 
  
  path_whales <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/5x5 Grid/Grid5km_whale.rds"
  
  path_Humpback_5km_long_monthly <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Input_Data/Humpback whale data/Forney et al./Humpback_5km_long_monthly.rds" 
  
  path_BlueWhale_5km_long_monthly <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Input_Data/Blue whale data/Overlay on 5km Grid/BlueWhale_5km_long_monthly.rds"
    
  path_effort_shift_scenariocomparisons_annual_statewide_df_n <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/annual_statewide_effort_shift_scenariocomparisons_n_2020-05-19.rds" # from effort_shift_comparison.R
  
  path_tradeoff_df_effort_shift_scenariocomparisons_n <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/tradeoff_df_effort_shift_scenariocomparisons_n_2020-05-19.rds"
  
  path_scenario_table_all <- "/Users/jameal.samhouri/Dropbox/Projects/In progress/RAIMBOW/raimbow/tradeoffs/Management scenarios/scenario_table_all_2020-06-09.rds"
  
  path_scenario_table_focal_scenarios <- "/Users/jameal.samhouri/Dropbox/Projects/In progress/RAIMBOW/raimbow/tradeoffs/Management scenarios/scenario_table_focal_scenarios_2020-06-12.rds"
  
  path_annual_statewide_df_focal_scenarios <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/annual_statewide_df_focal_scenarios_2020-06-12.rds" 
  
  path_df_tradeoff_focal_scenarios <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/df_tradeoff_focal_scenarios_2020-06-12.rds" 
  
  path_annual_statewide_df_focal_scenarios_small_vessels <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/annual_statewide_df_focal_scenarios_small_vessels_2020-06-16.rds"
  
  path_df_tradeoff_focal_scenarios_small_vessels <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/df_tradeoff_focal_scenarios_small_vessels_2020-06-16.rds"
  
  path_prioritizr_all <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/prioritizr/prioritzr_outputs_all.rds"
  

} else if (user == "SMW") {
  NULL
} else {
  stop("Invalid user")
}


```

# Figure 2. 

Changes over time (2009-2019) in (a) the number of confirmed whale entanglements along the US West Coast, (b) expected risk to blue and humpback whales, measured as spatial overlap (see Methods), from the California Dungeness crab fishery, and (c) revenue to the California Dungeness crab fishery.

## Entanglements
TS of ONLY CA commercial DCRB entanglements
```{r fig2, echo=FALSE}

entanglement_df_annual_complete <- read_csv(path_entanglement_file2)

entanglements_ts <- ggplot(entanglement_df_annual_complete %>% filter (Year > 2008), 
                           aes(x=Year, y=count, group=species, colour=species)) + # group=1 tells ggplot that there is only 1 group
  geom_point(size=2, position = position_dodge(0.4)) +
  geom_line(position = position_dodge(0.4)) + 
  ylab("Number of entanglements") +
  xlab("Year") +
  labs(subtitle = "Data from NMFS WRO") +
  #ggtitle("Number of Whale\nEntanglements,US West Coast") +
  scale_colour_viridis_d(begin=0.2, end=0.8) +
  scale_x_continuous(breaks=seq(2009, 2019 , 1),limits=c(2008,2020))+
  scale_y_continuous(breaks=seq(0, 25, 5),limits=c(0,25))+
  theme_classic() +
  #scale_fill_manual(values=c("lightskyblue", "coral")) +
  theme(legend.title = element_blank(),
        #title = element_text(size = 32),
        legend.text = element_text(size=14),
        axis.text.x = element_text(hjust = 1,size = 14, angle = 90),
        axis.text.y = element_text(size = 14),
        axis.title = element_text(size = 16),
        legend.position = c(0.2, 0.8)
        )
entanglements_ts 

png(paste0(path_figures, "/entanglements_ts.png"), width = 14, height = 10, units = "in", res = 300)
entanglements_ts
invisible(dev.off())


```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
