---
title: "Samhouri et al tradeoff analysis run analysis"
author: "Jameal Samhouri"
date: "11/4/2020"
output: 
  html_document:
      toc: true
      toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This document runs the analysis included in the Samhouri et al. tradeoff manuscript.

See pseudocode in "Full workflow for risk assessment and tradeoff analysis" google doc. https://docs.google.com/document/d/1T8S5rEOACMCeGa5efS5BnEQY4gOw-PbxmvILJvShsQo/edit


1. MAKE SCENARIOS
2. RUN SCENARIOS
3. MAKE DF'S FOR TIME SERIES OF RISK, REVENUE, AND LANDINGS
+ status quo as separate df, all scenarios in one df, and scale_compare df
+ repeat for sm vessels
4. CALCULATE SCENARIO RANKS
+ one df by year, one by period
+ repeat for statewide only and cenCA only scenarios
+ ranks based on hump risk, blwh risk, DCRB $, DCRB lbs, cost effectiveness for humps, cost effectiveness for blues
5. CALCULATE CHANGES IN RISK AND REVENUE BY PERIOD AND SCENARIO
+ repeat for statewide only and cenCA only scenarios
+ % change in each scenario's impact on risk and revenue between 2009-14 and 2014-18 periods
6. QUERY ENTANGLEMENT DATA
7. QUERY FISHING DATA
8. QUERY WHALE MODEL OUTPUTS
+ predicted occurrence/density by period, statewide and by region

* Fig 1 data. Whale and fishing data on 5km grid, averaged by time period
* Fig 2 data. Entanglement reports, DCRB revenue by crab year. whale risk is calculated in the scenario_analysis step
* Fig 3 data. Whale risk and DCRB revenue by scenario calculated in the scenario_analysis step
* Fig 4 data. Whale risk and DCRB revenue by scenario calculated in the scenario_analysis step

## Prep libraries and file paths
(no need to display messages)
```{r prep, include=FALSE}

# just getting libraries loaded and file paths sorted

library(tidyverse)
library(foreign) # read.dbf()
library(here)
library(lubridate)
library(sf)
library(viridis)
library(ggrepel)
library(magrittr)
library(ggerr)
library(scales)
library(maps)
library(rnaturalearth)
library(gridExtra)
library(ggpubr)
library(knitr)

source(here::here("User_script_local.R"))
if (user == "JS") {
  #flag.save = TRUE # option to save all outputs or not
  
  path_entanglement_file2 <- paste0(here::here(), "/tradeoffs/full_analysis_samhourietal/output_dfs/entanglement_df_annual_complete.csv")
  
  path_dcrb_vms_tix_analysis <- "~/Documents/RAIMBOW/Processed Data/VMS/vms_all_interpolated_w_grd_vlengths.RDS"
  
  path_dcrb_daily_all <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/VMS/CA_DCRB_vms_fishing_daily_2009-2019_all_vessels_regions_depths.RDS"
  
  path_season.st.date.key <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Input_Data/season_start_dates/start_dates_by_CA_region.rds"
  
  # path_Grid_5km_landerased <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/5x5 Grid/Grid_5km_landerased.RDATA"
  # 
  # path_figures <- "/Users/jameal.samhouri/Dropbox/Projects/In progress/RAIMBOW/raimbow/tradeoffs/Management scenarios/figures"
  # 
  # path_rds <- "/Users/jameal.samhouri/Dropbox/Projects/In progress/RAIMBOW/raimbow/tradeoffs/Management scenarios"
  # 
  # path_maps <- "/Users/jameal.samhouri/Dropbox/Projects/In progress/RAIMBOW/raimbow/tradeoffs/maps"
  # 
  # path_status_quo <- paste0("/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/scenario_output_dataframes/status_quo_risk_2009_2019_yr_mth_2020-06-14.rds") # from extract_status_quo_scenario_summary.R
  # 
  # path_status_quo_small_vessels <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/scenario_output_dataframes/status_quo_risk_small_vessels_2009_2019_yr_mth_2020-06-17.rds" # from extract_status_quo_scenario_summary_small.R 
  # 
  # path_whales <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/5x5 Grid/Grid5km_whale.rds"
  # 
  # path_Humpback_5km_long_monthly <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Input_Data/Humpback whale data/Forney et al./Humpback_5km_long_monthly.rds" 
  # 
  # path_BlueWhale_5km_long_monthly <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Input_Data/Blue whale data/Overlay on 5km Grid/BlueWhale_5km_long_monthly.rds"
  #   
  # path_effort_shift_scenariocomparisons_annual_statewide_df_n <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/annual_statewide_effort_shift_scenariocomparisons_n_2020-05-19.rds" # from effort_shift_comparison.R
  # 
  # path_tradeoff_df_effort_shift_scenariocomparisons_n <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/tradeoff_df_effort_shift_scenariocomparisons_n_2020-05-19.rds"
  # 
  # path_scenario_table_all <- "/Users/jameal.samhouri/Dropbox/Projects/In progress/RAIMBOW/raimbow/tradeoffs/Management scenarios/scenario_table_all_2020-06-09.rds"
  # 
  # path_scenario_table_focal_scenarios <- "/Users/jameal.samhouri/Dropbox/Projects/In progress/RAIMBOW/raimbow/tradeoffs/Management scenarios/scenario_table_focal_scenarios_2020-06-12.rds"
  # 
  # path_annual_statewide_df_focal_scenarios <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/annual_statewide_df_focal_scenarios_2020-06-12.rds" 
  # 
  # path_df_tradeoff_focal_scenarios <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/df_tradeoff_focal_scenarios_2020-06-12.rds" 
  # 
  # path_annual_statewide_df_focal_scenarios_small_vessels <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/annual_statewide_df_focal_scenarios_small_vessels_2020-06-16.rds"
  # 
  # path_df_tradeoff_focal_scenarios_small_vessels <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/df_tradeoff_focal_scenarios_small_vessels_2020-06-16.rds"
  # 
  # path_prioritizr_all <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/prioritizr/prioritzr_outputs_all.rds"
  

} else if (user == "SMW") {
  NULL
} else {
  stop("Invalid user")
}


```


# MAKE SCENARIOS
```{r make_scenarios, echo=FALSE}

source(paste0(here::here(), "/tradeoffs/Management scenarios/make_scenarios_table.R"))
# note season.st.key argument has to be defined separately from above

# scenario_table[1,]
# scenario_table_edr[1,]

# separate scenario tables need to be bound together.
# number the scenarios for easier reference. sq indicates status quo; redundant indicates scenario already represented and can be ignored; alt indicates scenario uses alt set of assumptions about what happens to effort (either remove or not removed and/or 10% v 100% of effort redistributed)
scenario_table_all <- scenario_table %>%
  bind_rows(scenario_table_edr) %>%
  mutate(
    scenario_number = c(
      "sq","1","2","3","6","8","4alt","7alt","9alt","5",
      "10","11","sq_redundant","1_redundant","2_redundant",
      "3_redundant","6_redundant","8_redundant","4","7","9",
      "5alt","10alt","11alt","12","13","14","15","16","17",
      "18","13alt","14alt","15alt","16alt","17alt","18alt"
    )
  )

# scenario_table_all[3,] == scenario_table_all[15,]

# convert factors to characters
scenario_table_all$delay_scenario <- as.character(scenario_table_all$delay_scenario)
scenario_table_all$delay_scenario <- as.character(scenario_table_all$delay_scenario)
scenario_table_all$closure_scenario <- as.character(scenario_table_all$closure_scenario)
scenario_table_all$closure_scenario <- as.character(scenario_table_all$closure_scenario)

glimpse(scenario_table_all)

write_csv(scenario_table_all,here::here(
  "tradeoffs",
  "Management scenarios",
  "scenario_table_all.csv"
          )
)

```

# RUN SCENARIOS
```{r run_scenarios, echo=FALSE}

### 1) Loop through scenarios of interest and create a list of output df's

# 1a) load function to shift fishing effort and daily fishing data
source(here::here(
  "tradeoffs",
  "Management scenarios",
  "Mgmt_scenarios_shift_effort.R")
)
x.orig <- read_rds(path_dcrb_daily_all)

# 1b) read in season start date key
# grab helper function. https://stackoverflow.com/questions/6364783/capitalize-the-first-letter-of-both-words-in-a-two-word-string
simpleCap <- function(x) {
  s <- sapply(strsplit(x, " "), function(i) i[[1]])
  paste(toupper(substring(s, 1, 1)), substring(s, 2),
        sep = "", collapse = " ")
}
season.st.date.key <- read_rds(path_season.st.date.key) %>% 
  mutate(crab_year = gsub("-", "_", .data$crab_season), 
         Region = unname(sapply(CA_region, simpleCap))) %>% 
  select(crab_year, Region, start_of_season_oneperc)

# 1c) shift effort

start.time <- Sys.time()

scenario.output.list <- lapply(1:nrow(scenario_table_all), function(i, scenario_table_input) { # for testing. nrow(scenario_table_all[1:2,])
  print(i)
  #browser()
  
  # i=1 # testing. no longer breaks because "At least one of delay.date or closure.date must not be NULL"
  # i=2 # testing. no longer breaks because when switch() is used and does not return NULL, it returns nothing
  
  scenario.output.df <- effort_mgmt(
    x = x.orig,
    
    season.st.key = season.st.date.key,
    
    preseason.days = scenario_table_all$preseason.days[i],
    
    season.st.backstop = if (scenario_table_all$season.st.backstop[i] == "NULL") {
      NULL
      }
    else {
      as.Date(scenario_table_all$season.st.backstop[i])
      },
    
    early.data.method = scenario_table_all$early.data.method[i], 
    
    delay.date = if (scenario_table_all$delay.date[i] == "NULL") {
      NULL
    } else {
      as.Date(scenario_table_all$delay.date[i])
    },
    
    delay.region = if (scenario_table_all$delay.region[i] == "NULL") NULL else scenario_table_all$delay.region[i],
    
    delay.method = if (scenario_table_all$delay.method[i] == "NULL") NULL else scenario_table_all$delay.method[i],
    
    delay.method.fidelity = scenario_table_all$delay.method.fidelity[i],
    
    closure.date = if (scenario_table_all$closure.date[i] == "NULL") NULL else as.Date(scenario_table_all$closure.date[i]),
    
    closure.region = if (scenario_table_all$closure.region[i] == "NULL") NULL else scenario_table_all$closure.region[i],
    
    closure.method = if (scenario_table_all$closure.method[i] == "NULL") NULL else scenario_table_all$closure.method[i],
    
    closure.redist.percent = scenario_table_all$closure.redist.percent[i],
    
    depth.shallow = if (scenario_table_all$depth.shallow[i] == "NULL") NULL else as.numeric(scenario_table_all$depth.shallow[i]), 
    
    depth.deep = if (scenario_table_all$depth.deep[i] == "NULL") NULL else as.numeric(scenario_table_all$depth.deep[i]),
    
    reduction.before.date = if (scenario_table_all$reduction.before.date[i] == "NULL") NULL else scenario_table_all$reduction.before.date[i],
    
    reduction.before.percent = 50,
    
    reduction.before.region = if (scenario_table_all$reduction.before.region[i] == "NULL") NULL else scenario_table_all$reduction.before.region[i],
    
    reduction.after.date = if (scenario_table_all$reduction.after.date[i] == "NULL") NULL else as.Date(scenario_table_all$reduction.after.date[i]),
    
    reduction.after.percent = 50,
    
    reduction.after.region = if (scenario_table_all$reduction.after.region[i] == "NULL") NULL else scenario_table_all$reduction.after.region[i],
    
    reduction.after.redist = if (scenario_table_all$reduction.after.redist[i] == "FALSE") FALSE else TRUE,
    
    reduction.after.redist.percent = scenario_table_all$reduction.after.redist.percent[i]
    
  )
  
}, scenario_table_input = scenario_table_all)

Sys.time() - start.time

save.image(paste0("/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/scenario_output_dataframes/scenario_output_",today(),".RData"))

### 2) Calculate and summarize risk for multiple scenarios, including status quo
### 3) make annual and tradeoff df's.

```

# QUERY ENTANGLEMENT DATA

## Entanglements by period and species
```{r entanglements, echo=FALSE}

entanglement_df_annual_complete <- read_csv(path_entanglement_file2)

entanglement_df_annual_complete %>% 
  mutate(
    period = case_when(
      Year == 2009 | Year == 2010 | Year == 2011 | Year == 2012 | Year == 2013 ~ "pre",
      Year == 2014 | Year == 2015 | Year == 2016 | Year == 2017 | Year == 2018 ~ "mhw",
      Year == 2019 ~ "post",
      TRUE ~ "OTHER"
    )
  ) %>%
  filter( Year >= 2009) %>%
  group_by(period, species) %>%
  summarise(
    total = sum(count)
  ) %>%
  arrange(
    species,
    factor(period, levels = c("pre", "mhw", "post"))
  ) %>%
  knitr::kable()

```

## Entanglements by period, all species combined
```{r entanglements2, echo=FALSE}

entanglement_df_annual_complete <- read_csv(path_entanglement_file2)

entanglement_df_annual_complete %>% 
  mutate(
    period = case_when(
      Year == 2009 | Year == 2010 | Year == 2011 | Year == 2012 | Year == 2013 ~ "pre",
      Year == 2014 | Year == 2015 | Year == 2016 | Year == 2017 | Year == 2018 ~ "mhw",
      Year == 2019 ~ "post",
      TRUE ~ "OTHER"
    )
  ) %>%
  filter( Year >= 2009) %>%
  group_by(period) %>%
  summarise(
    total = sum(count)
  ) %>%
  arrange(
    factor(period, levels = c("pre", "mhw", "post"))
  ) %>%
  knitr::kable()

```

# QUERY FISHING DATA

```{r fishing, include=FALSE}


# most of this code originally from query_fishing_data.R
# this takes a few min to read in
dcrb_vms_tix_analysis <- read_rds(path_dcrb_vms_tix_analysis)
glimpse(dcrb_vms_tix_analysis)

# check years included
unique(dcrb_vms_tix_analysis$year)

#####################################################
# filter the data using the control variables below
# years <- seq(2009,2019,1)

state_agency_code <- "C"
#state <- "CA"
target_rev <- "DCRB"
target_lbs <- "DCRB"
region_north <- c(1040, 1041, 1042)
vessel_size_category_break <- 40
winter_months <- c("November", "December", "January", "February", "March")
crab_months <- c(1:7,11:12)
removal_types <- c("COMMERCIAL (NON-EFP)", "COMMERCIAL(DIRECT SALES)", "UNKNOWN")
max_speed <- 4.11556
min_speed <- 0

#ports <- c("CCA", "ERA", "BGA","BDA", "SFA", "MRA", "MNA")
#region_south <- c(1035, 1036, 1037, 1038)
#####################################################

#####################################################

# subset the data based on above queries. CA only
dcrb_ca_vms_tix_analysis <- dcrb_vms_tix_analysis %>%
  filter(agency_code == state_agency_code) %>%
  #filter(agency_code == state_agency_code & STATE == state) %>%
  filter(TARGET_rev == target_rev | TARGET_lbs == target_lbs) %>%
  filter(removal_type_name %in% removal_types) %>%
  filter(CA_OFFSHOR != -999) %>%
  filter(DEPTH_CATM == "0-100m" | DEPTH_CATM == "100-150m") %>%
  #filter(NGDC_M <= 0 & NGDC_M >= -12000) %>% # considered adding 052220. also see "evaluate fishing depths.R". probably unnecessary because NGDC_M refers to centroid depth and we can match depths by 5km grid cell later
  filter(avg_speed_recalc <= max_speed & avg_speed_recalc >= min_speed) %>%
  #filter(is.na(in_port) == TRUE) %>% # only removes ~4000 records
  #filter(port_group_code %in% ports) %>%
  mutate(#GRID5KM_ID = as.character(GRID5KM_ID),
    Region = ifelse(CA_OFFSHOR %in% region_north,
                    "NorCA","CenCA"),
    BIA_mn_noNAs = ifelse(is.na(BIA_mn)==TRUE,0,BIA_mn),
    BIA_bm_noNAs = ifelse(is.na(BIA_bm)==TRUE,0,BIA_bm),
    #year = lubridate::year(westcoastdate_notime), # removed because it is included with the vessel_lengths step
    year_month = paste0(lubridate::year(westcoastdate_notime),"_", substr(lubridate::ymd(westcoastdate_notime),6,7)), # substr() ensures month is a 2 digit value
    month = lubridate::month(westcoastdate_notime, label=TRUE, abbr = FALSE),
    month_as_numeric = month(westcoastdate_notime),
    week_of_year = week(westcoastdate_notime),
    day_of_year = yday(westcoastdate_notime),
    season = as.character(ifelse(month %in% winter_months, "Winter", "Spring-Summer")),
    crab_year = ifelse(
      month_as_numeric >= 11, paste0(year,"_",1+year), paste0(year - 1,"_",year)
    ),
    Vessel_Size = as.character(ifelse(FINAL_LENGTH >= vessel_size_category_break, paste0(">=",vessel_size_category_break, " ft"),paste0("<",vessel_size_category_break, " ft"))),
    BIA_bm_or_mn = ifelse(BIA_bm_noNAs !=0 | BIA_mn_noNAs != 0, "Inside BIA","Outside BIA")#,
  ) %>%
  filter(westcoastdate_notime >= as.Date("2009-11-01") & westcoastdate_notime <= as.Date("2019-08-01")) %>%
  filter(month_as_numeric %in% crab_months) %>%
  dplyr::select(
    Rec_ID, VMS_RECNO, drvid, Vessel_Size,
    westcoastdate_notime, year, crab_year, year_month, month, month_as_numeric, week_of_year, day_of_year, season,
    GRID5KM_ID, BAND_25KM, BAND_50KM, CA_OFFSHOR, Region, BIA_mn_noNAs, BIA_bm_noNAs, BIA_bm_or_mn, pacfin_port_code, port_group_code, DEPTH_CATM, NGDC_M,
    TARGET_lbs, TARGET_rev, DCRB_lbs, DCRB_revenue
  )

#glimpse(dcrb_ca_vms_tix_analysis)
```

## All 10 Years
Number unique vessels  tix, geolocations, and total revenue and landings
(VMS-tracked only)

```{r fishing2, echo=FALSE}
# number of unique vessels
total_num_vessels <- length(unique(dcrb_ca_vms_tix_analysis$drvid)) #283

# number of unique tix
total_num_tix <- length(unique(dcrb_ca_vms_tix_analysis$Rec_ID)) #16182

# number of unique geolocations
total_num_locations <- length(unique(dcrb_ca_vms_tix_analysis$VMS_RECNO)) #369688

# total revenue
total_rev <- sum(dcrb_ca_vms_tix_analysis$DCRB_revenue) #6563123120

# total landings
total_landings <- sum(dcrb_ca_vms_tix_analysis$DCRB_lbs) #2234063912


data.frame(total_num_vessels,
           total_num_tix,
           total_num_locations,
           total_rev,
           total_landings) %>%
  knitr::kable()


```

## By period

Number unique vessels  tix, geolocations, and total revenue and landings by period
(VMS-tracked only)
```{r fishing3, echo=FALSE}

# summed revenue of VMS vessels 2009-14 vs 2014-18
#unique(dcrb_ca_vms_tix_analysis$crab_year)

period_summaries <- dcrb_ca_vms_tix_analysis %>%
  mutate(
    period = case_when(
      crab_year %in% c("2009_2010","2010_2011","2011_2012","2012_2013","2013_2014") ~ "pre",
      crab_year %in% c("2014_2015","2016_2017","2015_2016","2017_2018") ~ "MHW",
      TRUE ~ "post"
    ) 
  ) %>%
  group_by(period) %>%
  summarise(
    num_vessels = length(unique(drvid)),
    num_tix = length(unique(Rec_ID)),
    num_locations = length(unique(VMS_RECNO)),
    sum_DCRB_rev = sum(DCRB_revenue),
    sum_DCRB_lbs = sum(DCRB_lbs)
    
  ) %>%
  mutate(
    annual_rev = case_when(
      period == "MHW" ~ sum_DCRB_rev/4,
      period == "pre" ~ sum_DCRB_rev/5,
      period == "post" ~ sum_DCRB_rev/1
      #TRUE ~ "ERROR"
    )
  ) %>%
    arrange(
    factor(period, levels = c("pre", "MHW", "post"))
           ) 

annual_percent_rev_change_MHW_v_pre <- 100 * ( period_summaries$annual_rev[which(period_summaries$period == "MHW")] - period_summaries$annual_rev[which(period_summaries$period == "pre")] ) / period_summaries$annual_rev[which(period_summaries$period == "pre")]

period_summaries %>%
  knitr::kable()

print("annual_percent_rev_change_MHW_v_pre")
annual_percent_rev_change_MHW_v_pre


```

## By period and region

Number unique vessels  tix, geolocations, and total revenue and landings by period and region
(VMS-tracked only)
```{r fishing4, echo=FALSE}

# summed revenue of VMS vessels 2009-14 vs 2014-18
#unique(dcrb_ca_vms_tix_analysis$crab_year)

period_region_summaries <- dcrb_ca_vms_tix_analysis %>%
  mutate(
    period = case_when(
      crab_year %in% c("2009_2010","2010_2011","2011_2012","2012_2013","2013_2014") ~ "pre",
      crab_year %in% c("2014_2015","2016_2017","2015_2016","2017_2018") ~ "MHW",
      TRUE ~ "post"
    ) 
  ) %>%
  group_by(period, Region) %>%
  summarise(
    num_vessels = length(unique(drvid)),
    num_tix = length(unique(Rec_ID)),
    num_locations = length(unique(VMS_RECNO)),
    sum_DCRB_rev = sum(DCRB_revenue),
    sum_DCRB_lbs = sum(DCRB_lbs)
    
  ) %>%
  #group_by(Region) %>%
  mutate(
    annual_rev = case_when(
      period == "MHW" ~ sum_DCRB_rev/4,
      period == "pre" ~ sum_DCRB_rev/5,
      period == "post" ~ sum_DCRB_rev/1
      #TRUE ~ "ERROR"
    )
  ) %>%
    arrange(
      Region,
    factor(period, levels = c("pre", "MHW", "post"))
           ) 

annual_percent_rev_change_MHW_v_pre_region <- 100 * ( period_region_summaries$annual_rev[which(period_region_summaries$period == "MHW")] - period_region_summaries$annual_rev[which(period_region_summaries$period == "pre")] ) / period_region_summaries$annual_rev[which(period_region_summaries$period == "pre")]

period_region_summaries %>%
  knitr::kable()

print("annual_percent_rev_change_MHW_v_pre_region")
annual_percent_rev_change_MHW_v_pre_region


```

Total revenue and landings by crab year and period for all fish tix
(not yet analyzed)
