---
title: "Samhouri et al tradeoff analysis run analysis"
author: "Jameal Samhouri"
date: "11/4/2020 - 11/16/2020"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This document runs the analysis included in the Samhouri et al. tradeoff manuscript.

See pseudocode in "Full workflow for risk assessment and tradeoff analysis" google doc. https://docs.google.com/document/d/1T8S5rEOACMCeGa5efS5BnEQY4gOw-PbxmvILJvShsQo/edit

1. MAKE SCENARIOS
2. RUN SCENARIOS
3. MAKE DF'S FOR TIME SERIES OF RISK, REVENUE, AND LANDINGS
4. CALCULATE SCENARIO RANKS
5. CALCULATE CHANGES IN RISK AND REVENUE BY PERIOD AND SCENARIO
6. QUERY ENTANGLEMENT DATA
7. QUERY FISHING DATA
8. QUERY WHALE MODEL OUTPUTS

* Fig 1 data. Whale and fishing data on 5km grid, averaged by time period
* Fig 2 data. Entanglement reports, DCRB revenue by crab year. whale risk is calculated in the scenario_analysis step
* Fig 3 data. Whale risk and DCRB revenue by scenario calculated in the scenario_analysis step
* Fig 4 data. Whale risk and DCRB revenue by scenario calculated in the scenario_analysis step

## Prep libraries and file paths and functions
(no need to display messages)
```{r prep, include=FALSE}

# just getting libraries loaded and file paths sorted

library(tidyverse)
library(foreign) # read.dbf()
library(here)
library(lubridate)
library(sf)
library(viridis)
library(ggrepel)
library(magrittr)
library(ggerr)
library(scales)
library(maps)
library(rnaturalearth)
library(gridExtra)
library(ggpubr)
library(knitr)

source(here::here("User_script_local.R"))
if (user == "JS") {
  #flag.save = TRUE # option to save all outputs or not
  
  path_entanglement_file2 <- paste0(here::here(), "/tradeoffs/full_analysis_samhourietal/output_dfs/entanglement_df_annual_complete.csv")
  
  path_dcrb_vms_tix_analysis <- "~/Documents/RAIMBOW/Processed Data/VMS/vms_all_interpolated_w_grd_vlengths.RDS"
  
  path_dcrb_daily_all <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/VMS/CA_DCRB_vms_fishing_daily_2009-2019_all_vessels_regions_depths.RDS"
  
  path_season.st.date.key <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Input_Data/season_start_dates/start_dates_by_CA_region.rds"
  
  path_whales <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/5x5 Grid/Grid5km_whale.rds"
  
  path_Grid_5km_landerased <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/5x5 Grid/Grid_5km_landerased.RDATA"
  
  path_Humpback_5km_long_monthly <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Input_Data/Humpback whale data/Forney et al./Humpback_5km_long_monthly.rds"

  path_BlueWhale_5km_long_monthly <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Input_Data/Blue whale data/Overlay on 5km Grid/BlueWhale_5km_long_monthly.rds"
  
  path_grid_key <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/5x5 Grid/Grid5km_key.rds"
  
  #path_status_quo <- paste0("/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/scenario_output_dataframes/risk_5km_yr_mth_sq_all_2020-11-16.rds") # status_quo_risk_2009_2019_yr_mth_2020-06-14.rds was from extract_status_quo_scenario_summary.R
   
  # 
  # path_figures <- "/Users/jameal.samhouri/Dropbox/Projects/In progress/RAIMBOW/raimbow/tradeoffs/Management scenarios/figures"
  # 
  # path_rds <- "/Users/jameal.samhouri/Dropbox/Projects/In progress/RAIMBOW/raimbow/tradeoffs/Management scenarios"
  # 
  # path_maps <- "/Users/jameal.samhouri/Dropbox/Projects/In progress/RAIMBOW/raimbow/tradeoffs/maps"
  # 
  # path_status_quo <- paste0("/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/scenario_output_dataframes/status_quo_risk_2009_2019_yr_mth_2020-06-14.rds") # from extract_status_quo_scenario_summary.R
  # 
  # path_status_quo_small_vessels <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/scenario_output_dataframes/status_quo_risk_small_vessels_2009_2019_yr_mth_2020-06-17.rds" # from extract_status_quo_scenario_summary_small.R 
  # 
  # 
  #   
  # path_effort_shift_scenariocomparisons_annual_statewide_df_n <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/annual_statewide_effort_shift_scenariocomparisons_n_2020-05-19.rds" # from effort_shift_comparison.R
  # 
  # path_tradeoff_df_effort_shift_scenariocomparisons_n <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/tradeoff_df_effort_shift_scenariocomparisons_n_2020-05-19.rds"
  # 
  # path_scenario_table_all <- "/Users/jameal.samhouri/Dropbox/Projects/In progress/RAIMBOW/raimbow/tradeoffs/Management scenarios/scenario_table_all_2020-06-09.rds"
  # 
  # path_scenario_table_focal_scenarios <- "/Users/jameal.samhouri/Dropbox/Projects/In progress/RAIMBOW/raimbow/tradeoffs/Management scenarios/scenario_table_focal_scenarios_2020-06-12.rds"
  # 
  # path_annual_statewide_df_focal_scenarios <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/annual_statewide_df_focal_scenarios_2020-06-12.rds" 
  # 
  # path_df_tradeoff_focal_scenarios <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/df_tradeoff_focal_scenarios_2020-06-12.rds" 
  # 
  # path_annual_statewide_df_focal_scenarios_small_vessels <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/annual_statewide_df_focal_scenarios_small_vessels_2020-06-16.rds"
  # 
  # path_df_tradeoff_focal_scenarios_small_vessels <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/df_tradeoff_focal_scenarios_small_vessels_2020-06-16.rds"
  # 
  # path_prioritizr_all <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/prioritizr/prioritzr_outputs_all.rds"
  

} else if (user == "SMW") {
  NULL
} else {
  stop("Invalid user")
}


### Function to calculate crab year
func_crab_year <- function(z) {
  # z: character; year_month
  
  stopifnot(require(lubridate))
  z.yr <- as.numeric(substr(z, 1, 4))
  z.mon <- as.numeric(substr(z, 6, 7))
  
  ifelse(
    z.mon %in% 11:12, 
    paste(z.yr, z.yr + 1, sep = "_"), paste(z.yr - 1, z.yr, sep = "_")
  )
}

```


# 1. MAKE SCENARIOS
```{r make_scenarios, echo=FALSE}

source(paste0(here::here(), "/tradeoffs/Management scenarios/make_scenarios_table.R"))
# note season.st.key argument has to be defined separately from above

# scenario_table[1,]
# scenario_table_edr[1,]

# separate scenario tables need to be bound together.
# number the scenarios for easier reference. sq indicates status quo; redundant indicates scenario already represented and can be ignored; alt indicates scenario uses alt set of assumptions about what happens to effort (either remove or not removed and/or 10% v 100% of effort redistributed)
scenario_table_all <- scenario_table %>%
  bind_rows(scenario_table_edr) %>%
  mutate(
    scenario_number = c(
      "sq","1","2","3","6","8","4alt","7alt","9alt","5",
      "10","11","sq_redundant","1_redundant","2_redundant",
      "3_redundant","6_redundant","8_redundant","4","7","9",
      "5alt","10alt","11alt","12","13","14","15","16","17",
      "18","13alt","14alt","15alt","16alt","17alt","18alt"
    )
  )

# scenario_table_all[3,] == scenario_table_all[15,]

# convert factors to characters
scenario_table_all$delay_scenario <- as.character(scenario_table_all$delay_scenario)
scenario_table_all$delay_scenario <- as.character(scenario_table_all$delay_scenario)
scenario_table_all$closure_scenario <- as.character(scenario_table_all$closure_scenario)
scenario_table_all$closure_scenario <- as.character(scenario_table_all$closure_scenario)

#glimpse(scenario_table_all)

scenario_table_all %>% knitr::kable()

write_csv(scenario_table_all,here::here(
  "tradeoffs",
  "Management scenarios",
  "scenario_table_all.csv"
          )
)

```

# 2. RUN SCENARIOS

## Loop through scenarios of interest and create a list of output df's
```{r run_scenarios1, echo=FALSE}

# note this chunk was adapted from scenario_analysis_all_vessels.R

### 1) Loop through scenarios of interest and create a list of output df's

# 1a) load function to shift fishing effort and daily fishing data
source(here::here(
  "tradeoffs",
  "Management scenarios",
  "Mgmt_scenarios_shift_effort.R")
)
x.orig <- read_rds(path_dcrb_daily_all)

# 1b) read in season start date key
# grab helper function. https://stackoverflow.com/questions/6364783/capitalize-the-first-letter-of-both-words-in-a-two-word-string
simpleCap <- function(x) {
  s <- sapply(strsplit(x, " "), function(i) i[[1]])
  paste(toupper(substring(s, 1, 1)), substring(s, 2),
        sep = "", collapse = " ")
}
season.st.date.key <- read_rds(path_season.st.date.key) %>% 
  mutate(crab_year = gsub("-", "_", .data$crab_season), 
         Region = unname(sapply(CA_region, simpleCap))) %>% 
  select(crab_year, Region, start_of_season_oneperc)

# 1c) shift effort for scenarios of interest

# drop redundant scenarios

scenario_table_all_trim <- scenario_table_all %>%
  filter(!grepl('redundant', scenario_number))

scenario_table_all_trim %>% select(scenario_number, scenario_df_name) %>% knitr::kable()

start.time <- Sys.time()

scenario.output.list <- lapply(1:nrow(scenario_table_all_trim), function(i, scenario_table_input) { # for testing. nrow(scenario_table_all_trim[1:2,])
  print(i)
  #browser()
  
  # i=1 # testing. no longer breaks because "At least one of delay.date or closure.date must not be NULL"
  # i=2 # testing. no longer breaks because when switch() is used and does not return NULL, it returns nothing
  
  scenario.output.df <- effort_mgmt(
    x = x.orig,
    
    season.st.key = season.st.date.key,
    
    preseason.days = scenario_table_all_trim$preseason.days[i],
    
    season.st.backstop = if (scenario_table_all_trim$season.st.backstop[i] == "NULL") {
      NULL
      }
    else {
      as.Date(scenario_table_all_trim$season.st.backstop[i])
      },
    
    early.data.method = scenario_table_all_trim$early.data.method[i], 
    
    delay.date = if (scenario_table_all_trim$delay.date[i] == "NULL") {
      NULL
    } else {
      as.Date(scenario_table_all_trim$delay.date[i])
    },
    
    delay.region = if (scenario_table_all_trim$delay.region[i] == "NULL") NULL else scenario_table_all_trim$delay.region[i],
    
    delay.method = if (scenario_table_all_trim$delay.method[i] == "NULL") NULL else scenario_table_all_trim$delay.method[i],
    
    delay.method.fidelity = scenario_table_all_trim$delay.method.fidelity[i],
    
    closure.date = if (scenario_table_all_trim$closure.date[i] == "NULL") NULL else as.Date(scenario_table_all_trim$closure.date[i]),
    
    closure.region = if (scenario_table_all_trim$closure.region[i] == "NULL") NULL else scenario_table_all_trim$closure.region[i],
    
    closure.method = if (scenario_table_all_trim$closure.method[i] == "NULL") NULL else scenario_table_all_trim$closure.method[i],
    
    closure.redist.percent = scenario_table_all_trim$closure.redist.percent[i],
    
    depth.shallow = if (scenario_table_all_trim$depth.shallow[i] == "NULL") NULL else as.numeric(scenario_table_all_trim$depth.shallow[i]), 
    
    depth.deep = if (scenario_table_all_trim$depth.deep[i] == "NULL") NULL else as.numeric(scenario_table_all_trim$depth.deep[i]),
    
    reduction.before.date = if (scenario_table_all_trim$reduction.before.date[i] == "NULL") NULL else scenario_table_all_trim$reduction.before.date[i],
    
    reduction.before.percent = 50,
    
    reduction.before.region = if (scenario_table_all_trim$reduction.before.region[i] == "NULL") NULL else scenario_table_all_trim$reduction.before.region[i],
    
    reduction.after.date = if (scenario_table_all_trim$reduction.after.date[i] == "NULL") NULL else as.Date(scenario_table_all_trim$reduction.after.date[i]),
    
    reduction.after.percent = 50,
    
    reduction.after.region = if (scenario_table_all_trim$reduction.after.region[i] == "NULL") NULL else scenario_table_all_trim$reduction.after.region[i],
    
    reduction.after.redist = if (scenario_table_all_trim$reduction.after.redist[i] == "FALSE") FALSE else TRUE,
    
    reduction.after.redist.percent = scenario_table_all_trim$reduction.after.redist.percent[i]
    
  )
  
}, scenario_table_input = scenario_table_all_trim)

Sys.time() - start.time

# check to see how much nobase effort is dropped relative to total under sq

scenario.output.list[[1]] %>% 
  filter(date_past_region_end != TRUE) %>% 
  filter(nobase == FALSE) %>%
  group_by(crab_year, Region, year_month) %>% 
            summarise(num_records = n(), 
                      Num_DCRB_VMS_pings_sum = sum(Num_DCRB_VMS_pings), 
                      .groups = "drop") %>%
  filter(year_month == "2019_05")

# JS. update directory as needed
save.image(paste0("/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/scenario_output_dataframes/scenario_output_",today(),".RData"))

```


## Calculate and summarize risk for multiple scenarios, including status quo
```{r run_scenarios2, echo=FALSE}

# note this chunk was adapted from scenario_analysis_all_vessels.R

### 2) Calculate and summarize risk for multiple scenarios, including status quo

# 2a) grab shifted effort fishing data if not running full code above
# JS. update directory as needed 
# load(paste0("/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/scenario_output_dataframes/scenario_output_","2020-11-11",".RData"))

# 2b) grab whale data
x.whale <-read_rds(path_whales)

# 2c) Load and prep grid cell - area key
load(path_Grid_5km_landerased) # this is called grid.5km.lno
area.key <- grid.5km.lno %>% 
  st_drop_geometry() %>% 
  select(GRID5KM_ID, area_km_lno) %>% 
  distinct()

### 2d) grab fishing metrics range to make sure normalizations are conducted properly (deals with normalizing within/across years and regions issue)
CA_fishing_metrics_range_2009_2019 <- read_rds(here:: here(
  "grid-prep",
  "CA_fishing_metrics_range_2009_2019.rds")
)

### 2e) Calculate and summarize risk
source(here:: here(
"tradeoffs",
"Management scenarios",
"Mgmt_scenarios_risk.R")
)

# will need to re-run all of the above and below for sm and lg vessels

start.time <- Sys.time()

risk_out_list <- lapply(1:nrow(scenario_table_all_trim), function(i, scenario_table_input) { # for testing. nrow(scenario_table[1:3,])
  print(paste("Summarizing risk for Scenario", i))
  #browser()
  
  risk_out <- risk_mgmt(
    x = scenario.output.list[[i]] %>% filter(date_past_region_end != TRUE) %>% filter(nobase == FALSE), # can add filter for nobase == FALSE here too
    x.col = Num_DCRB_VMS_pings, 
    y = x.whale,
    risk.unit = "dens", 
    area.key = area.key,
    scale.list = CA_fishing_metrics_range_2009_2019, 
    ym.min = "2009_11", 
    ym.max = "2019_07"
  )
}, scenario_table_input = scenario_table_all_trim
)

Sys.time() - start.time
# Time difference of 19 secs

### save status quo scenario as its own df ###
#scenario_table_all_trim$scenario_df_name # use scenario_number == sq, which is 1

glimpse(risk_out_list[[1]])
risk_5km_yr_mth_sq_all <- risk_out_list[[1]]

# JS. update directory as needed
write_rds(risk_5km_yr_mth_sq_all, paste0("/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/scenario_output_dataframes/risk_5km_yr_mth_sq_all_",today(),".rds"))

# 2f) summarize risk by region

start.time <- Sys.time()
risk_out_summ_list <- lapply(1:nrow(scenario_table_all_trim), function(i, scenario_table_input) { # for testing. nrow(scenario_table[1:3,])
  print(paste("Summarizing risk for Scenario", i))
  #browser()
  
  risk_out_summ <- risk_mgmt_summ(
    x = risk_out_list[[i]], 
    summary.level = "Region"
  )
}, scenario_table_input = scenario_table_all_trim
)

Sys.time() - start.time
#Time difference of 5 secs

glimpse(risk_out_summ_list[[1]])

# JS. update directory as needed
save.image(paste0("/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/scenario_output_dataframes/scenario_output_risk_",today(),".RData"))

```


## Out of season records

Here is a table summarizing the number of records that got dropped because they occurred outside of the scheduled season end dates.

```{r run_scenarios3, echo=FALSE}
# track out of season records
season_keep_drop_df <- inseason_records <- outseason_records <- percent_outseason <- c()
for(i in 1:nrow(scenario_table_all_trim)){
  
inseason_records[i] <- nrow(scenario.output.list[[i]] %>% filter(date_past_region_end != TRUE))
outseason_records[i] <- nrow(scenario.output.list[[i]] %>% filter(date_past_region_end == TRUE))
percent_outseason[i] <- 100*(outseason_records[i]/nrow(scenario.output.list[[i]]))
season_keep_drop_df <- bind_rows(season_keep_drop_df,
                                 c(inseason = inseason_records[i],outseason = outseason_records[i],
                                   percent_outseason = percent_outseason[i])
)
}

season_keep_drop_df %>% knitr::kable()
```


# 3. MAKE DATA FRAMES

MAKE DF'S FOR TIME SERIES OF RISK, REVENUE, AND LANDINGS (annual values), AND FOR MAPS
+ status quo as separate df, all scenarios in one df, main text and supplement df's
+ repeat for sm vessels [ not yet coded ]

## Make annual and tradeoff df's, all scenarios.
```{r make_dfs1, echo=TRUE}

# note this chunk was adapted from scenario_analysis_all_vessels.R

### 1) make annual and tradeoff df's.

# 1a) # if needed, load data
# JS. update directory as needed
# load(paste0("/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/scenario_output_dataframes/scenario_output_risk_","2020-11-11",".RData"))

# 1b) make data frames
# could subset or map to region for regional summaries here
source(here::here(
"tradeoffs",
"Management scenarios",
"make_tradeoff_dataframes_function.R")
)

# note that this function writes rds files specific to JS directories (sorry!)
# also note that write_rds should be modified if running analysis for small or large vessels (~lines 119, 224)
# also also note that if sq scenario name changes  objects need to be redefined in the function (sorry again!)

load("/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/scenario_output_dataframes/scenario_output_risk_2020-11-20.RData")

start.time <- Sys.time()
tradeoff_df_function(
  risk_list = risk_out_summ_list,
  scenario_names_table = scenario_table_all_trim,
  annual_statewide_df_name = "annual_statewide_df", # this ensures df is available in the globalEnv after running function
  df_tradeoff_name = "df_tradeoff" # this ensures df is available in the globalEnv after running function
  )
Sys.time() - start.time
#Time difference of 0.24 secs

# is risk actually greater for whales under some scenarios? yes.
dim(df_tradeoff)
length(which(df_tradeoff$relative_hump_risk < 0)) # 28 out of 310
length(which(df_tradeoff$relative_blwh_risk < 0)) # 41 out of 310
length(which(df_tradeoff$relative_both_risk < 0)) # 34 out of 310

# is $ or pounds actually greater for the fishery under some scenarios? yes
length(which(df_tradeoff$relative_dollars > 100)) # 8 out of 310
length(which(df_tradeoff$relative_pounds > 100)) # 8 out of 310

```

## Make subset of main text vs supplement scenarios.
```{r make_dfs2, echo=FALSE}

# note this chunk was adapted from scenario_analysis_all_vessels.R

# give scenarios pretty names, add pre_post and year columns
scenario_names_key <- read_csv(here::here(
  "tradeoffs",
  "full_analysis_samhourietal",
  "scenario_names_key.csv"
))

annual_statewide_df %<>%
  left_join(
    scenario_names_key
  ) %<>%
  mutate(
  plotting_year = as.numeric(substr(crab_year,6,9)),
  pre_post = ifelse(plotting_year >=2015 & plotting_year <2019, "2014-2018",
                    ifelse(plotting_year <2015, "2009-2014", "2018-2019")
                    )
  )

scenario_table_all_trim %<>%
  left_join(
    scenario_names_key
  )

df_tradeoff %<>%
  left_join(
    scenario_names_key
  ) %<>%
  mutate(
  plotting_year = as.numeric(substr(crab_year,6,9)),
  pre_post = ifelse(plotting_year >=2015 & plotting_year <2019, "2014-2018",
                    ifelse(plotting_year <2015, "2009-2014", "2018-2019")
                    )
  )

# join tradeoff df to annual df to include relative changes
annual_statewide_df %<>%
  left_join(df_tradeoff)

# 1a) Subset to focal scenarios for main text

annual_statewide_df_focal_scenarios <- annual_statewide_df %>%
  filter(!grepl('alt', scenario_number))

# JS. update directory as needed
write_rds(annual_statewide_df_focal_scenarios, paste0("/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/annual_statewide_df_focal_scenarios_",today(),".rds"))

scenario_table_focal_scenarios <- scenario_table_all_trim %>%
  filter(!grepl('alt', scenario_number))

# JS. update directory as needed
write_rds(scenario_table_focal_scenarios, here::here(
  "tradeoffs",
  "Management scenarios",
  paste0("scenario_table_focal_scenarios_",today(),".rds"))
)

df_tradeoff_focal_scenarios <- df_tradeoff  %>%
  filter(!grepl('alt', scenario_number)) %>%
  mutate(
     scenario_type = factor(
      case_when(
      pretty_scenario_names == "Status Quo" ~ "Status Quo",
      stringr::str_detect(pretty_scenario_names, 'Early Closure')  & stringr::str_detect(pretty_scenario_names, 'Delay') ~ "Delay and Early Closure",
      stringr::str_detect(pretty_scenario_names, 'Early Closure')  & stringr::str_detect(pretty_scenario_names, 'Delay', negate = TRUE) ~ "Early Closure",
      stringr::str_detect(pretty_scenario_names, 'Early Closure', negate = TRUE)  & stringr::str_detect(pretty_scenario_names, 'Delay') ~ "Delayed Opening",
      stringr::str_detect(pretty_scenario_names, 'Depth Restriction') & stringr::str_detect(pretty_scenario_names, 'Effort Reduction') ~ "Depth and Gear Restrictions",
      stringr::str_detect(pretty_scenario_names, 'Depth Restriction') & stringr::str_detect(pretty_scenario_names, 'Effort Reduction', negate = TRUE) ~ "Depth Restriction",
      stringr::str_detect(pretty_scenario_names, 'Depth Restriction', negate = TRUE) & stringr::str_detect(pretty_scenario_names, 'Effort Reduction') ~ "Gear Restriction",
      TRUE ~ "Delay and Early Closure"
    ),
    levels = c("Status Quo", "Depth Restriction", "Gear Restriction",
               "Depth and Gear Restrictions", "Delayed Opening", "Early Closure",
               "Delay and Early Closure")),
    spatial_domain = factor(
      case_when(
      stringr::str_detect(pretty_scenario_names, 'Statewide') & stringr::str_detect(pretty_scenario_names, 'CenCA') ~ "Statewide/CenCA",
      stringr::str_detect(pretty_scenario_names, 'BIA') & stringr::str_detect(pretty_scenario_names, 'CenCA') ~ "CenCA/BIA",
      stringr::str_detect(pretty_scenario_names, 'Statewide') & stringr::str_detect(pretty_scenario_names, 'BIA') ~ "Statewide/BIA",
      stringr::str_detect(pretty_scenario_names, 'Statewide') ~ "Statewide",
      stringr::str_detect(pretty_scenario_names, 'CenCA') ~ "Central California",
      stringr::str_detect(pretty_scenario_names, 'BIA') ~ "BIA",
      TRUE ~ "Status Quo"
    ),
    levels= c("Status Quo","BIA","Central California","Statewide","Statewide/CenCA", "CenCA/BIA","Statewide/BIA")
  )
  )

# JS. update directory as needed
write_rds(df_tradeoff_focal_scenarios, paste0("/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/df_tradeoff_focal_scenarios_",today(),".rds"))

# 1b) Subset to focal scenarios for supplement

annual_statewide_df_supplement <- annual_statewide_df %>%
  filter(grepl('alt', scenario_number) | grepl('sq', scenario_number))

# JS. update directory as needed
write_rds(annual_statewide_df_supplement, paste0("/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/annual_statewide_df_supplement_",today(),".rds"))

scenario_table_supplement <- scenario_table_all_trim %>%
  filter(grepl('alt', scenario_number) | grepl('sq', scenario_number))

# JS. update directory as needed
write_rds(scenario_table_supplement, here::here(
  "tradeoffs",
  "Management scenarios",
  paste0("scenario_table_supplement_",today(),".rds"))
)

df_tradeoff_supplement <- df_tradeoff  %>%
  filter(grepl('alt', scenario_number) | grepl('sq', scenario_number)) %>%
  mutate(
     scenario_type = factor(
      case_when(
      pretty_scenario_names == "Status Quo" ~ "Status Quo",
      stringr::str_detect(pretty_scenario_names, 'Early Closure')  & stringr::str_detect(pretty_scenario_names, 'Delay') ~ "Delay and Early Closure",
      stringr::str_detect(pretty_scenario_names, 'Early Closure')  & stringr::str_detect(pretty_scenario_names, 'Delay', negate = TRUE) ~ "Early Closure",
      stringr::str_detect(pretty_scenario_names, 'Early Closure', negate = TRUE)  & stringr::str_detect(pretty_scenario_names, 'Delay') ~ "Delayed Opening",
      stringr::str_detect(pretty_scenario_names, 'Depth Restriction') & stringr::str_detect(pretty_scenario_names, 'Effort Reduction') ~ "Depth and Gear Restrictions",
      stringr::str_detect(pretty_scenario_names, 'Depth Restriction') & stringr::str_detect(pretty_scenario_names, 'Effort Reduction', negate = TRUE) ~ "Depth Restriction",
      stringr::str_detect(pretty_scenario_names, 'Depth Restriction', negate = TRUE) & stringr::str_detect(pretty_scenario_names, 'Effort Reduction') ~ "Gear Restriction",
      TRUE ~ "Delay and Early Closure"
    ),
    levels = c("Status Quo", "Depth Restriction", "Gear Restriction",
               "Depth and Gear Restrictions", "Delayed Opening", "Early Closure",
               "Delay and Early Closure")),
    spatial_domain = factor(
      case_when(
      stringr::str_detect(pretty_scenario_names, 'Statewide') & stringr::str_detect(pretty_scenario_names, 'CenCA') ~ "Statewide/CenCA",
      stringr::str_detect(pretty_scenario_names, 'BIA') & stringr::str_detect(pretty_scenario_names, 'CenCA') ~ "CenCA/BIA",
      stringr::str_detect(pretty_scenario_names, 'Statewide') & stringr::str_detect(pretty_scenario_names, 'BIA') ~ "Statewide/BIA",
      stringr::str_detect(pretty_scenario_names, 'Statewide') ~ "Statewide",
      stringr::str_detect(pretty_scenario_names, 'CenCA') ~ "Central California",
      stringr::str_detect(pretty_scenario_names, 'BIA') ~ "BIA",
      TRUE ~ "Status Quo"
    ),
        levels= c("Status Quo","BIA","Central California","Statewide","Statewide/CenCA", "CenCA/BIA","Statewide/BIA")
  )
  )

# JS. update directory as needed
write_rds(df_tradeoff_supplement, paste0("/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/df_tradeoff_supplement_",today(),".rds"))

# 1c) Subset to status quo scenario only

annual_statewide_df_sq <- annual_statewide_df %>%
  filter(grepl('sq', scenario_number))

# JS. update directory as needed
write_rds(annual_statewide_df_sq, paste0("/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/annual_statewide_df_sq_",today(),".rds"))

scenario_table_sq <- scenario_table_all_trim %>%
  filter(grepl('sq', scenario_number))

# JS. update directory as needed
write_rds(scenario_table_sq, here::here(
  "tradeoffs",
  "Management scenarios",
  paste0("scenario_table_sq_",today(),".rds"))
)

df_tradeoff_sq <- df_tradeoff  %>%
  filter(grepl('sq', scenario_number))

# JS. update directory as needed
write_rds(df_tradeoff_sq, paste0("/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/df_tradeoff_sq_",today(),".rds"))

```

## Make df for maps

This chunk makes "Data frame for Figure 1 Samhouri et al.csv", which gets handed off to Blake
```{r make_dfs3, echo=FALSE}

#1) read in data
#1a) fishing data, positive fishing grid cells and year_months only

# fishing data. whale predictions here only pertain to year_months in which the DCRB fishery was open and grid cells in which fishing was nonzero
risk_out_sq <- risk_5km_yr_mth_sq_all
#risk_out_sq <- read_rds(path_status_quo)
#glimpse(risk_out_sq)

# sanity check. QA/QC and make sure all of the year_month - GRID5KM_ID pairs in risk_out_sq are unique
# glimpse(risk_out_sq %>% dplyr::select(GRID5KM_ID, year_month) %>% distinct())
# with(risk_out_sq, table(year_month, GRID5KM_ID))

### Function to calculate crab year
func_crab_year <- function(z) {
  # z: character; year_month
  
  stopifnot(require(lubridate))
  z.yr <- as.numeric(substr(z, 1, 4))
  z.mon <- as.numeric(substr(z, 6, 7))
  
  ifelse(
    z.mon %in% 11:12, 
    paste(z.yr, z.yr + 1, sep = "_"), paste(z.yr - 1, z.yr, sep = "_")
  )
}

risk_out_sq_complete <-  risk_out_sq %>%
  select(year_month, GRID5KM_ID, Num_DCRB_VMS_pings) %>%
  complete(year_month, nesting(GRID5KM_ID), fill = list(Num_DCRB_VMS_pings = 0)) 

# table(risk_out_sq_complete$year_month) # check that all year_months have same number of grid cells

#1b) whale data, fishing grid cells only but includes year_months without fishing

# x.whale <-read_rds(path_whales) # this is whale predictions within DCRB fishing cells, but in contrast to risk_out_sq it includes predictions in DCRB fishing cells for all yr_mths and fishing grid cells, not just months and grid cells in which DCRB fishing occurred
# glimpse(x.whale)
# table(x.whale$year_month) # check that all year_months have same number of grid cells

# check to see if all grid cells appear in both risk_out_sq_complete and x.whale
# all(risk_out_sq_complete$GRID5KM_ID %in% x.whale$GRID5KM_ID)
# which(risk_out_sq_complete$GRID5KM_ID %in% x.whale$GRID5KM_ID == FALSE)
# which(x.whale$GRID5KM_ID %in% risk_out_sq_complete$GRID5KM_ID == FALSE)
whale.id <- sort(unique(x.whale$GRID5KM_ID))
fish.id <- sort(unique(risk_out_sq_complete$GRID5KM_ID))
# whale.id[which(whale.id %in% fish.id == FALSE)]
# fish.id[which(fish.id %in% whale.id == FALSE)]

# 2 grid cells are problematic. they are on the edge of the nonNA fishing cells. ignore and drop them for now
cells_to_drop <- c(whale.id[which(whale.id %in% fish.id == FALSE)], fish.id[which(fish.id %in% whale.id == FALSE)])


#1c) humpbacks, all GRID5KM_ID, all year_months
hump.all <- read_rds(path_Humpback_5km_long_monthly) %>% # hump predictions in all 5km cells, all year_months 
  mutate(
    year_month = paste0(year(date), "_", format.Date(date, "%m"))
  )
# glimpse(hump.all)
#table(hump.all$year_month)

#1d) blues, all GRID5KM_ID, all year_months
blue.all <- read_rds(path_BlueWhale_5km_long_monthly) %>% # blue predictions in all 5km cells. output from https://github.com/jameals/raimbow/blob/master/blue_whale/BW_preds_aggr.Rmd
  mutate(
    year_month = paste0(year(date), "_", format.Date(date, "%m"))
  )
# glimpse(blue.all)
# table(blue.all$year_month)

#2) group_by GRID5KM_ID, summarise above data by 2009-14, 2014-18, 2018-19, excluding Aug-Oct 

sq_DCRBcells_maps_df <- x.whale %>% # ensures that we have whale density predictions for all grid cells and year months, by=c("year_month", "GRID5KM_ID")
  full_join(risk_out_sq_complete, by = c("GRID5KM_ID", "year_month")) %>% 
  filter(!(is.na(Num_DCRB_VMS_pings)) | year_month %in% unique(risk_out_sq_complete$year_month)) %>% # if pings != NA or  yr_mth is in df_fishing, keep
  replace_na(list(Num_DCRB_VMS_pings=0)) %>%
  mutate(
    year = as.numeric(substr(year_month, 1,4)),
    month = as.numeric(substr(year_month,6,7)),
    crab_year = func_crab_year(year_month),
    plotting_year = as.numeric(substr(crab_year,6,9)),
    time_period = ifelse(plotting_year <2015, "2009-2014",
                         ifelse(
                           plotting_year > 2017, "2018-2019",
                           "2014-2018"
                         )),
    Humpback_dens_mean = Humpback_abund_mean / area_km_lno,
    Humpback_dens_se = Humpback_abund_se / area_km_lno
      ) %>%
  filter(!(GRID5KM_ID %in% cells_to_drop)) %>%
#glimpse(sq_DCRBcells_maps_df)
  select(year_month, crab_year, plotting_year, time_period, GRID5KM_ID, Num_DCRB_VMS_pings, Humpback_dens_mean, Blue_occurrence_mean) %>%
  group_by(GRID5KM_ID, time_period) %>%
  summarise(quantile = scales::percent(c(0.025, 0.25, 0.5, 0.75, 0.975)),
            Num_DCRB_VMS_pings = quantile(Num_DCRB_VMS_pings, c(0.025, 0.25, 0.5, 0.75, 0.975), na.rm=TRUE),
            Humpback_dens = quantile(Humpback_dens_mean, c(0.025, 0.25, 0.5, 0.75, 0.975), na.rm=TRUE),
            Blue_occurrence = quantile(Blue_occurrence_mean, c(0.025, 0.25, 0.5, 0.75, 0.975), na.rm=TRUE),
    .groups = "drop"
    ) %>%
  filter(quantile == "50%") %>% 
  arrange(GRID5KM_ID, time_period) %>%
  group_by(GRID5KM_ID) %>%
  mutate(
    diff_Num_DCRB_VMS_pings = Num_DCRB_VMS_pings - lag(Num_DCRB_VMS_pings, default = 0),
    diff_Humpback_dens = Humpback_dens - lag(Humpback_dens, default = 0),
    diff_Blue_occurrence = Blue_occurrence - lag(Blue_occurrence, default = 0)
  )

# set all diffs for 2009-2014 to NA because they are bogus
sq_DCRBcells_maps_df$diff_Num_DCRB_VMS_pings[which(sq_DCRBcells_maps_df$time_period == "2009-2014")] <- NA
sq_DCRBcells_maps_df$diff_Humpback_dens[which(sq_DCRBcells_maps_df$time_period == "2009-2014")] <- NA
sq_DCRBcells_maps_df$diff_Blue_occurrence[which(sq_DCRBcells_maps_df$time_period == "2009-2014")] <- NA
  

# glimpse(sq_DCRBcells_maps_df)
# View(sq_DCRBcells_maps_df)


# # check that each grid cell is represented in each year month
# table(sq_DCRBcells_maps_df$year_month) # there is one less grid cell in this df than in risk_out_sq_complete

# all(sq_DCRBcells_maps_df$GRID5KM_ID %in% whale.id)
# glimpse(sq_DCRBcells_maps_df[which(sq_DCRBcells_maps_df$GRID5KM_ID %in% whale.id == FALSE),])
# unique(sq_DCRBcells_maps_df[which(sq_DCRBcells_maps_df$GRID5KM_ID %in% whale.id == FALSE),'GRID5KM_ID'])
# all(sq_DCRBcells_maps_df$GRID5KM_ID %in% fish.id)
# glimpse(sq_DCRBcells_maps_df[which(sq_DCRBcells_maps_df$GRID5KM_ID %in% fish.id == FALSE),])
# unique(sq_DCRBcells_maps_df[which(sq_DCRBcells_maps_df$GRID5KM_ID %in% fish.id == FALSE),'GRID5KM_ID'])

# table(sq_DCRBcells_maps_df$GRID5KM_ID) 
# all(sq_DCRBcells_maps_df$GRID5KM_ID %in% risk_out_sq_complete$GRID5KM_ID)
# which(sq_DCRBcells_maps_df$GRID5KM_ID %in% risk_out_sq_complete$GRID5KM_ID == FALSE)
# glimpse(sq_DCRBcells_maps_df[587,]) # GRID5KM_ID is 65500
# which(risk_out_sq_complete$GRID5KM_ID == 65500)
# 
# all(risk_out_sq_complete$GRID5KM_ID %in% x.fish$GRID5KM_ID)
# all(x.whale$GRID5KM_ID %in% x.fish$GRID5KM_ID)
# all(risk_out_sq_complete$GRID5KM_ID %in% x.whale$GRID5KM_ID)
# which(risk_out_sq_complete$GRID5KM_ID %in% x.whale$GRID5KM_ID == FALSE)
# which(x.whale$GRID5KM_ID %in% risk_out_sq_complete$GRID5KM_ID == FALSE)

# # check that Aug-Oct excluded. yes
# unique(sq_DCRBcells_maps_df$year_month)
# 
# # check for NAs for pings. none
# which(is.na(sq_DCRBcells_maps_df$Num_DCRB_VMS_pings)) # none
# 
# # check for NAs for whales. some NAs but they are the same NAs as in x.whale (where blue and hump predictions do not overlap). ok to proceed
# length(which(is.na(sq_DCRBcells_maps_df$Humpback_abund_mean)))
# length(which(is.na(x.whale$Humpback_abund_mean)))
# unique(sq_DCRBcells_maps_df$GRID5KM_ID[which(is.na(sq_DCRBcells_maps_df$Humpback_dens))])
# unique(x.whale$GRID5KM_ID[which(is.na(x.whale$Humpback_abund_mean))])
# unique(sq_DCRBcells_maps_df$GRID5KM_ID[which(is.na(sq_DCRBcells_maps_df$Blue_occurrence))])
# unique(x.whale$GRID5KM_ID[which(is.na(x.whale$Blue_occurrence_mean))])


  #dplyr::filter(month < 8 | month > 10) %>% #only including crab season months
  
 
  

# sq_maps_df <- risk_out_sq %>%
#   full_join(x.whale) %>% #, by=c("year_month", "GRID5KM_ID")
#   mutate(
#     Humpback_dens_mean = Humpback_abund_mean / area_km_lno,
#     Humpback_dens_se = Humpback_abund_se / area_km_lno,
#     month = as.numeric(substr(year_month,6,7)),
#   ) %>%
#   dplyr::filter(month < 8 | month > 10) %>% #only including crab season months
#   select(year_month, GRID5KM_ID, Num_DCRB_VMS_pings, Humpback_dens_mean, Blue_occurrence_mean) %>%
#   group_by(GRID5KM_ID) %>%
#   summarise(quantile = scales::percent(c(0.025, 0.25, 0.5, 0.75, 0.975)),
#             Num_DCRB_VMS_pings = quantile(Num_DCRB_VMS_pings, c(0.025, 0.25, 0.5, 0.75, 0.975), na.rm=TRUE),
#             Humpback_dens = quantile(Humpback_dens_mean, c(0.025, 0.25, 0.5, 0.75, 0.975), na.rm=TRUE),
#             Blue_occurrence = quantile(Blue_occurrence_mean, c(0.025, 0.25, 0.5, 0.75, 0.975), na.rm=TRUE),
#     .groups = "drop"
#     )

#3) make df's sf objects and prep for plotting

#grd <- sf::read_sf("/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/5x5 Grid/regions_master_final_lamb.shp")

# load(path_Grid_5km_landerased) # this is called grid.5km.lno
# area.key <- grid.5km.lno %>% 
#   st_drop_geometry() %>% 
#   select(GRID5KM_ID, area_km_lno) %>% 
#   distinct()

# background map
# states_grid.5km.lno <- st_as_sf(map("state", fill = TRUE, plot=FALSE)) %>%
#   st_transform(st_crs(grid.5km.lno))

rmap.base <- c(
  st_geometry(ne_states(country = "United States of America", returnclass = "sf")),   ne_countries(scale = 10, continent = "North America", returnclass = "sf") %>%
    filter(admin %in% c("Canada", "Mexico")) %>%
    st_geometry() %>%
    st_transform(st_crs(grid.5km.lno))
)

#states <- st_as_sf(map("state", fill = TRUE, plot=FALSE)) %>% 
#  st_transform(st_crs(grd))

# join grd to median whales and pings. takes a moment
sq_maps_grid <- grid.5km.lno %>%
  left_join(sq_DCRBcells_maps_df) %>%
  filter(!(is.na(Num_DCRB_VMS_pings)))

# join grd to median whales and pings. takes a moment
# sq_maps_grd_hump <- grd %>%
#   left_join(sq_maps_df %>% filter(quantile == "50%")) %>%
#   filter(is.na(Humpback_dens) == FALSE)
  # filter(year == 2010) %>%
  # filter(is.na(H_Avg_Abund)) %>%
  # mutate(
  #   empty_cell = 1
  # )



# x.whale_mean <- x.whale %>%
#   group_by(GRID5KM_ID) %>%
#   summarise(
#     Humpback_abund_mean = mean(Humpback_abund_mean)
#   )
# 
# x.whale_grid <- grid.5km.lno %>%
#   left_join(x.whale_mean)
# 
# hump.all_mean <- hump.all %>%
#   group_by(GRID5KM_ID) %>%
#   summarise(
#     Humpback_dens_mean = mean(Humpback_dens_mean)
#   )
# 
# hump.all_grid <- grid.5km.lno %>%
#   left_join(hump.all_mean)


sq_bbox <- st_bbox(sq_maps_grid %>% 
                      st_as_sf()
)

sq_maps_grid_noNA <- sq_maps_grid %>%
  filter(!(is.na(time_period)))

# JS. update directory as needed
write_csv(sq_maps_grid_noNA,paste0("/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/", "Data frame for Figure 1 Samhouri et al.csv"))


```

## Make df for time series

This chunk makes a data frame for Figure 2 time series Samhouri et al
```{r make_dfs4, echo=FALSE}

#risk_out_sq <- risk_5km_yr_mth_sq_all
#risk_out_sq <- read_rds(path_status_quo)

# summarise cumulative sum of $ and risk by crab_year
annual_risk_out_sq <- risk_out_sq %>%
  mutate(
  #month = as.numeric(substr(year_month,6,7)),
  crab_year = func_crab_year(year_month),
  plotting_year = as.numeric(substr(crab_year,6,9))
  ) %>%
  group_by(crab_year, plotting_year, year_month) %>%
  summarise(
    sum_DCRB_rev = sum(DCRB_rev, na.rm=TRUE)/10^5,
    sum_n_risk_humpback = sum(n_risk_humpback, na.rm=TRUE),
    sum_n_risk_blue = sum(n_risk_blue, na.rm=TRUE),
    
    .groups = "drop"
    ) %>%
  group_by(crab_year, plotting_year) %>%
  summarise(
    quantile = scales::percent(c(0.025, 0.25, 0.5, 0.75, 0.975)),
    DCRB_rev = quantile(sum_DCRB_rev, c(0.025, 0.25, 0.5, 0.75, 0.975), na.rm=TRUE),
    n_risk_humpback = quantile(sum_n_risk_humpback, c(0.025, 0.25, 0.5, 0.75, 0.975), na.rm=TRUE),
    n_risk_blue = quantile(sum_n_risk_blue, c(0.025, 0.25, 0.5, 0.75, 0.975), na.rm=TRUE),
    
    .groups = "drop"
    ) %>%
  pivot_longer(
    cols = c(DCRB_rev, n_risk_humpback, n_risk_blue), 
    names_to = "response_metric", 
    values_to = "response"
  )
glimpse(annual_risk_out_sq)


# just peek at 2015-16 season, which only includes March-Jul. Zeroes in Nov-Feb are not included as part of average monthly risk.
# View(risk_out_sq %>%
#    mutate(
#    #month = as.numeric(substr(year_month,6,7)),
#    crab_year = func_crab_year(year_month),
#    plotting_year = as.numeric(substr(crab_year,6,9))
#    ) %>%
#    group_by(crab_year, plotting_year, year_month) %>%
#    summarise(
#      sum_DCRB_rev = sum(DCRB_rev, na.rm=TRUE)/10^5,
#      sum_n_risk_humpback = sum(n_risk_humpback, na.rm=TRUE),
#      sum_n_risk_blue = sum(n_risk_blue, na.rm=TRUE),
#      
#      .groups = "drop"
#      ) %>% filter(crab_year == "2015_2016"))

```



Other output df's of interest are:

- ALL SCENARIOS: annual_statewide_df

- FOCAL SCENARIOS: annual_statewide_df_focal_scenarios, scenario_table_focal_scenarios, df_tradeoff_focal_scenarios

- SUPPLEMENT SCENARIOS: annual_statewide_df_supplement, scenario_table_supplement, df_tradeoff_supplement

- STATUS QUO SCENARIOS: annual_statewide_df_sq, scenario_table_sq, df_tradeoff_sq, annual_risk_out_sq

There are also corresponding output .rds files

# 4. CALCULATE SCENARIO RANKS

CALCULATE SCENARIO RANKS
+ one df by year, one by period
+ repeat for statewide only and cenCA only scenarios
+ also make scale_compare df by period
+ ranks based on hump risk, blwh risk, DCRB $, DCRB lbs, cost effectiveness for humps, cost effectiveness for blues

## Calculate scenario ranks each year and each period, all scenarios
```{r rank_scenarios1, echo=FALSE}

# by year, ranked by various response variables
# larger values get higher ranks
# for relative whale metrics higher is greater risk reduction and therefore better
# for relative DCRB metrics higher is greater yield reduction and therefore worse

# tmp<- data.frame(yr=c(2009,2009,2010,2010,2011,2011),scenario=c("A","B","A","B","A","B"),rel_risk=c(30,80,40))
# tmp %>% group_by(yr) %>% mutate(rank_risk = row_number(rel_risk))

annual_df_ranks <- annual_statewide_df %>%
  group_by(crab_year) %>%
  mutate(
    relative_dollars_cost = 100 - relative_dollars,
    costs_benefits_hump = relative_dollars_cost / relative_hump_risk_n,
    costs_benefits_blwh = relative_dollars_cost / relative_blwh_risk_n,
    rank_hump = row_number(relative_hump_risk_n),
    rank_blwh = row_number(relative_blwh_risk_n),
    rank_both = row_number(relative_both_risk_n),
    rank_dollars = row_number(relative_dollars),
    rank_pounds = row_number(relative_pounds),
    rank_cbhump = row_number(costs_benefits_hump), # note these are messed up because negative values are bad, as they reflect risk that increased relative to status quo
    rank_cbblwh = row_number(costs_benefits_blwh) # note these are messed up because negative values are bad, as they reflect risk that increased relative to status quo
  ) %>%
  select(crab_year, plotting_year, pre_post, 
         scenario_number, pretty_scenario_names,
         rank_hump,rank_blwh,rank_both,rank_dollars,rank_pounds,rank_cbhump,rank_cbblwh,
         n_risk_humpback,relative_hump_risk_n,
         n_risk_blue,relative_blwh_risk_n,
         n_risk_both, relative_both_risk_n,
         DCRB_rev,relative_dollars,
         DCRB_lbs,relative_pounds,
         costs_benefits_hump, costs_benefits_blwh
  )

#glimpse(annual_df_ranks)

annual_df_ranks %>% 
  select(scenario_number, pretty_scenario_names,
         rank_hump, rank_blwh, rank_both, rank_dollars, rank_pounds) %>% 
  knitr::kable()

annual_df_ranks %>% 
  select(scenario_number, pretty_scenario_names,
         n_risk_humpback, n_risk_blue, n_risk_both, relative_hump_risk_n, relative_blwh_risk_n, relative_both_risk_n,
         DCRB_rev, DCRB_lbs, relative_dollars, relative_pounds,
         costs_benefits_hump, costs_benefits_blwh) %>% 
  knitr::kable()

# by period, ranked by various response variables
# larger values get higher ranks
# for relative whale metrics higher is greater risk reduction and therefore better
# for relative DCRB metrics higher is greater yield reduction and therefore worse
annual_df_ranks_period <- annual_statewide_df %>%
  group_by(crab_year) %>%
  mutate(
    relative_dollars_cost = 100 - relative_dollars,
    costs_benefits_hump = relative_dollars_cost / relative_hump_risk_n, # note these are messed up because negative values are bad, as they reflect risk that increased relative to status quo
    costs_benefits_blwh = relative_dollars_cost / relative_blwh_risk_n # note these are messed up because negative values are bad, as they reflect risk that increased relative to status quo
  ) %>%
  group_by(scenario_number, pretty_scenario_names, pre_post) %>%
  summarise(
    mean_n_risk_humpback = mean(n_risk_humpback),
    mean_relative_hump_risk_n = mean(relative_hump_risk_n),
    mean_n_risk_blue = mean(n_risk_blue),
    mean_relative_blwh_risk_n = mean(relative_blwh_risk_n),
    mean_n_risk_both = mean(n_risk_both),
    mean_relative_both_risk_n = mean(relative_both_risk_n),
    mean_DCRB_rev = mean(DCRB_rev),
    mean_relative_dollars = mean(relative_dollars),
    mean_DCRB_lbs = mean(DCRB_lbs),
    mean_relative_pounds = mean(relative_pounds),
    mean_cbhump = mean(costs_benefits_hump),
    mean_cbblwh = mean(costs_benefits_blwh),
    .groups = "drop"
  ) %>%
  group_by(pre_post) %>%
  mutate(
    rank_hump = row_number(mean_relative_hump_risk_n),
    rank_blwh = row_number(mean_relative_blwh_risk_n),
    rank_both = row_number(mean_relative_both_risk_n),
    rank_dollars = row_number(mean_relative_dollars),
    rank_pounds = row_number(mean_relative_pounds),
    rank_cbhump = row_number(mean_cbhump),
    rank_cbblwh = row_number(mean_cbblwh),
    .groups = "drop"
  ) %>%
  select(pre_post, 
         scenario_number, pretty_scenario_names,
         rank_hump,rank_blwh,rank_both,rank_dollars,rank_pounds,rank_cbhump,rank_cbblwh,
         mean_n_risk_humpback, mean_relative_hump_risk_n,
         mean_n_risk_blue, mean_relative_blwh_risk_n,
         mean_n_risk_both, mean_relative_both_risk_n,
         mean_DCRB_rev, mean_relative_dollars,
         mean_DCRB_lbs, mean_relative_pounds,
         mean_cbhump,
         mean_cbblwh
  )

#glimpse(annual_df_ranks_period)

# make a df sorted by rank risk reduction to humps
annual_df_ranks_period_hump <- annual_df_ranks_period %>%
  arrange(
    -mean_n_risk_humpback,
    -mean_n_risk_blue,
    -mean_n_risk_both,
    -mean_DCRB_rev,
    -mean_DCRB_lbs,
    .groups = "drop"
  ) %>%
  mutate(scenario_rank = row_number())

#glimpse(annual_df_ranks_period_hump)

#annual_df_ranks_period_hump %>% knitr::kable()

annual_df_ranks_period_hump %>% 
  select(pre_post, scenario_number, pretty_scenario_names,
         rank_hump, rank_blwh, rank_both, rank_dollars, rank_pounds) %>% 
  knitr::kable()

annual_df_ranks_period_hump %>% 
  select(pre_post, scenario_number, pretty_scenario_names,
         mean_n_risk_humpback, mean_n_risk_blue, mean_n_risk_both, mean_relative_hump_risk_n, mean_relative_blwh_risk_n, mean_relative_both_risk_n,
         mean_DCRB_rev, mean_DCRB_lbs, mean_relative_dollars, mean_relative_pounds,
         mean_cbhump, mean_cbblwh) %>% 
  knitr::kable()


write_csv(annual_df_ranks_period_hump, 
          here::here(
            "tradeoffs",
            "full_analysis_samhourietal",
            "output_dfs",
            "annual_df_ranks_period_hump.csv")
          )

```

## Extract scenario ranks each year and each period, focal scenarios
```{r rank_scenarios2, echo=FALSE}

# by year, ranked by various response variables
# larger values get higher ranks
# for relative whale metrics higher is greater risk reduction and therefore better
# for relative DCRB metrics higher is greater yield reduction and therefore worse

annual_df_focal_scenarios_ranks <- annual_statewide_df_focal_scenarios %>%
  group_by(crab_year) %>%
  mutate(
    relative_dollars_cost = 100 - relative_dollars,
    costs_benefits_hump = relative_dollars_cost / relative_hump_risk_n, # note these are messed up because negative values are bad, as they reflect risk that increased relative to status quo
    costs_benefits_blwh = relative_dollars_cost / relative_blwh_risk_n, # note these are messed up because negative values are bad, as they reflect risk that increased relative to status quo
    rank_hump = row_number(relative_hump_risk_n),
    rank_blwh = row_number(relative_blwh_risk_n),
    rank_both = row_number(relative_both_risk_n),
    rank_dollars = row_number(relative_dollars),
    rank_pounds = row_number(relative_pounds),
    rank_cbhump = row_number(costs_benefits_hump),
    rank_cbblwh = row_number(costs_benefits_blwh)
  ) %>%
  select(crab_year, plotting_year, pre_post, 
         scenario_number, pretty_scenario_names,
         rank_hump,rank_blwh,rank_both,rank_dollars,rank_pounds,
         n_risk_humpback,relative_hump_risk_n,
         n_risk_blue,relative_blwh_risk_n,
         n_risk_both, relative_both_risk_n,
         DCRB_rev,relative_dollars,
         DCRB_lbs,relative_pounds,
         costs_benefits_hump, costs_benefits_blwh
  )

#glimpse(annual_df_focal_scenarios_ranks)

annual_df_focal_scenarios_ranks %>% 
  select(scenario_number, pretty_scenario_names,
         rank_hump, rank_blwh, rank_both, rank_dollars, rank_pounds) %>% 
  knitr::kable()

annual_df_focal_scenarios_ranks %>% 
  select(scenario_number, pretty_scenario_names,
         n_risk_humpback, n_risk_blue, n_risk_both, relative_hump_risk_n, relative_blwh_risk_n, relative_both_risk_n,
         DCRB_rev, DCRB_lbs, relative_dollars, relative_pounds,
         costs_benefits_hump, costs_benefits_blwh) %>% 
  knitr::kable()
  

# by period, ranked by various response variables
# larger values get higher ranks
# for relative whale metrics higher is greater risk reduction and therefore better
# for relative DCRB metrics higher is greater yield reduction and therefore worse
annual_df_ranks_period_focal <- annual_statewide_df_focal_scenarios %>%
  group_by(crab_year) %>%
  mutate(
    relative_dollars_cost = 100 - relative_dollars,
    costs_benefits_hump = relative_dollars_cost / relative_hump_risk_n, # note these are messed up because negative values are bad, as they reflect risk that increased relative to status quo
    costs_benefits_blwh = relative_dollars_cost / relative_blwh_risk_n # note these are messed up because negative values are bad, as they reflect risk that increased relative to status quo
  ) %>%
  group_by(scenario_number, pretty_scenario_names, pre_post) %>%
  summarise(
    mean_n_risk_humpback = mean(n_risk_humpback),
    mean_relative_hump_risk_n = mean(relative_hump_risk_n),
    mean_n_risk_blue = mean(n_risk_blue),
    mean_relative_blwh_risk_n = mean(relative_blwh_risk_n),
    mean_n_risk_both = mean(n_risk_both),
    mean_relative_both_risk_n = mean(relative_both_risk_n),
    mean_DCRB_rev = mean(DCRB_rev),
    mean_relative_dollars = mean(relative_dollars),
    mean_DCRB_lbs = mean(DCRB_lbs),
    mean_relative_pounds = mean(relative_pounds),
    mean_cbhump = mean(costs_benefits_hump),
    mean_cbblwh = mean(costs_benefits_blwh),
    .groups = "drop"
  ) %>%
  group_by(pre_post) %>%
  mutate(
    rank_hump = row_number(mean_relative_hump_risk_n),
    rank_blwh = row_number(mean_relative_blwh_risk_n),
    rank_both = row_number(mean_relative_both_risk_n),
    rank_dollars = row_number(mean_relative_dollars),
    rank_pounds = row_number(mean_relative_pounds),
    rank_cbhump = row_number(mean_cbhump),
    rank_cbblwh = row_number(mean_cbblwh),
    .groups = "drop"
  ) %>%
  select(pre_post, 
         scenario_number, pretty_scenario_names,
         rank_hump,rank_blwh,rank_both,rank_dollars,rank_pounds,
         mean_n_risk_humpback, mean_relative_hump_risk_n,
         mean_n_risk_blue, mean_relative_blwh_risk_n,
         mean_n_risk_both, mean_relative_both_risk_n,
         mean_DCRB_rev, mean_relative_dollars,
         mean_DCRB_lbs, mean_relative_pounds,
         mean_cbhump,
         mean_cbblwh
  )

#glimpse(annual_df_ranks_period_focal)

# display a df showing relative changes
annual_df_ranks_period_focal %>% 
  select(pre_post, pretty_scenario_names, mean_relative_hump_risk_n, mean_relative_blwh_risk_n, mean_relative_dollars) %>%
  knitr::kable()

annual_df_ranks_period_focal %>% 
  select(pre_post, pretty_scenario_names, mean_relative_hump_risk_n, mean_relative_blwh_risk_n, mean_relative_dollars) %>%
  group_by(pre_post) %>%
  summarise(
    max_relative_hump_risk_n = max(mean_relative_hump_risk_n), 
    mean_relative_hump_risk_n = mean(mean_relative_hump_risk_n), 
    max_relative_blwh_risk_n = max(mean_relative_blwh_risk_n), 
    mean_relative_blwh_risk_n = mean(mean_relative_blwh_risk_n), 
    min_relative_dollars = min(mean_relative_dollars),
    mean_relative_dollars = mean(mean_relative_dollars)
  ) %>%
  knitr::kable()


# make a df sorted by rank risk reduction to humps
annual_df_ranks_period_focal_hump <- annual_df_ranks_period_focal %>%
  arrange(
    -mean_n_risk_humpback,
    -mean_n_risk_blue,
    -mean_n_risk_both,
    -mean_DCRB_rev,
    -mean_DCRB_lbs,
    .groups = "drop"
  ) %>%
  mutate(scenario_rank = row_number())

#glimpse(annual_df_ranks_period_focal_hump)

#annual_df_ranks_period_focal_hump %>% knitr::kable()

annual_df_ranks_period_focal_hump %>% 
  select(pre_post, scenario_number, pretty_scenario_names,
         rank_hump, rank_blwh, rank_both, rank_dollars, rank_pounds) %>% 
  knitr::kable()

annual_df_ranks_period_focal_hump %>% 
  select(pre_post, scenario_number, pretty_scenario_names,
         mean_n_risk_humpback, mean_n_risk_blue, mean_n_risk_both, mean_relative_hump_risk_n, mean_relative_blwh_risk_n, mean_relative_both_risk_n,
         mean_DCRB_rev, mean_DCRB_lbs, mean_relative_dollars, mean_relative_pounds,
         mean_cbhump, mean_cbblwh) %>% 
  knitr::kable()

write_csv(annual_df_ranks_period_focal_hump, 
          here::here(
            "tradeoffs",
            "full_analysis_samhourietal",
            "output_dfs",
            "annual_df_ranks_period_focal_hump.csv")
          )


```

## Extract scenario ranks each year and each period, statewide focal scenarios
```{r rank_scenarios3, echo=FALSE}

# by year, ranked by various response variables
# larger values get higher ranks
# for relative whale metrics higher is greater risk reduction and therefore better
# for relative DCRB metrics higher is greater yield reduction and therefore worse

annual_df_focal_scenarios_statewide_ranks <- annual_statewide_df_focal_scenarios %>%
  filter(stringr::str_detect(pretty_scenario_names, 'Statewide|Status Quo') ) %>%
  filter(!stringr::str_detect(pretty_scenario_names, 'CenCA|BIA')) %>%
  group_by(crab_year) %>%
  mutate(
    relative_dollars_cost = 100 - relative_dollars,
    costs_benefits_hump = relative_dollars_cost / relative_hump_risk_n, # note these are messed up because negative values are bad, as they reflect risk that increased relative to status quo
    costs_benefits_blwh = relative_dollars_cost / relative_blwh_risk_n, # note these are messed up because negative values are bad, as they reflect risk that increased relative to status quo
    rank_hump = row_number(relative_hump_risk_n),
    rank_blwh = row_number(relative_blwh_risk_n),
    rank_both = row_number(relative_both_risk_n),
    rank_dollars = row_number(relative_dollars),
    rank_pounds = row_number(relative_pounds),
    rank_cbhump = row_number(costs_benefits_hump),
    rank_cbblwh = row_number(costs_benefits_blwh)
  ) %>%
  select(crab_year, plotting_year, pre_post, 
         scenario_number, pretty_scenario_names,
         rank_hump,rank_blwh,rank_both,rank_dollars,rank_pounds,
         n_risk_humpback,relative_hump_risk_n,
         n_risk_blue,relative_blwh_risk_n,
         n_risk_both, relative_both_risk_n,
         DCRB_rev,relative_dollars,
         DCRB_lbs,relative_pounds,
         costs_benefits_hump, costs_benefits_blwh
  )

#glimpse(annual_df_focal_scenarios_statewide_ranks)

annual_df_focal_scenarios_statewide_ranks %>% 
  select(scenario_number, pretty_scenario_names,
         rank_hump, rank_blwh, rank_both, rank_dollars, rank_pounds) %>% 
  knitr::kable()

annual_df_focal_scenarios_statewide_ranks %>% 
  select(scenario_number, pretty_scenario_names,
         n_risk_humpback, n_risk_blue, n_risk_both, relative_hump_risk_n, relative_blwh_risk_n, relative_both_risk_n,
         DCRB_rev, DCRB_lbs, relative_dollars, relative_pounds,
         costs_benefits_hump, costs_benefits_blwh) %>% 
  knitr::kable()

annual_df_focal_scenarios_statewide_ranks %>%
  group_by(scenario_number, pretty_scenario_names) %>%
  summarise(
    mean_relative_hump_risk_n = mean(relative_hump_risk_n), 
    mean_relative_blwh_risk_n = mean(relative_blwh_risk_n), 
    mean_relative_dollars = mean(relative_dollars), 
    mean_relative_pounds = mean(relative_pounds)
  ) %>% 
  knitr::kable()

# by period, ranked by various response variables
# larger values get higher ranks
# for relative whale metrics higher is greater risk reduction and therefore better
# for relative DCRB metrics higher is greater yield reduction and therefore worse
annual_df_ranks_period_focal_statewide <- annual_statewide_df_focal_scenarios %>%
  filter(stringr::str_detect(pretty_scenario_names, 'Statewide|Status Quo') ) %>%
  filter(!stringr::str_detect(pretty_scenario_names, 'CenCA|BIA')) %>%
  group_by(crab_year) %>%
  mutate(
    relative_dollars_cost = 100 - relative_dollars,
    costs_benefits_hump = relative_dollars_cost / relative_hump_risk_n, # note these are messed up because negative values are bad, as they reflect risk that increased relative to status quo
    costs_benefits_blwh = relative_dollars_cost / relative_blwh_risk_n # note these are messed up because negative values are bad, as they reflect risk that increased relative to status quo
  ) %>%
  group_by(scenario_number, pretty_scenario_names, pre_post) %>%
  summarise(
    mean_n_risk_humpback = mean(n_risk_humpback),
    mean_relative_hump_risk_n = mean(relative_hump_risk_n),
    mean_n_risk_blue = mean(n_risk_blue),
    mean_relative_blwh_risk_n = mean(relative_blwh_risk_n),
    mean_n_risk_both = mean(n_risk_both),
    mean_relative_both_risk_n = mean(relative_both_risk_n),
    mean_DCRB_rev = mean(DCRB_rev),
    mean_relative_dollars = mean(relative_dollars),
    mean_DCRB_lbs = mean(DCRB_lbs),
    mean_relative_pounds = mean(relative_pounds),
    mean_cbhump = mean(costs_benefits_hump),
    mean_cbblwh = mean(costs_benefits_blwh),
    .groups = "drop"
  ) %>%
  group_by(pre_post) %>%
  mutate(
    rank_hump = row_number(mean_relative_hump_risk_n),
    rank_blwh = row_number(mean_relative_blwh_risk_n),
    rank_both = row_number(mean_relative_both_risk_n),
    rank_dollars = row_number(mean_relative_dollars),
    rank_pounds = row_number(mean_relative_pounds),
    rank_cbhump = row_number(mean_cbhump),
    rank_cbblwh = row_number(mean_cbblwh),
    .groups = "drop"
  ) %>%
  select(pre_post, 
         scenario_number, pretty_scenario_names,
         rank_hump,rank_blwh,rank_both,rank_dollars,rank_pounds,
         mean_n_risk_humpback, mean_relative_hump_risk_n,
         mean_n_risk_blue, mean_relative_blwh_risk_n,
         mean_n_risk_both, mean_relative_both_risk_n,
         mean_DCRB_rev, mean_relative_dollars,
         mean_DCRB_lbs, mean_relative_pounds,
         mean_cbhump,
         mean_cbblwh
  )

#glimpse(annual_df_ranks_period_focal_statewide)

# display a df showing relative changes
annual_df_ranks_period_focal_statewide %>% 
  select(pre_post, pretty_scenario_names, mean_relative_hump_risk_n, mean_relative_blwh_risk_n, mean_relative_dollars) %>%
  knitr::kable()

annual_df_ranks_period_focal_statewide %>% 
  select(pre_post, pretty_scenario_names, mean_relative_hump_risk_n, mean_relative_blwh_risk_n, mean_relative_dollars) %>%
  group_by(pre_post) %>%
  summarise(
    max_relative_hump_risk_n = max(mean_relative_hump_risk_n), 
    mean_relative_hump_risk_n = mean(mean_relative_hump_risk_n), 
    max_relative_blwh_risk_n = max(mean_relative_blwh_risk_n), 
    mean_relative_blwh_risk_n = mean(mean_relative_blwh_risk_n), 
    min_relative_dollars = min(mean_relative_dollars),
    mean_relative_dollars = mean(mean_relative_dollars)
  ) %>%
  knitr::kable()

annual_df_ranks_period_focal_statewide %>% 
  filter(stringr::str_detect(pretty_scenario_names, 'Early')) %>% 
  select(pre_post, pretty_scenario_names, mean_relative_hump_risk_n, mean_relative_blwh_risk_n, mean_relative_dollars) %>%
  group_by(pre_post) %>%
  summarise(
    early_only_max_relative_hump_risk_n = max(mean_relative_hump_risk_n), 
    early_only_mean_relative_hump_risk_n = mean(mean_relative_hump_risk_n), 
    early_only_max_relative_blwh_risk_n = max(mean_relative_blwh_risk_n), 
    early_only_mean_relative_blwh_risk_n = mean(mean_relative_blwh_risk_n), 
    early_only_min_relative_dollars = min(mean_relative_dollars),
    early_only_mean_relative_dollars = mean(mean_relative_dollars)
  ) %>%
  knitr::kable()


# make a df sorted by rank risk reduction to humps
annual_df_ranks_period_focal_statewide_hump <- annual_df_ranks_period_focal_statewide %>%
  arrange(
    -mean_n_risk_humpback,
    -mean_n_risk_blue,
    -mean_n_risk_both,
    -mean_DCRB_rev,
    -mean_DCRB_lbs,
    .groups = "drop"
  ) %>%
  mutate(scenario_rank = row_number())

#glimpse(annual_df_ranks_period_focal_statewide_hump)

#annual_df_ranks_period_focal_statewide_hump %>% knitr::kable()

annual_df_ranks_period_focal_statewide_hump %>% 
  select(pre_post, scenario_number, pretty_scenario_names,
         rank_hump, rank_blwh, rank_both, rank_dollars, rank_pounds) %>% 
  knitr::kable()

annual_df_ranks_period_focal_statewide_hump %>% 
  select(pre_post, scenario_number, pretty_scenario_names,
         mean_n_risk_humpback, mean_n_risk_blue, mean_n_risk_both, mean_relative_hump_risk_n, mean_relative_blwh_risk_n, mean_relative_both_risk_n,
         mean_DCRB_rev, mean_DCRB_lbs, mean_relative_dollars, mean_relative_pounds,
         mean_cbhump, mean_cbblwh) %>% 
  knitr::kable()

write_csv(annual_df_ranks_period_focal_statewide_hump, 
          here::here(
            "tradeoffs",
            "full_analysis_samhourietal",
            "output_dfs",
            "annual_df_ranks_period_focal_statewide_hump.csv")
          )

```

## Extract scenario ranks each year and each period, cenCA only focal scenarios
```{r rank_scenarios4, echo=FALSE}

# by year, ranked by various response variables
# larger values get higher ranks
# for relative whale metrics higher is greater risk reduction and therefore better
# for relative DCRB metrics higher is greater yield reduction and therefore worse

annual_df_focal_scenarios_cenCA_ranks <- annual_statewide_df_focal_scenarios %>%
  filter(stringr::str_detect(pretty_scenario_names, 'CenCA|Status Quo') ) %>%
  filter(!stringr::str_detect(pretty_scenario_names, 'Statewide|BIA')) %>%
  group_by(crab_year) %>%
  mutate(
    relative_dollars_cost = 100 - relative_dollars,
    costs_benefits_hump = relative_dollars_cost / relative_hump_risk_n, # note these are messed up because negative values are bad, as they reflect risk that increased relative to status quo
    costs_benefits_blwh = relative_dollars_cost / relative_blwh_risk_n, # note these are messed up because negative values are bad, as they reflect risk that increased relative to status quo
    rank_hump = row_number(relative_hump_risk_n),
    rank_blwh = row_number(relative_blwh_risk_n),
    rank_both = row_number(relative_both_risk_n),
    rank_dollars = row_number(relative_dollars),
    rank_pounds = row_number(relative_pounds),
    rank_cbhump = row_number(costs_benefits_hump),
    rank_cbblwh = row_number(costs_benefits_blwh)
  ) %>%
  select(crab_year, plotting_year, pre_post, 
         scenario_number, pretty_scenario_names,
         rank_hump,rank_blwh,rank_both,rank_dollars,rank_pounds,
         n_risk_humpback,relative_hump_risk_n,
         n_risk_blue,relative_blwh_risk_n,
         n_risk_both, relative_both_risk_n,
         DCRB_rev,relative_dollars,
         DCRB_lbs,relative_pounds,
         costs_benefits_hump, costs_benefits_blwh
  )

#glimpse(annual_df_focal_scenarios_cenCA_ranks)

annual_df_focal_scenarios_cenCA_ranks %>% 
  select(scenario_number, pretty_scenario_names,
         rank_hump, rank_blwh, rank_both, rank_dollars, rank_pounds) %>% 
  knitr::kable()

annual_df_focal_scenarios_cenCA_ranks %>% 
  select(scenario_number, pretty_scenario_names,
         n_risk_humpback, n_risk_blue, n_risk_both, relative_hump_risk_n, relative_blwh_risk_n, relative_both_risk_n,
         DCRB_rev, DCRB_lbs, relative_dollars, relative_pounds,
         costs_benefits_hump, costs_benefits_blwh) %>% 
  knitr::kable()

# by period, ranked by various response variables
# larger values get higher ranks
# for relative whale metrics higher is greater risk reduction and therefore better
# for relative DCRB metrics higher is greater yield reduction and therefore worse
annual_df_ranks_period_focal_cenCA <- annual_statewide_df_focal_scenarios %>%
  filter(stringr::str_detect(pretty_scenario_names, 'CenCA|Status Quo') ) %>%
  filter(!stringr::str_detect(pretty_scenario_names, 'Statewide|BIA')) %>%
  group_by(crab_year) %>%
  mutate(
    relative_dollars_cost = 100 - relative_dollars,
    costs_benefits_hump = relative_dollars_cost / relative_hump_risk_n, # note these are messed up because negative values are bad, as they reflect risk that increased relative to status quo
    costs_benefits_blwh = relative_dollars_cost / relative_blwh_risk_n # note these are messed up because negative values are bad, as they reflect risk that increased relative to status quo
  ) %>%
  group_by(scenario_number, pretty_scenario_names, pre_post) %>%
  summarise(
    mean_n_risk_humpback = mean(n_risk_humpback),
    mean_relative_hump_risk_n = mean(relative_hump_risk_n),
    mean_n_risk_blue = mean(n_risk_blue),
    mean_relative_blwh_risk_n = mean(relative_blwh_risk_n),
    mean_n_risk_both = mean(n_risk_both),
    mean_relative_both_risk_n = mean(relative_both_risk_n),
    mean_DCRB_rev = mean(DCRB_rev),
    mean_relative_dollars = mean(relative_dollars),
    mean_DCRB_lbs = mean(DCRB_lbs),
    mean_relative_pounds = mean(relative_pounds),
    mean_cbhump = mean(costs_benefits_hump),
    mean_cbblwh = mean(costs_benefits_blwh),
    .groups = "drop"
  ) %>%
  group_by(pre_post) %>%
  mutate(
    rank_hump = row_number(mean_relative_hump_risk_n),
    rank_blwh = row_number(mean_relative_blwh_risk_n),
    rank_both = row_number(mean_relative_both_risk_n),
    rank_dollars = row_number(mean_relative_dollars),
    rank_pounds = row_number(mean_relative_pounds),
    rank_cbhump = row_number(mean_cbhump),
    rank_cbblwh = row_number(mean_cbblwh),
    .groups = "drop"
  ) %>%
  select(pre_post, 
         scenario_number, pretty_scenario_names,
         rank_hump,rank_blwh,rank_both,rank_dollars,rank_pounds,
         mean_n_risk_humpback, mean_relative_hump_risk_n,
         mean_n_risk_blue, mean_relative_blwh_risk_n,
         mean_n_risk_both, mean_relative_both_risk_n,
         mean_DCRB_rev, mean_relative_dollars,
         mean_DCRB_lbs, mean_relative_pounds,
         mean_cbhump,
         mean_cbblwh
  )

#glimpse(annual_df_ranks_period_focal_cenCA)

# make a df sorted by rank risk reduction to humps
annual_df_ranks_period_focal_cenCA_hump <- annual_df_ranks_period_focal_cenCA %>%
  arrange(
    -mean_n_risk_humpback,
    -mean_n_risk_blue,
    -mean_n_risk_both,
    -mean_DCRB_rev,
    -mean_DCRB_lbs,
    .groups = "drop"
  ) %>%
  mutate(scenario_rank = row_number())

#glimpse(annual_df_ranks_period_focal_cenCA_hump)

#annual_df_ranks_period_focal_cenCA_hump %>% knitr::kable()

annual_df_ranks_period_focal_cenCA_hump %>% 
  select(pre_post, scenario_number, pretty_scenario_names,
         rank_hump, rank_blwh, rank_both, rank_dollars, rank_pounds) %>% 
  knitr::kable()

annual_df_ranks_period_focal_cenCA_hump %>% 
  select(pre_post, scenario_number, pretty_scenario_names,
         mean_n_risk_humpback, mean_n_risk_blue, mean_n_risk_both, mean_relative_hump_risk_n, mean_relative_blwh_risk_n, mean_relative_both_risk_n,
         mean_DCRB_rev, mean_DCRB_lbs, mean_relative_dollars, mean_relative_pounds,
         mean_cbhump, mean_cbblwh) %>% 
  knitr::kable()

write_csv(annual_df_ranks_period_focal_cenCA_hump, 
          here::here(
            "tradeoffs",
            "full_analysis_samhourietal",
            "output_dfs",
            "annual_df_ranks_period_focal_cenCA_hump.csv")
          )

```

## Extract scenario ranks each year and each period, supplement scenarios
```{r rank_scenarios5, echo=FALSE}

annual_df_supplement_scenarios_ranks <- annual_statewide_df_supplement %>%
  group_by(crab_year) %>%
  mutate(
    relative_dollars_cost = 100 - relative_dollars,
    costs_benefits_hump = relative_dollars_cost / relative_hump_risk_n, # note these are messed up because negative values are bad, as they reflect risk that increased relative to status quo
    costs_benefits_blwh = relative_dollars_cost / relative_blwh_risk_n, # note these are messed up because negative values are bad, as they reflect risk that increased relative to status quo
    rank_hump = row_number(relative_hump_risk_n),
    rank_blwh = row_number(relative_blwh_risk_n),
    rank_both = row_number(relative_both_risk_n),
    rank_dollars = row_number(relative_dollars),
    rank_pounds = row_number(relative_pounds),
    rank_cbhump = row_number(costs_benefits_hump),
    rank_cbblwh = row_number(costs_benefits_blwh)
  ) %>%
  select(crab_year, plotting_year, pre_post, 
         scenario_number, pretty_scenario_names,
         rank_hump,rank_blwh,rank_both,rank_dollars,rank_pounds,
         n_risk_humpback,relative_hump_risk_n,
         n_risk_blue,relative_blwh_risk_n,
         n_risk_both, relative_both_risk_n,
         DCRB_rev,relative_dollars,
         DCRB_lbs,relative_pounds,
         costs_benefits_hump, costs_benefits_blwh
  )

#glimpse(annual_df_supplement_scenarios_ranks)

annual_df_supplement_scenarios_ranks %>% 
  select(scenario_number, pretty_scenario_names,
         rank_hump, rank_blwh, rank_both, rank_dollars, rank_pounds) %>% 
  knitr::kable()

annual_df_supplement_scenarios_ranks %>% 
  select(scenario_number, pretty_scenario_names,
         n_risk_humpback, n_risk_blue, n_risk_both, relative_hump_risk_n, relative_blwh_risk_n, relative_both_risk_n,
         DCRB_rev, DCRB_lbs, relative_dollars, relative_pounds,
         costs_benefits_hump, costs_benefits_blwh) %>% 
  knitr::kable()

# by period, ranked by various response variables
# larger values get higher ranks
# for relative whale metrics higher is greater risk reduction and therefore better
# for relative DCRB metrics higher is greater yield reduction and therefore worse
annual_df_ranks_period_supplement <- annual_statewide_df_supplement %>%
  group_by(crab_year) %>%
  mutate(
    relative_dollars_cost = 100 - relative_dollars,
    costs_benefits_hump = relative_dollars_cost / relative_hump_risk_n, # note these are messed up because negative values are bad, as they reflect risk that increased relative to status quo
    costs_benefits_blwh = relative_dollars_cost / relative_blwh_risk_n # note these are messed up because negative values are bad, as they reflect risk that increased relative to status quo
  ) %>%
  group_by(scenario_number, pretty_scenario_names, pre_post) %>%
  summarise(
    mean_n_risk_humpback = mean(n_risk_humpback),
    mean_relative_hump_risk_n = mean(relative_hump_risk_n),
    mean_n_risk_blue = mean(n_risk_blue),
    mean_relative_blwh_risk_n = mean(relative_blwh_risk_n),
    mean_n_risk_both = mean(n_risk_both),
    mean_relative_both_risk_n = mean(relative_both_risk_n),
    mean_DCRB_rev = mean(DCRB_rev),
    mean_relative_dollars = mean(relative_dollars),
    mean_DCRB_lbs = mean(DCRB_lbs),
    mean_relative_pounds = mean(relative_pounds),
    mean_cbhump = mean(costs_benefits_hump),
    mean_cbblwh = mean(costs_benefits_blwh),
    .groups = "drop"
  ) %>%
  group_by(pre_post) %>%
  mutate(
    rank_hump = row_number(mean_relative_hump_risk_n),
    rank_blwh = row_number(mean_relative_blwh_risk_n),
    rank_both = row_number(mean_relative_both_risk_n),
    rank_dollars = row_number(mean_relative_dollars),
    rank_pounds = row_number(mean_relative_pounds),
    rank_cbhump = row_number(mean_cbhump),
    rank_cbblwh = row_number(mean_cbblwh),
    .groups = "drop"
  ) %>%
  select(pre_post, 
         scenario_number, pretty_scenario_names,
         rank_hump,rank_blwh,rank_both,rank_dollars,rank_pounds,
         mean_n_risk_humpback, mean_relative_hump_risk_n,
         mean_n_risk_blue, mean_relative_blwh_risk_n,
         mean_n_risk_both, mean_relative_both_risk_n,
         mean_DCRB_rev, mean_relative_dollars,
         mean_DCRB_lbs, mean_relative_pounds,
         mean_cbhump,
         mean_cbblwh
  )

#glimpse(annual_df_ranks_period_supplement)

# make a df sorted by rank risk reduction to humps
annual_df_ranks_period_supplement_hump <- annual_df_ranks_period_supplement %>%
  arrange(
    -mean_n_risk_humpback,
    -mean_n_risk_blue,
    -mean_n_risk_both,
    -mean_DCRB_rev,
    -mean_DCRB_lbs,
    .groups = "drop"
  ) %>%
  mutate(scenario_rank = row_number())

#glimpse(annual_df_ranks_period_supplement_hump)

#annual_df_ranks_period_supplement_hump %>% knitr::kable()

annual_df_ranks_period_supplement_hump %>% 
  select(pre_post, scenario_number, pretty_scenario_names,
         rank_hump, rank_blwh, rank_both, rank_dollars, rank_pounds) %>% 
  knitr::kable()

annual_df_ranks_period_supplement_hump %>% 
  select(pre_post, scenario_number, pretty_scenario_names,
         mean_n_risk_humpback, mean_n_risk_blue, mean_n_risk_both, mean_relative_hump_risk_n, mean_relative_blwh_risk_n, mean_relative_both_risk_n,
         mean_DCRB_rev, mean_DCRB_lbs, mean_relative_dollars, mean_relative_pounds,
         mean_cbhump, mean_cbblwh) %>% 
  knitr::kable()


write_csv(annual_df_ranks_period_supplement_hump, 
          here::here(
            "tradeoffs",
            "full_analysis_samhourietal",
            "output_dfs",
            "annual_df_ranks_period_supplement_hump.csv")
          )
```

## Extract scenario ranks each year and each period, statewide supplement scenarios
```{r rank_scenarios6, echo=FALSE}

# by year, ranked by various response variables
# larger values get higher ranks
# for relative whale metrics higher is greater risk reduction and therefore better
# for relative DCRB metrics higher is greater yield reduction and therefore worse

annual_df_supplement_scenarios_statewide_ranks <- annual_statewide_df_supplement %>%
  filter(stringr::str_detect(pretty_scenario_names, 'Statewide|Status Quo') ) %>%
  filter(!stringr::str_detect(pretty_scenario_names, 'CenCA|BIA')) %>%
  group_by(crab_year) %>%
  mutate(
    relative_dollars_cost = 100 - relative_dollars,
    costs_benefits_hump = relative_dollars_cost / relative_hump_risk_n, # note these are messed up because negative values are bad, as they reflect risk that increased relative to status quo
    costs_benefits_blwh = relative_dollars_cost / relative_blwh_risk_n, # note these are messed up because negative values are bad, as they reflect risk that increased relative to status quo
    rank_hump = row_number(relative_hump_risk_n),
    rank_blwh = row_number(relative_blwh_risk_n),
    rank_both = row_number(relative_both_risk_n),
    rank_dollars = row_number(relative_dollars),
    rank_pounds = row_number(relative_pounds),
    rank_cbhump = row_number(costs_benefits_hump),
    rank_cbblwh = row_number(costs_benefits_blwh)
  ) %>%
  select(crab_year, plotting_year, pre_post, 
         scenario_number, pretty_scenario_names,
         rank_hump,rank_blwh,rank_both,rank_dollars,rank_pounds,
         n_risk_humpback,relative_hump_risk_n,
         n_risk_blue,relative_blwh_risk_n,
         n_risk_both, relative_both_risk_n,
         DCRB_rev,relative_dollars,
         DCRB_lbs,relative_pounds,
         costs_benefits_hump, costs_benefits_blwh
  )

#glimpse(annual_df_supplement_scenarios_statewide_ranks)

annual_df_supplement_scenarios_statewide_ranks %>% 
  select(scenario_number, pretty_scenario_names,
         rank_hump, rank_blwh, rank_both, rank_dollars, rank_pounds) %>% 
  knitr::kable()

annual_df_supplement_scenarios_statewide_ranks %>% 
  select(scenario_number, pretty_scenario_names,
         n_risk_humpback, n_risk_blue, n_risk_both, relative_hump_risk_n, relative_blwh_risk_n, relative_both_risk_n,
         DCRB_rev, DCRB_lbs, relative_dollars, relative_pounds,
         costs_benefits_hump, costs_benefits_blwh) %>% 
  knitr::kable()

# by period, ranked by various response variables
# larger values get higher ranks
# for relative whale metrics higher is greater risk reduction and therefore better
# for relative DCRB metrics higher is greater yield reduction and therefore worse
annual_df_ranks_period_supplement_statewide <- annual_statewide_df_supplement %>%
  filter(stringr::str_detect(pretty_scenario_names, 'Statewide|Status Quo') ) %>%
  filter(!stringr::str_detect(pretty_scenario_names, 'CenCA|BIA')) %>%
  group_by(crab_year) %>%
  mutate(
    relative_dollars_cost = 100 - relative_dollars,
    costs_benefits_hump = relative_dollars_cost / relative_hump_risk_n, # note these are messed up because negative values are bad, as they reflect risk that increased relative to status quo
    costs_benefits_blwh = relative_dollars_cost / relative_blwh_risk_n # note these are messed up because negative values are bad, as they reflect risk that increased relative to status quo
  ) %>%
  group_by(scenario_number, pretty_scenario_names, pre_post) %>%
  summarise(
    mean_n_risk_humpback = mean(n_risk_humpback),
    mean_relative_hump_risk_n = mean(relative_hump_risk_n),
    mean_n_risk_blue = mean(n_risk_blue),
    mean_relative_blwh_risk_n = mean(relative_blwh_risk_n),
    mean_n_risk_both = mean(n_risk_both),
    mean_relative_both_risk_n = mean(relative_both_risk_n),
    mean_DCRB_rev = mean(DCRB_rev),
    mean_relative_dollars = mean(relative_dollars),
    mean_DCRB_lbs = mean(DCRB_lbs),
    mean_relative_pounds = mean(relative_pounds),
    mean_cbhump = mean(costs_benefits_hump),
    mean_cbblwh = mean(costs_benefits_blwh),
    .groups = "drop"
  ) %>%
  group_by(pre_post) %>%
  mutate(
    rank_hump = row_number(mean_relative_hump_risk_n),
    rank_blwh = row_number(mean_relative_blwh_risk_n),
    rank_both = row_number(mean_relative_both_risk_n),
    rank_dollars = row_number(mean_relative_dollars),
    rank_pounds = row_number(mean_relative_pounds),
    rank_cbhump = row_number(mean_cbhump),
    rank_cbblwh = row_number(mean_cbblwh),
    .groups = "drop"
  ) %>%
  select(pre_post, 
         scenario_number, pretty_scenario_names,
         rank_hump,rank_blwh,rank_both,rank_dollars,rank_pounds,
         mean_n_risk_humpback, mean_relative_hump_risk_n,
         mean_n_risk_blue, mean_relative_blwh_risk_n,
         mean_n_risk_both, mean_relative_both_risk_n,
         mean_DCRB_rev, mean_relative_dollars,
         mean_DCRB_lbs, mean_relative_pounds,
         mean_cbhump,
         mean_cbblwh
  )

#glimpse(annual_df_ranks_period_supplement_statewide)

# make a df sorted by rank risk reduction to humps
annual_df_ranks_period_supplement_statewide_hump <- annual_df_ranks_period_supplement_statewide %>%
  arrange(
    -mean_n_risk_humpback,
    -mean_n_risk_blue,
    -mean_n_risk_both,
    -mean_DCRB_rev,
    -mean_DCRB_lbs,
    .groups = "drop"
  ) %>%
  mutate(scenario_rank = row_number())

#glimpse(annual_df_ranks_period_supplement_statewide_hump)

#annual_df_ranks_period_supplement_statewide_hump %>% knitr::kable()

annual_df_ranks_period_supplement_statewide_hump %>% 
  select(pre_post, scenario_number, pretty_scenario_names,
         rank_hump, rank_blwh, rank_both, rank_dollars, rank_pounds) %>% 
  knitr::kable()

annual_df_ranks_period_supplement_statewide_hump %>% 
  select(pre_post, scenario_number, pretty_scenario_names,
         mean_n_risk_humpback, mean_n_risk_blue, mean_n_risk_both, mean_relative_hump_risk_n, mean_relative_blwh_risk_n, mean_relative_both_risk_n,
         mean_DCRB_rev, mean_DCRB_lbs, mean_relative_dollars, mean_relative_pounds,
         mean_cbhump, mean_cbblwh) %>% 
  knitr::kable()

write_csv(annual_df_ranks_period_supplement_statewide_hump, 
          here::here(
            "tradeoffs",
            "full_analysis_samhourietal",
            "output_dfs",
            "annual_df_ranks_period_supplement_statewide_hump.csv")
          )


```

## Extract scenario ranks each year and each period, cenCA only supplement scenarios
```{r rank_scenarios7, echo=FALSE}

# by year, ranked by various response variables
# larger values get higher ranks
# for relative whale metrics higher is greater risk reduction and therefore better
# for relative DCRB metrics higher is greater yield reduction and therefore worse

annual_df_supplement_scenarios_cenCA_ranks <- annual_statewide_df_supplement %>%
  filter(stringr::str_detect(pretty_scenario_names, 'CenCA|Status Quo') ) %>%
  filter(!stringr::str_detect(pretty_scenario_names, 'Statewide|BIA')) %>%
  group_by(crab_year) %>%
  mutate(
    relative_dollars_cost = 100 - relative_dollars,
    costs_benefits_hump = relative_dollars_cost / relative_hump_risk_n, # note these are messed up because negative values are bad, as they reflect risk that increased relative to status quo
    costs_benefits_blwh = relative_dollars_cost / relative_blwh_risk_n, # note these are messed up because negative values are bad, as they reflect risk that increased relative to status quo
    rank_hump = row_number(relative_hump_risk_n),
    rank_blwh = row_number(relative_blwh_risk_n),
    rank_both = row_number(relative_both_risk_n),
    rank_dollars = row_number(relative_dollars),
    rank_pounds = row_number(relative_pounds),
    rank_cbhump = row_number(costs_benefits_hump),
    rank_cbblwh = row_number(costs_benefits_blwh)
  ) %>%
  select(crab_year, plotting_year, pre_post, 
         scenario_number, pretty_scenario_names,
         rank_hump,rank_blwh,rank_both,rank_dollars,rank_pounds,
         n_risk_humpback,relative_hump_risk_n,
         n_risk_blue,relative_blwh_risk_n,
         n_risk_both, relative_both_risk_n,
         DCRB_rev,relative_dollars,
         DCRB_lbs,relative_pounds,
         costs_benefits_hump, costs_benefits_blwh
  )

#glimpse(annual_df_supplement_scenarios_cenCA_ranks)

annual_df_supplement_scenarios_cenCA_ranks %>% 
  select(scenario_number, pretty_scenario_names,
         rank_hump, rank_blwh, rank_both, rank_dollars, rank_pounds) %>% 
  knitr::kable()

annual_df_supplement_scenarios_cenCA_ranks %>% 
  select(scenario_number, pretty_scenario_names,
         n_risk_humpback, n_risk_blue, n_risk_both, relative_hump_risk_n, relative_blwh_risk_n, relative_both_risk_n,
         DCRB_rev, DCRB_lbs, relative_dollars, relative_pounds,
         costs_benefits_hump, costs_benefits_blwh) %>% 
  knitr::kable()

# by period, ranked by various response variables
# larger values get higher ranks
# for relative whale metrics higher is greater risk reduction and therefore better
# for relative DCRB metrics higher is greater yield reduction and therefore worse
annual_df_ranks_period_supplement_cenCA <- annual_statewide_df_supplement %>%
  filter(stringr::str_detect(pretty_scenario_names, 'CenCA|Status Quo') ) %>%
  filter(!stringr::str_detect(pretty_scenario_names, 'Statewide|BIA')) %>%
  group_by(crab_year) %>%
  mutate(
    relative_dollars_cost = 100 - relative_dollars,
    costs_benefits_hump = relative_dollars_cost / relative_hump_risk_n, # note these are messed up because negative values are bad, as they reflect risk that increased relative to status quo
    costs_benefits_blwh = relative_dollars_cost / relative_blwh_risk_n # note these are messed up because negative values are bad, as they reflect risk that increased relative to status quo
  ) %>%
  group_by(scenario_number, pretty_scenario_names, pre_post) %>%
  summarise(
    mean_n_risk_humpback = mean(n_risk_humpback),
    mean_relative_hump_risk_n = mean(relative_hump_risk_n),
    mean_n_risk_blue = mean(n_risk_blue),
    mean_relative_blwh_risk_n = mean(relative_blwh_risk_n),
    mean_n_risk_both = mean(n_risk_both),
    mean_relative_both_risk_n = mean(relative_both_risk_n),
    mean_DCRB_rev = mean(DCRB_rev),
    mean_relative_dollars = mean(relative_dollars),
    mean_DCRB_lbs = mean(DCRB_lbs),
    mean_relative_pounds = mean(relative_pounds),
    mean_cbhump = mean(costs_benefits_hump),
    mean_cbblwh = mean(costs_benefits_blwh),
    .groups = "drop"
  ) %>%
  group_by(pre_post) %>%
  mutate(
    rank_hump = row_number(mean_relative_hump_risk_n),
    rank_blwh = row_number(mean_relative_blwh_risk_n),
    rank_both = row_number(mean_relative_both_risk_n),
    rank_dollars = row_number(mean_relative_dollars),
    rank_pounds = row_number(mean_relative_pounds),
    rank_cbhump = row_number(mean_cbhump),
    rank_cbblwh = row_number(mean_cbblwh),
    .groups = "drop"
  ) %>%
  select(pre_post, 
         scenario_number, pretty_scenario_names,
         rank_hump,rank_blwh,rank_both,rank_dollars,rank_pounds,
         mean_n_risk_humpback, mean_relative_hump_risk_n,
         mean_n_risk_blue, mean_relative_blwh_risk_n,
         mean_n_risk_both, mean_relative_both_risk_n,
         mean_DCRB_rev, mean_relative_dollars,
         mean_DCRB_lbs, mean_relative_pounds,
         mean_cbhump,
         mean_cbblwh
  )

#glimpse(annual_df_ranks_period_supplement_cenCA)

# make a df sorted by rank risk reduction to humps
annual_df_ranks_period_supplement_cenCA_hump <- annual_df_ranks_period_supplement_cenCA %>%
  arrange(
    -mean_n_risk_humpback,
    -mean_n_risk_blue,
    -mean_n_risk_both,
    -mean_DCRB_rev,
    -mean_DCRB_lbs,
    .groups = "drop"
  ) %>%
  mutate(scenario_rank = row_number())

#glimpse(annual_df_ranks_period_supplement_cenCA_hump)

#annual_df_ranks_period_supplement_cenCA_hump %>% knitr::kable()

annual_df_ranks_period_supplement_cenCA_hump %>% 
  select(pre_post, scenario_number, pretty_scenario_names,
         rank_hump, rank_blwh, rank_both, rank_dollars, rank_pounds) %>% 
  knitr::kable()

annual_df_ranks_period_supplement_cenCA_hump %>% 
  select(pre_post, scenario_number, pretty_scenario_names,
         mean_n_risk_humpback, mean_n_risk_blue, mean_n_risk_both, mean_relative_hump_risk_n, mean_relative_blwh_risk_n, mean_relative_both_risk_n,
         mean_DCRB_rev, mean_DCRB_lbs, mean_relative_dollars, mean_relative_pounds,
         mean_cbhump, mean_cbblwh) %>% 
  knitr::kable()

write_csv(annual_df_ranks_period_supplement_cenCA_hump, 
          here::here(
            "tradeoffs",
            "full_analysis_samhourietal",
            "output_dfs",
            "annual_df_ranks_period_supplement_cenCA_hump.csv")
          )



```

## Make df's to compare scales of management in different periods

Focal scenarios for main text
```{r rank_scenarios, echo=FALSE}

### compare risk based on scale of mgmt scenario

# 1a) focal scenarios for main text
scale_compare_focal_scenarios <- annual_statewide_df_focal_scenarios %>%
  filter(!stringr::str_detect(pretty_scenario_names, 'Status Quo|BIA')) %>%
  filter(!stringr::str_detect(pretty_scenario_names, 'Delay CenCA, Early Closure Statewide')) %>% 
  filter(!stringr::str_detect(pretty_scenario_names, 'Delay Statewide, Early Closure CenCA')) %>%
  filter(!stringr::str_detect(pretty_scenario_names, 'CenCA 50% Effort Reduction, Statewide <30 fathom Depth Restriction Apr-Jul')) %>%
  mutate(
  
    scale = ifelse(delay.region == "All" | closure.region == "All" | reduction.after.region == "All",
                 "Statewide",
                 ifelse(delay.region == "CenCA" | closure.region == "CenCA" | reduction.after.region == "CenCA",
                        "Central California Only", NA)),
    
  scenario_type = factor(
      case_when(
      pretty_scenario_names == "Status Quo" ~ "Status Quo",
      stringr::str_detect(pretty_scenario_names, 'Early Closure')  & stringr::str_detect(pretty_scenario_names, 'Delay') ~ "Delay and Early Closure",
      stringr::str_detect(pretty_scenario_names, 'Early Closure')  & stringr::str_detect(pretty_scenario_names, 'Delay', negate = TRUE) ~ "Early Closure",
      stringr::str_detect(pretty_scenario_names, 'Early Closure', negate = TRUE)  & stringr::str_detect(pretty_scenario_names, 'Delay') ~ "Delayed Opening",
      stringr::str_detect(pretty_scenario_names, 'Depth Restriction') & stringr::str_detect(pretty_scenario_names, 'Effort Reduction') ~ "Depth and Gear Restrictions",
      stringr::str_detect(pretty_scenario_names, 'Depth Restriction') & stringr::str_detect(pretty_scenario_names, 'Effort Reduction', negate = TRUE) ~ "Depth Restriction",
      stringr::str_detect(pretty_scenario_names, 'Depth Restriction', negate = TRUE) & stringr::str_detect(pretty_scenario_names, 'Effort Reduction') ~ "Gear Restriction",
      TRUE ~ "Delay and Early Closure"
    ),
    levels = c("Status Quo", "Depth Restriction", "Gear Restriction",
               "Depth and Gear Restrictions", "Delayed Opening", "Early Closure",
               "Delay and Early Closure"))
  ) %>%
  group_by(pretty_scenario_names, pre_post, scale, scenario_type) %>%
  summarise(
    mean_n_risk_blue = mean(n_risk_blue),
    mean_n_risk_humpback = mean(n_risk_humpback),
    mean_n_risk_both = mean(n_risk_both),
    mean_DCRB_rev = mean(DCRB_rev),
    
    se_n_risk_blue = sd(n_risk_blue)/sqrt(n()),
    se_n_risk_humpback = sd(n_risk_humpback)/sqrt(n()),
    se_n_risk_both = sd(n_risk_both)/sqrt(n()),
    se_DCRB_rev = sd(DCRB_rev)/sqrt(n()),
    
    mean_relative_hump_risk_n = mean(relative_hump_risk_n),
    mean_relative_blwh_risk_n = mean(relative_blwh_risk_n),
    mean_relative_both_risk_n = mean(relative_both_risk_n),
    mean_relative_dollars = mean(relative_dollars),
    
    se_relative_hump_risk_n = sd(relative_hump_risk_n)/sqrt(n()),
    se_relative_blwh_risk_n = sd(relative_blwh_risk_n)/sqrt(n()),
    se_relative_both_risk_n = sd(relative_both_risk_n)/sqrt(n()),
    se_relative_dollars = sd(relative_dollars)/sqrt(n()),
    
    .groups = "drop"
  ) 
  
#glimpse(scale_compare_focal_scenarios)

# scale_compare_focal_scenarios %>% knitr::kable()

scale_compare_focal_scenarios %>% 
  select(pre_post, pretty_scenario_names, scale, scenario_type,
         mean_n_risk_humpback, mean_n_risk_blue, mean_n_risk_both, mean_relative_hump_risk_n, mean_relative_blwh_risk_n, mean_relative_both_risk_n,
         mean_DCRB_rev, mean_relative_dollars) %>%  #mean_DCRB_lbs, mean_relative_pounds,
  knitr::kable()

write_rds(scale_compare_focal_scenarios, 
          here::here(
            "tradeoffs",
            "Management scenarios",
            "scale_compare_all_vessels.rds")
          )
```


Focal scenarios for supplement
```{r rank_scenarios8, echo=FALSE}

# 1b) focal scenarios for supplement
scale_compare_supplement <- annual_statewide_df_supplement %>%
  filter(!stringr::str_detect(pretty_scenario_names, 'Status Quo|BIA')) %>%
  filter(!stringr::str_detect(pretty_scenario_names, 'Delay CenCA, Early Closure Statewide')) %>% 
  filter(!stringr::str_detect(pretty_scenario_names, 'Delay Statewide, Early Closure CenCA')) %>%
  filter(!stringr::str_detect(pretty_scenario_names, 'CenCA 50% Effort Reduction, Statewide <30 fathom Depth Restriction Apr-Jul')) %>%
  mutate(
  
    scale = ifelse(delay.region == "All" | closure.region == "All" | reduction.after.region == "All",
                 "Statewide",
                 ifelse(delay.region == "CenCA" | closure.region == "CenCA" | reduction.after.region == "CenCA",
                        "Central California Only", NA)),
    
  scenario_type = factor(
      case_when(
      pretty_scenario_names == "Status Quo" ~ "Status Quo",
      stringr::str_detect(pretty_scenario_names, 'Early Closure')  & stringr::str_detect(pretty_scenario_names, 'Delay') ~ "Delay and Early Closure",
      stringr::str_detect(pretty_scenario_names, 'Early Closure')  & stringr::str_detect(pretty_scenario_names, 'Delay', negate = TRUE) ~ "Early Closure",
      stringr::str_detect(pretty_scenario_names, 'Early Closure', negate = TRUE)  & stringr::str_detect(pretty_scenario_names, 'Delay') ~ "Delayed Opening",
      stringr::str_detect(pretty_scenario_names, 'Depth Restriction') & stringr::str_detect(pretty_scenario_names, 'Effort Reduction') ~ "Depth and Gear Restrictions",
      stringr::str_detect(pretty_scenario_names, 'Depth Restriction') & stringr::str_detect(pretty_scenario_names, 'Effort Reduction', negate = TRUE) ~ "Depth Restriction",
      stringr::str_detect(pretty_scenario_names, 'Depth Restriction', negate = TRUE) & stringr::str_detect(pretty_scenario_names, 'Effort Reduction') ~ "Gear Restriction",
      TRUE ~ "Delay and Early Closure"
    ),
    levels = c("Status Quo", "Depth Restriction", "Gear Restriction",
               "Depth and Gear Restrictions", "Delayed Opening", "Early Closure",
               "Delay and Early Closure"))
  ) %>%
  group_by(pretty_scenario_names, pre_post, scale, scenario_type) %>%
  summarise(
    mean_n_risk_blue = mean(n_risk_blue),
    mean_n_risk_humpback = mean(n_risk_humpback),
    mean_n_risk_both = mean(n_risk_both),
    mean_DCRB_rev = mean(DCRB_rev),
    
    se_n_risk_blue = sd(n_risk_blue)/sqrt(n()),
    se_n_risk_humpback = sd(n_risk_humpback)/sqrt(n()),
    se_n_risk_both = sd(n_risk_both)/sqrt(n()),
    se_DCRB_rev = sd(DCRB_rev)/sqrt(n()),
    
    mean_relative_hump_risk_n = mean(relative_hump_risk_n),
    mean_relative_blwh_risk_n = mean(relative_blwh_risk_n),
    mean_relative_both_risk_n = mean(relative_both_risk_n),
    mean_relative_dollars = mean(relative_dollars),
    
    se_relative_hump_risk_n = sd(relative_hump_risk_n)/sqrt(n()),
    se_relative_blwh_risk_n = sd(relative_blwh_risk_n)/sqrt(n()),
    se_relative_both_risk_n = sd(relative_both_risk_n)/sqrt(n()),
    se_relative_dollars = sd(relative_dollars)/sqrt(n()),
    
    .groups = "drop"
  ) 
  
#glimpse(scale_compare_supplement)

#scale_compare_supplement %>% knitr::kable()

scale_compare_supplement %>% 
  select(pre_post, pretty_scenario_names, scale, scenario_type,
         mean_n_risk_humpback, mean_n_risk_blue, mean_n_risk_both, mean_relative_hump_risk_n, mean_relative_blwh_risk_n, mean_relative_both_risk_n,
         mean_DCRB_rev, mean_relative_dollars) %>%  #mean_DCRB_lbs, mean_relative_pounds,
  knitr::kable()

write_rds(scale_compare_supplement, 
          here::here(
            "tradeoffs",
            "Management scenarios",
            "scale_compare_all_vessels_supplement.rds")
          )


```

# 5. CALCULATE CHANGES IN RISK AND REVENUE BY PERIOD AND SCENARIO
+ repeat for statewide only and cenCA only scenarios
+ % change in each scenario's impact on risk and revenue between 2009-14 and 2014-18 periods

## For each scenario, calculate % change in risk/revenue between periods, all scenarios

Focus on middle row for each scenario number, which reflects the percent improvement in risk/rev/lbs/ce relative to the status quo during the MHW 2014-18 compared to the pre-MHW period 2009-2014.

```{r change1, echo=FALSE}

# what is the % difference in risk, revenue, etc compared to SQ, between pre and MHW periods, under each scenario?
# i.e., are some scenarios disproportionately more effective during the MHW?

percent_diff_period_df <- annual_df_ranks_period_hump %>%
  arrange(scenario_number, pre_post) %>%
  group_by(scenario_number, pretty_scenario_names) %>%
  summarise(
    percent_diff_relative_hump_risk_n = 100 * (mean_relative_hump_risk_n-lag(mean_relative_hump_risk_n))/lag(mean_relative_hump_risk_n),
    percent_diff_relative_blwh_risk_n = 100 * (mean_relative_blwh_risk_n-lag(mean_relative_blwh_risk_n))/lag(mean_relative_blwh_risk_n),
    percent_diff_relative_blwh_risk_n = 100 * (mean_relative_both_risk_n-lag(mean_relative_both_risk_n))/lag(mean_relative_both_risk_n),
    percent_diff_relative_dollars = 100 * (mean_relative_dollars-lag(mean_relative_dollars))/lag(mean_relative_dollars),
    percent_diff_relative_pounds = 100 * (mean_relative_pounds-lag(mean_relative_pounds))/lag(mean_relative_pounds),
    percent_diff_cbhump = 100 * (mean_cbhump-lag(mean_cbhump))/lag(mean_cbhump),
    percent_diff_cbblwh = 100 * (mean_cbblwh-lag(mean_cbblwh))/lag(mean_cbblwh),
    .groups = "drop") %>%
  bind_cols(annual_df_ranks_period_hump %>% select(scenario_number, pre_post) %>% arrange(scenario_number, pre_post) %>% select(-scenario_number))

#glimpse(percent_diff_period_df)

percent_diff_period_df %>% knitr::kable()

write_csv(percent_diff_period_df, 
          here::here(
            "tradeoffs",
            "Management scenarios",
            "percent_diff_period_df.csv")
          )

```

## For each scenario, calculate % change in risk/revenue between periods, statewide focal scenarios only
```{r change2, echo=FALSE}

# what is the % difference in risk, revenue, etc compared to SQ, between pre and MHW periods, under each scenario?
# i.e., are some scenarios disproportionately more effective during the MHW?

percent_diff_period_focal_statewide_df <- annual_df_ranks_period_focal_statewide_hump %>%
  arrange(scenario_number, pre_post) %>%
  group_by(scenario_number, pretty_scenario_names) %>%
  summarise(
    percent_diff_relative_hump_risk_n = 100 * (mean_relative_hump_risk_n-lag(mean_relative_hump_risk_n))/lag(mean_relative_hump_risk_n),
    percent_diff_relative_blwh_risk_n = 100 * (mean_relative_blwh_risk_n-lag(mean_relative_blwh_risk_n))/lag(mean_relative_blwh_risk_n),
    percent_diff_relative_blwh_risk_n = 100 * (mean_relative_both_risk_n-lag(mean_relative_both_risk_n))/lag(mean_relative_both_risk_n),
    percent_diff_relative_dollars = 100 * (mean_relative_dollars-lag(mean_relative_dollars))/lag(mean_relative_dollars),
    percent_diff_relative_pounds = 100 * (mean_relative_pounds-lag(mean_relative_pounds))/lag(mean_relative_pounds),
    percent_diff_cbhump = 100 * (mean_cbhump-lag(mean_cbhump))/lag(mean_cbhump),
    percent_diff_cbblwh = 100 * (mean_cbblwh-lag(mean_cbblwh))/lag(mean_cbblwh),
    .groups = "drop") %>%
  bind_cols(annual_df_ranks_period_focal_statewide_hump %>% select(scenario_number, pre_post) %>% arrange(scenario_number, pre_post) %>% select(-scenario_number))

#glimpse(percent_diff_period_focal_statewide_df)

percent_diff_period_focal_statewide_df %>% knitr::kable()

write_csv(percent_diff_period_focal_statewide_df, 
          here::here(
            "tradeoffs",
            "Management scenarios",
            "percent_diff_period_focal_statewide_df.csv")
          )

```

## For each scenario, calculate % change in risk/revenue between periods, cenCA focal scenarios only
```{r change3, echo=FALSE}

# what is the % difference in risk, revenue, etc compared to SQ, between pre and MHW periods, under each scenario?
# i.e., are some scenarios disproportionately more effective during the MHW?

percent_diff_period_focal_cenCA_df <- annual_df_ranks_period_focal_cenCA_hump %>%
  arrange(scenario_number, pre_post) %>%
  group_by(scenario_number, pretty_scenario_names) %>%
  summarise(
    percent_diff_relative_hump_risk_n = 100 * (mean_relative_hump_risk_n-lag(mean_relative_hump_risk_n))/lag(mean_relative_hump_risk_n),
    percent_diff_relative_blwh_risk_n = 100 * (mean_relative_blwh_risk_n-lag(mean_relative_blwh_risk_n))/lag(mean_relative_blwh_risk_n),
    percent_diff_relative_blwh_risk_n = 100 * (mean_relative_both_risk_n-lag(mean_relative_both_risk_n))/lag(mean_relative_both_risk_n),
    percent_diff_relative_dollars = 100 * (mean_relative_dollars-lag(mean_relative_dollars))/lag(mean_relative_dollars),
    percent_diff_relative_pounds = 100 * (mean_relative_pounds-lag(mean_relative_pounds))/lag(mean_relative_pounds),
    percent_diff_cbhump = 100 * (mean_cbhump-lag(mean_cbhump))/lag(mean_cbhump),
    percent_diff_cbblwh = 100 * (mean_cbblwh-lag(mean_cbblwh))/lag(mean_cbblwh),
    .groups = "drop") %>%
  bind_cols(annual_df_ranks_period_focal_cenCA_hump %>% select(scenario_number, pre_post) %>% arrange(scenario_number, pre_post) %>% select(-scenario_number))

#glimpse(percent_diff_period_focal_cenCA_df)

percent_diff_period_focal_cenCA_df %>% knitr::kable()

write_csv(percent_diff_period_focal_cenCA_df, 
          here::here(
            "tradeoffs",
            "Management scenarios",
            "percent_diff_period_focal_cenCA_df.csv")
          )

```

# 6. QUERY ENTANGLEMENT DATA

## Entanglements by period and species
```{r entanglements, echo=FALSE}

entanglement_df_annual_complete <- read_csv(path_entanglement_file2)

entanglement_df_annual_complete %>% 
  mutate(
    period = case_when(
      Year == 2009 | Year == 2010 | Year == 2011 | Year == 2012 | Year == 2013 ~ "pre",
      Year == 2014 | Year == 2015 | Year == 2016 | Year == 2017 | Year == 2018 ~ "mhw",
      Year == 2019 ~ "post",
      TRUE ~ "OTHER"
    )
  ) %>%
  filter( Year >= 2009) %>%
  group_by(period, species) %>%
  summarise(
    total = sum(count)
  ) %>%
  arrange(
    species,
    factor(period, levels = c("pre", "mhw", "post"))
  ) %>%
  knitr::kable()

```

## Entanglements by period, all species combined
```{r entanglements2, echo=FALSE}

entanglement_df_annual_complete <- read_csv(path_entanglement_file2)

entanglement_df_annual_complete %>% 
  mutate(
    period = case_when(
      Year == 2009 | Year == 2010 | Year == 2011 | Year == 2012 | Year == 2013 ~ "pre",
      Year == 2014 | Year == 2015 | Year == 2016 | Year == 2017 | Year == 2018 ~ "mhw",
      Year == 2019 ~ "post",
      TRUE ~ "OTHER"
    )
  ) %>%
  filter( Year >= 2009) %>%
  group_by(period) %>%
  summarise(
    total = sum(count)
  ) %>%
  arrange(
    factor(period, levels = c("pre", "mhw", "post"))
  ) %>%
  knitr::kable()

```

# 7. QUERY FISHING DATA

```{r fishing, include=FALSE}


# most of this code originally from query_fishing_data.R
# this takes a few min to read in
dcrb_vms_tix_analysis <- read_rds(path_dcrb_vms_tix_analysis)
glimpse(dcrb_vms_tix_analysis)

# check years included
unique(dcrb_vms_tix_analysis$year)

#####################################################
# filter the data using the control variables below
# years <- seq(2009,2019,1)

state_agency_code <- "C"
#state <- "CA"
target_rev <- "DCRB"
target_lbs <- "DCRB"
region_north <- c(1040, 1041, 1042)
vessel_size_category_break <- 40
winter_months <- c("November", "December", "January", "February", "March")
crab_months <- c(1:7,11:12)
removal_types <- c("COMMERCIAL (NON-EFP)", "COMMERCIAL(DIRECT SALES)", "UNKNOWN")
max_speed <- 4.11556
min_speed <- 0

#ports <- c("CCA", "ERA", "BGA","BDA", "SFA", "MRA", "MNA")
#region_south <- c(1035, 1036, 1037, 1038)
#####################################################

#####################################################

# subset the data based on above queries. CA only
dcrb_ca_vms_tix_analysis <- dcrb_vms_tix_analysis %>%
  filter(agency_code == state_agency_code) %>%
  #filter(agency_code == state_agency_code & STATE == state) %>%
  filter(TARGET_rev == target_rev | TARGET_lbs == target_lbs) %>%
  filter(removal_type_name %in% removal_types) %>%
  filter(CA_OFFSHOR != -999) %>%
  filter(DEPTH_CATM == "0-100m" | DEPTH_CATM == "100-150m") %>%
  #filter(NGDC_M <= 0 & NGDC_M >= -12000) %>% # considered adding 052220. also see "evaluate fishing depths.R". probably unnecessary because NGDC_M refers to centroid depth and we can match depths by 5km grid cell later
  filter(avg_speed_recalc <= max_speed & avg_speed_recalc >= min_speed) %>%
  #filter(is.na(in_port) == TRUE) %>% # only removes ~4000 records
  #filter(port_group_code %in% ports) %>%
  mutate(#GRID5KM_ID = as.character(GRID5KM_ID),
    Region = ifelse(CA_OFFSHOR %in% region_north,
                    "NorCA","CenCA"),
    BIA_mn_noNAs = ifelse(is.na(BIA_mn)==TRUE,0,BIA_mn),
    BIA_bm_noNAs = ifelse(is.na(BIA_bm)==TRUE,0,BIA_bm),
    #year = lubridate::year(westcoastdate_notime), # removed because it is included with the vessel_lengths step
    year_month = paste0(lubridate::year(westcoastdate_notime),"_", substr(lubridate::ymd(westcoastdate_notime),6,7)), # substr() ensures month is a 2 digit value
    month = lubridate::month(westcoastdate_notime, label=TRUE, abbr = FALSE),
    month_as_numeric = month(westcoastdate_notime),
    week_of_year = week(westcoastdate_notime),
    day_of_year = yday(westcoastdate_notime),
    season = as.character(ifelse(month %in% winter_months, "Winter", "Spring-Summer")),
    crab_year = ifelse(
      month_as_numeric >= 11, paste0(year,"_",1+year), paste0(year - 1,"_",year)
    ),
    Vessel_Size = as.character(ifelse(FINAL_LENGTH >= vessel_size_category_break, paste0(">=",vessel_size_category_break, " ft"),paste0("<",vessel_size_category_break, " ft"))),
    BIA_bm_or_mn = ifelse(BIA_bm_noNAs !=0 | BIA_mn_noNAs != 0, "Inside BIA","Outside BIA")#,
  ) %>%
  filter(westcoastdate_notime >= as.Date("2009-11-01") & westcoastdate_notime <= as.Date("2019-08-01")) %>%
  filter(month_as_numeric %in% crab_months) %>%
  dplyr::select(
    Rec_ID, VMS_RECNO, drvid, Vessel_Size,
    westcoastdate_notime, year, crab_year, year_month, month, month_as_numeric, week_of_year, day_of_year, season,
    GRID5KM_ID, BAND_25KM, BAND_50KM, CA_OFFSHOR, Region, BIA_mn_noNAs, BIA_bm_noNAs, BIA_bm_or_mn, pacfin_port_code, port_group_code, DEPTH_CATM, NGDC_M,
    TARGET_lbs, TARGET_rev, DCRB_lbs, DCRB_revenue
  )

#glimpse(dcrb_ca_vms_tix_analysis)
```

## All 10 Years
Number unique vessels  tix, geolocations, and total revenue and landings
(VMS-tracked only)

```{r fishing2, echo=FALSE}
# number of unique vessels
total_num_vessels <- length(unique(dcrb_ca_vms_tix_analysis$drvid)) #283

# number of unique tix
total_num_tix <- length(unique(dcrb_ca_vms_tix_analysis$Rec_ID)) #16182

# number of unique geolocations
total_num_locations <- length(unique(dcrb_ca_vms_tix_analysis$VMS_RECNO)) #369688

# total revenue
total_rev <- sum(dcrb_ca_vms_tix_analysis$DCRB_revenue) #6563123120

# total landings
total_landings <- sum(dcrb_ca_vms_tix_analysis$DCRB_lbs) #2234063912


data.frame(total_num_vessels,
           total_num_tix,
           total_num_locations,
           total_rev,
           total_landings) %>%
  knitr::kable()


```

## By crab year

Number unique vessels  tix, geolocations, and total revenue and landings by crab year
(VMS-tracked only)
```{r fishing3, echo=FALSE}

# summed revenue of VMS vessels each crab_year
#unique(dcrb_ca_vms_tix_analysis$crab_year)

crabyear_summaries <- dcrb_ca_vms_tix_analysis %>%
  mutate(
  plotting_year = as.numeric(substr(crab_year,6,9))
  ) %>%
  group_by(crab_year, plotting_year) %>%
  summarise(
    num_vessels = length(unique(drvid)),
    num_tix = length(unique(Rec_ID)),
    num_locations = length(unique(VMS_RECNO)),
    sum_DCRB_rev = sum(DCRB_revenue),
    sum_DCRB_lbs = sum(DCRB_lbs)
    
  ) 

crabyear_summaries %>%
  knitr::kable()

write_csv(crabyear_summaries, 
          here::here(
            "tradeoffs",
            "full_analysis_samhourietal",
            "output_dfs",
            "crabyear_summaries.csv"
          )
)

crabyear_summaries %>%
  summarise(
    mean_vessels = mean(num_vessels),
    min_vessels = min(num_vessels),
    max_vessels = max(num_vessels)
  ) %>%
  knitr::kable()


```

## By period

Number unique vessels  tix, geolocations, and total revenue and landings by period
(VMS-tracked only)
```{r fishing4, echo=FALSE}

# summed revenue of VMS vessels 2009-14 vs 2014-18
#unique(dcrb_ca_vms_tix_analysis$crab_year)

period_summaries <- dcrb_ca_vms_tix_analysis %>%
  mutate(
    period = case_when(
      crab_year %in% c("2009_2010","2010_2011","2011_2012","2012_2013","2013_2014") ~ "pre",
      crab_year %in% c("2014_2015","2016_2017","2015_2016","2017_2018") ~ "MHW",
      TRUE ~ "post"
    ) 
  ) %>%
  group_by(period) %>%
  summarise(
    num_vessels = length(unique(drvid)),
    num_tix = length(unique(Rec_ID)),
    num_locations = length(unique(VMS_RECNO)),
    sum_DCRB_rev = sum(DCRB_revenue),
    sum_DCRB_lbs = sum(DCRB_lbs)
    
  ) %>%
  mutate(
    annual_rev = case_when(
      period == "MHW" ~ sum_DCRB_rev/4,
      period == "pre" ~ sum_DCRB_rev/5,
      period == "post" ~ sum_DCRB_rev/1
      #TRUE ~ "ERROR"
    )
  ) %>%
    arrange(
    factor(period, levels = c("pre", "MHW", "post"))
           ) 

annual_percent_rev_change_MHW_v_pre <- 100 * ( period_summaries$annual_rev[which(period_summaries$period == "MHW")] - period_summaries$annual_rev[which(period_summaries$period == "pre")] ) / period_summaries$annual_rev[which(period_summaries$period == "pre")]

period_summaries %>%
  knitr::kable()

print("annual_percent_rev_change_MHW_v_pre")
annual_percent_rev_change_MHW_v_pre


```

## By period and region

Number unique vessels  tix, geolocations, and total revenue and landings by period and region
(VMS-tracked only)
```{r fishing5, echo=FALSE}

# summed revenue of VMS vessels 2009-14 vs 2014-18
#unique(dcrb_ca_vms_tix_analysis$crab_year)

period_region_summaries <- dcrb_ca_vms_tix_analysis %>%
  mutate(
    period = case_when(
      crab_year %in% c("2009_2010","2010_2011","2011_2012","2012_2013","2013_2014") ~ "pre",
      crab_year %in% c("2014_2015","2016_2017","2015_2016","2017_2018") ~ "MHW",
      TRUE ~ "post"
    ) 
  ) %>%
  group_by(period, Region) %>%
  summarise(
    num_vessels = length(unique(drvid)),
    num_tix = length(unique(Rec_ID)),
    num_locations = length(unique(VMS_RECNO)),
    sum_DCRB_rev = sum(DCRB_revenue),
    sum_DCRB_lbs = sum(DCRB_lbs)
    
  ) %>%
  #group_by(Region) %>%
  mutate(
    annual_rev = case_when(
      period == "MHW" ~ sum_DCRB_rev/4,
      period == "pre" ~ sum_DCRB_rev/5,
      period == "post" ~ sum_DCRB_rev/1
      #TRUE ~ "ERROR"
    )
  ) %>%
    arrange(
      Region,
    factor(period, levels = c("pre", "MHW", "post"))
           ) 

annual_percent_rev_change_MHW_v_pre_region <- 100 * ( period_region_summaries$annual_rev[which(period_region_summaries$period == "MHW")] - period_region_summaries$annual_rev[which(period_region_summaries$period == "pre")] ) / period_region_summaries$annual_rev[which(period_region_summaries$period == "pre")]

period_region_summaries %>%
  knitr::kable()

print("annual_percent_rev_change_MHW_v_pre_region")
annual_percent_rev_change_MHW_v_pre_region


```


Fish ticket IDs and Vessels landing crab after Jun 30 in CenCA and after Jul 15 in NorCA
Code creates output csv (table redacated because it contains confidential informatoion)
```{r fishing 6, echo=FALSE}

dcrb_ca_vms_tix_analysis_inseason <-  dcrb_ca_vms_tix_analysis %>%
  mutate(
    monthday = format(westcoastdate_notime, "%m%d"),
    in_season = case_when(
      Region == "CenCA" & monthday >= "1112" | monthday <= "0630" ~ "Y",
      Region == "NorCA" & monthday >= "1128" | monthday <= "0715" ~ "Y",
      TRUE ~ "N"
    )
  )

dcrb_ca_vms_tix_analysis_inseason_unique <- dcrb_ca_vms_tix_analysis_inseason %>%
  filter(in_season != "Y") %>%
  distinct(crab_year, westcoastdate_notime, Region, Rec_ID, drvid)

dcrb_ca_vms_tix_analysis_inseason_unique %>% select(-drvid) %>%
  knitr::kable()

write_csv(dcrb_ca_vms_tix_analysis_inseason_unique %>% select(-drvid),
          here::here(
            "tradeoffs",
            "full_analysis_samhourietal",
            "output_dfs",
            "dcrb_ca_vms_tix_analysis_outseason_unique.csv"
          )
)
  

```

Total revenue and landings by crab year and period for all fish tix
(not yet analyzed)

# 8. QUERY WHALE MODEL OUTPUTS
+ predicted occurrence/density by period, statewide and by region
[ not yet included]

````{r whales, echo=FALSE}

grid.key <- read_rds(path_grid_key)

# dont worry about NAs. see end of chunk make_dfs3 where we check for NAs for whales. some NAs but they are the same NAs as in x.whale (where blue and hump predictions do not overlap). ok to proceed
# sq_maps_grid_noNA %>%
#   summarise(
#     hump_na = length(which(is.na(Humpback_dens))),
#     blue_na = length(which(is.na(Blue_occurrence)))
#   )
# which(is.na(sq_maps_grid_noNA$Humpback_dens)

# coastwide
sq_maps_grid_noNA %>%
  left_join(grid.key, by = "GRID5KM_ID") %>% 
  group_by(time_period) %>%
  summarise(
    mean_hump_dens = mean(Humpback_dens, na.rm=TRUE),
    median_hump_dens = median(Humpback_dens, na.rm=TRUE),
    mean_blue_occur = mean(Blue_occurrence, na.rm=TRUE),
    median_blue_occur = median(Blue_occurrence, na.rm=TRUE),
    mean_pings = mean(Num_DCRB_VMS_pings, na.rm=TRUE),
    median_pings = median(Num_DCRB_VMS_pings, na.rm=TRUE)
  ) %>%
  arrange(
    time_period
  ) %>%
  st_drop_geometry %>%
  knitr::kable()

# by region
sq_maps_grid_noNA %>%
  left_join(grid.key, by = "GRID5KM_ID") %>% 
  mutate(
    Region = ifelse(Region == "OR", "NorCA", Region)
    ) %>%
  group_by(time_period, Region) %>%
  summarise(
    mean_hump_dens = mean(Humpback_dens, na.rm=TRUE),
    median_hump_dens = median(Humpback_dens, na.rm=TRUE),
    mean_blue_occur = mean(Blue_occurrence, na.rm=TRUE),
    median_blue_occur = median(Blue_occurrence, na.rm=TRUE),
    mean_pings = mean(Num_DCRB_VMS_pings, na.rm=TRUE),
    median_pings = median(Num_DCRB_VMS_pings, na.rm=TRUE)
  ) %>%
  arrange(
    Region, time_period
  ) %>%
  st_drop_geometry %>%
  knitr::kable()

```
