---
title: "Samhouri et al tradeoff analysis run analysis"
author: "Jameal Samhouri"
date: "11/4/2020"
output: 
  html_document:
      toc: true
      toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This document runs the analysis included in the Samhouri et al. tradeoff manuscript.

See pseudocode in "Full workflow for risk assessment and tradeoff analysis" google doc. https://docs.google.com/document/d/1T8S5rEOACMCeGa5efS5BnEQY4gOw-PbxmvILJvShsQo/edit


1) MAKE SCENARIOS
RUN SCENARIOS
MAKE DF'S FOR TIME SERIES OF RISK, REVENUE, AND LANDINGS
- status quo as separate df, all scenarios in one df, and scale_compare df
- repeat for sm vessels
CALCULATE SCENARIO RANKS
- one df by year, one by period
- repeat for statewide only and cenCA only scenarios
- ranks based on hump risk, blwh risk, DCRB $, DCRB lbs, cost effectiveness for humps, cost effectiveness for blues
CALCULATE CHANGES IN RISK AND REVENUE BY PERIOD AND SCENARIO
- repeat for statewide only and cenCA only scenarios
- % change in each scenario's impact on risk and revenue between 2009-14 and 2014-18 periods
QUERY FISHING DATA

QUERY ENTANGLEMENT DATA
QUERY WHALE MODEL OUTPUTS
- predicted occurrence/density by period, statewide and by region

Fig 1 data. Whale and fishing data on 5km grid, averaged by time period
Fig 2 data. Entanglement reports, DCRB revenue by crab year. whale risk is calculated in the scenario_analysis step
Fig 3 data. Whale risk and DCRB revenue by scenario calculated in the scenario_analysis step
Fig 4 data. Whale risk and DCRB revenue by scenario calculated in the scenario_analysis step

## Prep libraries and file paths

```{r prep, include=FALSE}

# just getting libraries loaded and file paths sorted

library(tidyverse)
library(foreign) # read.dbf()
library(here)
library(lubridate)
library(sf)
library(viridis)
library(ggrepel)
library(magrittr)
library(ggerr)
library(scales)
library(maps)
library(rnaturalearth)
library(gridExtra)
library(ggpubr)
library(knitr)

source(here::here("User_script_local.R"))
if (user == "JS") {
  #flag.save = TRUE # option to save all outputs or not
  
  path_entanglement_file2 <- paste0(here::here(), "/tradeoffs/full_analysis_samhourietal/output_dfs/entanglement_df_annual_complete.csv")
  
  path_Grid_5km_landerased <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/5x5 Grid/Grid_5km_landerased.RDATA"
  
  path_figures <- "/Users/jameal.samhouri/Dropbox/Projects/In progress/RAIMBOW/raimbow/tradeoffs/Management scenarios/figures"
  
  path_rds <- "/Users/jameal.samhouri/Dropbox/Projects/In progress/RAIMBOW/raimbow/tradeoffs/Management scenarios"
  
  path_maps <- "/Users/jameal.samhouri/Dropbox/Projects/In progress/RAIMBOW/raimbow/tradeoffs/maps"
  
  path_status_quo <- paste0("/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/scenario_output_dataframes/status_quo_risk_2009_2019_yr_mth_2020-06-14.rds") # from extract_status_quo_scenario_summary.R
  
  path_status_quo_small_vessels <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/scenario_output_dataframes/status_quo_risk_small_vessels_2009_2019_yr_mth_2020-06-17.rds" # from extract_status_quo_scenario_summary_small.R 
  
  path_whales <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/5x5 Grid/Grid5km_whale.rds"
  
  path_Humpback_5km_long_monthly <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Input_Data/Humpback whale data/Forney et al./Humpback_5km_long_monthly.rds" 
  
  path_BlueWhale_5km_long_monthly <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Input_Data/Blue whale data/Overlay on 5km Grid/BlueWhale_5km_long_monthly.rds"
    
  path_effort_shift_scenariocomparisons_annual_statewide_df_n <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/annual_statewide_effort_shift_scenariocomparisons_n_2020-05-19.rds" # from effort_shift_comparison.R
  
  path_tradeoff_df_effort_shift_scenariocomparisons_n <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/tradeoff_df_effort_shift_scenariocomparisons_n_2020-05-19.rds"
  
  path_scenario_table_all <- "/Users/jameal.samhouri/Dropbox/Projects/In progress/RAIMBOW/raimbow/tradeoffs/Management scenarios/scenario_table_all_2020-06-09.rds"
  
  path_scenario_table_focal_scenarios <- "/Users/jameal.samhouri/Dropbox/Projects/In progress/RAIMBOW/raimbow/tradeoffs/Management scenarios/scenario_table_focal_scenarios_2020-06-12.rds"
  
  path_annual_statewide_df_focal_scenarios <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/annual_statewide_df_focal_scenarios_2020-06-12.rds" 
  
  path_df_tradeoff_focal_scenarios <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/df_tradeoff_focal_scenarios_2020-06-12.rds" 
  
  path_annual_statewide_df_focal_scenarios_small_vessels <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/annual_statewide_df_focal_scenarios_small_vessels_2020-06-16.rds"
  
  path_df_tradeoff_focal_scenarios_small_vessels <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/df_tradeoff_focal_scenarios_small_vessels_2020-06-16.rds"
  
  path_prioritizr_all <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/prioritizr/prioritzr_outputs_all.rds"
  

} else if (user == "SMW") {
  NULL
} else {
  stop("Invalid user")
}


```

# QUERY ENTANGLEMENT DATA

Entanglements by period and species
```{r entanglements, echo=FALSE}

entanglement_df_annual_complete <- read_csv(path_entanglement_file2)

entanglement_df_annual_complete %>% 
  mutate(
    period = case_when(
      Year == 2009 | Year == 2010 | Year == 2011 | Year == 2012 | Year == 2013 ~ "pre",
      Year == 2014 | Year == 2015 | Year == 2016 | Year == 2017 | Year == 2018 ~ "mhw",
      Year == 2019 ~ "post",
      TRUE ~ "OTHER"
    )
  ) %>%
  filter( Year >= 2009) %>%
  group_by(period, species) %>%
  summarise(
    total = sum(count)
  ) %>%
  arrange(
    species,
    factor(period, levels = c("pre", "mhw", "post"))
  ) %>%
  knitr::kable()

```

Entanglements by period, all species combined
```{r entanglements, echo=FALSE}

entanglement_df_annual_complete <- read_csv(path_entanglement_file2)

entanglement_df_annual_complete %>% 
  mutate(
    period = case_when(
      Year == 2009 | Year == 2010 | Year == 2011 | Year == 2012 | Year == 2013 ~ "pre",
      Year == 2014 | Year == 2015 | Year == 2016 | Year == 2017 | Year == 2018 ~ "mhw",
      Year == 2019 ~ "post",
      TRUE ~ "OTHER"
    )
  ) %>%
  filter( Year >= 2009) %>%
  group_by(period) %>%
  summarise(
    total = sum(count)
  ) %>%
  arrange(
    factor(period, levels = c("pre", "mhw", "post"))
  ) %>%
  knitr::kable()

```

# QUERY FISHING DATA

- number vessels, unique tix/geolocations, revenue and landings by period for vms 
- total revenue by crab year and period for all fish tix
- pings by period, statewide and by region
```{r fishing, echo=FALSE}

# pick up here. include correct file paths above and then start making data frames here

```
