---
title: "Entanglements - risk value"
author: "Sam Woodman"
date: '`r format(Sys.Date(), "%d %B %Y")`'
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Goal: Examine monthly risk histograms relative to risk for entanglement reports, using lookback window to account for the uncertainty in when the actual entanglement occurred.

```{r, message=FALSE}
library(gridExtra)
library(lubridate)
library(readxl)
library(tidyverse)

source("Whale_risk_timeseries_funcs.R")
source("User_script_local.R")

if (user == "JS") {
  file.ent
  
} else if (user == "SMW") {
  file.ent <- "../raimbow-local/Data/Entanglements/Entanglements_known_gear_set_location.xlsx"
  path.rdata <- "../raimbow-local/RDATA_files/"
  path.plots1 <- "../raimbow-local/Plots/Entanglements_risk/Entanglements_timeseries/"
  path.plots2 <- "../raimbow-local/Plots/Entanglements_risk/Entanglements_county_timeseries/"
  
} else {
  stop("User not recognized")
}

# Options: "ca-" for CA- regions, or "county" for county-level
reg.flag <- "ca-"
hist.width <- 0.05
```

## Data import and processing

Read in entanglement data, and select relevant columns. Then if necessary, assign each entanglement to a region, using Gear_Set_County to determine region.

```{r}
### Read in and process entanglement data
x.orig <- read_xlsx(file.ent)

x <- x.orig %>% 
  select(CaseID, Year, Month, Report_Lat, Report_Long, Report_County, 
         Gear_Set_County, Gear_Set_Location_State, Timing)


# Get year-month factor levels
ym.levels <- tibble(
  year = c(rep(2009, 2), sapply(2010:2017, rep, 12), rep(2018, 7)),
  month = sprintf("%02d", c(11, 12, rep(1:12, 8), 1:7))
) %>% 
  mutate(ym = paste(year, month, sep = "-"))


# Set variables depending on region level
stopifnot(reg.flag %in% c("ca-", "county"))
if (reg.flag == "ca-") {
  # Large regions
  load(paste0(path.rdata, "Whale_risk_timeseries.Rdata"))
  path.plots <- path.plots1
  fig.h <- 8
  y.max.ent <- 15
  
  reg.list <- list(
    "Del_Norte" = "CA-N", "Humboldt" = "CA-N", "Mendocino" = "CA-N", 
    "Sonoma" = "CA-Cen", "Marin" = "CA-Cen", "San_Francisco" = "CA-Cen",
    "San_Mateo" = "CA-Cen", "Santa_Cruz" = "CA-Cen", "Monterey" = "CA-Cen", 
    "San_Luis_Obispo" = "CA-SCen", "Santa_Barbara" = "CA-SCen", 
    "Ventura" = "CA-S", "Los_Angeles" = "CA-S", "Orange" = "CA-S", 
    "San_Diego" = "CA-S"
  ) # Monterey definitely has some area in CA-SCen too
  
  county.reg <- data.frame(region = t(as.data.frame(reg.list)), 
                           stringsAsFactors = FALSE) %>% 
    rownames_to_column(var = "county") %>% 
    mutate(county = gsub("_", " ", county, ))
  
  reg.levels <- c("WA", "OR", "CA-N", "CA-Cen", "CA-SCen")
  x <- left_join(x, county.reg, by = c("Gear_Set_County" = "county"))
  
  
} else if (reg.flag == "county") {
  # County-level regions
  load(paste0(path.rdata, "Whale_risk_county_timeseries.Rdata"))
  path.plots <- path.plots2
  fig.h <- 10
  y.max.ent <- 10
  
  county.no <- c("Ventura", "Los_Angeles", "Orange", "San_Diego")
  reg.levels <- levels(all.df.summ$region)[!(levels(all.df.summ$region) %in% county.no)]
  x <- x %>% mutate(region = gsub(" ", "_", Gear_Set_County))
  
  
} else {
  stop("reg.flag must be eone of \"ca-\" and \"county\"")
}

# Finish entanglement record prep
x <- x %>% 
  mutate(Gear_Set_County = as.factor(Gear_Set_County), 
         region = factor(region, levels = reg.levels), 
         ym = factor(paste(Year, sprintf("%02d", Month), sep = "-"), levels = ym.levels$ym), 
         ym_date = parse_date_time(ym, "ym")) %>% 
  filter(!is.na(ym))

reg.plot <- sort(unique(as.character(x$region)))
```

Make time series of the entanglement reports

```{r, fig.height=4, fig.width=10}
x.max <- length(levels(x$ym))
x.lab.idx <- seq(1, to = x.max, by = 3)
x.lab <- sort(levels(x$ym))[x.lab.idx]
vert.lines <- seq(0.5, to = x.max, by = 12)

gg0 <- ggplot(x, aes(x = ym, fill = region)) + 
  geom_bar() + 
  geom_vline(xintercept = vert.lines, col = "black", lwd = 0.35)

if (reg.flag == "ca-") {
  gg0 <- gg0 + scale_fill_brewer(palette = "Set1", name = "Region", drop = FALSE)
} else if (reg.flag == "county") {
  gg0 <- gg0 + scale_fill_discrete(name = "region") #, drop = FALSE)
}

gg0 + ggtitle("CA DC confirmed and known-set-location entanglement reports") +
  xlab("Year-Month") +
  ylab("Entanglements") +
  scale_x_discrete(drop = FALSE, breaks = x.lab) +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 90, vjust = 0.4),
        legend.justification = "top")

ggsave(paste0(path.plots, "Ent_report_timeseries.png"))
```

## Compare entanglement reports with risk time series values

Now we load the time series data, and plot histograms of the risk time series values (densities) by region.

```{r, fig.width=14, fig.height=fig.h}
### Add entanglement data to time series data
# Time series data loaded above
all.df.summ.join <- all.df.summ %>% 
  mutate(region = factor(region, levels = levels(x$region))) %>% 
  select(ym, region, risk_sum_dens)


# Plot histogram of time series risk values
facet.nrow <- case_when(reg.flag == "ca-" ~ 1, reg.flag == "county" ~ 3)
ggplot(all.df.summ, aes(x = risk_sum_dens)) +
  geom_histogram(binwidth = hist.width, boundary = 0) + 
  facet_wrap(facets = vars(region), nrow = facet.nrow) +
  xlab("Risk (whales*VMS pings / km2)") + 
  ggtitle("Risk time series")

ggsave(paste0(path.plots, "Risk_hist_all.png"))
```

Next we determine the corresponding risk value for each entanglement report, treating year-months in which no entanglements were reported as zero entanglement reports. Then we can make the same plots, but color-coded to indicate if there was an entanglement report during that year-month.

```{r, fig.width=14, fig.height=fig.h}
x.join <- x %>% 
  select(ym, region, CaseID) %>% 
  left_join(all.df.summ.join, by = c("ym", "region"))

# Contains all risk values and all risk-entanglement report pairs, separately
y <- all.df.summ.join %>% 
  mutate(CaseID = NA) %>% 
  bind_rows(x.join) %>% 
  mutate(e_logical = factor(ifelse(is.na(CaseID), "All", "Entanglements")), 
         ym = factor(ym), 
         region = factor(region, levels(x$region))) %>% 
  arrange(region, ym)

# Plot histogram of time series risk values, color-coded by whether or not there was an entanglement
y %>% 
  filter(region %in% reg.plot) %>%
  mutate(y_max = ifelse(e_logical == "All", 90, y.max.ent)) %>% 
  ggplot(aes(x = risk_sum_dens)) + #, fill = e_logical
  geom_histogram(binwidth = hist.width, position = "dodge", boundary = 0) +
  facet_grid(cols = vars(region), rows = vars(e_logical), 
             scales = "free_y", space = "free_y") +
  geom_blank(aes(y = y_max)) + 
  # facet_wrap(facets = vars(region), nrow = facet.nrow) +
  scale_y_continuous(breaks = seq(0, 100, by = 5)) +
  guides(fill = guide_legend(title = "Risk values")) + 
  xlab("Risk (whales*VMS pings / km2)") + 
  ggtitle("Regional (monthly) risk - all risk values vs risk values for entanglement reports - original (actual) report timing")

ggsave(paste0(path.plots, "Risk_hist_orig.png"))
```

## Compare entanglement reports with risk time series values - pt2

Now, we wish to take into account the fact that we do not know exactly when the entanglement occurred. To do this, we use a 'lookback window', meaning we look back at the maximum risk value that occurred in that region within the past 1 month, 2 months, etc. 

TODO: we also want to look at the sum of the risk values that occurred in that region within the lookback window, but not sure yet how to make a meaningful plot.

```{r, fig.width=14, fig.height=fig.h}
# Set size of lookback window
# lookback.mons <- 3
facet.nrow2 <- case_when(reg.flag == "ca-" ~ 1, reg.flag == "county" ~ 2)

# Loop through possible lookback windows
for (lookback.mons in 0:12) {
  # Prep entanglement records - add idx column
  x.pre <- x %>% 
    select(region, ym_date) %>% 
    mutate(ent_idx = seq_along(ym_date))
  
  # Get all ym values in lookback window, determine which ym value has max 
  x2 <- lapply(0:lookback.mons, function(i) {
    x.pre %>% 
      mutate(ym_date_curr = ym_date %m-% months(i), 
             ym_curr = substr(as.character(ym_date_curr), 1, 7))
  }) %>% 
    bind_rows() %>% 
    select(region, ym = ym_curr, ent_idx) %>% 
    left_join(all.df.summ.join, by = c("ym", "region")) %>% 
    group_by(ent_idx) %>% 
    summarise(r_max = max(risk_sum_dens, na.rm = TRUE), 
              r_sum = sum(risk_sum_dens, na.rm = TRUE), 
              ym_max = ym[which.max(risk_sum_dens)]) %>% 
    ungroup() %>% 
    right_join(x.pre, by = "ent_idx") %>% 
    select(ent_idx, ym = ym_max, region, r_max, r_sum)
  
  # Summarise entanglement count by max ym
  x2.join <- x2 %>% 
    select(ym, region, ent_idx) %>% 
    left_join(all.df.summ.join, by = c("ym", "region"))
  
  y2 <- all.df.summ.join %>% 
    mutate(ent_idx = NA) %>% 
    bind_rows(x2.join) %>% 
    mutate(e_logical = factor(ifelse(is.na(ent_idx), "All", "Entanglements")), 
           ym = factor(ym), 
           region = factor(region, levels(x$region))) %>% 
    arrange(region, ym)
  
  # Plot
  gg2 <- y2 %>% 
    filter(region %in% reg.plot) %>%
    mutate(y_max = ifelse(e_logical == "All", 90, 15)) %>% 
    ggplot(aes(x = risk_sum_dens)) + #, fill = e_logical)) + 
    geom_histogram(binwidth = hist.width, position = "dodge", boundary = 0) +
    facet_grid(cols = vars(region), rows = vars(e_logical), 
               scales = "free_y", space = "free_y") +
    # facet_wrap(facets = vars(region), nrow = facet.nrow2) +
    geom_blank(aes(y = y_max)) + 
    scale_fill_discrete(drop = FALSE, name = "Risk values") + 
    scale_y_continuous(breaks = seq(0, 100, by = 5)) +
    xlab("Risk (whales*VMS pings / km2)") + 
    ggtitle(paste("Regional (monthly) risk - all risk values vs risk values for entanglement reports - ", 
                  lookback.mons, 
                  "month lookback window, max value within window"))
  
  
  # Save plot
  file.plot <- paste0(path.plots, "Risk_hist_lookback_", lookback.mons, "mons.png")
  ggsave(file.plot, gg2)
  
  # print(gg2)
}
```

Experimental: for each (some?) lookback window, how far back is the max risk?

```{r}
x.out <- x.orig %>% 
  left_join(select(x, CaseID, region), by = "CaseID")

for(lookback.mons in seq(0, 12, by = 3)) {
  # Prep entanglement records - add ent_idx column
  x.pre <- x %>% 
    mutate(ent_idx = seq_along(ym_date))
  
  # Get all ym values in lookback window, determine which ym value has max 
  x2 <- lapply(0:lookback.mons, function(i) {
    x.pre %>% 
      mutate(ym_date_curr = ym_date %m-% months(i), 
             ym_curr = substr(as.character(ym_date_curr), 1, 7))
  }) %>% 
    bind_rows() %>% 
    select(region, ym = ym_curr, ent_idx) %>% 
    left_join(all.df.summ.join, by = c("ym", "region")) %>% 
    group_by(ent_idx) %>% 
    summarise(r_max = max(risk_sum_dens, na.rm = TRUE), 
              r_sum = sum(risk_sum_dens, na.rm = TRUE), 
              ym_max = ym[which.max(risk_sum_dens)]) %>% 
    ungroup() %>% 
    right_join(x.pre, by = "ent_idx") %>% 
    select(CaseID, ym_max) %>% 
    set_names("CaseID", paste0("ym_max", lookback.mons))
  
  x.out <- left_join(x.out, x2, by = "CaseID")
}

write.csv(x.out, file = paste0(path.plots, "Ent_ymmax.csv"), row.names = FALSE)

table(x.out$ym_max0, as.character(x.out$region))
table(x.out$ym_max3, as.character(x.out$region))
table(x.out$ym_max6, as.character(x.out$region))
table(x.out$ym_max9, as.character(x.out$region))
table(x.out$ym_max12, as.character(x.out$region))
```
