---
title: "Humpback risk - management scenarios"
author: "Sam Woodman"
date: '`r format(Sys.Date(), "%d %B %Y")`'
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The goal of this document is to examine the change in risk under different management scenarios, e.g. early or late season closures.

```{r prep, message = FALSE}
library(scales)
library(tidyverse)

load("../raimbow-local/RDATA_files/Whale_risk_timeseries.Rdata")
# source("Whale_risk_mgt_funcs.R")

flag.save <- TRUE
path.plots <- "../raimbow-local/Plots/Management/"
```

To do this, we first must summarize the original risk (i.e., with no management scenarios) by region and fishing season

```{r orig}
x <- all.df.summ %>% 
  arrange(yr, mon, region) %>% 
  mutate(fish_season = cumsum(mon == 11 & region == "WA")) %>% 
  select(region, ym, yr, mon, fish_season, everything()) %>% 
  arrange(region, yr, mon)

x.orig <- x %>% 
  group_by(region, fish_season) %>% 
  summarise(yr = min(yr), 
            risk_total_season = sum(risk_sum_total), 
            risk_dens_season = sum(risk_sum_dens)) %>% 
  ungroup()

head(x.orig)
```

## Make function(s) and assign vals

Objects used throughout

```{r}
xlab.all   <- "Fishing season start year (November of start year to October of next year)"
ylab.dens  <- "Risk (whales*VMS pings / km2)"
ylab.total <- "Risk (whlaes*VMS pings)"
```

Functions used throughout this document

```{r funcs}
# Sum risk after filtering for desired months
mng_risk_sum <- function(x1, x2, mon.keep) {
  x1 %>% 
    filter(mon %in% mon.keep) %>% 
    group_by(region, fish_season) %>% 
    summarise(risk_total_season = sum(risk_sum_total), 
              risk_dens_season = sum(risk_sum_dens)) %>% 
    ungroup() %>% 
    mutate(risk_total_ratio = risk_total_season / x2$risk_total_season, 
           risk_dens_ratio = risk_dens_season / x2$risk_dens_season, 
           risk_total_diff = risk_total_season - x2$risk_total_season, 
           risk_dens_diff = risk_dens_season - x2$risk_dens_season)
}

# Plot of percentage of risk remaining for the season after applying scenario
mng_plot_perc <- function(x1, x2, x.title, x.xlab, x.ylab) {
  ggplot(x1, aes(fish_season, risk_dens_ratio, colour = region, group = region)) + 
    geom_point() + 
    geom_path() + 
    scale_colour_brewer(palette = "Set1", name = "Region") + 
    ggtitle(x.title) + 
    xlab(x.xlab) + 
    ylab(x.ylab) + 
    coord_cartesian(ylim = c(0, 1)) +
    scale_x_continuous(breaks = seq_along(unique(x2$fish_season)), 
                       labels = unique(x2$yr))
}


# Plot of risk removed by management scenario
mng_plot_diff <- function(x1, x2, x.title, x.xlab, x.ylab) {
  ggplot(x1, aes(fish_season, risk_dens_diff, colour = region, group = region)) + 
    geom_point() + 
    geom_path() + 
    scale_colour_brewer(palette = "Set1", name = "Region") + 
    ggtitle(x.title) + 
    xlab(x.xlab) + 
    ylab(x.ylab) + 
    coord_cartesian(ylim = c(-1, 0)) +
    scale_x_continuous(breaks = seq_along(unique(x2$fish_season)), 
                       labels = unique(x2$yr))
}


# Plot of summed risk density for the year with and without management scenario
mng_plot_bar <- function(x1, x2, v1, x.title, x.ylab) {
  v1 <- enquo(v1)
  
  x.new <- bind_rows(
    x2 %>% select(region, fish_season, risk_total_season, risk_dens_season), 
    x1 %>% select(region, fish_season, risk_total_season, risk_dens_season)
  ) %>% 
    mutate(orig = factor(c(rep("Original", nrow(x.orig)), rep("Closure", nrow(x.late)))))
  
  
  ggplot(x.new, aes(fish_season, !!v1, fill = orig)) +
    geom_col(position = "dodge") +
    facet_wrap(facets = vars(region), nrow = 2) +
    guides(fill = guide_legend(title = "Scenario")) +
    ggtitle(x.title) +
    xlab(xlab.all) +
    ylab(x.ylab) +
    scale_x_continuous(breaks = seq_along(unique(x2$fish_season)),
                       labels = unique(x2$yr)) + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.4))
}

# Plot of summed risk for the year without management scenario, and with orig value as dot above
mng_plot_bardot <- function(x1, x2, v1, x.title, x.ylab) {
  v1 <- enquo(v1)
  
  x.new <- bind_rows(
    x2 %>% select(region, fish_season, risk_total_season, risk_dens_season), 
    x1 %>% select(region, fish_season, risk_total_season, risk_dens_season)
  ) %>% 
    mutate(orig = factor(c(rep("Original", nrow(x.orig)), rep("Closure", nrow(x.late)))))
  
  ggplot() +
    geom_col(data = x1, aes(fish_season, !!v1, fill = "mng")) +
    geom_point(data = x.orig, aes(fish_season, !!v1, colour = "orig")) + 
    facet_wrap(facets = vars(region), nrow = 2) +
    scale_fill_manual(name = "Risk", 
                      values = c('mng' = hue_pal()(1)), 
                      labels = "Management") + 
    scale_color_manual(name = "", 
                       values = c('orig' = 'black'), 
                       labels = "Original") + 
    ggtitle(x.title) +
    xlab(xlab.all) +
    ylab(x.ylab) +
    scale_x_continuous(breaks = seq_along(unique(x2$fish_season)),
                       labels = unique(x2$yr)) + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.4))
}
```

## Late season closure, e.g. CA-CBD settlement

This scenario is meant to test how risk would have changed if the CA-CBD settlement, closing the DC fishery April 15, had been in place beginning in 2009. Because we are working with monthly-summarized data, we will consider the fishery to be open through April.

```{r late}
x.late <- mng_risk_sum(x, x.orig, c(11, 12, 1:4))

### Line plot(s)
mng_plot_perc(
  x.late, x.orig, 
  x.title = "Percentage of risk remaining with closing DC fishery May 1", 
  x.xlab = xlab.all, x.ylab = "Percentage"
)
if (flag.save) ggsave(filename = paste0(path.plots, "Mn_DC_close05.png"))

mng_plot_diff(
  x.late, x.orig, 
  x.title = "Difference in risk for fishing season with closing DC fishery May 1", 
  x.xlab = xlab.all, x.ylab = ylab.dens
)
if (flag.save) ggsave(filename = paste0(path.plots, "Mn_DC_close05_diff.png"))


### Bar plot(s)
# mng_plot_bar(
#   x.late, x.orig, risk_dens_season, 
#   "Summed risk (density) - closing DC fishery May 1", ylab.dens
# )
# if (flag.save) ggsave(filename = paste0(path.plots, "Mn_DC_close05_bar.png"))
# 
# mng_plot_bar(
#   x.late, x.orig, risk_total_season, 
#   "Summed risk (total) - closing DC fishery May 1", ylab.total
# )
# if (flag.save) ggsave(filename = paste0(path.plots, "Mn_DC_close05_bar_total.png"))

mng_plot_bardot(
  x.late, x.orig, risk_dens_season, 
  "Summed risk (density) - closing DC fishery May 1", ylab.dens
)
if (flag.save) ggsave(filename = paste0(path.plots, "Mn_DC_close05_bardot.png"))

mng_plot_bardot(
  x.late, x.orig, risk_total_season, 
  "Summed risk (total) - closing DC fishery May 1", ylab.total
)
if (flag.save) ggsave(filename = paste0(path.plots, "Mn_DC_close05_bardot_total.png"))
```

Close April 1

```{r late2}
x.late <- mng_risk_sum(x, x.orig, c(11, 12, 1:3))

### Line plot(s)
mng_plot_perc(
  x.late, x.orig, 
  x.title = "Percentage of risk remaining with closing DC fishery April 1", 
  x.xlab = xlab.all, x.ylab = "Percentage"
)
if (flag.save) ggsave(filename = paste0(path.plots, "Mn_DC_close04.png"))

mng_plot_diff(
  x.late, x.orig, 
  x.title = "Difference in risk for fishing season with closing DC fishery April 1", 
  x.xlab = xlab.all, x.ylab = ylab.dens
)
if (flag.save) ggsave(filename = paste0(path.plots, "Mn_DC_close04_diff.png"))


### Bar plot(s)
# mng_plot_bar(
#   x.late, x.orig, risk_dens_season, 
#   "Summed risk (density) - closing DC fishery April 1", ylab.dens
# )
# if (flag.save) ggsave(filename = paste0(path.plots, "Mn_DC_close04_bar.png"))
# 
# mng_plot_bar(
#   x.late, x.orig, risk_total_season, 
#   "Summed risk (total) - closing DC fishery April 1", ylab.total
# )
# if (flag.save) ggsave(filename = paste0(path.plots, "Mn_DC_close04_bar_total.png"))

mng_plot_bardot(
  x.late, x.orig, risk_dens_season, 
  "Summed risk (density) - closing DC fishery April 1", ylab.dens
)
if (flag.save) ggsave(filename = paste0(path.plots, "Mn_DC_close04_bardot.png"))

mng_plot_bardot(
  x.late, x.orig, risk_total_season, 
  "Summed risk (total) - closing DC fishery April 1", ylab.total
)
if (flag.save) ggsave(filename = paste0(path.plots, "Mn_DC_close04_bardot_total.png"))
```

The only fishing season in which this closure had a strong impact on regions in which most confirmed entanglements have happened is 2015-16, when the fishery was closed unitl March and thus 

## Early season closure, e.g. 2019-20 fishing season

Opening fishery Dec 1

```{r early1}
x.early1 <- mng_risk_sum(x, x.orig, c(12, 1:10))

### Line plots
mng_plot_perc(
  x.early1, x.orig, 
  x.title = "Percentage of risk remaining with opening DC fishery Dec 1", 
  x.xlab = xlab.all, x.ylab = "Percentage"
)
if (flag.save) ggsave(filename = paste0(path.plots, "Mn_DC_open12.png"))

mng_plot_diff(
  x.early1, x.orig, 
  x.title = "Difference in risk for fishing season with opening DC fishery Dec 1", 
  x.xlab = xlab.all, x.ylab = ylab.dens
)
if (flag.save) ggsave(filename = paste0(path.plots, "Mn_DC_open12_diff.png"))


### Bar plot(s)
# mng_plot_bar(
#   x.early1, x.orig, risk_dens_season, 
#   "Summed risk (density) - opening DC fishery Dec 1", ylab.dens
# )
# if (flag.save) ggsave(filename = paste0(path.plots, "Mn_DC_open12_bar.png"))
# 
# mng_plot_bar(
#   x.early1, x.orig, risk_total_season, 
#   "Summed risk (total) - opening DC fishery Dec 1", ylab.total
# )
# if (flag.save) ggsave(filename = paste0(path.plots, "Mn_DC_open12_bar_total.png"))

mng_plot_bardot(
  x.early1, x.orig, risk_dens_season, 
  "Summed risk (density) - opening DC fishery Dec 1", ylab.dens
)
if (flag.save) ggsave(filename = paste0(path.plots, "Mn_DC_open12_bardot.png"))

mng_plot_bardot(
  x.early1, x.orig, risk_total_season, 
  "Summed risk (total) - opening DC fishery Dec 1", ylab.total
)
if (flag.save) ggsave(filename = paste0(path.plots, "Mn_DC_open12_bardot_total.png"))
```

Opening fishery Jan 1

```{r early2}
x.early2 <- mng_risk_sum(x, x.orig, c(1:10))

### Line plots
mng_plot_perc(
  x.early2, x.orig, 
  x.title = "Percentage of risk remaining with opening DC fishery Jan 1", 
  x.xlab = xlab.all, x.ylab = "Percentage"
)
if (flag.save) ggsave(filename = paste0(path.plots, "Mn_DC_open01.png"))

mng_plot_diff(
  x.early2, x.orig, 
  x.title = "Difference in risk for fishing season with opening DC fishery Jan 1", 
  x.xlab = xlab.all, x.ylab = ylab.dens
)
if (flag.save) ggsave(filename = paste0(path.plots, "Mn_DC_open01_diff.png"))


### Bar plot(s)
# mng_plot_bar(
#   x.early2, x.orig, risk_dens_season, 
#   "Summed risk (density) - opening DC fishery Jan 1", ylab.dens
# )
# if (flag.save) ggsave(filename = paste0(path.plots, "Mn_DC_open01_bar.png"))
# 
# mng_plot_bar(
#   x.early2, x.orig, risk_total_season, 
#   "Summed risk (total) - opening DC fishery Jan 1", ylab.total
# )
# if (flag.save) ggsave(filename = paste0(path.plots, "Mn_DC_open01_bar_total.png"))

mng_plot_bardot(
  x.early2, x.orig, risk_dens_season, 
  "Summed risk (density) - opening DC fishery Jan 1", ylab.dens
)
if (flag.save) ggsave(filename = paste0(path.plots, "Mn_DC_open01_bardot.png"))

mng_plot_bardot(
  x.early2, x.orig, risk_total_season, 
  "Summed risk (total) - opening DC fishery Jan 1", ylab.total
)
if (flag.save) ggsave(filename = paste0(path.plots, "Mn_DC_open01_bardot_total.png"))
```
