---
title: "Humpback risk - management scenarios - county"
author: "Sam Woodman"
date: '`r format(Sys.Date(), "%d %B %Y")`'
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The goal of this document is to examine the change in risk under different management scenarios, e.g. early or late season closures.

```{r prep, message = FALSE}
library(tidyverse)

load("../raimbow-local/RDATA_files/Whale_risk_county_timeseries.Rdata")

flag.save <- TRUE
path.plots <- "../raimbow-local/Plots/Management/"
```

To do this, we first must summarize the original risk (i.e., with no management scenarios) by region and fishing season

```{r orig}
county.rm <- c("Ventura", "Los_Angeles", "Orange", "San_Diego")
county.levels <- levels(all.df.summ$region)[!(levels(all.df.summ$region) %in% county.rm)]

x <- all.df.summ %>% 
  mutate(region = factor(region, levels = county.levels)) %>% 
  arrange(yr, mon, region) %>% 
  mutate(fish_season = cumsum(mon == 11 & region == "WA")) %>% 
  select(region, ym, yr, mon, fish_season, everything()) %>% 
  arrange(region, yr, mon)

x.orig <- x %>% 
  group_by(region, fish_season) %>% 
  summarise(yr = min(yr), 
            risk_total_season = sum(risk_sum_total), 
            risk_dens_season = sum(risk_sum_dens)) %>% 
  ungroup()

head(x.orig)
```

## Late season closure, e.g. CA-CBD settlement

This scenario is meant to test how risk would have changed if the CA-CBD settlement, closing the DC fishery April 15, had been in place beginning in 2009. Because we are working with monthly-summarized data, we will consider the fishery to be open through April.

```{r late}
x.late <- x %>% 
  filter(mon %in% c(11, 12, 1:4)) %>% 
  group_by(region, fish_season) %>% 
  summarise(risk_total_season = sum(risk_sum_total), 
            risk_dens_season = sum(risk_sum_dens)) %>% 
  ungroup() %>% 
  mutate(risk_total_ratio = risk_total_season / x.orig$risk_total_season, 
         risk_dens_ratio = risk_dens_season / x.orig$risk_dens_season)

ggplot(x.late, aes(fish_season, risk_dens_ratio, colour = region, group = region)) + 
  geom_point() + 
  geom_path() + 
  # scale_colour_brewer(palette = "Set1", name = "Region") + 
  ggtitle("Percentage of risk remaining with closing DC fishery May 1") + 
  xlab("Fishing season start year") + 
  ylab("Percentage") + 
  coord_cartesian(ylim = c(0, 1)) +
  scale_x_continuous(breaks = seq_along(unique(x.orig$fish_season)), 
                     labels = unique(x.orig$yr))

if (flag.save) ggsave(filename = paste0(path.plots, "county_Mn_DC_close05.png"))
```

The only fishing season in which this closure had a strong impact on regions in which most confirmed entanglements have happened is 2015-16, when the fishery was closed unitl March and thus 

## Early season closure, e.g. 2019-20 fishing season

Opening fishery Dec 1

```{r early1}
x.early1 <- x %>% 
  filter(mon %in% c(12, 1:10)) %>% 
  group_by(region, fish_season) %>% 
  summarise(risk_total_season = sum(risk_sum_total), 
            risk_dens_season = sum(risk_sum_dens)) %>% 
  ungroup() %>% 
  mutate(risk_total_ratio = risk_total_season / x.orig$risk_total_season, 
         risk_dens_ratio = risk_dens_season / x.orig$risk_dens_season)

ggplot(x.early1, aes(fish_season, risk_dens_ratio, colour = region, group = region)) + 
  geom_point() + 
  geom_path() + 
  # scale_colour_brewer(palette = "Set1", name = "Region") + 
  ggtitle("Percentage of risk remaining with opening DC fishery Dec 1") + 
  xlab("Fishing season start year") + 
  ylab("Percentage") + 
  coord_cartesian(ylim = c(0, 1)) +
  scale_x_continuous(breaks = seq_along(unique(x.orig$fish_season)), 
                     labels = unique(x.orig$yr))

if (flag.save) ggsave(filename = paste0(path.plots, "county_Mn_DC_open12.png"))
```

Opening fishery Jan 1

```{r early2}
x.early2 <- x %>% 
  filter(mon %in% c(1:10)) %>% 
  group_by(region, fish_season) %>% 
  summarise(risk_total_season = sum(risk_sum_total), 
            risk_dens_season = sum(risk_sum_dens)) %>% 
  ungroup() %>% 
  mutate(risk_total_ratio = risk_total_season / x.orig$risk_total_season, 
         risk_dens_ratio = risk_dens_season / x.orig$risk_dens_season)

ggplot(x.early2, aes(fish_season, risk_dens_ratio, colour = region, group = region)) + 
  geom_point() + 
  geom_path() + 
  # scale_colour_brewer(palette = "Set1", name = "Region") + 
  ggtitle("Percentage of risk remaining with opening DC fishery Jan 1") + 
  xlab("Fishing season start year") + 
  ylab("Percentage") + 
  coord_cartesian(ylim = c(0, 1)) +
  scale_x_continuous(breaks = seq_along(unique(x.orig$fish_season)), 
                     labels = unique(x.orig$yr))

if (flag.save) ggsave(filename = paste0(path.plots, "county_Mn_DC_open01.png"))
```
