---
title: "Calculate risk"
author: "Jameal Samhouri"
date: "6/2/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(lubridate)
library(here)
library(sf)
knitr::opts_chunk$set(echo = TRUE)
library(rnaturalearth)
```

# Purpose

The purpose of this script is to calculate risk to blue and humpback whales due to Dungeness crab fishing in Washington state. 

It relies on antecedent scripts that process raw logbook data, and that join whale model outputs to a custom 5km grid developed by Blake Feist.

Risk calculations follow those in Samhouri et al. (in review)

# Read in data
```{r, read_data}

# read in status quo fishing data -- note new version adj_summtraps_2013_2020.rds
# JS 
x.orig.noinfo <- read_rds("/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Logbook-VMS/WA logbooks - mapping for CP/adj_summtraps.rds")
#x.orig.noinfo <- read_rds(here::here('wdfw','data','adj_summtraps_2013_2020.rds'))
glimpse(x.orig.noinfo) 


# grab whale data -- how has Grid5km_whale_WA.rds been filtered? grinds in WA with fishing?
# JS 
x.whale <-readRDS("/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/5x5 Grid/Grid5km_whale_WA.rds")
#x.whale <- read_rds(here::here('wdfw','data','Grid5km_whale.rds'))
#x.whale_WA <- read_rds(here::here('wdfw','data','Grid5km_whale_WA.rds'))
x.whale.bw <- read_rds(here::here('wdfw','data','BlueWhale_5km_long_monthly_2019Aug_2021Sep.rds')) #--> this has not been filtered for WA or OR

# Load and prep grid cell - area key
load("/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/5x5 Grid/Grid_5km_landerased.RDATA")
#load("E:\\Leena\\Documents\\Projects\\raimbow\\wdfw\\data\\Grid_5km_landerased.RDATA")
#path.grid.5km.lno <- "C:/Users/Leena.Riekkola/Projects/NOAA data/maps_ts_whales/data/Grid_5km_landerased.rds"
#grid.5km.lno <- readRDS(path.grid.5km.lno)
area.key <- grid.5km.lno %>% 
  st_drop_geometry() %>% 
  select(GRID5KM_ID, area_km_lno) %>% 
  distinct()

                          

```

# Read in functions

```{r, functions}

# 1) helper  function
simpleCap <- function(x) {
  s <- sapply(strsplit(x, " "), function(i) i[[1]])
  paste(toupper(substring(s, 1, 1)), substring(s, 2),
        sep = "", collapse = " ")
}

# 2) shift effort function
source("tradeoffs/Management scenarios/Mgmt_scenarios_shift_effort.R")

# 3) calculate risk function
source("tradeoffs/Management scenarios/Mgmt_scenarios_risk.R")

```

# Make season start date key and a fishing metrics range file

```{r, make_keys}

# read in season start date key. we need to make a file like this that is WA-specific. see add_day_of_season.Rmd

# JS 
season.st.date.key <- readRDS("/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Input_Data/season_start_dates/start_dates_by_CA_region.rds") %>% 
  mutate(crab_year = gsub("-", "_", .data$crab_season), 
         Region = unname(sapply(CA_region, simpleCap))) %>% 
  select(crab_year, Region, start_of_season_oneperc)

### Calculate and summarize risk for sq scenario with normalized outputs
# we need to make a file like this that is WA-specific. see prep_data_for_scenario_df_function.R
CA_fishing_metrics_range_2009_2019 <- read_rds(here:: here(
  "grid-prep",
  "CA_fishing_metrics_range_2009_2019.rds")
)

```

# Simple status quo risk

For status quo risk, we can simply join the fishing and whale data frames and multiply
```{r, join_fishing_whales}

# make avg value for traps for each yr month
x.orig.noinfo2 <- x.orig.noinfo %>%
  group_by(season_month, GRID5KM_ID, grd_x, grd_y, AREA) %>% 
  summarise( 
    number_obs = n(), #no. of grid cells in that season_month that had traps in them 
    # sum_M1_tottraps = sum(M1_tottraps), 
    # sum_M2_tottraps = sum(weighted_traps), 
    mean_M1_trapdens = mean(M1_trapdens), 
    mean_M2_trapdens = mean(M2_trapdens), 
    # M1_sdtrapdens = sd(M1_trapdens), 
    # M2_sdtrapdens = sd(M2_trapdens)
  )

# make column for year month for fishing data
## step 1, make some columns that we can use
x.orig.noinfo3 <- x.orig.noinfo2 %>%
  separate(season_month, into = c("season", "month_name"), sep = "_") %>%
  separate(season, into = c("yr_start", "yr_end"), sep = "-") %>%
  mutate(month = match(month_name, month.name)) %>% #month becomes one digit number
  mutate(month = sprintf("%02d", as.numeric(month))) #change month to two digit number
## step 2, grab yr_start for December effort and make a year_month column in a new df
df1 <- x.orig.noinfo3 %>%
  filter(month_name=='December')
df1 <- df1 %>%
  mutate(year_month = paste0(yr_start,"_",month))
## step 3, grab yr_end for non-December effort and make a year_month column in a new df
df2 <- x.orig.noinfo3 %>%
  filter(month_name !='December')
df2 <- df2 %>%
  mutate(year_month = paste0(yr_end,"_",month))
# squish the December and non-December df's together  
x.orig.noinfo4 <- rbind(df1,df2)


x.whale.bw2 <- x.whale.bw %>%
  separate(date, into = c("year", "month", "day"), sep = "-") %>%
  mutate(year_month = paste0(year,"_",month))


# join the whale and fishing data
joined_df <- x.orig.noinfo4 %>%
  #left_join(x.whale,by=c("year_month","GRID5KM_ID"))
  left_join(x.whale.bw2,by=c("year_month","GRID5KM_ID"))


# check how many NAs are there
length(which(is.na(joined_df))) # 6887, that is a lot. let's find them
# let's focus on NA in the columns we really really need
length(which(is.na(joined_df$GRID5KM_ID))) # 0, good
length(which(is.na(joined_df$mean_M1_trapdens))) # 0, good
length(which(is.na(joined_df$mean_M2_trapdens))) # 0, good
length(which(is.na(joined_df$Blue_occurrence_mean))) # 1675
length(which(is.na(joined_df$Humpback_abund_mean))) # 182

# so the NA problems are with the whale data, but i suspect that those are grid cells we never had whale predictions for (there were 140 cells like that). see grid.5km.na from Grid5km_raimbow_prep.Rmd (pasted below)

length(unique(joined_df$GRID5KM_ID[which(is.na(joined_df$Blue_occurrence_mean))])) # 192, more than the 140

# multiply Blue_occurrence_mean or Humpback_abund_mean by effort to get raw risk
joined_df <- joined_df %>%
  mutate(
    blue_risk_M1 = Blue_occurrence_mean * mean_M1_trapdens,
    blue_risk_M2 = Blue_occurrence_mean * mean_M2_trapdens,
    hump_risk_M1 = Humpback_abund_mean * mean_M1_trapdens,
    hump_risk_M2 = Humpback_abund_mean * mean_M2_trapdens
  )

write_rds(joined_df, "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Logbook-VMS/WA logbooks - mapping for CP/simple_sq_risk_df.rds")

# unique(grid.5km.na$GRID5KM_ID)
#   [1] 117642 118300 118630 118960 119290 119620 120942 122261 122589 122591 122592 122919
#  [13] 122920 125550 125551 125552 125553 125554 125555 125556 125557 125880 125881 125882
#  [25] 125883 125884 125885 125886 126209 126210 126211 126212 126213 126214 126215 126216
#  [37] 126538 126539 126540 126541 126542 126543 126544 126545 126546 126868 126869 126870
#  [49] 126871 126872 126873 126874 126875 127197 127198 127199 127200 127201 127202 127203
#  [61] 127204 127527 127528 127529 127530 127531 127532 127533 127534 127857 127858 127859
#  [73] 127860 127861 127862 127863 127864 127865 128186 128187 128188 128189 128190 128191
#  [85] 128192 128193 128194 128516 128517 128518 128519 128520 128521 128522 128523 128846
#  [97] 128847 128848 128849 128850 128851 129176 129177 129178 129179 129180 129181 129506
# [109] 129507 129508 129509 129510 129837 129838 129839 129840 129841 130167 130168 130169
# [121] 130170 130497 130498 130499 130827 130828 130829 131157 131158 131159 131488 131818
# [133] 131819 132148 132149 132478 132479 132809 132810 133140

```

## Summarize simple status quo risk

```{r, summarize_sq_risk}

joined_df <- read_rds("/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Logbook-VMS/WA logbooks - mapping for CP/simple_sq_risk_df.rds") %>%
  mutate(
    season = case_when(
      month_name == "December" | month_name == "January" | month_name == "February" | month_name == "March" | month_name == "April" ~ "Winter",
      month_name == "May" | month_name == "June" | month_name == "July" | month_name == "August" | month_name == "September" ~ "Spring/Summer"
    )
  )
glimpse(joined_df)
  
# make a df for dec-april of each crab season

joined_df_winter <- joined_df %>%
  filter(season == "Winter") %>%
  group_by(yr_end, GRID5KM_ID) %>%
  summarise(
    Blue_occurrence_mean = mean(Blue_occurrence_mean),
    Humpback_abund_mean = mean(Humpback_abund_mean),
    mean_M1_trapdens = mean(mean_M1_trapdens),
    mean_M2_trapdens = mean(mean_M2_trapdens),
    blue_risk_M1 = mean(blue_risk_M1),
    blue_risk_M2 = mean(blue_risk_M2),
    hump_risk_M1 = mean(hump_risk_M1),
    hump_risk_M2 = mean(hump_risk_M2)
    ) %>%
  mutate(
    season = "Winter"
  )
glimpse(joined_df_winter)
  

# make a df for may-sept of each crab season

joined_df_sprsum <- joined_df %>%
  filter(season != "Winter") %>%
  group_by(yr_end, GRID5KM_ID) %>%
  summarise(
    Blue_occurrence_mean = mean(Blue_occurrence_mean),
    Humpback_abund_mean = mean(Humpback_abund_mean),
    mean_M1_trapdens = mean(mean_M1_trapdens),
    mean_M2_trapdens = mean(mean_M2_trapdens),
    blue_risk_M1 = mean(blue_risk_M1),
    blue_risk_M2 = mean(blue_risk_M2),
    hump_risk_M1 = mean(hump_risk_M1),
    hump_risk_M2 = mean(hump_risk_M2)
    ) %>%
  mutate(
    season = "Spring/Summer"
  )
glimpse(joined_df_sprsum)

# make a single seasonal df
names(joined_df_sprsum) == names(joined_df_winter)
joined_df_season <- rbind(joined_df_winter,joined_df_sprsum)
glimpse(joined_df_season)

write_rds(joined_df_season, "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Logbook-VMS/WA logbooks - mapping for CP/simple_sq_risk_df_annual_byseason.rds")


```

# Produce a data frame that is easy to apply the risk calculation function to

This chunk calculates effort in every grid cell 

```{r, shift_effort}
##### summarize effort for sq scenario

# STATUS QUO
scenario.output.df.noinfo.sq <- effort_mgmt(
  x = x.orig.noinfo,
  season.st.key = season.st.date.key, 
  preseason.days = 3,
  season.st.backstop = NULL, 
  early.data.method = "remove", 
  delay.date = NULL,
  delay.region = NULL,
  delay.method = NULL,
  delay.method.fidelity = NULL,
  closure.date = NULL,
  closure.region = NULL,
  closure.method = NULL,
  closure.redist.percent = 100,
  depth.shallow = NULL, 
  depth.deep = NULL, 
  reduction.before.date = NULL,
  reduction.before.percent = 50,
  reduction.before.region = NULL,
  reduction.after.date = NULL,
  reduction.after.percent = 50,
  reduction.after.region = NULL
)


```

# Calculate risk and write out data frames

Calculate risk
```{r, calculate}

risk_out_sq <- risk_mgmt(
  x = scenario.output.df.noinfo.sq, 
  x.col = Num_DCRB_VMS_pings, 
  y = x.whale,
  risk.unit = "dens", 
  area.key = area.key,
  scale.list = CA_fishing_metrics_range_2009_2019, 
  ym.min = "2009_11", 
  ym.max = "2019_07"
)
glimpse(risk_out_sq)

range(risk_out_sq$Num_DCRB_VMS_pings) # max value for pings

```

Write to files
```{r, write}

# write out for use in downstream mapping and figure plotting by yr_mth
# (we previously used this with tradeoff figures Rmd)
write_rds(risk_out_sq, paste0("/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/scenario_output_dataframes/status_quo_risk_2009_2019_yr_mth_",today(),".rds"))

risk_out_sq_list_by_yr_mth <- risk_out_sq %>% split(.$year_month)

# write out for use in downstream mapping and figure plotting for each yr_mth separately
# (we previously used this with prioritizr)
# write_rds(risk_out_sq_list_by_yr_mth, paste0("/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/scenario_output_dataframes/status_quo_risk_2009_2019_list_by_yr_mth_",today(),".rds"))

```