---
title: "Calculate risk"
author: "Jameal Samhouri"
date: "6/2/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(lubridate)
library(here)
library(sf)
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

The purpose of this script is to calculate risk to blue and humpback whales due to Dungeness crab fishing in Washington state. 

It relies on antecedent scripts that process raw logbook data, and that join whale model outputs to a custom 5km grid developed by Blake Feist.

Risk calculations follow those in Samhouri et al. (in review)

# Read in data
```{r, read_data}

# read in status quo fishing data
# JS 
x.orig.noinfo <- read_rds("/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Logbook-VMS/WA logbooks - mapping for CP/adj_summtraps.rds")
glimpse(x.orig.noinfo) # there is no column M2_tottraps, but it is the same as weighted_traps


# grab whale data
# JS 
x.whale <-readRDS("/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/5x5 Grid/Grid5km_whale.rds")

# Load and prep grid cell - area key
load("/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/5x5 Grid/Grid_5km_landerased.RDATA")
area.key <- grid.5km.lno %>% 
  st_drop_geometry() %>% 
  select(GRID5KM_ID, area_km_lno) %>% 
  distinct()

                          

```

# Read in functions

```{r, functions}

# 1) helper  function
simpleCap <- function(x) {
  s <- sapply(strsplit(x, " "), function(i) i[[1]])
  paste(toupper(substring(s, 1, 1)), substring(s, 2),
        sep = "", collapse = " ")
}

# 2) shift effort function
source("tradeoffs/Management scenarios/Mgmt_scenarios_shift_effort.R")

# 3) calculate risk function
source("tradeoffs/Management scenarios/Mgmt_scenarios_risk.R")

```

# Make season start date key and a fishing metrics range file

```{r, make_keys}

# read in season start date key. we need to make a file like this that is WA-specific. see add_day_of_season.Rmd

# JS 
season.st.date.key <- readRDS("/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Input_Data/season_start_dates/start_dates_by_CA_region.rds") %>% 
  mutate(crab_year = gsub("-", "_", .data$crab_season), 
         Region = unname(sapply(CA_region, simpleCap))) %>% 
  select(crab_year, Region, start_of_season_oneperc)

### Calculate and summarize risk for sq scenario with normalized outputs
# we need to make a file like this that is WA-specific. see prep_data_for_scenario_df_function.R
CA_fishing_metrics_range_2009_2019 <- read_rds(here:: here(
  "grid-prep",
  "CA_fishing_metrics_range_2009_2019.rds")
)

```

# Simple status quo risk

For status quo risk, we can simply join the fishing and whale data frames and multiply
```{r, join_fishing_whales}

# make column for year month for fishing data, make avg value for traps for each yr month
x.orig.noinfo2 <- x.orig.noinfo %>%
  group_by(season_month, GRID5KM_ID, grd_x, grd_y, AREA) %>% 
  summarise( 
    number_obs = n(), #no. of grid cells in that season_month that had traps in them 
    # sum_M1_tottraps = sum(M1_tottraps), 
    # sum_M2_tottraps = sum(weighted_traps), 
    mean_M1_trapdens = mean(M1_trapdens), 
    mean_M2_trapdens = mean(M2_trapdens), 
    # M1_sdtrapdens = sd(M1_trapdens), 
    # M2_sdtrapdens = sd(M2_trapdens)
  ) %>%
  mutate(
    year_month = # translate season_month
  )

joined_df <- x.orig.noinfo2 %>%
  left_join(x.whale)

# multiply Blue_occurrence_mean or Humpback_abund_mean by effort to get raw risk

```

# Produce a data frame that is easy to apply the risk calculation function to

This chunk calculates effort in every grid cell 

```{r, shift_effort}
##### summarize effort for sq scenario

# STATUS QUO
scenario.output.df.noinfo.sq <- effort_mgmt(
  x = x.orig.noinfo,
  season.st.key = season.st.date.key, 
  preseason.days = 3,
  season.st.backstop = NULL, 
  early.data.method = "remove", 
  delay.date = NULL,
  delay.region = NULL,
  delay.method = NULL,
  delay.method.fidelity = NULL,
  closure.date = NULL,
  closure.region = NULL,
  closure.method = NULL,
  closure.redist.percent = 100,
  depth.shallow = NULL, 
  depth.deep = NULL, 
  reduction.before.date = NULL,
  reduction.before.percent = 50,
  reduction.before.region = NULL,
  reduction.after.date = NULL,
  reduction.after.percent = 50,
  reduction.after.region = NULL
)


```

# Calculate risk and write out data frames

Calculate risk
```{r, calculate}

risk_out_sq <- risk_mgmt(
  x = scenario.output.df.noinfo.sq, 
  x.col = Num_DCRB_VMS_pings, 
  y = x.whale,
  risk.unit = "dens", 
  area.key = area.key,
  scale.list = CA_fishing_metrics_range_2009_2019, 
  ym.min = "2009_11", 
  ym.max = "2019_07"
)
glimpse(risk_out_sq)

range(risk_out_sq$Num_DCRB_VMS_pings) # max value for pings

```

Write to files
```{r, write}

# write out for use in downstream mapping and figure plotting by yr_mth
# (we previously used this with tradeoff figures Rmd)
write_rds(risk_out_sq, paste0("/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/scenario_output_dataframes/status_quo_risk_2009_2019_yr_mth_",today(),".rds"))

risk_out_sq_list_by_yr_mth <- risk_out_sq %>% split(.$year_month)

# write out for use in downstream mapping and figure plotting for each yr_mth separately
# (we previously used this with prioritizr)
# write_rds(risk_out_sq_list_by_yr_mth, paste0("/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/scenario_output_dataframes/status_quo_risk_2009_2019_list_by_yr_mth_",today(),".rds"))

```