---
title: "Figure 2-16 Method"
author: "Victoria"
date: "June 9, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## packages

```{r}
library(odbc) # this is needed to connect to WAFT
library(dplyr) # includes dplyr and other packages
library(dbplyr)
library(tidyr)
library(ggplot2)
library(data.table)
library("xlsx")
library(writexl)
library(tidyverse)
library(DBI)
```

## connect to waft
## you probably have a different way of doing this

```{r}
con_waft <- dbConnect(odbc::odbc(), 
                      driver = "SQL Server", 
                      server = "busprod.dfw.wa.lcl\\busprod", 
                      database = "FishTicket")

waft <- con_waft %>% 
  tbl(in_schema("Reporting", "vFishTicketFlat"))
```

# define coastal areas so we're only looking at the coastal Dungeness fishery
# create crab_db to pull out only coastal crab fish tickets, and have crabseason column to describe the fishery season rather than the calendar year in which the crab was caught
```{r}
coastal_areas=c("59A2","60A1","60B ","59A1","60A2","60C ","60D ","61  ", "62  ", "63  ")

crab_db=waft %>% filter(MarketSpeciesCategoryName == "DUNGENESS CRAB" & FisherTypeID == 1 & CatchAreaCode%in%coastal_areas)%>% mutate(crabseason=if_else(LandingMonth<=9, BatchYear-1, BatchYear)) %>% select(FishTicketNumber, LandingMonth, crabseason, BatchYear, FisherLicenseID, PortDescription, DaysFishedCount, VesselID, CatchAreaCode, RoundPoundQuantity, BuyerLicenseID) %>% arrange(BatchYear, LandingMonth) %>% collect()%>% mutate(crabseason=paste(crabseason, crabseason+1, sep="-")) %>% collect()
```
Above is the crab database with the crabseason column added to help when working with the data as fishery year rather than calendar year. Now, the database can be queried by crab season. It also does not include the pot limit column because this information is not included on fish tickets. 

#Below I will read in the current license list and join it with the fish ticket data so I can query for max number of lines in the water (using participation of each fisher and each fisher's respective pot limit). I also retained the vessel length, but this will only be useful for looking at current data (because it can change frequently and the historical fish ticket data will not account for this)

#information on the license list can change frequently, this is old code from when I first started to work with the data. You would need to swap the file name and path for the license list file Robert sent you to get the most up-to-date estimates.
```{r}
licenses=fread("H:/Crab Stuff/Crab Mgmt/11.26.19 License List.csv")
licensesabbrev=licenses %>% select(`License ID`, `Crab Pot Limit`, `Vessel One Length`) %>% collect()
crabpot_db=crab_db %>% inner_join(licensesabbrev, by=c("FisherLicenseID"="License ID"))
```

##then I looked for the number of pots that are in the water during any given month, going back to the 2008-09 crab season.

##I don't have much experience using R to develop visualizations, so I wrote the data into an Excel file, then graphed it there.

```{r}
HistoricPotsinWater=crabpot_db %>% filter(BatchYear>=2009 | (BatchYear==2008 & LandingMonth>9)) %>% group_by(crabseason, LandingMonth, `Crab Pot Limit`) %>% summarise(numberoflicenses=n_distinct(FisherLicenseID)) %>% mutate(PotsFished=sum(numberoflicenses * `Crab Pot Limit`)) %>% select(crabseason, LandingMonth, PotsFished) %>% distinct() %>% collect()

HistoricPotsinWater %>% spread(LandingMonth, PotsFished)

write_xlsx(x = HistoricPotsinWater %>% spread(LandingMonth, PotsFished), path = "HistoricPotsinWater.xlsx", col_names = TRUE)
```

