---
title: "Scenario analysis for tri-states webinar"
author: "Jameal Samhouri"
date: "2/18/2020"
output: html_document
---

The focus of this analysis is to understand how alternative management scenarios, applied retrospectively during the years 2009-2018, may have affected risk of entanglement for blue and humpback whales, and risk to the Dungeness crab (DCRB) fishery in terms of lost revenue.

This code relies on VMS locations of DCRB boats in CA from 2009-18, each of which has been assigned a predicted occurrence of blue whales (based on Abrahms et al. 2019) and predicted density of humpback whales (based on Forney et al. in prep Model 1). 

Note that for DCRB fishing data, 2009 is truncated to Nov/Dec 2009 (and) before april 1 in 2009 refers to Nov-Dec 2009), which are actually part of the 2009-2010 crab season. Similarly, 2018 is truncated to Jan-Jul 2018, the 2017-18 crab season.

Dependencies:
create_Rdata.Rmd, VMS_extracted_metrics.R, Match processed VMS data to fish ticket landings and revenue.Rmd, Make confidential data summarized by grid cell 2009-18.Rmd: these files create a df that represents DCRB fishing activity in CA (filtered by depth and speed), joined to various grids and to Abrahms et al blwh predictions

get_humpback_predictions.R: generates monthly summaries of humpback whale densities predicted by Forney et al in prep Model 1, and produces an output file entitled "Humpback whale abundance monthly abundance predictions 2009-2018.csv"

Simple early closure scenario analysis.Rmd: generates Scenario_Analysis_Data_2009_2018.RData, which contains full.df, the data frame we start with below.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Prep for running code
Clear workspace
```{r, echo=FALSE}
rm(list=ls())
```
<br>

Install packages
```{r, echo=FALSE}
library(foreign)
library(lubridate)
library(tidyverse)
library(reshape2)
library(scales)
library(zoo)
library(ggrepel)
library(sf)
library(data.table)
library(wesanderson)
library(viridis)
library(here)
library(ggerr)
```
<br>

### Read in data and pull out initial data frame
```{r}
load("~/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/Scenario_Analysis_Data_2009_2018.RData")

View(full.df)
names(full.df)

dat <- full.df  %>% #filter(year %in% c(2012,2013)) %>% 
  dplyr::select(crab.year,year,month,GRID5KM_ID,Region,BIA_mn_noNAs,BIA_bm_noNAs,dollars_DCRB,Num_DCRB_VMS_pings,Blue_occurrence,H_Avg_Abund)

```
<br>


### Create df for each scenario (15 total + status quo) 

Spatial:
Statewide
cenCA only
all BIAs

Temporal: months for closures
One-month delay in opening --> 11,12
Two-month delay in opening --> 11, 12
Early closure --> months 4,5,6,7


```{r}

# Make sure that there is a record for all grid cells for all months, and fill in missing ping values with zeroes. this is necessary because we need to "look back" (use lags) in the next step

### Make calculations
# Total dollars_DCRB, total pings, total number of grid cells, ping density, mean whale abundance, and risk metrics

df <- dat
scenario_name <- "1_Month_Delay"
spatial_domain <- "Region"
scenario_lag <- 0

# doesn't output df the way i want
scenario_summary_function <- function(df, spatial_domain, scenario_lag, scenario_name) {
  #df.tmp <- paste0(scenario_name,"_dat")
  enquo_spatial_domain <- enquo(spatial_domain) # https://dplyr.tidyverse.org/articles/programming.html
  df %>%
    group_by(!!enquo_spatial_domain) %>%
    summarise(npings=sum(Num_DCRB_VMS_pings,na.rm=TRUE))
}
scenario_summary_function(dat,"Region",0,"1_Month_Delay")
    
    complete(nesting(!!enquo_spatial_domain,GRID5KM_ID),crab.year,month=1:12,fill=list(dollars_DCRB=0, Num_DCRB_VMS_pings=0)) %>% 
    # lag number of pings by 1 month to simulate redistribution
  group_by(GRID5KM_ID) %>% 
  arrange(year,month) %>% 
  mutate(
    npings_scenario=lag(Num_DCRB_VMS_pings,scenario_lag),
    dollars_scenario=lag(dollars_DCRB,scenario_lag)
    ) %>%
  
  
}



```
<br>




Include new variables relative risk and relative revenue reduction
Glue together into tradeoff df
```{r}

# Match months of the year to months of the crab season (starting in November)
month_of_season_match <- tibble(month=c(1:7,11,12),month_of_season=c(3:9,1:2))

dat.setup <- dat %>% 
  left_join(month_of_season_match) 

```
<br>

Make tradeoff plots
```{r}



```
<br>

