---
title: "Scenario analysis for tri-states webinar"
author: "Jameal Samhouri"
date: "2/18/2020"
output: html_document
---

The focus of this analysis is to understand how alternative management scenarios, applied retrospectively during the years 2009-2018, may have affected risk of entanglement for blue and humpback whales, and risk to the Dungeness crab (DCRB) fishery in terms of lost revenue.

This code relies on VMS locations of DCRB boats in CA from 2009-18, each of which has been assigned a predicted occurrence of blue whales (based on Abrahms et al. 2019) and predicted density of humpback whales (based on Forney et al. in prep Model 1). 

Note that for DCRB fishing data, 2009 is truncated to Nov/Dec 2009 (and) before april 1 in 2009 refers to Nov-Dec 2009), which are actually part of the 2009-2010 crab season. Similarly, 2018 is truncated to Jan-Jul 2018, the 2017-18 crab season.

Dependencies:
create_Rdata.Rmd, VMS_extracted_metrics.R, Match processed VMS data to fish ticket landings and revenue.Rmd, Make confidential data summarized by grid cell 2009-18.Rmd: these files create a df that represents DCRB fishing activity in CA (filtered by depth and speed), joined to various grids and to Abrahms et al blwh predictions

get_humpback_predictions.R: generates monthly summaries of humpback whale densities predicted by Forney et al in prep Model 1, and produces an output file entitled "Humpback whale abundance monthly abundance predictions 2009-2018.csv"

Simple early closure scenario analysis.Rmd: generates Scenario_Analysis_Data_2009_2018.RData, which contains full.df, the data frame we start with below.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Prep for running code
Clear workspace
```{r, echo=FALSE}
rm(list=ls())
```
<br>

Install packages
```{r, echo=FALSE}
library(foreign)
library(lubridate)
library(tidyverse)
library(reshape2)
library(scales)
library(zoo)
library(ggrepel)
library(sf)
library(data.table)
library(wesanderson)
library(viridis)
library(here)
library(ggerr)
```
<br>

### Read in data and pull out initial data frame
```{r}
load("~/Documents/RAIMBOW/Processed Data/Samhouri et al. whales risk/Output_Data/Scenario_Analysis_Data_2009_2018.RData")

View(full.df)
names(full.df)

dat <- full.df  %>% #filter(year %in% c(2012,2013)) %>% 
  dplyr::select(crab.year,year,month,GRID5KM_ID,Region,BIA_mn_noNAs,BIA_bm_noNAs,dollars_DCRB,lbs_DCRB,Num_DCRB_VMS_pings,Blue_occurrence,H_Avg_Abund) %>%
  mutate(BIA_bm_or_mn = ifelse(BIA_bm_noNAs !=0 | BIA_mn_noNAs != 0, "Inside BIA","Outside BIA")) 

```
<br>

### Make table of scenarios

Spatial:
Statewide
cenCA only
all BIAs

Temporal: months for closures
Marine life delay in opening --> to Dec 15
Early closure --> months 4,5,6,7

```{r}

# make table of scenarios
delay_scenarios <- c(
  "No_Delay",
  "Statewide_Marine_Life_Delay",
  "CenCA_Marine_Life_Delay"
)

closure_scenarios <- c(
  "No_Early_Closure",
  "Statewide_Early_Closure",
  "CenCA_Early_Closure",
  "BIA_Early_Closure"
)
  
scenario_table <- expand.grid(
  "delay_scenario" = delay_scenarios,
  "closure_scenario" = closure_scenarios
) 
  # slice(-3, .preserve=TRUE) %>% # remove BIA delay
  # slice(-8, .preserve=TRUE) # remove BIA squeeze

# scenario_table <- add_row(scenario_table,
#                           spatial="Status_Quo",
#                           temporal="Status_Quo"
#                           )

#add columns
scenario_table <- scenario_table %>%
  mutate(
    scenario_df_name = paste(delay_scenario,closure_scenario,sep="_"),
    delay_time_scenario = ifelse(delay_scenario != "No_Delay", "Dec-15", NA),
    delay_domain_scenario = ifelse(delay_scenario != "No_Delay", substr(delay_scenario,1,5), NA),
    closure_time_scenario = ifelse(closure_scenario != "No_Early_Closure", "Spring-Summer", NA),
    closure_domain_scenario =  ifelse(closure_scenario != "No_Early_Closure", substr(closure_scenario,1,3), NA)
  ) %>%
  dplyr::select(scenario_df_name, delay_scenario, closure_scenario, delay_time_scenario, delay_domain_scenario, closure_time_scenario, closure_domain_scenario)


# write_rds(scenario_table, here::here(
#   "tradeoffs",
#   "Management scenarios",
#   "scenario_table.RDS"
#   )
# )
# 
# scenario_table <- read_rds(here::here(
#   "tradeoffs",
#   "Management scenarios",
#   "scenario_table.RDS"
#   )
# )

```
<br>

### Create df for each scenario (11 total + status quo) 


```{r}

# Set up tibble to match months of the year to months of the crab season (starting in November)
month_of_season_match <- tibble(month=c(1:7,11,12),month_of_season=c(3:9,1:2))

# function to return scenario specific data frames

# 022520: need to consider having 1 function to introduce temporal lags, 1 for spatial lags,  1 for summarizing each scenario, 1 for making tradeoff df, others for making plots

#https://tidyeval.tidyverse.org/sec-up-to-speed.html#quote-and-unquote
scenario_summary_function <- function(df, spatial_domain, scenario_lag, scenario_df_name) {
  
  sym_spatial_domain <- sym(spatial_domain)
  
  out <- df %>%
    
    # Make sure that there is a record for all grid cells for all months, and fill in missing ping values with zeroes. this is necessary because we need to "look back" (use lags) in the next step
    complete(nesting(!!sym_spatial_domain,GRID5KM_ID,crab.year),month=1:12,fill=list(dollars_DCRB=0, Num_DCRB_VMS_pings=0)) %>%
    ungroup() %>% 
    
    left_join(month_of_season_match,by='month') %>% 
    
    # lag number of pings by scenario_lag to simulate redistribution
    group_by(GRID5KM_ID,crab.year) %>% 
    mutate(
      npings_scenario=lag(Num_DCRB_VMS_pings,scenario_lag,order_by=month_of_season),
      dollars_scenario=lag(dollars_DCRB,scenario_lag,order_by=month_of_season),
      lbs_scenario=lag(lbs_DCRB,scenario_lag,order_by=month_of_season),
      ) %>% 
    # removes all of the empty rows that were created during complete
    filter_at(vars(lbs_DCRB,dollars_DCRB,Num_DCRB_VMS_pings,npings_scenario,dollars_scenario,lbs_scenario),any_vars(.>0)) %>%
    # remove any effort in nov or dec
  mutate(npings_scenario=ifelse(scenario_lag == 1 & month==11,0,npings_scenario),
         dollars_scenario=ifelse(scenario_lag == 1 & month==11,0,dollars_scenario),
         lbs_scenario=ifelse(scenario_lag == 1 & month==11,0,lbs_scenario),
         npings_scenario=ifelse(scenario_lag == 2 & month%in% c(11,12),0,npings_scenario),
         dollars_scenario=ifelse(scenario_lag == 2 & month%in% c(11,12),0,dollars_scenario),
         lbs_scenario=ifelse(scenario_lag == 2 & month%in% c(11,12),0,lbs_scenario)
         ) %>% 
  filter(!is.na(npings_scenario),!is.na(dollars_scenario),!is.na(lbs_scenario)) %>% 
  ungroup() %>% 
  group_by(!!sym_spatial_domain,year,month) %>% 
  summarise(totpings=sum(npings_scenario,na.rm=T),
            n_cells=n_distinct(GRID5KM_ID,na.rm = T),
            ping_density=totpings/n_cells,
            totdollars=sum(dollars_scenario,na.rm=T),
            totlbs=sum(lbs_scenario,na.rm=T),
            
            # mean and sd humpback abundance (could do sum or max?)
            mean_H_Avg_Abund=mean(H_Avg_Abund,na.rm=T),
            sd_H_Avg_Abund = sd(mean_H_Avg_Abund, na.rm=TRUE),
            
            # mean and sd blue occurrence (could do sum or max?)
            mean_Blue_occurrence = mean(mean_Blue_occurrence, na.rm=TRUE),
            sd_mean_Blue_occurrence = sd(mean_Blue_occurrence, na.rm=TRUE),
            
            # risk as mean abundance times total pings divided by number of grid cells
            risk=sum(npings_scenario*H_Avg_Abund,na.rm=T)/n_cells
            ) %>% 
  ungroup() %>% 
  mutate(scenario=scenario_df_name) # substr(scenario_df_name,1,nchar(scenario_df_name)-4) # -4 removes "_dat" 
  
assign(paste(scenario_df_name,"_dat",sep="_"),out,envir=.GlobalEnv)

}

# uncomment 4 lines below to interact with function above
# df <- dat
# scenario_df_name <- "One_Month_Delay_dat"
# spatial_domain <- "Region"
# scenario_lag <- 1

# try function on 1 scenario
# scenario_summary_function(dat,"Region",0,"One_Month_Delay_dat")
# glimpse(One_Month_Delay_dat)

# for loop to create scenario df's

start.time <- Sys.time()
for(i in 1:nrow(scenario_table)) {
  
  # pick up here. need if_else to translate scenario table values for spatial to , eg, "", "Region", BIAs and lag = 0,1,2, April1orAfter, etc
  
  scenario_summary_function(dat,scenario_table$spatial_domain[i],scenario_table$scenario_lag[i],scenario_table$scenario_df_name[i])
  
  print(paste("made df for scenario",i))
}
Sys.time() - start.time

```
<br>

### Make calculations and create tradeoff df

Total dollars_DCRB, total pings, total number of grid cells, ping density, mean whale abundance, and risk metrics
```{r}


```
<br>

Include new variables relative risk and relative revenue reduction
Glue together into tradeoff df
```{r}


```
<br>

Make tradeoff plots
```{r}



```
<br>

