---
title: "Tradeoff manuscript whale time series figures"
author: "Sam Woodman"
date: "10/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This document generates whale prediction time series plots for Jameal's tradeoff manuscript. It duplicates a fair bit of the work done in '2_Whale_risk.Rmd' and '3_Whale_risk_timeseries.Rmd', but also does it for blue whales and does it without needing to worry about carrying along the risk metrics/calculations and thus the duplication was worth having this document be self-contained.

We can use 'Grid5km_whale.rds' because we're only making time series plots for CA.

```{r intro, message=FALSE}
library(gridExtra)
library(lubridate)
library(sf)
library(tidyverse)


source(here::here("User_script_local.R"))

if (user == "JS") {
  
  file.grid.region <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/5x5 Grid/Grid5km_key.rds"
  file.whale.data <- "/Users/jameal.samhouri/Documents/RAIMBOW/Processed Data/5x5 Grid/Grid5km_whale.rds"
  
} else if (user == "SMW") {
  flag.save <- FALSE
  
  # file.data.bmpreds <- "C:/SMW/RAIMBOW/raimbow-local/Outputs/BlueWhale_5km_long_monthly.rds"
  # file.data.mnpreds <- "C:/SMW/RAIMBOW/raimbow-local/Outputs/Humpback_5km_long_monthly.rds"
  file.grid.region <- "C:/SMW/RAIMBOW/raimbow-local/RDATA_files/Grid5km_key.rds"
  file.whale.data <- "C:/SMW/RAIMBOW/raimbow-local/RDATA_files/Grid5km_whale.rds"
  
} else {
  stop("User not recognized")
}
```

## Process whale predictions

```{r}
grid.key <- readRDS(file.grid.region)
x.orig <- readRDS(file.whale.data) 

x <- x.orig %>% 
  left_join(grid.key, by = "GRID5KM_ID") %>% 
  mutate(Humpback_dens = Humpback_abund_mean / area_km_lno, 
         Region = factor(Region, levels = c("NorCA", "CenCA")))

stopifnot(!anyNA(x$Region))


# Quantiles - from https://www.tidyverse.org/blog/2020/03/dplyr-1-0-0-summarise/
# Note: could swap these out to use normalized values instead
quibble2 <- function(x, q = c(0.025, 0.25, 0.50, 0.75, 0.975)) {
  tibble("{{ x }}" := quantile(x, q, na.rm = TRUE), "{{ x }}_q" := q)
}


bm.q <- x %>% 
  group_by(year_month, Region) %>% 
  summarise(quibble2(Blue_occurrence_mean), 
            .groups = "drop") %>% 
  mutate(Blue_occurrence_mean = unname(Blue_occurrence_mean)) %>% 
  tidyr::pivot_wider(names_from = Blue_occurrence_mean_q, names_prefix = "bm_occ_", 
                     values_from = Blue_occurrence_mean)

mn.q <- x %>% 
  group_by(year_month, Region) %>% 
  summarise(quibble2(Humpback_dens), 
            .groups = "drop") %>% 
  mutate(Humpback_dens = unname(Humpback_dens)) %>% 
  tidyr::pivot_wider(names_from = Humpback_dens_q, names_prefix = "mn_dens_", 
                     values_from = Humpback_dens)


bm.q.ca <- x %>% 
  group_by(year_month) %>% 
  summarise(quibble2(Blue_occurrence_mean), 
            .groups = "drop") %>% 
  mutate(Blue_occurrence_mean = unname(Blue_occurrence_mean)) %>% 
  tidyr::pivot_wider(names_from = Blue_occurrence_mean_q, names_prefix = "bm_occ_", 
                     values_from = Blue_occurrence_mean)

mn.q.ca <- x %>% 
  group_by(year_month) %>% 
  summarise(quibble2(Humpback_dens), 
            .groups = "drop") %>% 
  mutate(Humpback_dens = unname(Humpback_dens)) %>% 
  tidyr::pivot_wider(names_from = Humpback_dens_q, names_prefix = "mn_dens_", 
                     values_from = Humpback_dens)




x.summ <- x %>% 
  group_by(year_month, Region) %>% 
  summarise(bm_occ_mean = mean(Blue_occurrence_mean, na.rm = TRUE),
            mn_dens_mean = mean(Humpback_dens, na.rm = TRUE), 
            bm_normalized_med = median(normalized_blue, na.rm = TRUE),
            mn_normalized_med = median(normalized_humpback, na.rm = TRUE),
            .groups = "drop") %>% 
  left_join(bm.q, by = c("year_month", "Region")) %>% 
  left_join(mn.q, by = c("year_month", "Region"))

x.summ.ca <- x %>% 
  group_by(year_month) %>% 
  summarise(bm_occ_mean = mean(Blue_occurrence_mean, na.rm = TRUE),
            mn_dens_mean = mean(Humpback_dens, na.rm = TRUE),
            bm_normalized_med = median(normalized_blue, na.rm = TRUE),
            mn_normalized_med = median(normalized_humpback, na.rm = TRUE),
            .groups = "drop") %>% 
  mutate(Region = factor("CA")) %>% 
  left_join(bm.q.ca, by = "year_month") %>% 
  left_join(mn.q.ca, by = "year_month")



x.max <- length(unique(x.summ$year_month))
x.lab.idx <- seq(1, to = x.max, by = 3)
x.lab <- sort(unique(x.summ$year_month))[x.lab.idx]
vert.lines <- seq(0.5, to = x.max, by = 12)
# vert.lines2 <- seq(6.5, to = x.max, by = 12)
```

Function for generating time series plots

```{r}
raimbow_ca_ts <- function(obj.df, x.aes.x, y, plot.main = NULL, y.lab = NULL, filter.reg.idx, 
                          month.flag = FALSE, y.range = NULL) {
  ### Inputs
  # obj.df: summary of risk, etc, by month and region
  # y: unquoted name of value to be plotted
  # wa.flag: logical indicating whether WA values should be plotted
  # filter.reg.idx: index number indicating regions to be filtered for
  
  x.aes.x <- enquo(x.aes.x)
  y <- enquo(y)
  
  x.breaks <- if (month.flag) month.abb else x.lab
  if (is.null(y.range)) y.range <- c(min(select(obj.df, !!y)), max(select(obj.df, !!y)))
  # y.max <- max(select(obj.df, !!y))
  
  tmp <- obj.df %>% 
    mutate(DC_season = factor(1)) %>% 
    # ggplot(aes(!!x.aes.x, !!y, colour = Region, group = Region, linetype = DC_season)) +
    ggplot(aes(!!x.aes.x, !!y, group = Region)) +
    geom_point() + 
    geom_line()
  
  if (!month.flag)  tmp <- tmp + 
    geom_vline(xintercept = vert.lines, col = "grey", lwd = 0.35) #+
  # geom_vline(xintercept = vert.lines2, col = "green", lwd = 0.35) 
  
  
  tmp + 
    # scale_colour_manual(values = cols.vals, drop = TRUE, name = "Region") +
    # scale_colour_brewer(palette = "Set1", name = "Region", drop = FALSE) +
    # guides(linetype = guide_legend(title = "DC season", label = FALSE,
    #                                override.aes = list(colour = "black"))) +
    
    # coord_cartesian(xlim = c(-1, x.max)) + 
    scale_x_discrete(breaks = x.breaks) +
    xlab (NULL) + 
    ylab(y.lab) + 
    ylim(y.range[1], y.range[2]) + 
    # ylim(0, y.max) + 
    ggtitle(plot.main) + 
    theme_classic() +
    theme(plot.title = element_text(hjust = 0.5), 
          axis.text.x = element_text(angle = 90, vjust = 0.4), 
          legend.justification = "top")
}


###############################################################################
raimbow_ca_ts_facets <- function(x.data, x.aes.x, x.aes.y, y.lab, title, 
                                 month.flag = FALSE, y.range = NULL) {
  x.aes.x <- enquo(x.aes.x)
  x.aes.y <- enquo(x.aes.y)
  x.breaks <- if (month.flag) month.abb else x.lab
  if (is.null(y.range)) 
    y.range <- c(min(select(x.data, !!x.aes.y)), max(select(x.data, !!x.aes.y)))
  
  tmp <- ggplot(
    data = x.data, 
    aes(!!x.aes.x, !!x.aes.y, group = Region)
  ) +
    geom_point(size = 1) +
    geom_line()
  
  if (!month.flag) tmp <- tmp + 
    geom_vline(xintercept = vert.lines, col = "grey", lwd = 0.35) 
  
  # geom_vline(xintercept = vert.lines, col = "grey", lwd = 0.35) +
  # geom_vline(xintercept = c(2014.5, 2018.5), linetype=2) +
  tmp + 
    facet_grid(rows = vars(Region),
               scales = "free_y",
               labeller = as_labeller(c(
                 `CenCA` = "Central CA",
                 `NorCA` = "Northern CA"
               )),
               switch = "y"
    ) + 
    # scale_x_continuous(breaks = seq(2010, 2019),
    #                    labels = str_replace(unique(annual_risk_out_sq$crab_year),"_","-")) + 
    ylim(y.range[1], y.range[2]) + 
    ylab(y.lab) + 
    xlab("") +
    scale_x_discrete(breaks = x.breaks) +
    ggtitle(title) + 
    theme_classic() +
    theme(legend.title = element_blank(),
          #title = element_text(size = 26),
          # legend.text = element_text(size = 20),
          legend.position = c(.15, .85),
          plot.title = element_text(hjust = 0.5), 
          # axis.text.x = element_text(angle = 90, vjust = 0.4), 
          axis.text.x = element_text(hjust = 1, angle = 60),
          # axis.text.y = element_text(size = 14),
          # axis.title = element_text(size = 16),
          # strip.text = element_text(size=14),
          strip.background = element_blank(),
          strip.placement = "left"
    )
}
```

## Blue whale time series

```{r}
raimbow_ca_ts(
  x.summ.ca, year_month, bm_normalized_med, plot.main = "Blue whales - California - median - normalized",
  y.lab = "Whales / km2", filter.reg.idx = c(1)
)

raimbow_ca_ts(
  x.summ.ca, year_month, bm_occ_0.5, plot.main = "Blue whales - California - median",
  y.lab = "Whales / km2", filter.reg.idx = c(1)
)

raimbow_ca_ts_facets(
  x.summ, year_month, bm_occ_0.5, 
  "Probability of occurrence", "Blue whales - median"
)

# ggsave()
```

## Humpback time series

```{r}
raimbow_ca_ts(
  x.summ.ca, year_month, mn_normalized_med, plot.main = "Humpback whales - California - median - normalized",
  y.lab = "Whales / km2", filter.reg.idx = c(1)
)

raimbow_ca_ts(
  x.summ.ca, year_month, mn_dens_0.5, plot.main = "Humpback whales - California - median",
  y.lab = "Whales / km2", filter.reg.idx = c(1)
)
raimbow_ca_ts_facets(
  x.summ, year_month, mn_dens_0.5, "Whales / km2", "Humpback whales - median"
)
```

## Seasonality 2014-2018

First we must generate data frames with averages by month for our desired time periods. Of note - the year-months start January of 2009 and run through July 2019, so there are not equal number of records for all months.

NOTE: These averages are calculated using the original data (one value per grid cell per month), not the median values for each year-month

```{r}
# TODO by Jameal: update y value if needed, and update plot titles and file names
x.month <- x %>%  
  mutate(year = as.numeric(substr(year_month, 1, 4)), 
         month = as.numeric(substr(year_month, 6, 7)), 
         ym_date = as.Date(paste(year, month, "01", sep = "-"))) %>% 
  filter(between(ym_date, as.Date("2009-11-01"), as.Date("2014-10-15")))
# filter(between(ym_date, as.Date("2014-11-01"), as.Date("2018-10-15")))
# filter(between(ym_date, as.Date("2018-11-01"), as.Date("2019-10-15")))

range(x.month$ym_date)
```

This code should be able to be run with no alterations

```{r}
source(here::here("tradeoffs", "whale_predictions_timeseries_helper.R"), local = TRUE)
# ###############################################################################
# bm.q.month <- x.month %>% 
#   group_by(month, Region) %>% 
#   summarise(quibble2(Blue_occurrence_mean), 
#             .groups = "drop") %>% 
#   mutate(Blue_occurrence_mean = unname(Blue_occurrence_mean)) %>% 
#   tidyr::pivot_wider(names_from = Blue_occurrence_mean_q, names_prefix = "bm_occ_", 
#                      values_from = Blue_occurrence_mean)
# 
# mn.q.month <- x.month %>% 
#   group_by(month, Region) %>% 
#   summarise(quibble2(Humpback_dens), 
#             .groups = "drop") %>% 
#   mutate(Humpback_dens = unname(Humpback_dens)) %>% 
#   tidyr::pivot_wider(names_from = Humpback_dens_q, names_prefix = "mn_dens_", 
#                      values_from = Humpback_dens)
# 
# 
# bm.q.ca.month <- x.month %>% 
#   group_by(month) %>% 
#   summarise(quibble2(Blue_occurrence_mean), 
#             .groups = "drop") %>% 
#   mutate(Blue_occurrence_mean = unname(Blue_occurrence_mean)) %>% 
#   tidyr::pivot_wider(names_from = Blue_occurrence_mean_q, names_prefix = "bm_occ_", 
#                      values_from = Blue_occurrence_mean)
# 
# mn.q.ca.month <- x.month %>% 
#   group_by(month) %>% 
#   summarise(quibble2(Humpback_dens), 
#             .groups = "drop") %>% 
#   mutate(Humpback_dens = unname(Humpback_dens)) %>% 
#   tidyr::pivot_wider(names_from = Humpback_dens_q, names_prefix = "mn_dens_", 
#                      values_from = Humpback_dens)
# 
# 
# 
# ###############################################################################
# x.summ.month <- x.month %>%   
#   group_by(month, Region) %>%
#   summarise(bm_occ_mean = mean(Blue_occurrence_mean, na.rm = TRUE),
#             mn_dens_mean = mean(Humpback_dens, na.rm = TRUE),
#             bm_normalized_med = median(normalized_blue, na.rm = TRUE),
#             mn_normalized_med = median(normalized_humpback, na.rm = TRUE),
#             .groups = "drop") %>%
#   left_join(bm.q.month, by = c("month", "Region")) %>%
#   left_join(mn.q.month, by = c("month", "Region")) %>% 
#   mutate(month = factor(month.abb[month], levels = month.abb))
# 
# x.summ.ca.month <- x.month %>%   
#   group_by(month) %>%
#   summarise(bm_occ_mean = mean(Blue_occurrence_mean, na.rm = TRUE),
#             mn_dens_mean = mean(Humpback_dens, na.rm = TRUE),
#             bm_normalized_med = median(normalized_blue, na.rm = TRUE),
#             mn_normalized_med = median(normalized_humpback, na.rm = TRUE),
#             .groups = "drop") %>%
#   left_join(bm.q.ca.month, by = c("month")) %>%
#   left_join(mn.q.ca.month, by = c("month")) %>% 
#   mutate(Region = "CA", 
#          month = factor(month.abb[month], levels = month.abb))
```

Here the user must update the y values, plot titles, and file names

```{r}
# Plot

raimbow_ca_ts(
  x.summ.ca.month, month, mn_normalized_med, plot.main = "Humpback whale - normalized median - Nov 2009 to Oct 2014", 
  y.lab = "Whales / km2", filter.reg.idx, month.flag = TRUE
)
# ggsave(filename.out, plot = month.plot)

raimbow_ca_ts_facets(
  x.summ.month, month, mn_dens_0.5, "Whales / km2", "Humpback whale - median - Nov 2009 to Oct 2014", 
  month.flag = TRUE
)
# ggsave(filename.out, plot = month.plot)

raimbow_ca_ts(
  x.summ.ca.month, month, bm_normalized_med, "Blue whale - normalized median - Nov 2009 to Oct 2014", 
  y.lab = "Probability of occurrence", filter.reg.idx, month.flag = TRUE
)
# ggsave(filename.out, plot = month.plot)

raimbow_ca_ts_facets(
  x.summ.month, month, bm_occ_0.5, "Whales / km2", "Blue whale - median - Nov 2009 to Oct 2014", 
  month.flag = TRUE
)
# ggsave(filename.out, plot = month.plot)
```

## Seasonality 2014-2018

```{r}
x.month <- x %>%  
  mutate(year = as.numeric(substr(year_month, 1, 4)), 
         month = as.numeric(substr(year_month, 6, 7)), 
         ym_date = as.Date(paste(year, month, "01", sep = "-"))) %>% 
  # filter(between(ym_date, as.Date("2009-11-01"), as.Date("2014-10-15")))
  filter(between(ym_date, as.Date("2014-11-01"), as.Date("2018-10-15")))
# filter(between(ym_date, as.Date("2018-11-01"), as.Date("2019-10-15")))

range(x.month$ym_date)


source(here::here("tradeoffs", "whale_predictions_timeseries_helper.R"), local = TRUE)



raimbow_ca_ts(
  x.summ.ca.month, month, mn_normalized_med, plot.main = "Humpback whale - normalized median - Nov 2014 to Oct 2018", 
  y.lab = "Whales / km2", filter.reg.idx, month.flag = TRUE
)
# ggsave(filename.out, plot = month.plot)

raimbow_ca_ts_facets(
  x.summ.month, month, mn_dens_0.5, "Whales / km2", "Humpback whale - median - Nov 2014 to Oct 2018", 
  month.flag = TRUE
)
raimbow_ca_ts_facets(
  x.summ.month, month, mn_dens_mean, "Whales / km2", "Humpback whale - mean - Nov 2014 to Oct 2018", 
  month.flag = TRUE
)
# ggsave(filename.out, plot = month.plot)

raimbow_ca_ts(
  x.summ.ca.month, month, bm_normalized_med, "Blue whale - normalized median - Nov 2014 to Oct 2018", 
  y.lab = "Probability of occurrence", filter.reg.idx, month.flag = TRUE
)
# ggsave(filename.out, plot = month.plot)

raimbow_ca_ts_facets(
  x.summ.month, month, bm_occ_0.5, "Whales / km2", "Blue whale - median - Nov 2014 to Oct 2018", 
  month.flag = TRUE
)
# ggsave(filename.out, plot = month.plot)
```

## Seasonality 2018-2019

```{r}
x.month <- x %>%  
  mutate(year = as.numeric(substr(year_month, 1, 4)), 
         month = as.numeric(substr(year_month, 6, 7)), 
         ym_date = as.Date(paste(year, month, "01", sep = "-"))) %>% 
  # filter(between(ym_date, as.Date("2009-11-01"), as.Date("2014-10-15")))
  # filter(between(ym_date, as.Date("2014-11-01"), as.Date("2018-10-15")))
  filter(between(ym_date, as.Date("2018-11-01"), as.Date("2019-10-15")))

range(x.month$ym_date)


source(here::here("tradeoffs", "whale_predictions_timeseries_helper.R"), local = TRUE)



raimbow_ca_ts(
  x.summ.ca.month, month, mn_normalized_med, plot.main = "Humpback whale - normalized median - Nov 2018 to Oct 2019", 
  y.lab = "Whales / km2", filter.reg.idx, month.flag = TRUE
)
# ggsave(filename.out, plot = month.plot)

raimbow_ca_ts_facets(
  x.summ.month, month, mn_dens_0.5, "Whales / km2", "Humpback whale - median - Nov 2018 to Oct 2019", 
  month.flag = TRUE
)
# ggsave(filename.out, plot = month.plot)

raimbow_ca_ts(
  x.summ.ca.month, month, bm_normalized_med, "Blue whale - normalized median - Nov 2018 to Oct 2019", 
  y.lab = "Probability of occurrence", filter.reg.idx, month.flag = TRUE
)
# ggsave(filename.out, plot = month.plot)

raimbow_ca_ts_facets(
  x.summ.month, month, bm_occ_0.5, "Whales / km2", "Blue whale - median - Nov 2018 to Oct 2019", 
  month.flag = TRUE
)
# ggsave(filename.out, plot = month.plot)
```
