---
title: "Overlay blue whale predictions"
author: "Sam Woodman"
date: '`r format(Sys.Date(), "%d %B %Y")`'
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The goal of this document is to aggregate the overlaid blue whale predictions by month, calculating both the mean and SD values for each grid cell for each year-month from Nov 2009 to the most recent predictions. When rerunning this file after adding new predictions, the user must update `date.max`. This value should be the first day of the last month with predictions to ensure that only months with complete sets of predictions are included.

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)

source(here::here("User_script_local.R"))
source(here::here("whalepreds_aggregate", "Whalepreds_aggregate.R"), local = TRUE, echo = FALSE)

if (user == "JS") {
  
} else if (user == "SMW") {
  path.bwpreds.rds <- "C:/SMW/RAIMBOW/raimbow-local/Outputs/BlueWhale_5km_wide_daily.rds"

  file.out <- "C:/SMW/RAIMBOW/raimbow-local/Outputs/BlueWhale_5km_long_monthly.rds"
}


#Leena
path.bwpreds.rds <- "C:/Users/Leena.Riekkola/Projects/NOAA data/maps_ts_whales/data/BlueWhale_5km_wide_daily.rds"

file.out <- "C:/Users/Leena.Riekkola/Projects/NOAA data/maps_ts_whales/data/BlueWhale_5km_long_monthly_2019Aug_2021Sep.rds"
```

## Aggregate by month and tidy

User TODO: update `date.max`. This section can take a few minutes to run

```{r}
bm.raw <- readRDS(path.bwpreds.rds)
tail(names(bm.raw))

#date.max <- as.Date("2019-08-01")
date.max <- as.Date("2021-09-30") #max date in new data pull is 2021-09-30
#range.dates <- seq(from = as.Date("2009-01-01"), to = date.max, by = "months")
range.dates <- seq(from = as.Date("2019-08-16"), to = date.max, by = "months") #the first date of the new data pull is 2019-08-16

bm.aggr <- whalepreds_aggregate(
  #bm.raw, 3:ncol(bm.raw), 4:13, aggr.level = NULL, range.dates = range.dates, #need to change the indices of characters in column names that specify the date 
  bm.raw, 3:ncol(bm.raw), 18:27, aggr.level = NULL, range.dates = range.dates, #there were some gaps in the bw output, which causes a warning message here
  se.calc = TRUE
) %>% 
  set_names(c("GRID5KM_ID", "area_km_lno", paste0("Bm_", names(.)[-c(1:2)]))) %>% 
  set_names(gsub("Avg_user_", "", names(.))) %>% 
  set_names(gsub("user_", "", names(.)))

# # Sam's musings
# bm.raw.long <- bm.raw %>% 
#   gather(key = "key", value = "Bm_occur", -GRID5KM_ID, -area_km_lno)
# bm.aggr2 <- bm.raw.long %>% 
#   mutate(date = ymd(substr(key, 4, 13)), 
#          yr = year(date), 
#          mon = month(date)) %>% 
#   group_by(yr, mon)
```

Now we make the data long, extract means and SEs, ...

```{r}
bm.proc <- bm.aggr %>% 
  gather(key = "key", value = "value", -GRID5KM_ID, -area_km_lno) %>% 
  mutate(type = ifelse(grepl("SE", key), "se", "pred"), ##what is grepl??
         date = lubridate::ymd(ifelse(type == "se", substr(key, 7, 16), substr(key, 4, 13))))


bm.proc.long <- bm.proc %>% 
  select(-key) %>%
  pivot_wider(names_from = type, values_from = value) %>% 
  rename(Blue_occurrence_mean = pred, Blue_occurrence_se = se)

#in previous data pull the date was always the first of a month. But in new data pull the first file was 2019-08-16
```

...and save

```{r}
saveRDS(bm.proc.long, file = file.out)
```
