---
title: "Overlay blue whale predictions"
author: "Sam Woodman"
date: '`r format(Sys.Date(), "%d %B %Y")`'
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this document, blue whale predictions from Abrahms et al. 2019 are overlaid onto the 5km equal area grid using the 'overlay' method (areal interpolation). This is done to minimize the amount of information lost, particularly in coastal grid cells. Processing includes reading in the original raster predictions, converting these rasters to sf objects, and then doing the overlay.

```{r, message=FALSE}
library(dplyr)
# devtools::install_github("smwoodman/eSDM") #Currently need dev version of eSDM
library(eSDM)
library(sf)
library(stars)


source(here::here("User_script_local.R"))
if (user == "JS") {
  
} else if (user == "SMW") {
  path.bw.preds <- "C:/SMW/RAIMBOW/raimbow-local/Data/Blue whale predictions/"
  path.grid.5km.lno <- "C:/SMW/RAIMBOW/raimbow-local/RDATA_files/Grid_5km_landerased.rds"
  
  file.out.tmp <- "C:/SMW/RAIMBOW/raimbow-local/RDATA_files/Bm_tmp_over.rds"
  
  file.out.csv <- "C:/SMW/RAIMBOW/raimbow-local/Outputs/BlueWhale_5km_wide_daily.csv"
  file.out.rds <- "C:/SMW/RAIMBOW/raimbow-local/Outputs/BlueWhale_5km_wide_daily.rds"
}
```

## Import blue whale predictions

First we must import the blue whale predictions, expected to be from 2009 to present. These predictions have been saved as rasters; we read them in as stars objects because it is more straightforward to convert stars objects to sf objects

```{r}
file.count <- length(list.files(path.bw.preds))
files.grd1 <- list.files(path.bw.preds, pattern = ".grd")
files.grd <- files.grd1[!grepl("aux", files.grd1)]
table(substr(files.grd, 1, 4)) #See the number of files for each year

## IMPORTANT NOTES: for read_star() to work, the source folder MUST contain both .grd AND .gri files 
## ALSO: when specifying file path to path.bw.preds, you MUST to have "/" at the end for the below paste0 to join folder and file names correctly
list.stars <- lapply(paste0(path.bw.preds, files.grd), read_stars) #~30s
names(list.stars) <- paste0("Bm_", files.grd)

stopifnot( #unnecessary sanity check
  all.equal(list.stars[[1]], read_stars(paste(path.bw.preds, files.grd[1], sep = "/")))
)
```

After importing, we convert the stars objects to sf objects

```{r}
bm.stars <- do.call(c, list.stars) #convert to actual stars object
bm.sf <- st_as_sf(bm.stars) %>% #~20s
  purrr::set_names(gsub("[.]", "_", substr(names(.), 1, 13))) %>% 
  st_set_crs(4326)
```

## Overlay and write output

The overlay takes ~1.5hrs on Sam's computer; it takes so much longer than the humpback overlay because the blue whale predictions go so much farther offshore. If desired, you can copy the code sections from this document into an R script to run as a job in the background.

```{r}
# Load 5km ea grid, with land erased
grid.5km.lno <- readRDS(path.grid.5km.lno)

# Do overlay
preds.names <- names(st_drop_geometry(bm.sf))
tmp.over <- overlay_sdm(st_geometry(grid.5km.lno), bm.sf, preds.names, overlap.perc = 0)

# saveRDS(tmp.over, file = file.out.tmp)
# tmp.over <- readRDS(file.out.tmp)

bm.preds.lno.5kmover <- tmp.over %>% 
  mutate(GRID5KM_ID = grid.5km.lno$GRID5KM_ID, 
         area_km_lno = as.numeric(units::set_units(st_area(geometry), "km^2"))) %>% 
  select(GRID5KM_ID, area_km_lno, starts_with("Bm_"))

# # Sanity check
# identical(st_geometry(bm.preds.lno.5kmover), st_geometry(grid.5km.lno))
# eSDM::model_abundance(bm.sf, "Bm_2009_01_05")
# eSDM::model_abundance(bm.preds.lno.5kmover, "Bm_2009_01_05")

# # Shows that all columns have NAs in the same rows
# x.df <- st_drop_geometry(bm.preds.lno.5kmover) %>% select(starts_with("Bm_"))
# x.na.which <- purrr::map(x.df, function(i) which(is.na(i)))
# x.temp <- vapply(
#   x.na.which, function(i, j) {identical(i, j)}, as.logical(1),
#   j = x.na.which[[1]])
# all(x.temp)

# Remove rows with all NAs
bm.preds.lno.5kmover.nona <- bm.preds.lno.5kmover %>% 
  filter(!is.na(Bm_2009_01_01)) %>% 
  st_drop_geometry()
```

Write to output. Also save as an RDS for the sake of reading it back in later (aka faster)

```{r}
readr::write_csv(bm.preds.lno.5kmover.nona, path = file.out.csv)
saveRDS(bm.preds.lno.5kmover.nona, file = file.out.rds)
```
