---
title: "Pre-processing OR logbook data"
author: "Leena"
date: "11/08/2021"
output: html_document
---

# Pre-processing OR logbook data so that the code used for WDFW logbook analysis and mapping can be used for OR as well
# The code here are based on code files received from WDFW:
# 'Read and prep crab logbook data_2020-08-28.Rmd' and
# 'Create-and-save-GIS-sf_and_gdb.Rmd'

```{r}
library(tidyverse)
library(here)
library(data.table)
library(stringr)
library(magrittr)
library(lubridate)
library(sf)

```
# Read in data

In the logbook data folder received from ODFW, there was a file Samhouri_CrabLogData_sent030620.xlsx
This file has logbook data covering seasons from 2007-2008 to 2017-2018
and it was saved as csv prior to loading into R

```{r}
logs <- read_csv(here::here('wdfw','data','OR','Samhouri_CrabLogData_sent030620.csv'),col_types = 'cdccddddddddddccc')
#fields such as vessel ID is often a  number, but want to read it in as a character

#rename some variables to match WDFW data
logs %<>% 
  rename(season=Season, Vessel=VessID, SetDate=Date, Depth_fth=Depth, PotsFished=Pots, adj_lbs=AdjLbs, adj_val=AdjVal, LATBEG=BegLat, LONGBEG=Beglon, LATEND=EndLat, LONGEND=EndLon)

glimpse(logs)
#note that not all columns exist in both WA and OR logbook data
```
# Format SetDate column
```{r}
logs %<>%
  mutate(SetDate=as.Date(SetDate,"%d-%b-%y"))

```
# Add an ID

Each row in the OR logbook data should reflect an individual string-line of pots
First add an arbitrary ID value starting from 1 for the entire data set
Then create a 'SetID' that incorporates the season in question, and the ID number
The value of the SetID doesn't matter that much, but we want to make sure each string-line has a unique ID

```{r}
logs$IDsets <- seq.int(nrow(logs))

logs$SetID <- paste0(logs$season, "_", logs$IDsets)

```
#The following is from 'Create-and-save-GIS-sf_and_gdb.Rmd' from WDFW
this Rmd was ran up to line 58 (in the original Rmd from WDFW) for both WA and OR logbook data

# Stack Coordinates
the code is from WDFW and uses data.table - convert to data.table first to run the code
```{r}
setDT(logs) 

log_coords <- rbind(logs[, .(lat = LATBEG,
               lon = LONGBEG,
               coord_type = "begin"), keyby = SetID],
      logs[, .(lat = LATEND,
               lon = LONGEND,
               coord_type = "end"), keyby = SetID])

log_coords <- log_coords[complete.cases(log_coords), ]

log_coords$lon <- log_coords$lon*(-1)
```
# Rejoin Attributes

This is an inner join using standard data.table syntax (i.e. select rows of logs, filtering on SetIDs that appear in log_coords, and nomatch=NULL tells it to drop rows with no matches.)
```{r}
logs_sf_pts <- logs[log_coords, on = "SetID", nomatch = NULL]
```
## we only ran 'Create-and-save-GIS-sf_and_gdb.Rmd'on WA logs data up to this point (i.e. did not save GIS files)

# OPTIONAL
consider removing any records for whom the spatial information is unreasonable (variable `SpatialFlag==T`)?
Alternatively, leave in for now and remove later, also these points should get removed in script 1 when removing points too deep or on land

The WA logbook coding from WDFW added start dates into the df, but these were not used in analysis and mapping, therefore not adding those in for OR

```{r}

```
# Couple additional check from 'Read and prep crab logbook data_2020-08-28.Rmd'
## Check for errors

There should only be two years within a season. 
2015-2016 and 2017-2018 only have data in the latter years - Fits because season opening date was only in January

First, get the season start and ending years. Test to see if the Set Date year matches one of those years.

The OR logs did not come with landing dates, but they could probably be retrieved using FishTicket numbers
```{r}
#logs_sf_pts[, .(year = unique(year(SetDate))), 
#     keyby = season][, dcast(.SD, season ~ year)]

#logs_sf_pts[, c("start_year", "end_year") := tstrsplit(season, "-")
#     ][, `:=`(start_year = as.numeric(start_year), 
#              end_year = as.numeric(end_year))]

#check_setdate <- logs_sf_pts[year(SetDate) != start_year &
#       year(SetDate) != end_year, ]

#check_setdate #empty
```
#Export as csv
```{r}
#Originally exported data here, but needed to join with permit data first
# fwrite(logs_sf_pts, paste0(here::here(), 
#                     "/wdfw/", "/data/", "/OR/", 
#                     "ODFW-Dcrab-logbooks-compiled_stackcoords_2007-2018", 
#                     "_", 
#                     Sys.Date(), ".csv"))

```

#JOINING PERMIT DATA
```{r}
# From data.table to data.frame
setDF(logs_sf_pts)
# for some reason joining didn't work if read in the ODFW-Dcrab-logbooks-compiled_stackcoords_2007-2018.csv that was saved, but worked on the logs_sf_pts dataframe (which was saved as ODFW-Dcrab-logbooks-compiled_stackcoords_2007-2018.csv)


library(fuzzyjoin)
# OR permits: 
# in folder from ODFW had OregonCrabPermitData2007-2019.xlsx, saved as csv 
# Read in and join license & pot limit info
OR_pot_limit_info_raw <- read_csv(here::here('wdfw', 'data', 'OR', 'OregonCrabPermitData2007-2019.csv'))

OR_pot_limit_info <- OR_pot_limit_info_raw %>% 
  rename(Vessel = Docnum,
         PermitNumber = Number)

OR_pot_limit_info %<>%
  mutate(Begindate=as.Date(Begindate,"%m/%d/%Y"),
         Enddate=as.Date(Enddate,"%m/%d/%Y"))

OR_pot_limit_info %>% distinct(Potlimit) # 500, 300, 200
# OR permits - a permit can change from vessel to vessel sometimes 
# but does the pot limit for a given permit number stay the same?
test <- OR_pot_limit_info %>%                              
  group_by(PermitNumber) %>%
  summarise(count = n_distinct(Potlimit))
# Yes, except for 2 instances: Permit Numbers 96125 and 96262 have 2 unique pot limit values
cases <- OR_pot_limit_info %>% 
  filter(PermitNumber == 96125 | PermitNumber == 96262)
# 96125: for 12 years pot limit is 300, but in 2014 it is 500 - assume mistake for now
# 96262: for 12 years pot limit is 300, but in 2008 it is 200 - assume mistake for now, also possibly outside years of interest anyway
OR_pot_limit_info %<>%
  mutate(Potlimit = ifelse(PermitNumber == 96125 | PermitNumber == 96262, 300, Potlimit))

OR_pot_limit_info_v2 <- OR_pot_limit_info %>% 
  filter(Year >= 2013) %>% 
  select(PermitNumber, Vessel, Begindate, Enddate, Potlimit)

# joining permit data
# doing it 'manually' for now while figuring out how to code it more efficiently

#2013-2014 season - 12mins to run with permit data filtered to post 2013
ODFW_Dcrab_logbooks_20132014 <- logs_sf_pts %>% 
  filter(season=='2013-2014')
tm <- proc.time()
ODFW_Dcrab_logbooks_20132014_joined <- fuzzy_left_join(
  ODFW_Dcrab_logbooks_20132014, OR_pot_limit_info_v2,
  by = c(
    "Vessel" = "Vessel",
    "SetDate" = "Begindate",
    "SetDate" = "Enddate"
  ),
  match_fun = list(`==`, `>=`, `<=`)
)
proc.time()-tm
write_csv(ODFW_Dcrab_logbooks_20132014_joined,here::here('wdfw', 'data', 'OR', "ODFW-Dcrab-logbooks-compiled_stackcoords_license_2013-2014.csv"))


# 2014-2015 season - 12mins to run with permit data filtered to post 2013
ODFW_Dcrab_logbooks_20142015 <- logs_sf_pts %>% 
  filter(season=='2014-2015')
tm <- proc.time()
ODFW_Dcrab_logbooks_20142015_joined <- fuzzy_left_join(
  ODFW_Dcrab_logbooks_20142015, OR_pot_limit_info_v2,
  by = c(
    "Vessel" = "Vessel",
    "SetDate" = "Begindate",
    "SetDate" = "Enddate"
  ),
  match_fun = list(`==`, `>=`, `<=`)
)
proc.time()-tm
write_csv(ODFW_Dcrab_logbooks_20142015_joined,here::here('wdfw', 'data', 'OR', "ODFW-Dcrab-logbooks-compiled_stackcoords_license_2014-2015.csv"))


# 2015-2016 season - 15mins to run with permit data filtered to post 2013
ODFW_Dcrab_logbooks_20152016 <- logs_sf_pts %>% 
  filter(season=='2015-2016')
tm <- proc.time()
ODFW_Dcrab_logbooks_20152016_joined <- fuzzy_left_join(
  ODFW_Dcrab_logbooks_20152016, OR_pot_limit_info_v2,
  by = c(
    "Vessel" = "Vessel",
    "SetDate" = "Begindate",
    "SetDate" = "Enddate"
  ),
  match_fun = list(`==`, `>=`, `<=`)
)
proc.time()-tm
write_csv(ODFW_Dcrab_logbooks_20152016_joined,here::here('wdfw', 'data', 'OR', "ODFW-Dcrab-logbooks-compiled_stackcoords_license_2015-2016.csv"))


# 2016-2017 season - 14mins to run with permit data filtered to post 2013
ODFW_Dcrab_logbooks_20162017 <- logs_sf_pts %>% 
  filter(season=='2016-2017')
tm <- proc.time()
ODFW_Dcrab_logbooks_20162017_joined <- fuzzy_left_join(
  ODFW_Dcrab_logbooks_20162017, OR_pot_limit_info_v2,
  by = c(
    "Vessel" = "Vessel",
    "SetDate" = "Begindate",
    "SetDate" = "Enddate"
  ),
  match_fun = list(`==`, `>=`, `<=`)
)
proc.time()-tm
write_csv(ODFW_Dcrab_logbooks_20162017_joined,here::here('wdfw', 'data', 'OR', "ODFW-Dcrab-logbooks-compiled_stackcoords_license_2016-2017.csv"))


# 2017-2018 season - 14mins to run with permit data filtered to post 2013
ODFW_Dcrab_logbooks_20172018 <- logs_sf_pts %>% 
  filter(season=='2017-2018')
tm <- proc.time()
ODFW_Dcrab_logbooks_20172018_joined <- fuzzy_left_join(
  ODFW_Dcrab_logbooks_20172018, OR_pot_limit_info_v2,
  by = c(
    "Vessel" = "Vessel",
    "SetDate" = "Begindate",
    "SetDate" = "Enddate"
  ),
  match_fun = list(`==`, `>=`, `<=`)
)
proc.time()-tm
write_csv(ODFW_Dcrab_logbooks_20172018_joined,here::here('wdfw', 'data', 'OR', "ODFW-Dcrab-logbooks-compiled_stackcoords_license_2017-2018.csv"))


#2011-2012 season -- 4mins to run with subset permit data - oddly quick
OR_pot_limit_info_v3 <- OR_pot_limit_info %>% 
  filter(Year >= 2011 & Year < 2013) %>% 
  select(PermitNumber, Vessel, Begindate, Enddate, Potlimit)

ODFW_Dcrab_logbooks_20112012 <- logs_sf_pts %>% 
  filter(season=='2011-2012')
tm <- proc.time()
ODFW_Dcrab_logbooks_20112012_joined <- fuzzy_left_join(
  ODFW_Dcrab_logbooks_20112012, OR_pot_limit_info_v3,
  by = c(
    "Vessel" = "Vessel",
    "SetDate" = "Begindate",
    "SetDate" = "Enddate"
  ),
  match_fun = list(`==`, `>=`, `<=`)
)
proc.time()-tm
write_csv(ODFW_Dcrab_logbooks_20112012_joined,here::here('wdfw', 'data', 'OR', "ODFW-Dcrab-logbooks-compiled_stackcoords_license_2011-2012.csv"))

#2012-2013 season -- 4mins to run with subset permit data - oddly quick
OR_pot_limit_info_v3 <- OR_pot_limit_info %>% 
  filter(Year >= 2012 & Year < 2014) %>% 
  select(PermitNumber, Vessel, Begindate, Enddate, Potlimit)

ODFW_Dcrab_logbooks_20122013 <- logs_sf_pts %>% 
  filter(season=='2012-2013')
tm <- proc.time()
ODFW_Dcrab_logbooks_20122013_joined <- fuzzy_left_join(
  ODFW_Dcrab_logbooks_20122013, OR_pot_limit_info_v3,
  by = c(
    "Vessel" = "Vessel",
    "SetDate" = "Begindate",
    "SetDate" = "Enddate"
  ),
  match_fun = list(`==`, `>=`, `<=`)
)
proc.time()-tm
write_csv(ODFW_Dcrab_logbooks_20122013_joined,here::here('wdfw', 'data', 'OR', "ODFW-Dcrab-logbooks-compiled_stackcoords_license_2012-2013.csv"))

##################
#logbook years when 100% logs entered

#2010-2011 season -- last season with full 100% logbook entry - 13mins to run with subset permit data 
OR_pot_limit_info_v3 <- OR_pot_limit_info %>% 
  filter(Year >= 2010 & Year < 2012) %>% 
  select(PermitNumber, Vessel, Begindate, Enddate, Potlimit)

ODFW_Dcrab_logbooks_20102011 <- logs_sf_pts %>% 
  filter(season=='2010-2011')
tm <- proc.time()
ODFW_Dcrab_logbooks_20102011_joined <- fuzzy_left_join(
  ODFW_Dcrab_logbooks_20102011, OR_pot_limit_info_v3,
  by = c(
    "Vessel" = "Vessel",
    "SetDate" = "Begindate",
    "SetDate" = "Enddate"
  ),
  match_fun = list(`==`, `>=`, `<=`)
)
proc.time()-tm
write_csv(ODFW_Dcrab_logbooks_20102011_joined,here::here('wdfw', 'data', 'OR', "ODFW-Dcrab-logbooks-compiled_stackcoords_license_2010-2011.csv"))


#2009-2010 season -- 13mins to run with subset permit data 
OR_pot_limit_info_v3 <- OR_pot_limit_info %>% 
  filter(Year >= 2009 & Year < 2011) %>% 
  select(PermitNumber, Vessel, Begindate, Enddate, Potlimit)

ODFW_Dcrab_logbooks_20092010 <- logs_sf_pts %>% 
  filter(season=='2009-2010')
tm <- proc.time()
ODFW_Dcrab_logbooks_20092010_joined <- fuzzy_left_join(
  ODFW_Dcrab_logbooks_20092010, OR_pot_limit_info_v3,
  by = c(
    "Vessel" = "Vessel",
    "SetDate" = "Begindate",
    "SetDate" = "Enddate"
  ),
  match_fun = list(`==`, `>=`, `<=`)
)
proc.time()-tm
write_csv(ODFW_Dcrab_logbooks_20092010_joined,here::here('wdfw', 'data', 'OR', "ODFW-Dcrab-logbooks-compiled_stackcoords_license_2009-2010.csv"))


#2008-2009 season -- 10mins to run with subset permit data 
OR_pot_limit_info_v3 <- OR_pot_limit_info %>% 
  filter(Year >= 2008 & Year < 2010) %>% 
  select(PermitNumber, Vessel, Begindate, Enddate, Potlimit)

ODFW_Dcrab_logbooks_20082009 <- logs_sf_pts %>% 
  filter(season=='2008-2009')
tm <- proc.time()
ODFW_Dcrab_logbooks_20082009_joined <- fuzzy_left_join(
  ODFW_Dcrab_logbooks_20082009, OR_pot_limit_info_v3,
  by = c(
    "Vessel" = "Vessel",
    "SetDate" = "Begindate",
    "SetDate" = "Enddate"
  ),
  match_fun = list(`==`, `>=`, `<=`)
)
proc.time()-tm
write_csv(ODFW_Dcrab_logbooks_20082009_joined,here::here('wdfw', 'data', 'OR', "ODFW-Dcrab-logbooks-compiled_stackcoords_license_2008-2009.csv"))

#2007-2008 season -- 10mins to run with subset permit data 
OR_pot_limit_info_v3 <- OR_pot_limit_info %>% 
  filter(Year >= 2007 & Year < 2009) %>% 
  select(PermitNumber, Vessel, Begindate, Enddate, Potlimit)

ODFW_Dcrab_logbooks_20072008 <- logs_sf_pts %>% 
  filter(season=='2007-2008')
tm <- proc.time()
ODFW_Dcrab_logbooks_20072008_joined <- fuzzy_left_join(
  ODFW_Dcrab_logbooks_20072008, OR_pot_limit_info_v3,
  by = c(
    "Vessel" = "Vessel",
    "SetDate" = "Begindate",
    "SetDate" = "Enddate"
  ),
  match_fun = list(`==`, `>=`, `<=`)
)
proc.time()-tm
write_csv(ODFW_Dcrab_logbooks_20072008_joined,here::here('wdfw', 'data', 'OR', "ODFW-Dcrab-logbooks-compiled_stackcoords_license_2007-2008.csv"))
```
#Read in individual season files with permit info
```{r}
# Identify all csv files in folder, Store all files in list

#error because in 2009-2010 season 'vessel number' column has all numbers (double) while in other seasons it is character
logs_all_raw <- list.files(path = "E:/Leena/Documents/Projects/raimbow/wdfw/data/OR/seasons with permit info", pattern = "*.csv", full.names = TRUE) %>% 
  lapply(read_csv) %>% 
  # Combine data sets into one
  bind_rows                                                        
#glimpse(logs_all_raw)

#Now read in and join 2009-2010
logs20092010 <- read_csv(here::here('wdfw','data','OR','seasons with permit info','2009-2010',  'ODFW-Dcrab-logbooks-compiled_stackcoords_license_2009-2010.csv'),col_types = 'cdcDddddddddddlccdcddcdcDDd')
logs_all_raw <- rbind(logs_all_raw, logs20092010)

logs_all <- logs_all_raw %>% 
  select(-Vessel.y,-Begindate,-Enddate) %>% 
  rename(Vessel=Vessel.x) %>% 
  mutate(season = factor(season)) %>% 
  arrange(season)
glimpse(logs_all)


```
#Investigate and fix NAs with permit data

```{r}
numNA <- logs_all %>% 
  summarise(across(everything(),~sum(is.na(.x))))
# There are 7 instances of NA in Permit Number/Pot Limit for 2013-2018
# There are 81 instances when 2010-2011 and 2013-2018 seasons 
# There are 173 instances when have seasons 2007-2008 to 2010-2011 AND 2013-2014 to 2017-2018
# There are 183 instances when have seasons 2007-2018 

logs_NA <- logs_all %>% 
  filter(is.na(PermitNumber)|is.na(Potlimit))
#2013-2018: All instances of NAs are for vessel ID 549931
# Looked through the Permit Info file, and in all instances the vessel had logbook data few days prior to its Permit being active/valid. The vessel always had the same permit number 96231 and pot limit of 200
#2010-2011: more cases but same story, the Set Date is a few days before the Begin Date of the permit 
#2007-2010: few instance where no vessel number provided in logbook data -- those will have to be deleted
  # cases where Set Date is a few days before the Begin Date of the permit
  # for vessel ID OR438ABD in 2008-09 season: fishing in logs in January 2009, but permit (96386) only start in April 2009, until then the permit 96386 belonged to vessel ID OR377YF - which was owned by the same person as OR438ABD
#in 2011-2012: no NAs, in 2012-2013: few cases where Set Date is a few days before the Begin Date of the permit 

#-----------------------------------------------------
#Fix all NAs

#2013-2018 data only
# logs_all_permits <-  logs_all %>% 
#   mutate(PermitNumber = ifelse(is.na(PermitNumber), 96231, PermitNumber),
#          Potlimit = ifelse(is.na(Potlimit), 200, Potlimit))




#grab records where data is not missing and remove any cases where no Vessel ID was provided
logs_not_NA <- logs_all %>% 
  filter(!is.na(Vessel)) %>% 
  filter(!is.na(PermitNumber)|!is.na(Potlimit))



#For NA cases, remove any cases where no Vessel ID was provided, then input correct PotLimit values for cases where they were missing
logs_NA_fixed <-  logs_NA %>% 
  filter(!is.na(Vessel)) %>% 
  mutate(
    Potlimit = case_when(
      #2020-2011, 2013-2018
      Vessel == 1037811 | Vessel == 554887 | Vessel == 539240 ~ 500,
      Vessel == 254696 | Vessel == 539216 ~ 300,
      Vessel == "OR001UA" | Vessel == 549931 ~ 200,
      #2007-2010
      Vessel == 500657 | Vessel == 544609 ~ 500,
      Vessel == 547465 | Vessel == 226744 | Vessel == 520984 ~ 300,
      Vessel == "OR438ABD" | Vessel == 578558 ~ 200,
      #2012-2013
      Vessel == 976374 ~ 500,
      Vessel == 539216 ~ 300
    )
  )

#Join all back together to a dataset that has PotLimit for all instances
logs_all_permits <- rbind(logs_not_NA, logs_NA_fixed)
```
# Export as csv
```{r}
write_csv(logs_all_permits,here::here('wdfw', 'data', 'OR', "ODFW-Dcrab-logbooks-compiled_stackcoords_license_2013-2018_2021-08-17.csv"))

write_csv(logs_all_permits,here::here('wdfw', 'data', 'OR', "ODFW-Dcrab-logbooks-compiled_stackcoords_license_2007-2018_20210830.csv"))
```

