---
title: "Whale risk time series - county"
author: "Sam Woodman"
date: '`r format(Sys.Date(), "%d %B %Y")`'
output:
    html_document: default
    pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

README: This document is the same as 'Whale_risk_timeseries.Rmd', except using counties for regions and removing the anomaly section.

This document summarizes and plots time series of humpback whale risk of entanglement, as well as the humpback predictions and fishing measure values. It loads values generated in 'Whale_risk.Rmd', and sums them by region. These summed values, as well as their densities, are then plotted. In addition, we plot the monthly ratio (x / x_mean) and z score ((x - x_mean) / x_sd), where the mean and sd are calculated by region/month and using only values from the 2009 - 2013 fishing seasons (Nov 2009 - Oct 2013).

The whale risk file, e.g. 'Whale_risk_long_nona.Rdata', is created in 'Whale_risk.Rmd'.

This analysis currently is done using all CA data, and non-confidential OR and WA fishing data.

```{r, message=FALSE}
library(gridExtra)
library(lubridate)
library(sf)
library(tidyverse)

source(here::here("humpback_risk/funcs_whale_risk_timeseries.R"))
fig.h <- 14
```

Set file paths depending on user. The user should also specifies `flag.save`, i.e. if, when knitting, the plots are saved to a file as well as being displayed in the knitted file.

```{r}
source(here::here("User_script_local.R"))
if (user == "JS") {
  
} else if (user == "SMW") {
  flag.save <- TRUE
  file.landerased.rds <- "C:/SMW/RAIMBOW/raimbow-local/RDATA_files/Grid_5km_landerased.rds"
  file.risk <- "C:/SMW/RAIMBOW/raimbow-local/RDATA_files/Whale_risk_long_nona.Rdata"
  file.grid.region <- "C:/SMW/RAIMBOW/raimbow-local/RDATA_files/Grid_region_county.Rdata"
  
  path.plots <- "C:/SMW/RAIMBOW/raimbow-local/Plots/Whale_risk_timeseries/"
  file.out.timeseries <- "C:/SMW/RAIMBOW/raimbow-local/RDATA_files/Whale_risk_county_timeseries.Rdata"
  
} else {
  stop("User not recognized")
}
```

## Data processing

Load humpback, fishing, and risk absolute values from "2_Whale_risk.Rmd", and the grid cell ID - region key from "2b_Grid_region.R"

```{r}
load(file.risk)
load(file.grid.region)
```

We can visualize regions

```{r,fig.width=5, fig.height=6}
library(rnaturalearth)
rmap.base <- c(
  st_geometry(ne_states(country = "United States of America", returnclass = "sf")), 
  ne_countries(scale = 10, continent = "North America", returnclass = "sf") %>% 
    filter(admin %in% c("Canada", "Mexico")) %>% 
    st_geometry()
)

grid.5km.lno <- readRDS(file.landerased.rds)
grid.region.sf <- grid.region.county %>% 
  mutate(region = factor(region, levels = reg.names.county)) %>% 
  left_join(grid.5km.lno, by = "GRID5KM_ID") %>% 
  st_sf()
plot(grid.region.sf["region"], main = "Region", axes = TRUE, border = NA, 
     xaxt = "n", key.length = 1, key.width = lcm(5), reset = FALSE)
sf::.degAxis(1, at = c(-126, -122, -118))
plot(rmap.base, add = TRUE, col = "tan", border = "black")
graphics::box()
```

We want to sum the values by region and time period. We also calculate the densities, using either the area of cells with non-NA values for 1) that specific month or 2) the whole time series. We remove region CA-S because of the lack of fishing (and thus risk) in that region

```{r}
# Join regions by grid cell IDs
all.df <- all.df %>% 
  left_join(grid.region.county, by = "GRID5KM_ID") %>% 
  filter(!(region %in% c("Ventura", "Los_Angeles", "Orange", "San_Diego")))

# Summarize
all.df.summ <- all.df %>% 
  filter(region != "CA-S") %>% 
  group_by(region, ym) %>% 
  summarise(yr = unique(yr), 
            mon = unique(mon), 
            mn_sum_abund = sum(mn_abund, na.rm = TRUE), 
            vms_sum_pings = sum(vms_pings, na.rm = TRUE), 
            risk_sum_total = sum(risk_total, na.rm = TRUE), 
            area_sum = sum(area_km_lno), 
            area_sum_mon1 = sum(area_km_lno[!is.na(mn_abund)]), 
            area_sum_mon2 = sum(area_km_lno[!is.na(risk_total)])) %>% 
  ungroup() %>% 
  mutate(mn_sum_dens = mn_sum_abund / area_sum, 
         vms_sum_dens = vms_sum_pings / area_sum, 
         risk_sum_dens = risk_sum_total / area_sum, 
         mn_sum_dens_mon = mn_sum_abund / area_sum_mon1, 
         vms_sum_dens_mon = vms_sum_pings / area_sum_mon2, 
         risk_sum_dens_mon = risk_sum_total / area_sum_mon2) %>% 
  replace_na(list(vms_sum_dens_mon = 0, risk_sum_dens_mon = 0))

if (flag.save) save(all.df, all.df.summ, file = file.out.timeseries)

all.df.summ %>% 
  filter(!duplicated(region)) %>% 
  select(region, area_sum)
```

## Plot

Once we have done all ofthe above data processing, we can generate plots. First, total sum plots:

```{r, fig.height=fig.h, fig.width=10}
p1 <- raimbow_ggplot_region(
  all.df.summ, risk_sum_total, plot.main = "Risk", 
  y.lab = "Whales * VMS pings"
)
p2 <- raimbow_ggplot_region(
  all.df.summ, mn_sum_abund, plot.main = "Humpback whales", 
  y.lab = "Whales"
)
p3 <- raimbow_ggplot_region(
  all.df.summ, vms_sum_pings, plot.main = "Fishing", 
  y.lab = "VMS pings"
)

p123 <- grid.arrange(p1, p2, p3, nrow = 3)

if (flag.save) 
  ggsave(filename = "County_linear_humpback_risk_regional.png", p123, path = path.plots, 
         height = fig.h, width = 10)
```

Next we plot the risk, humpback, and fishing densities, calculated using the area of cells with non-NA values for the whole time series

```{r, fig.height=fig.h, fig.width=10}
p1 <- raimbow_ggplot_region(
  all.df.summ, risk_sum_dens, plot.main = "Risk", 
  y.lab = "Whales * VMS pings / km2"
)
p2 <- raimbow_ggplot_region(
  all.df.summ, mn_sum_dens, plot.main = "Humpback whales", 
  y.lab = "Whales / km2"
)
p3 <- raimbow_ggplot_region(
  all.df.summ, vms_sum_dens, plot.main = "Fishing", 
  y.lab = "VMS pings / km2"
)

p123 <- grid.arrange(p1, p2, p3, nrow = 3)

if (flag.save)
  ggsave(filename = "County_linear_humpback_risk_regional_areanorm.png", p123, path = path.plots, 
         height = fig.h, width = 10)
```

Finally, we plot the risk, humpback, and fishing densities, calculated using the area of cells with non-NA values for that specific month

```{r, fig.height=fig.h, fig.width=10}
p1 <- raimbow_ggplot_region(
  all.df.summ, risk_sum_dens_mon, plot.main = "Risk", 
  y.lab = "Whales * VMS pings / km2"
)
p2 <- raimbow_ggplot_region(
  all.df.summ, mn_sum_dens_mon, plot.main = "Humpback whales", 
  y.lab = "Whales / km2"
)
p3 <- raimbow_ggplot_region(
  all.df.summ, vms_sum_dens_mon, plot.main = "Fishing", 
  y.lab = "VMS pings / km2"
)

p123 <- grid.arrange(p1, p2, p3, nrow = 3)

if (flag.save) 
  ggsave(filename = "County_linear_humpback_risk_regional_monthlydens.png", p123, path = path.plots, 
         height = fig.h, width = 10)
```
