---
title: "Whale risk time series - county"
author: "Sam Woodman"
date: '`r format(Sys.Date(), "%d %B %Y")`'
output:
    html_document: default
    pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (Sys.info()["nodename"] == "SWC-SWOODMAN-L") knitr::opts_knit$set(root.dir = "../")
```

## Introduction

README: This document is the same as 'Whale_risk_timeseries.Rmd', except using counties for regions and removing the anomaly section.

This document summarizes and plots time series of humpback whale risk of entanglement, as well as the humpback predictions and fishing measure values. It loads values generated in 'Whale_risk.Rmd', and converts the humpback and risk densities to absolute values (abundance and abundance * VMS measure, respectively) in order to sum values by region. These summed values, as well as their densities, are then plotted. In addition, we plot the monthly ratio (x / x_mean) and z score ((x - x_mean) / x_sd), where the mean and sd are calculated by region/month and using only values from the 2009 - 2013 fishing seasons (Nov 2009 - Oct 2013).

The whale risk file, e.g. 'Whale_risk.Rdata', is created in 'Whale_risk.Rmd' from the [raimbow-whaleRisk GitHub repo](https://github.com/smwoodman/raimbow-whaleRisk). 

This analysis currently is done using all CA data, and non-confidential OR and WA fishing data.

```{r, message=FALSE}
library(gridExtra)
library(lubridate)
library(sf)
library(tidyverse)

source("humpback_risk/funcs_whale_risk_timeseries.R")
fig.h <- 14
```

Set file paths depending on user. The user should also specifies `save.flag`, i.e. if, when knitting, the plots are saved to a file as well as being displayed in the knitted file.

```{r}
source("User_script_local.R")
if (user == "JS") {
  save.flag <- FALSE
  file.landerased <- ""
  file.risk <- ""
  
  path.plots <- ""
  path.rdata <- ""
  
} else if (user == "SMW") {
  save.flag <- FALSE
  file.landerased <- "../raimbow-local/RDATA_files/Grid_5km_landerased.RDATA"
  file.risk <- "../raimbow-local/RDATA_files/Whale_risk.Rdata"
  
  path.plots <- "../raimbow-local/Plots/Whale_risk_timeseries/"
  path.rdata <- "../raimbow-local/RDATA_files/"
  
  
} else {
  stop("User not recognized")
}
```

## Prep

Load humpback, fishing, and risk values, and convert humpback and risk densities to total values. 

```{r}
load(file.risk) # Created in Whale_risk.Rmd; Mn and risk values are densities

mult.func <- function(x, y) {x * y}
grid.area <- humpback.all.dens %>% select(GRID5KM_ID, area_km_lno)

humpback.abund <- humpback.all.dens %>% 
  mutate_at(vars(starts_with("Mn_")), mult.func, y = grid.area$area_km_lno)
# fish.all.num is already non-density values
risk.total <- risk.all.dens %>% 
  mutate_at(vars(starts_with("Mn_DC_risk_")), mult.func, y = grid.area$area_km_lno)
```

Next, we remove the rows with all NA risk values before plotting

```{r}
### Monthly
risk.nona.any <- apply(select(risk.total, starts_with("Mn_DC_risk")), 1, function(i) any(!is.na(i))) 
fish.nona.any <- apply(select(fish.all.num, starts_with("DC_")), 1, function(i) any(!is.na(i)))

# Fishing data has smaller spatial footprint, and thus fish NAs should match risk NAs
#   However, it shouldn't be a show-stopper if this isn't true
identical(risk.nona.any, fish.nona.any)

# Will do all plotting with nona data frames for ease of computation, 
#   but mostly for summing data only across cells that have risk
risk.nona <- risk.total[risk.nona.any, ]
humpback.nona <- humpback.abund[risk.nona.any, ]
fish.nona <- fish.all.num[risk.nona.any, ]
rm(risk.all.dens, risk.total, humpback.abund, humpback.all.dens, fish.all.num)
```

## Data processing

The first step is to make the whale, VMS, and risk data frames 'long', meaning there is one row for each grid cell for each time period

```{r}
h.df <- humpback.nona %>% 
  gather(col_name, mn_abund, 
         -GRID5KM_ID, -LONGITUDE, -LATITUDE, -area_km_lno) %>% 
  mutate(yr = as.numeric(substr(col_name, 4, 7)), 
         mon = as.numeric(substr(col_name, 9, 10)), 
         ym = paste(yr, sprintf("%02d", mon), sep = "-")) %>% 
  select(GRID5KM_ID, LONGITUDE, LATITUDE, area_km_lno, ym, yr, mon, mn_abund)

f.df <- fish.nona %>% 
  gather(col_name, vms_pings, 
         -GRID5KM_ID, -LONGITUDE, -LATITUDE, -area_km_lno) %>% 
  mutate(yr = as.numeric(substr(col_name, 4, 7)), 
         mon = as.numeric(substr(col_name, 9, 10))) %>% 
  select(-col_name)

r.df <- risk.nona %>% 
  gather(col_name, risk_total, 
         -GRID5KM_ID, -LONGITUDE, -LATITUDE, -area_km_lno) %>% 
  mutate(yr = as.numeric(substr(col_name, 12, 15)), 
         mon = as.numeric(substr(col_name, 17, 18))) %>% 
  select(-col_name)

# Sanity check
col.same <- c("GRID5KM_ID", "LONGITUDE", "LATITUDE", "area_km_lno", "yr", "mon")
stopifnot(
  all.equal(h.df[, col.same], f.df[, col.same]), 
  all.equal(h.df[, col.same], r.df[, col.same])
)

# 
all.df <- h.df %>% 
  left_join(f.df, by = col.same) %>% 
  left_join(r.df, by = col.same)
rm(humpback.nona, fish.nona, risk.nona)
```

Next, we determine the region of each grid cell. In this document, region is county-level for CA counties!

```{r}
# From Entanglement_gridID.Rmd
reg.bound <- c(
  42, 41.5, 40, 38.83, 38.3, 
  37.8, 37.7, 37.1, 36.83, 35.8, 
  35, 34.375, 34, 33.75, 33.35, 
  32.5
)

reg.list <- list(
  "Del_Norte"       = c(reg.bound[1], reg.bound[2]), 
  "Humboldt"        = c(reg.bound[2], reg.bound[3]), 
  "Mendocino"       = c(reg.bound[3], reg.bound[4]), 
  "Sonoma"          = c(reg.bound[4], reg.bound[5]), 
  "Marin"           = c(reg.bound[5], reg.bound[6]), 
  "San_Francisco"   = c(reg.bound[6], reg.bound[7]), 
  "San_Mateo"       = c(reg.bound[7], reg.bound[8]), 
  "Santa_Cruz"      = c(reg.bound[8], reg.bound[9]), 
  "Monterey"        = c(reg.bound[9], reg.bound[10]), 
  "San_Luis_Obispo" = c(reg.bound[10], reg.bound[11]), 
  "Santa_Barbara"   = c(reg.bound[11], reg.bound[12]), 
  "Ventura"         = c(reg.bound[12], reg.bound[13]), 
  "Los_Angeles"     = c(reg.bound[13], reg.bound[14]), 
  "Orange"          = c(reg.bound[14], reg.bound[15]), 
  "San_Diego"       = c(reg.bound[15], reg.bound[16])
)

reg.names <- names(reg.list)

# Add in WA and OR
reg.bound <- rev(c(50, 46.25, reg.bound))
reg.names <- rev(c("WA", "OR", reg.names))
stopifnot(length(reg.bound) == length(reg.names) + 1)

grid.region <- all.df %>%
  filter(!duplicated(GRID5KM_ID)) %>%
  arrange(GRID5KM_ID) %>%
  mutate(region.idx = findInterval(LATITUDE, reg.bound, left.open = TRUE),
         region = factor(reg.names[region.idx], levels = rev(reg.names))) %>%
  select(GRID5KM_ID, region)
```

```{r}
# reg.bound <- c(32.5, 34.4, 36.3, 38.76683, 42, 46.25, 50)
# reg.names <- c("CA-S", "CA-SCen", "CA-Cen", "CA-N", "OR", "WA")
# stopifnot(length(reg.names) == length(reg.bound) - 1)
# 
# grid.region <- all.df %>%
#   filter(!duplicated(GRID5KM_ID)) %>%
#   arrange(GRID5KM_ID) %>%
#   mutate(region.idx = findInterval(LATITUDE, reg.bound, left.open = TRUE),
#          region = factor(reg.names[region.idx], levels = rev(reg.names))) %>%
#   select(GRID5KM_ID, region)
```

We can visualize regions

```{r,fig.width=4, fig.height=6}
library(rnaturalearth)
rmap.base <- c(
  st_geometry(ne_states(country = "United States of America", returnclass = "sf")), 
  ne_countries(scale = 10, continent = "North America", returnclass = "sf") %>% 
    filter(admin %in% c("Canada", "Mexico")) %>% 
    st_geometry()
)

load(file.landerased)
grid.region.sf <- grid.region %>% 
  mutate(region = factor(region, levels = reg.names)) %>% 
  left_join(grid.5km.lno, by = "GRID5KM_ID") %>% 
  st_sf()
plot(grid.region.sf["region"], main = "Region", axes = TRUE, border = NA, 
     xaxt = "n", key.length = 1, key.width = lcm(3.2), reset = FALSE)
sf::.degAxis(1, at = c(-126, -122, -118))
plot(rmap.base, add = TRUE, col = "tan", border = "black")
```

We want to sum the values by region and time period. We also calculate the densities, using either the area of cells with non-NA values for 1) that specific month or 2) the whole time series. We remove region CA-S because of the lack of fishing (and thus risk) in that region

```{r}
all.df <- all.df %>% 
  left_join(grid.region, by = "GRID5KM_ID") %>% 
  filter(!(region %in% c("Ventura", "Los_Angeles", "Orange", "San_Diego")))

all.df.summ <- all.df %>% 
  filter(region != "CA-S") %>% 
  group_by(region, ym) %>% 
  summarise(yr = unique(yr), 
            mon = unique(mon), 
            mn_sum_abund = sum(mn_abund, na.rm = TRUE), 
            vms_sum_pings = sum(vms_pings, na.rm = TRUE), 
            risk_sum_total = sum(risk_total, na.rm = TRUE), 
            area_sum = sum(area_km_lno), 
            area_sum_mon1 = sum(area_km_lno[!is.na(mn_abund)]), 
            area_sum_mon2 = sum(area_km_lno[!is.na(risk_total)])) %>% 
  ungroup() %>% 
  mutate(mn_sum_dens = mn_sum_abund / area_sum, 
         vms_sum_dens = vms_sum_pings / area_sum, 
         risk_sum_dens = risk_sum_total / area_sum, 
         mn_sum_dens_mon = mn_sum_abund / area_sum_mon1, 
         vms_sum_dens_mon = vms_sum_pings / area_sum_mon2, 
         risk_sum_dens_mon = risk_sum_total / area_sum_mon2) %>% 
  replace_na(list(vms_sum_dens_mon = 0, risk_sum_dens_mon = 0))

# if (save.flag) save(all.df, all.df.summ, file = paste0(path.rdata, "Whale_risk_county_timeseries.Rdata"))

all.df.summ %>% 
  filter(!duplicated(region)) %>% 
  select(region, area_sum)
```

## Plot

Once we have done all ofthe above data processing, we can generate plots. First, total sum plots:

```{r, fig.height=fig.h, fig.width=10}
p1 <- raimbow_ggplot_region(
  all.df.summ, risk_sum_total, plot.main = "Risk", 
  y.lab = "Whales * VMS pings"
)
p2 <- raimbow_ggplot_region(
  all.df.summ, mn_sum_abund, plot.main = "Humpback whales", 
  y.lab = "Whales"
)
p3 <- raimbow_ggplot_region(
  all.df.summ, vms_sum_pings, plot.main = "Fishing", 
  y.lab = "VMS pings"
)

p123 <- grid.arrange(p1, p2, p3, nrow = 3)

if (save.flag) 
  ggsave(filename = "Linear_humpback_risk_regional.png", p123, path = path.plots, 
         height = 10, width = 10)
```

Next we plot the risk, humpback, and fishing densities, calculated using the area of cells with non-NA values for the whole time series

```{r, fig.height=fig.h, fig.width=10}
p1 <- raimbow_ggplot_region(
  all.df.summ, risk_sum_dens, plot.main = "Risk", 
  y.lab = "Whales * VMS pings / km2"
)
p2 <- raimbow_ggplot_region(
  all.df.summ, mn_sum_dens, plot.main = "Humpback whales", 
  y.lab = "Whales / km2"
)
p3 <- raimbow_ggplot_region(
  all.df.summ, vms_sum_dens, plot.main = "Fishing", 
  y.lab = "VMS pings / km2"
)

p123 <- grid.arrange(p1, p2, p3, nrow = 3)

if (save.flag)
  ggsave(filename = "Linear_humpback_risk_regional_areanorm.png", p123, path = path.plots, 
         height = 10, width = 10)
```

Finally, we plot the risk, humpback, and fishing densities, calculated using the area of cells with non-NA values for that specific month

```{r, fig.height=fig.h, fig.width=10}
p1 <- raimbow_ggplot_region(
  all.df.summ, risk_sum_dens_mon, plot.main = "Risk", 
  y.lab = "Whales * VMS pings / km2"
)
p2 <- raimbow_ggplot_region(
  all.df.summ, mn_sum_dens_mon, plot.main = "Humpback whales", 
  y.lab = "Whales / km2"
)
p3 <- raimbow_ggplot_region(
  all.df.summ, vms_sum_dens_mon, plot.main = "Fishing", 
  y.lab = "VMS pings / km2"
)

p123 <- grid.arrange(p1, p2, p3, nrow = 3)

if (save.flag) 
  ggsave(filename = "Linear_humpback_risk_regional_monthlydens.png", p123, path = path.plots, 
         height = 10, width = 10)
```
